### 2022-12-20 ms1_data_analysis.Rmd
Add log ratio processing scripts. Not integrated yet with the markdown page.

---
title: "Data Analysis"
author: "Piotr Szefer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

## General effects of biotic factors

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)

source('code/log_ratio_extraction.R')
source('code/log_ratio_tests.R')

# Overdispersion test for the density in the fungicide treatment
lrdata <- DENSITY.LR
lvT <- leveneTest(lratio ~ site, data = lrdata %>% filter(treatment == "f"))
lvtpval <- lvT$`Pr(>F)`[1]

```

### 2022-12-20 code/git_setup.R

@@ -1,5 +1,5 @@
 # PAT 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
-
+# usethis::create_github_token()
 # gitcreds::gitcreds_set()
 
 # https://happygitwithr.com/https-pat.html

### 2022-12-20 code/log_ratio_extraction.R

@@ -1,0 +1,147 @@
+# rm(list = ls())
+library(dplyr)
+
+# 1. Read main data-set ----
+# elev.dat <- read.csv("data/elevation_vegetation_data.csv")
+source("code/data_processing.R")
+
+ # 2. Filter data ----
+
+# 2.1 All plants withouth Musa ----
+# ed.nomusa <- BIODATA %>%
+#   filter(!grepl("musa", 
+#                unified_names, 
+#                ignore.case = T))
+
+# 2.1 Only woody plants ----
+# ed.woody <- BIODATA %>%
+#   filter(life.form == "woody")
+
+# 3. For each garden calculate log ratios ----
+
+# Makes a copy of the original dataset
+# dataset <- ed.nomusa
+
+# Calcualtions of wood
+# dataset <- ed.woody
+
+# dataset$treatment <- as.factor(dataset$treatment)
+
+# gdat <- dataset %>%
+#   group_by(site, garden, treatment) %>%
+#   summarise(bio = sum(tot_bio))
+
+# Treatments used for the altitudinal comparison
+# low herbivory and chlorkill are not included as the former was 
+# used only in lowest elevation (Wanang) and the latter in NUmba and Yawan
+
+# alt.comp.treat <- c("p","i","f","h","ch")
+
+# lrdata <- data.frame()
+
+# for (gard in unique(dataset$garden)){
+#   
+#   print(gard)
+#   
+#   # Dataset only for control plot
+#   based <- gdat %>%
+#     filter(garden == gard, treatment == "c")
+#   
+#   # Dataset for each treatment
+#   gdatf <- gdat %>%
+#     filter(garden == gard, treatment %in% alt.comp.treat) %>% #
+#     mutate(cval = based$bio) %>%
+#     
+#     # log ratio is Factor(present)/Factor(absent)
+#     # Thus in the case of excluding factors (i.e. insecticide, exclosure)
+#     # and fungicide) we would divide control value over the treatment value.
+#     # In case of herbivore addition we would divide treatment 
+#     # (addition of herbivores) over the control.
+#     
+#     mutate(lratio = ifelse(treatment %in% c("h"), 
+#                            log(bio/cval), 
+#                            log(cval/bio)))
+# 
+#   lrdata <- rbind(lrdata, gdatf)
+#   
+#   }
+
+# Change the order of site factors in the lrdata
+# lrdata$site <- factor(lrdata$site, levels = c("wanang", "numba","yawan"))
+
[…]

### 2022-12-20 code/log_ratio_tests.R

@@ -1,0 +1,395 @@
+# This file is obsolete. Delete and use only for the test.
+
+# rm(list = ls())
+# source("code/log_ratio_extraction.R")
+
+library(ggplot2)
+library(emmeans)
+library(ggpubr)
+library(nlme)
+library(car)
+library(multcompView)
+
+
+# Statistical test
+
+# 1) Compare relative strengths of individual factors at each site
+
+# 1.1) On raw biomass data glmm with garden as a random factor (in different script!)
+
+# 2) strength changing with elevation
+lrdata <- BIOMASS.LR
+treatment_code <- "f"
+
+calculateTest <- function(lrdata, treatment_code = "f"){
+  # Specific wrapper around testing procedure
+
+  # Filter data
+  trtdata  <- lrdata %>%
+    filter(treatment == treatment_code)
+
+  ltest.p <- leveneTest(lratio ~ site, data =  trtdata)
+  btest.p <- bartlett.test(lratio ~ site, data =  trtdata)
+  print(ltest.p)
+  print(btest.p)
+  
+  # If the varianvces are unequal then perform GLS
+  if(ltest.p$`Pr(>F)`[1] < 0.05){
+    
+    trtlm <- gls(lratio ~ 0 + site, data = trtdata,
+                 weights = varIdent(form = ~1|site))
+    summary(trtlm)
+    
+    print("Model runned")
+    
+    # Tukey post-hoc
+    emm <- emmeans::emmeans(object = trtlm, 
+                   specs = "site", 
+                   adjust = "wald")
+    
+    print("Post hoc runned")
+    
+    zero.dif <- ifelse(summary(trtlm)$tTable[,4] > 0.05, "","*")
+    print("zero.dif assigned")
+    
+    mod_means <- multcomp::cld(object = emm,
+                               Letters = letters)$.group
+    
+    paste(mod_means,zero.dif, sep = "") -> res
+    
+  }else{
+    trtlm <- lm(lratio ~ 0 + site, data = trtdata)
+    
+    # Tukey post-hoc
+    emm <- emmeans(object = trtlm, 
+                   specs = "site", 
+                   adjust = "wald")
+    
+    zero.dif <- ifelse(summary(trtlm)$coefficients[,4] > 0.05, "","*")
+    mod_means <- multcomp::cld(object = emm,
[…]

### 2022-12-20 code/log_ratio_extraction.R
Log ratio extraction working.

@@ -1,143 +1,10 @@
-# rm(list = ls())
-library(dplyr)
-
 # 1. Read main data-set ----
-# elev.dat <- read.csv("data/elevation_vegetation_data.csv")
-source("code/data_processing.R")
+source("code/prepare_data.R")
 
- # 2. Filter data ----
-
-# 2.1 All plants withouth Musa ----
-# ed.nomusa <- BIODATA %>%
-#   filter(!grepl("musa", 
-#                unified_names, 
-#                ignore.case = T))
-
-# 2.1 Only woody plants ----
-# ed.woody <- BIODATA %>%
-#   filter(life.form == "woody")
-
-# 3. For each garden calculate log ratios ----
-
-# Makes a copy of the original dataset
-# dataset <- ed.nomusa
-
-# Calcualtions of wood
-# dataset <- ed.woody
-
-# dataset$treatment <- as.factor(dataset$treatment)
-
-# gdat <- dataset %>%
-#   group_by(site, garden, treatment) %>%
-#   summarise(bio = sum(tot_bio))
-
-# Treatments used for the altitudinal comparison
-# low herbivory and chlorkill are not included as the former was 
-# used only in lowest elevation (Wanang) and the latter in NUmba and Yawan
-
-# alt.comp.treat <- c("p","i","f","h","ch")
-
-# lrdata <- data.frame()
-
-# for (gard in unique(dataset$garden)){
-#   
-#   print(gard)
-#   
-#   # Dataset only for control plot
-#   based <- gdat %>%
-#     filter(garden == gard, treatment == "c")
-#   
-#   # Dataset for each treatment
-#   gdatf <- gdat %>%
-#     filter(garden == gard, treatment %in% alt.comp.treat) %>% #
-#     mutate(cval = based$bio) %>%
-#     
-#     # log ratio is Factor(present)/Factor(absent)
-#     # Thus in the case of excluding factors (i.e. insecticide, exclosure)
-#     # and fungicide) we would divide control value over the treatment value.
-#     # In case of herbivore addition we would divide treatment 
-#     # (addition of herbivores) over the control.
-#     
-#     mutate(lratio = ifelse(treatment %in% c("h"), 
-#                            log(bio/cval), 
-#                            log(cval/bio)))
-# 
-#   lrdata <- rbind(lrdata, gdatf)
-#   
-#   }
-
-# Change the order of site factors in the lrdata
[…]

### 2022-12-20 code/log_ratio_tests.R

@@ -1,4 +1,6 @@
-# This file is obsolete. Delete and use only for the test.
+# This file is obsolete. Use only for the test
+
+# This may be used to produce labels for the log-ratio comparison figure.
 
 # rm(list = ls())
 # source("code/log_ratio_extraction.R")
@@ -21,68 +23,9 @@ library(multcompView)
 lrdata <- BIOMASS.LR
 treatment_code <- "f"
 
-calculateTest <- function(lrdata, treatment_code = "f"){
-  # Specific wrapper around testing procedure
-
-  # Filter data
-  trtdata  <- lrdata %>%
-    filter(treatment == treatment_code)
-
-  ltest.p <- leveneTest(lratio ~ site, data =  trtdata)
-  btest.p <- bartlett.test(lratio ~ site, data =  trtdata)
-  print(ltest.p)
-  print(btest.p)
-  
-  # If the varianvces are unequal then perform GLS
-  if(ltest.p$`Pr(>F)`[1] < 0.05){
-    
-    trtlm <- gls(lratio ~ 0 + site, data = trtdata,
-                 weights = varIdent(form = ~1|site))
-    summary(trtlm)
-    
-    print("Model runned")
-    
-    # Tukey post-hoc
-    emm <- emmeans::emmeans(object = trtlm, 
-                   specs = "site", 
-                   adjust = "wald")
-    
-    print("Post hoc runned")
-    
-    zero.dif <- ifelse(summary(trtlm)$tTable[,4] > 0.05, "","*")
-    print("zero.dif assigned")
-    
-    mod_means <- multcomp::cld(object = emm,
-                               Letters = letters)$.group
-    
-    paste(mod_means,zero.dif, sep = "") -> res
-    
-  }else{
-    trtlm <- lm(lratio ~ 0 + site, data = trtdata)
-    
-    # Tukey post-hoc
-    emm <- emmeans(object = trtlm, 
-                   specs = "site", 
-                   adjust = "wald")
-    
-    zero.dif <- ifelse(summary(trtlm)$coefficients[,4] > 0.05, "","*")
-    mod_means <- multcomp::cld(object = emm,
-                               Letters = letters)$.group
-    
-    paste(mod_means,zero.dif, sep = "") -> res
-  }
-  
[…]

### 2022-12-20 code/prepare_data.R

@@ -6,7 +6,7 @@
 library(dplyr)
 
 # Custom made function to create site x species matrix for ordination
-source("code/contingencyTable.R")
+source("code/specific_functions.R")
 
 # Read the data
 vegetation.data <- read.csv('data/main.data.w.traits.csv')
@@ -21,29 +21,26 @@ data <- vegetation.data %>%
          treatment %in% c("c","h","i","p","f")) %>%
   mutate(plotid = paste(site, garden, treatment, sep = "_")) # create an original 
 
-# Check the data dimension
-summary(data)
-dim(data)
-summary(data$tot_bio)
-
-# Prepare data for the LRR plot (Figure 1 in the MS)
-# 
+# Prepare tables for the LRR plot (Figure 1 in the MS)
 DENSITY <- data %>%
-  # filter((unified_names %in% wpnames) & !is.na(dbh)) %>%
-  group_by(site, garden, treatment, unified_names)
-
+  mutate(garden = paste(substr(site,1,1),garden, sep=""))%>%
+  group_by(site,garden,treatment)%>%
+  summarise(tot.stem.no = sum(no_stems))
 
 BIOMASS <- data %>%
+  mutate(garden = paste(substr(site,1,1),garden, sep=""))%>%
   group_by(site,garden,treatment)%>%
   summarise(tot_bio = sum(tot_bio))
 
 RICHNESS <- data %>%
+  mutate(garden = paste(substr(site,1,1),garden, sep=""))%>%
   group_by(site, garden, treatment) %>%
   summarise(sp_no = length(unique(unified_names)),
             simpson = vegan::diversity(tot_bio, index = "invsimpson"))
 
 COMMDATA <- contingencyTable2(data,
                               "unified_names", "plotid", "tot_bio")
+
 COMMVARS <- data %>%
   group_by(site,garden, treatment, plotid) %>%
   summarise()

### 2022-12-20 code/specific_functions.R

@@ -1,0 +1,149 @@
+# Define functions specific for the data that is used for this study
+
+contingencyTable2 <- function(dataset, ROW, COL, VALUE, rm.null=TRUE){
+  
+  # Takes a data frame and creates a matrix with specified
+  # row labels and column labels populated with a sum of specified VALUE
+  # Used to summarise 
+  
+  # Get rid of the empty factors
+  dataset[, colnames(dataset) == ROW] <- as.character(dataset[, colnames(dataset) == ROW])
+  dataset[, colnames(dataset) == COL] <- as.character(dataset[, colnames(dataset) == COL])
+  # Make a table, get rid of the empty rows and columns
+  plants <- table(dataset[, colnames(dataset) == ROW], dataset[, colnames(dataset) == COL])
+  if(rm.null){
+    plants <- plants[rowSums(plants) != 0, colSums(plants) != 0]
+  }
+  
+  # See where to insert values
+  allSpecCodes <- colnames(plants)
+  allPlotCodes <- rownames(plants)
+  entries <- which(plants != 0, arr.ind = TRUE)
+  # Loop through the entries and insert values
+  for (entry in 1:dim(entries)[1]){
+    plot <- entries[entry,1]
+    plant <- entries[entry,2]
+    specCode <- allSpecCodes[plant]
+    plotCode <- allPlotCodes[plot]
+    res <- dataset[dataset[,ROW] == plotCode & dataset[,COL] == specCode,VALUE]
+    plants[plot,plant] <- sum(res, na.rm = TRUE)
+  }
+  plants[is.na(plants)] <- 0
+  
+  # Change the table to a matrix or data.frame
+  mat_a <- matrix(0, nrow = dim(plants)[1], ncol = dim(plants)[2])
+  colnames(mat_a) <- colnames(plants)
+  rownames(mat_a) <- rownames(plants)
+  for (row in 1:dim(plants)[1]){
+    for (col in 1:dim(plants)[2])
+      mat_a[row,col] <- plants[row,col]
+  }
+  
+  #return(plants)
+  return(mat_a)
+}
+
+getLogRatio <- function (dataset,
+                         effect,
+                         alt.comp.treat = c("p","i","f","h")){
+  
+  # What this function does?...
+  
+  lrdata <- data.frame()
+  
+  dset <- dataset
+  
+  names(dset)[grepl(effect, names(dset))] <- "value"
+  
+  for (gard in unique(dset$garden)){
+    
+    print(gard)
[…]

### 2022-12-21 code/log_ratio_extraction.R
Extraction and test code intrroduced into the log_ratio_extraction.R script

@@ -1,6 +1,13 @@
 # 1. Read main data-set ----
 source("code/prepare_data.R")
 
+library(emmeans)
+# library(ggpubr)
+library(nlme)
+library(car)
+# library(multcompView)
+
+
 BIOMASS.LR <- getLogRatio(BIOMASS, 
                            effect = "tot.bio")
 
@@ -14,3 +21,48 @@ RICHNESS.LR <- getLogRatio(RICHNESS,
                             effect = "sp_no")
 
 # 2. Perform statistical set and output labels for the figure
+
+# Biomasss
+f.test <- runTest(BIOMASS.LR, treatment_code = "f", "biomass")
+# ch.test <- runTest(BIOMASS.LR, treatment_code = "ch")
+i.test <- runTest(BIOMASS.LR, treatment_code = "i","biomass")
+p.test <- runTest(BIOMASS.LR, treatment_code = "p","biomass")
+h.test <- runTest(BIOMASS.LR, treatment_code = "h","biomass")
+biotest_labels <- c(f.test[[1]], i.test[[1]], p.test[[1]], h.test[[1]])
+biotest_table <- rbind(f.test[[2]], i.test[[2]], p.test[[2]], h.test[[2]])
+
+# Diversity
+f.test <- runTest(DIVERSITY.LR, treatment_code = "f", "diversity")
+# ch.test <- runTest(DIVERSITY.LR, treatment_code = "ch")
+i.test <- runTest(DIVERSITY.LR, treatment_code = "i","diversity")
+p.test <- runTest(DIVERSITY.LR, treatment_code = "p","diversity")
+h.test <- runTest(DIVERSITY.LR, treatment_code = "h","diversity")
+divtest_labels <- c(f.test[[1]], i.test[[1]], p.test[[1]], h.test[[1]])
+divtest_table <- rbind(f.test[[2]], i.test[[2]], p.test[[2]], h.test[[2]])
+
+# Richness
+f.test <- runTest(RICHNESS.LR, treatment_code = "f", "richness")
+# ch.test <- runTest(RICHNESS.LR, treatment_code = "ch")
+i.test <- runTest(RICHNESS.LR, treatment_code = "i","richness")
+p.test <- runTest(RICHNESS.LR, treatment_code = "p", "richness")
+h.test <- runTest(RICHNESS.LR, treatment_code = "h", "richness")
+richtest_labels <- c(f.test[[1]], i.test[[1]], p.test[[1]], h.test[[1]])
[…]

### 2022-12-21 code/log_ratio_tests.R

@@ -3,7 +3,7 @@
 # This may be used to produce labels for the log-ratio comparison figure.
 
 # rm(list = ls())
-# source("code/log_ratio_extraction.R")
+source("code/log_ratio_extraction.R")
 
 library(ggplot2)
 library(emmeans)
@@ -15,44 +15,19 @@ library(multcompView)
 
 # Statistical test
 
-# 1) Compare relative strengths of individual factors at each site
+# lrdata <- BIOMASS.LR
+# treatment_code <- "f"
+# 
 
-# 1.1) On raw biomass data glmm with garden as a random factor (in different script!)
-
-# 2) strength changing with elevation
-lrdata <- BIOMASS.LR
-treatment_code <- "f"
-
-
-
-# runTest <- function(lrdata, treatment_code = "f"){
-#   # Specific wrapper around testing procedure
-#   
-#   # Filter data
-#   trtdata  <- lrdata %>%
-#     filter(treatment == treatment_code)
-#   
-#   # Calculate model
-#   # No intercept model
-#   trtlmw <- lm(lratio ~ 1, data = trtdata %>% filter(site == "wanang"))
-#   trtlmn <- lm(lratio ~ 1, data = trtdata %>% filter(site == "numba"))
-#   trtlmy <- lm(lratio ~ 1, data = trtdata %>% filter(site == "yawan"))
-#   
-#   # Tukey post-hoc
-#   # emm <- emmeans(trtlm, "site", correction = "tukey")
-#   
-#   return(c())
-#   
-# }
 
 # 2.1) Tests for individual treatments
 
 # Biomasss
-f.test <- runTest(BIOMASS.LR, treatment_code = "f")
+f.test <- runTest(BIOMASS.LR, treatment_code = "f", "biomass")
 # ch.test <- runTest(BIOMASS.LR, treatment_code = "ch")
-i.test <- runTest(BIOMASS.LR, treatment_code = "i")
-p.test <- runTest(BIOMASS.LR, treatment_code = "p")
-h.test <- runTest(BIOMASS.LR, treatment_code = "h")
+i.test <- runTest(BIOMASS.LR, treatment_code = "i","biomass")
+p.test <- runTest(BIOMASS.LR, treatment_code = "p","biomass")
+h.test <- runTest(BIOMASS.LR, treatment_code = "h","biomass")
 
 
 

### 2022-12-21 code/prepare_data.R

@@ -5,6 +5,7 @@
 
 library(dplyr)
 
+
 # Custom made function to create site x species matrix for ordination
 source("code/specific_functions.R")
 

### 2022-12-21 code/specific_functions.R

@@ -88,7 +88,7 @@ getLogRatio <- function (dataset,
   
 }
 
-runTest <- function(lrdata, treatment_code = "f"){
+runTest <- function(lrdata, treatment_code = "f", label){
   # Specific wrapper around testing procedure
   
   # Filter data
@@ -97,19 +97,23 @@ runTest <- function(lrdata, treatment_code = "f"){
   
   ltest.p <- leveneTest(lratio ~ site, data =  trtdata)
   btest.p <- bartlett.test(lratio ~ site, data =  trtdata)
-  print(ltest.p)
-  print(btest.p)
   
   # If the varianvces are unequal then perform GLS
   if(ltest.p$`Pr(>F)`[1] < 0.05){
-    
+    variance <- 'non-homogenous'
     trtlm <- gls(lratio ~ 0 + site, data = trtdata,
                  weights = varIdent(form = ~1|site))
     summary(trtlm)
     
+    aovtrt <- anova(trtlm)
+    testStat <- aovtrt$`F-value`
+    testP <- aovtrt$`p-value`
+    
     print("Model runned")
     
     # Tukey post-hoc
+    
+    print("Stage1")
     emm <- emmeans::emmeans(object = trtlm, 
                             specs = "site", 
                             adjust = "wald")
@@ -125,7 +129,12 @@ runTest <- function(lrdata, treatment_code = "f"){
     paste(mod_means,zero.dif, sep = "") -> res
     
   }else{
+    variance <- 'homogenous'
     trtlm <- lm(lratio ~ 0 + site, data = trtdata)
+    
+    aovtrt <- anova(trtlm)
+    testStat <- aovtrt$`F value`
+    testP <- aovtrt$`Pr(>F)`
     
     # Tukey post-hoc
     emm <- emmeans(object = trtlm, 
@@ -139,11 +148,20 @@ runTest <- function(lrdata, treatment_code = "f"){
     paste(mod_means,zero.dif, sep = "") -> res
   }
   
+  
+  
+  test.df <- data.frame(testType = c("Levene Test", "Bartlett Test", paste(variance, 'anova')),
+                        statName = c("F-value", "Bartlett's K-squared", "F-value"),
+                        testStat = c(ltest.p$`F value`[1], as.numeric(btest.p$statistic),testStat[1]),
+                        pVal = c(ltest.p$`Pr(>F)`[1], btest.p$p.value[1], testP[1]),
+                        treatmentCode = treatment_code,
+                        caselabel = label)
+  
[…]

### 2022-12-22 ms1_data_analysis.Rmd
Panel 1 works fine.

@@
 ```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
-library(dplyr)
-library(ggplot2)
-
 source('code/log_ratio_extraction.R')
-source('code/log_ratio_tests.R')
 
@@
 
+sjPlot::tab_df(lrr_test_df)
+
 ```
+
+Description...
+
+```{r}
+library(ggplot2)
+# library(GGally)
+# library(lme4)
+# library(lmerTest)
+# library(ggpmisc)
+# library(tibble)
+
+# bio, dens, div, rich
+panel_data <- data.frame()
+
+dats <- list(biomass = BIOMASS.LR, 
+             denssity = DENSITY.LR, 
+             diversity = DIVERSITY.LR, 
+             richness = RICHNESS.LR)
+
+# Stack the data and rename descriptor names
+for (file in names(dats)){
+  print(file)
+  dat <- dats[[file]]
+  dat$descriptor <- ifelse(grepl("bio", file), "Biomass",
+                           ifelse(grepl("dens", file), "Woody plant density",
+                                  ifelse(grepl("div", file),"Diversity", "Richness")))
+  dat <- dat[, c("site", "garden", "treatment", "lratio","descriptor")]
+  panel_data <- rbind(dat, panel_data)
+}
+
+# Make a renamed factor variable for the treatments
+
+panel_data$ftreatment <- panel_data$treatment
+
+# Fix treatment names
+panel_data$ftreatment[panel_data$ftreatment == "i"] <- "Insects"
+panel_data$ftreatment[panel_data$ftreatment == "p"] <- "Birds, bats and ants"
+panel_data$ftreatment[panel_data$ftreatment == "f"] <- "Fungi"
+panel_data$ftreatment[panel_data$ftreatment == "h"] <- "Increased herbivory"
+panel_data$ftreatment <- factor(panel_data$ftreatment,
+                           levels = c("Fungi",
+                                      "Insects",
+                                      "Birds, bats and ants",
+                                      "Increased herbivory"
+                                      ))
+
+# Set the transparency value
+aph <- 0.5
+
+# Create a factor variable for sites
+panel_data$fsite <- factor(panel_data$site, labels = c("Wanang",
+                                      "Numba",
+                                      "Yawan"))
+
+# Display panel plot
+ggplot(panel_data, aes(x = fsite, y = lratio))+
+  geom_jitter(width = 0.1, alpha = 0.1, size = 2)+
+
+  # Group means
+  stat_summary(fun = mean, na.rm = T,
+               geom = "point", size = 2,
+               alpha = aph)+
+
+  stat_summary(fun.data = "mean_cl_boot",
+               alpha = aph) +
+
+  # Annotations
+  geom_text(data = panel_data %>%
+              group_by(ftreatment, fsite) %>%
+              summarise(mlr = mean(lratio)),
+            aes(x = fsite, y = mlr + 1.5, group = ftreatment,
+                label = lrr_panel_labs),
+            size = 3)+
+
+  theme_bw()+
+  theme(strip.text = element_text(size =5.5),
+        axis.text.x = element_text(angle = 45,
+                                   hjust = 1))+
+  geom_hline(yintercept = 0, lty = 2)+
+  ylab("Log response ratio")+xlab("")+
+  facet_grid(descriptor ~ ftreatment)
+
+```

### 2022-12-22 code/log_ratio_extraction.R

@@ -2,9 +2,11 @@
 source("code/prepare_data.R")
 
 library(emmeans)
-# library(ggpubr)
+library(lsmeans)
 library(nlme)
 library(car)
+
+# library(ggpubr)
 # library(multcompView)
 
 
@@ -24,35 +26,32 @@ RICHNESS.LR <- getLogRatio(RICHNESS,
 
 # Biomasss
 f.test <- runTest(BIOMASS.LR, treatment_code = "f", "biomass")
-# ch.test <- runTest(BIOMASS.LR, treatment_code = "ch")
-i.test <- runTest(BIOMASS.LR, treatment_code = "i","biomass")
-p.test <- runTest(BIOMASS.LR, treatment_code = "p","biomass")
-h.test <- runTest(BIOMASS.LR, treatment_code = "h","biomass")
+i.test <- runTest(BIOMASS.LR, treatment_code = "i", "biomass")
+p.test <- runTest(BIOMASS.LR, treatment_code = "p", "biomass")
+h.test <- runTest(BIOMASS.LR, treatment_code = "h", "biomass")
 biotest_labels <- c(f.test[[1]], i.test[[1]], p.test[[1]], h.test[[1]])
 biotest_table <- rbind(f.test[[2]], i.test[[2]], p.test[[2]], h.test[[2]])
 
 # Diversity
 f.test <- runTest(DIVERSITY.LR, treatment_code = "f", "diversity")
-# ch.test <- runTest(DIVERSITY.LR, treatment_code = "ch")
-i.test <- runTest(DIVERSITY.LR, treatment_code = "i","diversity")
-p.test <- runTest(DIVERSITY.LR, treatment_code = "p","diversity")
-h.test <- runTest(DIVERSITY.LR, treatment_code = "h","diversity")
+i.test <- runTest(DIVERSITY.LR, treatment_code = "i", "diversity")
+p.test <- runTest(DIVERSITY.LR, treatment_code = "p", "diversity")
+h.test <- runTest(DIVERSITY.LR, treatment_code = "h", "diversity")
 divtest_labels <- c(f.test[[1]], i.test[[1]], p.test[[1]], h.test[[1]])
 divtest_table <- rbind(f.test[[2]], i.test[[2]], p.test[[2]], h.test[[2]])
 
 # Richness
 f.test <- runTest(RICHNESS.LR, treatment_code = "f", "richness")
-# ch.test <- runTest(RICHNESS.LR, treatment_code = "ch")
[…]

### 2022-12-22 code/log_ratio_tests.R

@@ -1,313 +1,0 @@
-# This file is obsolete. Use only for the test
-
-# This may be used to produce labels for the log-ratio comparison figure.
-
-# rm(list = ls())
-source("code/log_ratio_extraction.R")
-
-library(ggplot2)
-library(emmeans)
-library(ggpubr)
-library(nlme)
-library(car)
-library(multcompView)
-
-
-# Statistical test
-
-# lrdata <- BIOMASS.LR
-# treatment_code <- "f"
-# 
-
-
-# 2.1) Tests for individual treatments
-
-# Biomasss
-f.test <- runTest(BIOMASS.LR, treatment_code = "f", "biomass")
-# ch.test <- runTest(BIOMASS.LR, treatment_code = "ch")
-i.test <- runTest(BIOMASS.LR, treatment_code = "i","biomass")
-p.test <- runTest(BIOMASS.LR, treatment_code = "p","biomass")
-h.test <- runTest(BIOMASS.LR, treatment_code = "h","biomass")
-
-
-
-biotest_labels <- c(f.test, i.test, p.test, h.test)
-# biotest_labels <- c("","","","a*","b*","a","a*","a" ,"","","","a*","b","b")
-
-# Diversity
-f.test <- runTest(DIVERSITY.LR, treatment_code = "f")
-# ch.test <- runTest(DIVERSITY.LR, treatment_code = "ch")
-i.test <- runTest(DIVERSITY.LR, treatment_code = "i")
-p.test <- runTest(DIVERSITY.LR, treatment_code = "p")
-h.test <- runTest(DIVERSITY.LR, treatment_code = "h")
-
-
-divtest_labels <- c(f.test, i.test, p.test, h.test)
-# c("","","","","","a","a","a*" ,"a","a","a*","","","")
-
-# Richness
-f.test <- runTest(RICHNESS.LR, treatment_code = "f")
-# ch.test <- runTest(RICHNESS.LR, treatment_code = "ch")
-i.test <- runTest(RICHNESS.LR, treatment_code = "i")
-p.test <- runTest(RICHNESS.LR, treatment_code = "p")
-h.test <- runTest(RICHNESS.LR, treatment_code = "h")
-
-richtest_labels <- c(f.test, i.test, p.test, h.test)
-
-
-# Density
-# f.test <- runTest(DENSITY.LR, treatment_code = "f") # unequal variances
-f.test <- c(" a*", " a",  " a" )
[…]

### 2022-12-22 code/lrr_panel_figure.R

@@ -1,0 +1,256 @@
+library(ggplot2)
+library(GGally)
+# library(lme4)
+# library(lmerTest)
+# library(ggpmisc)
+# library(tibble)
+
+
+source('code/log_ratio_extraction.R')
+
+# 0. Data preparation ----
+
+# Data comes from the log_ratio_extraction.R file
+
+# bio, dens, div, rich
+panel_data <- data.frame()
+
+dats <- list(biomass = BIOMASS.LR, 
+             denssity = DENSITY.LR, 
+             diversity = DIVERSITY.LR, 
+             richness = RICHNESS.LR)
+
+# Stack the data and rename descriptor names
+
+for (file in names(dats)){
+  print(file)
+  dat <- dats[[file]]
+  dat$descriptor <- ifelse(grepl("bio", file), "Biomass",
+                           ifelse(grepl("dens", file), "Woody plant density",
+                                  ifelse(grepl("div", file),"Diversity", "Richness")))
+  dat <- dat[, c("site", "garden", "treatment", "lratio","descriptor")]
+  panel_data <- rbind(dat, panel_data)
+}
+
+# Remove chlorkill treatment
+
+pd <- panel_data %>% 
+  filter(treatment != "ch")
+
+# Make a renamed factor variable for the treatments
+
+pd$ftreatment <- pd$treatment
+
+# Fix treatment names
+pd$ftreatment[pd$ftreatment == "i"] <- "Insects"
+pd$ftreatment[pd$ftreatment == "p"] <- "Birds, bats and ants"
+pd$ftreatment[pd$ftreatment == "f"] <- "Fungi"
+pd$ftreatment[pd$ftreatment == "h"] <- "Increased herbivory"
+
+pd$ftreatment <- factor(pd$ftreatment,
+                           levels = c("Fungi",
+                                      "Insects",
+                                      "Birds, bats and ants",
+                                      "Increased herbivory"
+                                      ))
+
+# Add elevation variable
+pd$elev <- ifelse(pd$site == "wanang", 200, 
+                  ifelse(pd$site == "numba", 750, 1900))
+
+pd$elev2 <- pd$elev^2
+
+# Set the transparency value
+aph <- 0.5
+
+# Create a factor variable for sites
+pd$fsite <- factor(pd$site, labels = c("Wanang",
+                                      "Numba",
[…]

### 2022-12-22 code/specific_functions.R

@@ -88,6 +88,8 @@ getLogRatio <- function (dataset,
   
 }
 
+# lrdata <- DENSITY.LR
+# treatment_code = "p"
 runTest <- function(lrdata, treatment_code = "f", label){
   # Specific wrapper around testing procedure
   
@@ -101,9 +103,8 @@ runTest <- function(lrdata, treatment_code = "f", label){
   # If the varianvces are unequal then perform GLS
   if(ltest.p$`Pr(>F)`[1] < 0.05){
     variance <- 'non-homogenous'
-    trtlm <- gls(lratio ~ 0 + site, data = trtdata,
+    trtlm <- nlme::gls(lratio ~ 0 + site, data = trtdata,
                  weights = varIdent(form = ~1|site))
-    summary(trtlm)
     
     aovtrt <- anova(trtlm)
     testStat <- aovtrt$`F-value`
@@ -114,19 +115,19 @@ runTest <- function(lrdata, treatment_code = "f", label){
     # Tukey post-hoc
     
     print("Stage1")
-    emm <- emmeans::emmeans(object = trtlm, 
-                            specs = "site", 
-                            adjust = "wald")
+    # emm <- emmeans(object = trtlm, 
+    #                         specs = "site",
+    #                adjust = "mvt")
     
-    print("Post hoc runned")
+    print("Post hoc not runned")
     
     zero.dif <- ifelse(summary(trtlm)$tTable[,4] > 0.05, "","*")
     print("zero.dif assigned")
     
-    mod_means <- multcomp::cld(object = emm,
-                               Letters = letters)$.group
+    # mod_means <- multcomp::cld(object = emm,
+    #                            Letters = letters)$.group
     
-    paste(mod_means,zero.dif, sep = "") -> res
+    paste("a",zero.dif, sep = "") -> res
     
   }else{
     variance <- 'homogenous'
@@ -165,3 +166,4 @@ runTest <- function(lrdata, treatment_code = "f", label){
   return(list(res, test.df))
   
 }
+

### 2023-01-05 ms1_data_analysis.Rmd
added figure comaparing raw descriptors between sites

@@
 
-## General effects of biotic factors
+## Experimental design
 
 ```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
+# The source file loads the data preparation script, where the raw data is read and summarised for further analyses
 source('code/log_ratio_extraction.R')
 
-# Overdispersion test for the density in the fungicide treatment
+# Over-dispersion test for the density in the fungicide treatment
 lrdata <- DENSITY.LR
@@
 lvtpval <- lvT$`Pr(>F)`[1]
+```
 
-sjPlot::tab_df(lrr_test_df)
+We collected data on biomass (total of `r round(sum(BIOMASS$tot_bio), 0)` kg), density, and plant traits: area lost to herbivores, water content, leaf dry matter content and specific leaf area for `r length(unique(data$unified_names))` woody plant species from three tropical forest sites on elevations spanning `r 1900 - 200` m from 200 to 1900 m a.s.l.
+
+Figure 1. Meadian, 1st, second quartiles, 1.5 IRQ distanceT and outliers in biomass, richness diversity and density at each location/elevation, but some Yawan sites were extremely productive.
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
+
+library(ggplot2)
+library(dplyr)
+
+elev.bind <- rbind(BIOMASS %>% dplyr::select(site, garden, treatment, tot_bio), 
+                   RICHNESS %>% dplyr::select(site, garden, treatment, sp_no), 
+                   RICHNESS %>% dplyr::select(site, garden, treatment, simpson), 
+                   DENSITY %>% dplyr::select(site, garden, treatment, tot.stem.no))
+
+findoutlier <- function(x) {
+  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
+}
+
+elevdata <- elev.bind %>% 
+  mutate(value = ifelse(!is.na(tot_bio), tot_bio, 
+                        ifelse(!is.na(sp_no), sp_no, 
+                               ifelse(!is.na(simpson), simpson, tot.stem.no)))) %>%
+  mutate(descriptor = ifelse(!is.na(tot_bio), "Total biomass [kg]", 
+                        ifelse(!is.na(sp_no), "Number of species", 
+                               ifelse(!is.na(simpson), "Simpson's diversity", "Abundance")))) %>%
+  dplyr::select(site, garden, treatment, value, descriptor) %>%
+  mutate(fsite = factor(site, levels = c("wanang", "numba","yawan"))) %>%
+  arrange(descriptor, fsite, treatment, garden)
+
+elevdataT <- elevdata %>% 
+  mutate(valueTrans = ifelse(descriptor == "Simpson's diversity", 
+                             value, log(value)))
+
+outliers <- c()
+for (desc in unique(elevdataT$descriptor)) {
+  for (st in unique(elevdataT$site)) {
+    for (trt in unique(elevdataT$treatment)) {
+
+      print(paste(desc,st,trt))
+
+      outliers <- c(outliers,findoutlier(pull(elevdataT %>% filter(descriptor == desc,
+                          site == st,
+                          treatment == trt), valueTrans)))
+
+    }
+
+  }
+}
+elevdataT$outlierBool <- outliers 
+elevdataT <- elevdataT %>% mutate(outlier = ifelse(outlierBool, garden, ""))
+
+ggplot(elevdataT, aes(y = valueTrans, x = fsite, label = garden))+
+  geom_boxplot()+
+  geom_text(aes(label=outlier), hjust = 1.2, vjust = -0.0005, na.rm=TRUE)+
+  facet_grid(vars(descriptor), vars(treatment), scales = "free")+
+  theme_bw()
 
@@
 
-Description...
+Looking at the control plots Numba (the mid-elevation in our study) has the highest abundance and number of species.
+Predictably productivity decreases with elevation.
+Control plots in Numba are more variable than in other two elevations.
+Yawan has one site with lowest abundance (garden number seven).However no significant differences in variance were detected (Levene's and Bartellet's tests)
 
-```{r}
+## General effects of biotic factors
+
+We calculated log-response ratios for plant biomass, species richness, species diversity and stem density to be able to compare the effect sizes at individual elevations.
+
+Table 1. Test results for over-dispersion all descriptors at three studied elevations.
+
+```{r, echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
+sjPlot::tab_df(lrr_test_df)
+```
+
+There is generally little evidence for the over-dispersion in the LRR values at individual elevations.
+Only woody stem density for the predator exclusion showed some evidence of over-dispersion.
+However, results of a GLS model accounting for over-dispersion (using 'varIdent(form = ~1|site)' as weights in the GLS function) were very similar to those obtained from a simple LM model.
+
+```{r,  echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
+# Results of a GLS model with over-dispersion vs LM model for density LRR in the predator exclusion treatment.
+
+trtdata <-  trtdata  <- lrdata %>%
+    filter(treatment == 'p')
+
+trtlm <- nlme::gls(lratio ~ 0 + site, data = trtdata,
+                 weights = varIdent(form = ~1|site))
+
+    aovtrt <- anova(trtlm)
+    testStat <- aovtrt$`F-value`
+    testP <- aovtrt$`p-value`
+
+    print("Model runned")
+
+    # Tukey post-hoc
+    print("Stage1")
+    emm <- emmeans(object = trtlm,
+                            specs = "site",
+                   adjust = "mvt")
+
+    print("Post hoc not runned")
+
+    zero.dif <- ifelse(summary(trtlm)$tTable[,4] > 0.05, "","*")
+    print("zero.dif assigned")
+
+    # mod_means <- multcomp::cld(object = emm,
+    #                            Letters = letters)$.group
+
+    paste("a",zero.dif, sep = "") -> res
+
+```
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
 library(ggplot2)

### 2023-01-05 ms1_data_analysis.Rmd
'impact of treatments' section ready

@@
 # Over-dispersion test for the density in the fungicide treatment
-lrdata <- DENSITY.LR
-lvT <- leveneTest(lratio ~ site, data = lrdata %>% filter(treatment == "f"))
-lvtpval <- lvT$`Pr(>F)`[1]
+# lrdata <- DENSITY.LR
+# lvT <- leveneTest(lratio ~ site, data = lrdata %>% filter(treatment == "f"))
+# lvtpval <- lvT$`Pr(>F)`[1]
 ```
@@
 We collected data on biomass (total of `r round(sum(BIOMASS$tot_bio), 0)` kg), density, and plant traits: area lost to herbivores, water content, leaf dry matter content and specific leaf area for `r length(unique(data$unified_names))` woody plant species from three tropical forest sites on elevations spanning `r 1900 - 200` m from 200 to 1900 m a.s.l.
+Low elevation is Wanang (200 m a.s.l.), Mid-elevation is Numba (700 m a.s.l.), and High elevation is Yawan (1900 m a.s.l.).
+We refer to the sites as Low, Mid- and High elevation throughout the text.
 
-Figure 1. Meadian, 1st, second quartiles, 1.5 IRQ distanceT and outliers in biomass, richness diversity and density at each location/elevation, but some Yawan sites were extremely productive.
+Figure 1. Median, 1st, second quadrilles, 1.5 IRQ distance and outliers in biomass, richness diversity and density at each location/elevation, but some Yawan sites were extremely productive.
 
@@
 
-There is generally little evidence for the over-dispersion in the LRR values at individual elevations.
-Only woody stem density for the predator exclusion showed some evidence of over-dispersion.
-However, results of a GLS model accounting for over-dispersion (using 'varIdent(form = ~1|site)' as weights in the GLS function) were very similar to those obtained from a simple LM model.
+```{r,  echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
 
-```{r,  echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
 # Results of a GLS model with over-dispersion vs LM model for density LRR in the predator exclusion treatment.
-
-trtdata <-  trtdata  <- lrdata %>%
+trtdata <-  trtdata  <- DENSITY.LR %>%
     filter(treatment == 'p')
 
-trtlm <- nlme::gls(lratio ~ 0 + site, data = trtdata,
+trtlm <- lm(lratio ~ 0 + site, data = trtdata)
+
+trtgls <- nlme::gls(lratio ~ 0 + site, data = trtdata,
                  weights = varIdent(form = ~1|site))
 
+summary(trtlm)
+summary(trtgls)
+
 aovtrt <- anova(trtlm)
+
 testStat <- aovtrt$`F-value`
@@
 
-    print("Model runned")
-
 # Tukey post-hoc
-    print("Stage1")
 emm <- emmeans(object = trtlm,
@@
 
-    print("Post hoc not runned")
+mod_means <- multcomp::cld(object = emm,
+                           Letters = letters)$.group
 
-    zero.dif <- ifelse(summary(trtlm)$tTable[,4] > 0.05, "","*")
-    print("zero.dif assigned")
+zero.dif.lm <- ifelse(summary(trtlm)$coefficients[,4] > 0.05, "","*")
+zero.dif.gls <- ifelse(summary(trtgls)$tTable[,4] > 0.05, "","*")
 
-    # mod_means <- multcomp::cld(object = emm,
-    #                            Letters = letters)$.group
+lmres <- as_tibble(summary(trtlm)$coefficients)
+lmres <- rename(lmres, Estimate = 'Estimate', 
+               `Standard Error` = `Std. Error`, 
+               `T-value` =`t value`,
+               `P-value` = `Pr(>|t|)`)
 
-    paste("a",zero.dif, sep = "") -> res
+glsres <- as_tibble(summary(trtgls)$tTable)
+glsres <- rename(glsres,
+                 Estimate = 'Value', 
+                 `Standard Error` = "Std.Error", 
+                 `T-value` = "t-value", 
+                 `P-value` = 'p-value')
+
+model_comp <- rbind(lmres, glsres)
+model_comp$Site <- rep(c("Wanang", "Numba", "Yawan"), 2)
+model_comp$Model <- rep(c("LM", "GLS"), each = 3)
 
@@
 
-```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
+Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation.
+
+```{r}
+sjPlot::tab_df(model_comp)
+```
+
+There is generally little evidence for the over-dispersion in the LRR values at individual elevations.
+Only woody stem density for the predator exclusion showed some evidence of over-dispersion.
+Including this in statistical modle (using 'varIdent(form = ~1|site)' as weights in the GLS function) showed significant difference of LRR in Yawan, that was not observed for LM model.
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 1.
+Magnitude and direction of fungi, insects and their predators on biomass, richness, diversity and density of woody plants along elevation gradient: Wanang (200 m a.s.l.); Numba (750 m a.s.l.); Yawan (1900 m a.s.l.).
+Log response ratio is a log-ratio of a given descriptor value from plots where a biotic factor was present to where it was absent.
+Grey points represent empirical values.
+Letters indicate statistically significant ($\\alpha$ = 0.05) differences for pairwise comparisons between elevations within a treatment-descriptor combination with Tukey correction for multiplicity.
+For better readability letters are given only in cases where any of the mean were either different form each other or different from zero.
+Zero effect are indicated with a dashed line.
+Stars in the upper index indicate significant ($\\alpha$ = 0.05) differences of the effect mean from zero (significance of the effect at a given site)", fig.width= 8, fig.height=8}
+
 library(ggplot2)
-# library(GGally)
-# library(lme4)
-# library(lmerTest)
-# library(ggpmisc)
-# library(tibble)
 
@@
 ```
+
+## Impact of treatments
+
+Positive effect of pathogenic fungi on richness was found at low and high elevations (Fig. 1) and on density at low elevation.
+
+Insects reduced biomass only at the mid-elevation, increased diversity at high and richness at high and low elevation.
+
+```{r}
+
+# Check whether the RICHNESS.LR for predators is correct.
+# trtdata <-  trtdata  <- RICHNESS.LR %>%
+#     filter(treatment == 'p')
+
+# trtlm <- lm(lratio ~ 0 + site, data = trtdata)
+
+# summary(trtlm)
+
+# Tukey post-hoc
+# emm <- emmeans(object = trtlm,
+#                specs = "site",
+#                adjust = "mvt")
+
+# mod_means <- multcomp::cld(object = emm,
+#                            Letters = letters)$.group
+
+```
+At high elevation predator removal caused an increase in diversity, richness and density of woody plants.
+
+Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.
+
+## LRR correlations
+
+In case of species richness only effect of insect and herbivore addition and removal had strong significant correlation with each other.
+Herbivore addition and removal were negatively correlated as expected.
+If there is a compensatory effect there should be no correlation (because when we remove one factor the other takes over and equalize the effect).
+Positive correlation rules out compensation.
+All positive correlations suggest bottom-up control.

### 2023-01-05 code/specific_functions.R

@@ -149,8 +149,6 @@ runTest <- function(lrdata, treatment_code = "f", label){
     paste(mod_means,zero.dif, sep = "") -> res
   }
   
-  
-  
   test.df <- data.frame(testType = c("Levene Test", "Bartlett Test", paste(variance, 'anova')),
                         statName = c("F-value", "Bartlett's K-squared", "F-value"),
                         testStat = c(ltest.p$`F value`[1], as.numeric(btest.p$statistic),testStat[1]),

### 2023-01-18 ms1_data_analysis.Rmd
Adding food chain model with Rmd

@@
 
-```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure S1 Characteistics"}
 
@@
 
-Positive effect of pathogenic fungi on richness was found at low and high elevations (Fig. 1) and on density at low elevation.
-
+Positive effect of pathogenic fungi on richness was found at low and high elevations (Fig.
+1) and on density at low elevation.
 Insects reduced biomass only at the mid-elevation, increased diversity at high and richness at high and low elevation.
+At high elevation predator removal caused an increase in diversity, richness and density of woody plants.Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.
 
@@
 ```
-At high elevation predator removal caused an increase in diversity, richness and density of woody plants.
-
-Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.
-
 ## LRR correlations
 
-In case of species richness only effect of insect and herbivore addition and removal had strong significant correlation with each other.
+Differences of LRR averages from zero may suggest that the effects are random, generally weak and noisy.
+We now look at the LRR correlation for different treatments to probe some possible mechanisms of the biotic factors' action.
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 2. ...", fig.width= 10, fig.height= 10}
+
+# library(dplyr)
+# library(ggplot2)
+# library(GGally)
+# library(lme4)
+# library(lmerTest)
+# library(ggpmisc)
+# library(tibble)
+
+# LRR correlation plot
+
+ftreats <- unique(panel_data$ftreatment)
+
+fdat <- panel_data %>%
+  filter(ftreatment == ftreats[1])%>%
+  arrange(descriptor, site, garden)
+
+hdat <- panel_data %>%
+  filter(ftreatment == ftreats[2])%>%
+  arrange(descriptor, site, garden)
+
+idat <- panel_data %>%
+  filter(ftreatment == ftreats[3])%>%
+  arrange(descriptor, site, garden)
+
+pdat <- panel_data %>%
+  filter(ftreatment == ftreats[4])%>%
+  arrange(descriptor, site, garden)
+
+# chdat <- panel_data %>%
+#   filter(ftreatment == ftreats[5])%>%
+#   arrange(descriptor, site, garden)
+
+# Create one dataset
+pdpairs <- panel_data %>%
+  group_by(descriptor, site, garden) %>%
+  summarise() # Summarise based on what??????
+
+# Add columns with LRRs
+pdpairs$f <- fdat$lratio
+pdpairs$h <- hdat$lratio
+pdpairs$i <- idat$lratio
+pdpairs$p <- pdat$lratio
+
+biolrr <- pdpairs %>% filter(descriptor == "Biomass")
+biocont <- BIOMASS  %>% filter(treatment == "c") %>%
+  arrange(as.character(site), garden)
+
+cpdata <- data.frame(descriptor = rep(pdpairs$descriptor,2),
+                     site = rep(pdpairs$site,2),
+                     garden = rep(pdpairs$garden),
+                     x = c(pdpairs$h,
+                           pdpairs$p),
+                     y = rep(pdpairs$i, 2),
+                     response = rep(c("Elevated herbivory effect",
+                                      "Top predators effect"), 
+                                    each = length(pdpairs$descriptor)))
+
+lrr.cor.test.data <- c()
+
+for (resp in unique(cpdata$response))   {
+  for (desc in unique(cpdata$descriptor)){
+    for (loc in unique(cpdata$site))     {
+
+      print(paste(resp, desc,loc))
+
+      lrrtd <- cpdata %>%
+        filter(response == resp,
+               descriptor == desc,
+               site == loc)
+
+      # np.test.p <- cor.test(lrrtd$x, lrrtd$y, method = c('pearson'))$p.value
+      # Test for noramlity
+      print(shapiro.test(lrrtd$x))
+      print(shapiro.test(lrrtd$y))
+
+      lm.test.p <- summary(lm(lrrtd$y~lrrtd$x, data = lrrtd))$coefficients[2,4]
+
+      print(lm.test.p)
+
+      lrr.res <- data.frame(response = resp,
+                            descriptor = desc,
+                            site = loc,
+                            non.param.p = ifelse(lm.test.p<0.05,"signif","ns"))
+      lrr.cor.test.data <- rbind(lrr.cor.test.data, lrr.res)
+
+    }
+  }
+}
+
+lrr.cor.test.data -> lctd
+
+cpdata$sign <- ""
+
+sign.dat <- lctd[lctd$non.param.p == "signif",]
+
+for(row in 1:dim(sign.dat)[1]){
+  resp <- sign.dat[row, ]$response
+  desc <- sign.dat[row, ]$descriptor
+  loc <- sign.dat[row, ]$site
+
+  c1 <- cpdata$response == resp
+  c2 <- cpdata$descriptor == desc
+  c3 <- cpdata$site == loc
+
+  cpdata[c1 & c2 & c3, ]$sign <- "signif"
+}
+
+ggplot(cpdata, aes(x = x, y = y)) +
+  geom_point(size = 2, alpha = 0.4)+
+  stat_smooth(data = cpdata %>%
+                filter(sign == "signif"),
+              method ="lm")+
+  geom_hline(yintercept = 0)+
+  geom_vline(xintercept = 0)+
+  ylab("Effect size of insects")+
+  xlab("Effect size")+
+  theme_bw()+
+  facet_grid(response*descriptor~site)
+
+```
+
+#### OBSERVATIONS
+
+Correlation between LRR of insect effects were not very common and were observed in `r 9/(3*8) * 100` % of cases.
+Correlations were more often found in Yawan (five out of eight studied correlations) than in Numba (four out of eight) and no significant correlation was found in Wanang.
+If we also include correlations between predator exclusion (P) and P+A this ratio gets significantly higher for Yawan (9 out of 12 correlations).
+
+All observed correlations of the top-predators effects were positively correlated with insect effects, whereas all effects of herbivore addition were negatively correlated with the insect effects.
+
+#### EXPLANATION
+
+If a plant community has a potential to be strongly negatively affected by insects (negative LRR of insects, strong top-down control), then the P+A would result in weaker or positive effects of additional generalist herbivores.
+Top down control seem to be stronger in the highlands and we see more frequent correlation of the LRR's.
+(Model where Intra-guild predators are being strongly limited by top predators.)
+
+What exactly happened after addition can however only be speculated here.
+In case of a  strong bottom-up control should not result in any sort of correlation like in Wanang.
+What I think happened was that spiders could be more active in the exclosures which resulted in their effects being similar to that of a simple exclusion.
+
+A few aspects of insect effects on plants can be important here: 
+If community is saturated with herbivores (Konno) then addition may have negative results.
+Herbivore resource limitation type (relative or absolute). Feeding mode change, adaptive trade off, risk avoidance. All these can only be speculated.
+
+Generalists are worse in consuming any individual plant biomass, therefore they may have positive effects on plants, when natural insect communities have negative effects.
+It is possible that addition of herbivores cause som disruption in the functioning of the insect community effect on vegetation.
+There were no correlation between factors’ effects on vegetation at the low elevation. Herbivore addition and removal were not correlated as expected by the strong bottom-up control.
+
+Simulation showed that it cannot be simply an effect of the effect-abundance relationship between abundance of insects and their effect (quadratic and linear function result in positive relationship).
+
+Top-down effects of predators and their positive correlation with insect effects can be explained well by the bottom-up control of insects by plants.
+It was also assumed in the above simulations.
+Empirically we do not observe any correlation between insect abundance and strength of the effect (lrr_simualtions.R at the end).
+There is also no correlation between richness of a community and its LRR for any of the treatments (lrr_simulations.R).
+
+#### SOME DISCLAIMERS
+
+It is noteworthy that plots where herbivores were added were also covered with nets to prevent increased predation, wich is known to be correlated with increased leaf damage (Katka).
+Therefore it cannot be interpreted directly as a simple increase in herbivory pressure even though this is what should be happening assuming simple chain model and top-down control by top-predators.
+We do not know the faith of the added predators. It is most likely that they did not cause a permanent increase in the species abundance (graph here). However, herbivore damage in general was...
+
+### Predator exclusion and predator exclusion with insect addition
+
+The hypothesis that I have stated above, about increased activity of spiders in exclosures is here being investigated further.
+I want to see whether the addition of herbivores has a positive correlation with the regular exclosures.
+Positive correlation would suggest that the herbivore addition had little to no effect, and that spiders responding to a lack of predators would control the herbivore population even with boosted abundances.
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 2.
+Log-Response Ratio correlations for insects and fungi effects on bimoass, diversity adn density of woody plants"}
+
+fidata <- data.frame(descriptor = pdpairs$descriptor,
+                     site = pdpairs$site,
+                     garden = rep(pdpairs$garden),
+                     x = pdpairs$p,
+                     y = pdpairs$h)
+
+lrr.cor.test.data <- c()
+
+# for (resp in unique(fidata$response))   {
+  for (desc in unique(fidata$descriptor)){
+    for (loc in unique(fidata$site))     {
+
+      print(paste(desc,loc))
+
+      lrrtd <- fidata %>%
+        filter(descriptor == desc,
+               site == loc)
+
+      # Test for normality
+      print(shapiro.test(lrrtd$x))
+      print(shapiro.test(lrrtd$y))
+
+      lm.test.p <- summary(lm(lrrtd$y~lrrtd$x, data = lrrtd))$coefficients[2,4]
+
+      print(lm.test.p)
+
+      lrr.res <- data.frame(descriptor = desc,
+                            site = loc,
+                            non.param.p = ifelse(lm.test.p<0.05,"signif","ns"))
+      lrr.cor.test.data <- rbind(lrr.cor.test.data, lrr.res)
+
+    }
+  }
+# }
+
+lrr.cor.test.data -> lctd
+
+fidata$sign <- ""
+
+sign.dat <- lctd[lctd$non.param.p == "signif",]
+
+for(row in 1:dim(sign.dat)[1]){
+  # resp <- sign.dat[row, ]$response
+  desc <- sign.dat[row, ]$descriptor
+  loc <- sign.dat[row, ]$site
+
+  # c1 <- cpdata$response == resp
+  c2 <- fidata$descriptor == desc
+  c3 <- fidata$site == loc
+
+  fidata[c2 & c3, ]$sign <- "signif"
+}
+
+ggplot(fidata, aes(x = x, y = y)) +
+  geom_point(size = 2, alpha = 0.4)+
+  stat_smooth(data = fidata %>%
+                filter(sign == "signif"),
+              method ="lm")+
+  geom_hline(yintercept = 0)+
+  geom_vline(xintercept = 0)+
+  xlab("Predator removal")+
+  ylab("Predator removal and insect addition")+
+  theme_bw()+
+  facet_grid(descriptor~site)
+
+```
+From the figure we see that there is usually no correlation in Wanang and Numba but it is always present in Yawan.
+Moreover it is not positive but negative.
+This means that when we observer negative effect of predator removal on biomass we can expect to have positive effect of insect addition to the exclosure.
+In cases where the effect of predators is positive, we would get a negative effect of insect addition.
+
+### Fungi vs insect correlation
+No correlation should be observed in case of compensatory effect
+In case of species richness only effects of insect, herbivore addition and herbivore removal were correlated.
 Herbivore addition and removal were negatively correlated as expected.
@@
 Positive correlation rules out compensation.
-All positive correlations suggest bottom-up control.
+A positive correlations would suggest bottom-up control.
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 2.
+Log-Response Ratio correlations for insects and fungi effects on bimoass, diversity adn density of woody plants"}
+
+fidata <- data.frame(descriptor = pdpairs$descriptor,
+                     site = pdpairs$site,
+                     garden = rep(pdpairs$garden),
+                     x = pdpairs$i,
+                     y = pdpairs$f)
+
+lrr.cor.test.data <- c()
+
+# for (resp in unique(fidata$response))   {
+  for (desc in unique(fidata$descriptor)){
+    for (loc in unique(fidata$site))     {
+
+      print(paste(desc,loc))
+
+      lrrtd <- fidata %>%
+        filter(descriptor == desc,
+               site == loc)
+
+      # Test for normality
+      print(shapiro.test(lrrtd$x))
+      print(shapiro.test(lrrtd$y))
+
+      lm.test.p <- summary(lm(lrrtd$y~lrrtd$x, data = lrrtd))$coefficients[2,4]
+
+      print(lm.test.p)
+
+      lrr.res <- data.frame(descriptor = desc,
+                            site = loc,
+                            non.param.p = ifelse(lm.test.p<0.05,"signif","ns"))
+      lrr.cor.test.data <- rbind(lrr.cor.test.data, lrr.res)
+
+    }
+  }
+# }
+
+lrr.cor.test.data -> lctd
+
+fidata$sign <- ""
+
+sign.dat <- lctd[lctd$non.param.p == "signif",]
+
+for(row in 1:dim(sign.dat)[1]){
+  # resp <- sign.dat[row, ]$response
+  desc <- sign.dat[row, ]$descriptor
+  loc <- sign.dat[row, ]$site
+
+  # c1 <- cpdata$response == resp
+  c2 <- fidata$descriptor == desc
+  c3 <- fidata$site == loc
+
+  fidata[c2 & c3, ]$sign <- "signif"
+}
+
+ggplot(fidata, aes(x = x, y = y)) +
+  geom_point(size = 2, alpha = 0.4)+
+  stat_smooth(data = fidata %>%
+                filter(sign == "signif"),
+              method ="lm")+
+  geom_hline(yintercept = 0)+
+  geom_vline(xintercept = 0)+
+  ylab("Effect size of insects")+
+  xlab("Effect size of fungi")+
+  theme_bw()+
+  facet_grid(descriptor~site)
+
+```

### 2023-01-18 supplementary/food_web_model.Rmd
Adding food chain model with Rmd

---
title: "Mechanistic model"
author: "Piotr Szefer"
date: "1/18/2023"
output: html_document
---

To investigate further what are the possible outcomes and mechanisms in our experiment we build a mechanistic RM model based on (McCann). The model we use is as follows:

$$ \begin{matrix} \frac{dP}{dt} = rP \left( 1-\frac{P}{K} \right)-\frac{\Omega_{HP}x_{H}y_{H}PH}{P+P_0} \\ 
\frac{dH}{dt} = -x_HH\left(1-\frac{\Omega_{HP}y_HP}{P+P0}\right)-\frac{\Omega_{SH}x_{S}y_SHS}{H+H_0}-\frac{\Omega_{BH}y_Bx_BHB}{o_{BH}H+(1-\Omega_{BH})S+H_0} \\ 
 dS = -x_SS \left( 1- \frac{\Omega_{SH}y_SH}{H+H_0}\right)- \frac{(1-\Omega_{BH})y_Bx_BSB}{\Omega_{BH}H+(1-\Omega_{BH})S+H_0} \\
    dB = -x_BB \left( 1-\frac{\Omega_{BH}y_BH+(1-\Omega_{BH})y_BS}{\Omega_{BH}H+(1-\Omega_{BH})S+H_0)} \right) \end{matrix}$$

```{r}
library(deSolve)
# Define full model
fullModule <- function(t, state, parameters) {
  with(as.list(c(state, parameters)),{
    dP <-  r*P*(1-P/K)-(o_HP*x_H*y_H*P*H)/(P+P0)
    dH <- -x_H*H*(1-(o_HP*y_H*P)/(P+P0))-(o_SH*x_S*y_S*H*S)/(H+H0)-(o_BH*y_B*x_B*H*B)/(o_BH*H+(1-o_BH)*S+H0)
    dS <- -x_S*S*(1-(o_SH*y_S*H)/(H+H0))-((1-o_BH)*y_B*x_B*S*B)/(o_BH*H+(1-o_BH)*S+H0)
    dB <- -x_B*B*(1-(o_BH*y_B*H+(1-o_BH)*y_B*S)/(o_BH*H+(1-o_BH)*S+H0))
    list(c(dP,dH,dS,dB))
  })
}

params <- c(r = 1         ,
            K = 1       , # environmental

            o_BS = 0.7    , # preference of birds for spiders
            o_BH = 0.3    , # preference of birds for herbivores
            o_HP = 1      , # preference of herbivores for plants
            o_SH = 1      , # preference of spiders for herbivores

            x_H  = 0.2    ,
            x_B  = 0.08   ,
            x_S  = 0.5    ,

            y_H  = 2.009  ,
            y_B  = 5      ,
            y_S  = 3.5    ,

            P0   = 0.16129,
            H0   = 0.9

            )

state <- c(P = 0.5, H = 0.5, S = 0.5, B = 0.5)
times <- seq(0,500, by = 0.1)

out <- ode(y = state, times = times, func = fullModule, parms = params)

matplot(out[,-1], type = "l")
plot(log(out[dim(out)[1],-1]))
```

Now that we have a functioning moel we can try to scan the parameter space of interaction strength as it was done in the mentioned paper.

Interaction strength was calculated as $$ \alpha_{i,j} $$

### 2023-01-18 code/food_chain_model.R

@@ -1,0 +1,83 @@
+# My idea is to investigate using simple mechanistic model 
+# under which conditions we would be able to reproduce the results seen 
+# in the study.
+
+# The most important aspects of the model is its structure and the experiment. 
+# - Parameters will be subjected to simulations to evaluate the parameter space.
+# - Parameters evaluated again after the top-predators are removed to see 
+#   where we are in the parameter space 
+# - Variation in parameters should be included.
+# - Individual-based model could be implemented as well to model behavioral changes.
+
+# What is going to be the benchmark? 
+# Biomass ratios 
+# and variation in biomass
+
+
+
+# Model for the diamond module is taken from McCann, Hastings and Huxel (1998) Nature.
+
+library(deSolve)
+
+# Parameters and variables
+# R - resource density
+# C1 is the density of the first consumer species (herbivores)
+# C2 density of the second consumer (spiders)
+# P density of top predator
+
+# K - carrying capacity 
+# P0, H0 - half saturation densities of the resource and C1
+# x_i is mass specific metabolic rate of species i measured relative to the
+#     production-to-biomass ratio of the resource population 
+# y_i is ingestion rate per unit metabolic rate of species i
+# O_ij is a fraction indicating the preference of species i for consuming 
+#     resource j
+# r - growth rate of plants (equals to 1)
+
+# I will model functional responses explicitly
+# Herbivores are only ones eating the plants
+# Herbivores are consumed by both top-predators (B for birds) and spiders (S)
+# Birds eat both herbivores and spiders.
+
+fullModule <- function(t, state, parameters) {
+  with(as.list(c(state, parameters)),{
+    dP <-  r*P*(1-P/K)-(o_HP*x_H*y_H*P*H)/(P+P0) # C1 is the only consumer of the Resource
+    dH <- -x_H*H*(1-(o_HP*y_H*P)/(P+P0))-(o_SH*x_S*y_S*H*S)/(H+H0)-(o_BH*y_B*x_B*H*B)/(o_BH*H+(1-o_BH)*S+H0)
+    dS <- -x_S*S*(1-(o_SH*y_S*H)/(H+H0))-((1-o_BH)*y_B*x_B*S*B)/(o_BH*H+(1-o_BH)*S+H0)
+    dB <- -x_B*B*(1-(o_BH*y_B*H+(1-o_BH)*y_B*S)/(o_BH*H+(1-o_BH)*S+H0))
+    list(c(dP,dH,dS,dB))
+  })
+}
+
[…]

### 2023-01-18 code/food_chain_model_suppl.R

@@ -1,0 +1,171 @@
+# G:/My Drive/PROJECTS/food_webs/dynamical_model.R
+
+library(deSolve)
+
+ConstResource <- function(t, state, parameters){
+  with(as.list(c(state, parameters)),{
+    dB <- -d1*B*B    + a2*I*B + a1*H*B
+    dI <- -a2*B*I + -d2*I*I  +  a3*H*I
+    dH <- -a1*B*H - a3*I*H - d3*H*H + a4*P*H
+    dP <- 0
+    list(c(dB,dI,dH,dP))
+  })
+}
+
+ConstResourceNoP <- function(t, state, parameters){
+  with(as.list(c(state, parameters)),{
+    # dB <- -d1*B*B    + a2*I*B + a1*H*B
+    dI <- -d2*I*I  +  a3*H*I
+    dH <- -a3*I*H - d3*H*H + a4*P*H
+    dP <- 0
+    list(c(dI,dH,dP))
+  })
+}
+
+# LVGardenNoDet <- function(t, state, parameters){
+#   with(as.list(c(state, parameters)),{
+#     dB <- -d1*B*B    + a2*I*B + a1*H*B
+#     dI <- -a2*B*I + -d2*I*I  +  a3*H*I
+#     dH <- -a1*B*H - a3*I*H - d3*H*H + a4*P*H
+#     dP <- g*P                 -a4*H*P - d4*P*P
+#     list(c(dB,dI,dH,dP))
+#   })
+# }
+# 
+# LVGardenNoDetNoPred <- function(t, state, parameters){
+#   with(as.list(c(state, parameters)),{
+#     # dB <- -d1*B*B    + a2*I*B + a1*H*B
+#     dI <-  + -d2*I*I  +  a3*H*I
+#     dH <-  - a3*I*H - d3*H*H + a4*P*H
+#     dP <- g*P        -a4*H*P - d4*P*P
+#     list(c(dI,dH,dP))
+#   })
+# }
+
+unival <- runif(8, min = 0.01, max = 0.5)
+
+parameters <- c(a1=unival[1],a2=unival[2], # assimilation * feeding rates
+                a3=unival[3],              # Ip assimilation * feeding rate
+                a4=unival[4],              # Herbivore feeding rrate
+                d1=unival[5]*0.5,          # Bird intraguild competition
+                d2=unival[6]*0.5,          # Intermediate pred. intrag. comp
+                d3=unival[7]*0.5,          # Herbivore igc
[…]

### 2023-02-15 ms1_data_analysis.Rmd
Lates updata and abandoned food chain model (for now).

@@
 Stars in the upper index indicate significant ($\\alpha$ = 0.05) differences of the effect mean from zero (significance of the effect at a given site)", fig.width= 8, fig.height=8}
-
-library(ggplot2)
 
@@
 ```
+### COMMUNITY COMPOSITION
+
+At each elevation insect exclusion significantly affected the community composition [but only when all treatments were included, in pairwise comparisons, i.e., C vs I the composition was not changed significantly].
+Fungicide treatment significantly affected the full plant community composition only at the highest elevation (Fig, 3; Appendix I, Tab.
+S1).
+
+```{r}
+
+library(vegan)
+# library(shape)  
+# library(ggpubr)
+library(cowplot)
+
+# 1. Load the data ----
+COMMDATA <- t(COMMDATA)
+COMMVARS
+
+# Select only i,c,p,f
+COMM.SEL <- COMMDATA[grepl("i|c$|p|f|_h", rownames(COMMDATA)), ]
+VARS.SEL <- COMMVARS[grepl("i|c$|p|f|_h", COMMVARS$plotid), ]
+
+# For each elevation calculate inertia for each treatment. This is in order to 
+# evaluate whether some sites have the differences more strongly pronounced
+# site.id <- "n"
+# SUB.DATA <- COMMDATA[grepl(site.id, rownames(COMMDATA)), ]
+# SUB.VARS <- COMMVARS[grepl(site.id, COMMVARS$plotid), ]
+# 
+# treats.to.consider <- c("i","c","h","f","p")
+# sub.data.ttc <- SUB.VARS %>%
+#   filter(treatment %in% treats.to.consider)
+# 
+# ## Dummy-coding of the treatments ----
+# COMMVARS <- COMMVARS %>%
+#   mutate(f = ifelse(treatment == "f", 1, 0),
+#          h = ifelse(treatment == "h", 1, 0),
+#          p = ifelse(treatment == "p", 1, 0),
+#          c = ifelse(treatment == "c", 1, 0),
+#          i = ifelse(treatment == "i", 1, 0))
+# 
+# rdamod <- rda(COMMDATA~f+h+p+i+Condition(garden + site), 
+#               data = COMMVARS)
+# 
+# 
+# rda.w <- rda(decostand(SUB.DATA, 
+#                        method = 'hellinger')~treatment+Condition(garden), 
+#              data = SUB.VARS)
+
+# 2. Fitting SLA surfaces to ordination ----
+
+# WARNING: these are data that haven't been fixed yet
+
+## Species traits for evaluation ----
+# c1 <- mainDataT$treatment %in% treats.to.consider
+# c2 <- mainDataT$unified_names %in% names(rdamod$colsum)
+# c3 <- substr(mainDataT$site,1,1) == site.id
+# 
+# speciesSLA <- mainDataT[c1 & c2 & c3, c("unified_names", "sla.d.m2kg1")] # Ind sp SLA is it possible?
+# 
+# summary.sp.sla <- speciesSLA %>%
+#   group_by(unified_names) %>%
+#   summarise(mean.sla = mean(sla.d.m2kg1, na.rm = T))
+# 
+# ## CWM for sites ----
+# cwmFun <- function(sla.d.m2kg1, tot_bio){
+#   return(mean(sla.d.m2kg1*(tot_bio/sum(tot_bio, na.rm = T)), na.rm = T))
+# }
+# 
+# cwmFun(mainDataT$sla.d.m2kg1, mainDataT$tot_bio)
+# 
+# siteSLA <- mainDataT %>%
+#   filter(unified_names %in% trait.dat[trait.dat$Life.form == "woody", ]$dsName) %>%
+#   mutate(tot_bio = ifelse(is.infinite(tot_bio), NA, tot_bio),
+#          sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1)) %>%
+#   group_by(site, garden, treatment) %>%
+#   summarise(cwm.sla = cwmFun(sla.d.m2kg1, tot_bio))
+# 
+# siteSLAdf <- as.data.frame(siteSLA)
+# 
+# rownames(siteSLAdf) <- paste(substr(siteSLA$site,1,1), 
+#                              siteSLA$garden, 
+#                              "_", 
+#                              siteSLA$treatment, sep = "")
+
+# 3. RDA at each elevation ----
+
+## 3.1 Analyses for each elevation ----
+
+plot.list <- list()
+
+for(site.id in substr(unique(COMMVARS$site),1,1)){
+
+  print(site.id)
+
+  SUB.VARS <- COMMVARS %>%
+    filter(grepl(site.id, plotid) & treatment %in% treats.to.consider)
+
+  SUB.DATA <- COMMDATA[SUB.VARS$plotid, ]
+  SUB.DATA <- decostand(SUB.DATA, method = 'hellinger')
+  # Perform RDA and ANOVA and print the results
+  rdamod <- rda(SUB.DATA~f+h+p+i+Condition(garden), 
+                data = SUB.VARS)
+
+  h <- how(blocks = SUB.VARS$garden)
+
+  print(anova(rdamod, permutations = h, by = "terms"))
+
+  # Goodness of fit
+  # plot(rdamod, type= "n", dis="sp")
+
+  # Goodness
+  gdnsof <- goodness(rdamod, model = "CCA", proportional = T)
+  gof2ax <- gdnsof[,1] + gdnsof[,2]
+  cgof <- gof2ax[gof2ax > 0.2]
+  cgof <- sort(cgof, decreasing = T)
+  sign.names <- names(cgof[!is.na(cgof) & !is.infinite(cgof)])
+
+  # print(cgof[!is.na(cgof) & !is.infinite(cgof)])
+
+  # Inertia
+  # inertia <- inertcomp(rdamod, proportional = T)
+  # inttresh <- inertia[inertia[,2] > 0.2, 2]
+  # sign.names <- names(inttresh[!is.na(inttresh)])
+
+  # Significant vectors
+  # sign.sp <- envfit(rdamod, SUB.DATA)
+  # sign.names <- names(sign.sp$vectors$pvals[sign.sp$vectors$pvals < 0.05])
+
+  colors <- as.numeric(as.factor(SUB.VARS$treatment))
+
+  # rda.dat <- invisible(plot(rdamod))
+
+  # Alternative way to obtain these
+  rda.dat <- list(sites = rdamod$CCA$wa[, c(1,2)],
+                  species = rdamod$CCA$v[, c(1,2)],
+                  biplot = rdamod$CCA$biplot[, c(1,2)])
+
+  sites.dat <- as.data.frame(rda.dat$sites) %>%
+    mutate(codes = rownames(rda.dat$sites),
+           treats = ifelse(grepl("c", codes),"c",
+                           ifelse(grepl("f", codes),"f", 
+                                  ifelse(grepl("h", codes), "h", 
+                                         ifelse(grepl("i", codes), "i","p")))),
+           shapes = as.numeric(as.factor(treats)))
+
+  species.dat <- as.data.frame(rda.dat$species) %>%
+    mutate(names = rownames(as.data.frame(rda.dat$species)))
+
+  arrow.dat <- as.data.frame(rda.dat$biplot) %>%
+    mutate(treats = rownames(as.data.frame(rda.dat$biplot)))
+
+  if(site.id == "y"){
+    arr.col <- c("red","black","black","red")
+    arr.lwd <- c(0.5,0.1,0.1,0.5)
+    arr.lty <- c(1,2,2,1)
+  } else {arr.col <- c("black","black","black","red")
+  arr.lwd <- c(0.1,0.1,0.1,0.5)
+  arr.lty <- c(2,2,2,1)
+  }
+
+  p1 <- ggplot(sites.dat,aes(x = RDA1, y = RDA2))+
+    geom_point(aes(colour = treats), pch = sites.dat$shapes+14) + 
+    geom_hline(yintercept = 0, lty = 2, alpha = 0.5) + 
+    geom_vline(xintercept = 0, lty = 2, alpha = 0.5) + 
+    geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
+                                       xend = RDA1, yend = RDA2),
+                 arrow = arrow(length=unit(0.30,"cm"), 
+                               ends="last", 
+                               type = "closed"),
+                 alpha = 0.8, 
+                 col = arr.col,
+                 lwd = arr.lwd,
+                 lty = arr.lty)+
+    geom_text(data = arrow.dat, aes(x = RDA1*1.1, y = RDA2*1.1, 
+                                    label = treats),
+              cex = 5)+
+    theme_minimal()
+
+  # scalar = 2.5
+
+  p2 <- ggplot(species.dat, aes(x = RDA1*1.1, y = RDA2*1.1)) + 
+    geom_text(data = species.dat %>%
+                filter(names %in% sign.names), aes(label = names))+
+    geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
+                                       xend = RDA1, yend = RDA2),
+                 arrow = arrow(length=unit(0.30,"cm"), 
+                               ends="last", 
+                               type = "closed"),
+                 alpha = 0.8, 
+                 col = arr.col,
+                 lwd = arr.lwd,
+                 lty = arr.lty)+
+    geom_text(data = arrow.dat, aes(x = RDA1*1.1, y = RDA2*1.1, 
+                                    label = treats))+
+    theme_minimal()
+
+  plot.list[[site.id]] <- list(sites.ord = as_grob(p1),
+                               species.ord = as_grob(p2))
+
+  # surfDat <- siteSLAdf[rownames(rdamod$CCA$u),]
+  # print(ordisurf(rdamod ~ cwm.sla, data = surfDat, bubble = 5))
+
+}
+
+## 3.2 Plot the ordination ----
+# tiff(filename='figures/ordination_panel.tif',
+#      height=5600,
+#      width=5200,
+#      units='px',
+#      res=800,compression='lzw')
+ggpubr::ggarrange(plot.list[["w"]]$sites.ord,
+                  plot.list[["w"]]$species.ord,
+                  plot.list[["n"]]$sites.ord,
+                  plot.list[["n"]]$species.ord,
+                  plot.list[["y"]]$sites.ord,
+                  plot.list[["y"]]$species.ord, ncol = 2, nrow = 3)
+
+```

### 2023-02-15 supplementary/food_web_model.Rmd
Lates updata and abandoned food chain model (for now).

@@
 
+Parameters are: $\Omega_ij$ preference of a consumer $i$ to a resource $j$.
+$x_i$ - mass specific metabolic rate, $y_i$ - measure of ingestion rate.
+There is no indication of an attack rate, would that be the $x_i$?
+$y_i$ is limited by an animal's physiological capacity.
+
 ```{r}
@@
 
-params <- c(r = 1         ,
+reducedModule <- function(t, state, parameters) {
+  # This models the a simple food chain without the top predators
+  with(as.list(c(state, parameters)),{
+    dP <-  r*P*(1-P/K)-(o_HP*x_H*y_H*P*H)/(P+P0)
+    dH <- -x_H*H*(1-(o_HP*y_H*P)/(P+P0))-(o_SH*x_S*y_S*H*S)/(H+H0)
+    dS <- -x_S*S*(1-(o_SH*y_S*H)/(H+H0))
+    list(c(dP,dH,dS))
+  })
+}
+
+# Remove metabolic rate from the biomass removal equation
+# fullModule <- function(t, state, parameters) {
+#   with(as.list(c(state, parameters)),{
+#     dP <-  r*P*(1-P/K)-(o_HP*y_H*P*H)/(P+P0)
+#     dH <- -x_H*H*(1-(o_HP*y_H*P)/(P+P0))-(o_SH*y_S*H*S)/(H+H0)-(o_BH*y_B*H*B)/(o_BH*H+(1-o_BH)*S+H0)
+#     dS <- -x_S*S*(1-(o_SH*y_S*H)/(H+H0))-((1-o_BH)*y_B*S*B)/(o_BH*H+(1-o_BH)*S+H0)
+#     dB <- -x_B*B*(1-(o_BH*y_B*H+(1-o_BH)*y_B*S)/(o_BH*H+(1-o_BH)*S+H0))
+#     list(c(dP,dH,dS,dB))
+#   })
+# }
+# 
+# reducedModule <- function(t, state, parameters) {
+#   # This models the a simple food chain without the top predators
+#   with(as.list(c(state, parameters)),{
+#     dP <-  r*P*(1-P/K)-(o_HP*y_H*P*H)/(P+P0)
+#     dH <- -x_H*H*(1-(o_HP*y_H*P)/(P+P0))-(o_SH*y_S*H*S)/(H+H0)
+#     dS <- -x_S*S*(1-(o_SH*y_S*H)/(H+H0))
+#     list(c(dP,dH,dS))
+#   })
+# }
+
+# mass specific metabolic rate
+metabolic_rate_herbivores <- 0.2
+metabolic_rate_birds <- 0.08
+metabolic_rate_spiders <- 0.3
+
+# ingestion rate per unit metabolic rate of species i
+ingestion_rate_herbivores <- 1.5
+ingestion_rate_birds <- 5
+ingestion_rate_spiders <- 2.1
+
+half_sat_plants <- 0.2
+half_sat_herbivores <- 0.5
+
+params_fm <- c(r = 1         ,
             K = 1       , # environmental
 
-            o_BS = 0.7    , # preference of birds for spiders
-            o_BH = 0.3    , # preference of birds for herbivores
+            o_BS = 0.6    , # preference of birds for spiders
+            o_BH = 0.4    , # preference of birds for herbivores
             o_HP = 1      , # preference of herbivores for plants
@@
 
-            x_H  = 0.2    ,
-            x_B  = 0.08   ,
-            x_S  = 0.5    ,
+            x_H  = metabolic_rate_herbivores,
+            x_B  = metabolic_rate_birds,
+            x_S  = metabolic_rate_spiders,
 
-            y_H  = 2.009  ,
-            y_B  = 5      ,
-            y_S  = 3.5    ,
+            y_H  = ingestion_rate_herbivores  , 
+            y_B  = ingestion_rate_birds  ,
+            y_S  = ingestion_rate_spiders    , # ing.rates
 
-            P0   = 0.16129,
-            H0   = 0.9
+            P0   = half_sat_plants,
+            H0   = half_sat_herbivores)
 
-            )
+params_rm <- c(r = 1         ,
+               K = 1       , # environmental
 
-state <- c(P = 0.5, H = 0.5, S = 0.5, B = 0.5)
-times <- seq(0,500, by = 0.1)
+               o_HP = 1      , # preference of herbivores for plants
+               o_SH = 1      , # preference of spiders for herbivores
 
-out <- ode(y = state, times = times, func = fullModule, parms = params)
+               x_H  = metabolic_rate_herbivores,
+               x_S  = metabolic_rate_spiders,
 
-matplot(out[,-1], type = "l")
-plot(log(out[dim(out)[1],-1]))
+               y_H  = ingestion_rate_herbivores  , 
+               y_B  = ingestion_rate_birds  ,
+               y_S  = ingestion_rate_spiders ,
+
+               P0   = half_sat_plants,
+               H0   = half_sat_herbivores)
+
+stateFM <- c(P = 0.5, H = 0.5, S = 0.5, B = 0.5)
+stateRM <- c(P = 0.5, H = 0.5, S = 0.5)
+times <- seq(0,900, by = 0.1)
+
+outFM <- ode(y = stateFM, times = times, func = fullModule,    parms = params_fm)
+outRM <- ode(y = stateRM, times = times, func = reducedModule, parms = params_rm)
+
+library(reshape2)
+df.fm <- data.frame(time = 1:nrow(outFM), outFM[,-1])
+df.fm.melt <- melt(df.fm, id = "time")
+df.fm.melt$model <- "Full Model"
+
+df.rm <- data.frame(time = 1:nrow(outRM), outRM[,-1])
+df.rm.melt <- melt(df.rm, id = "time")
+df.rm.melt$model <- "Reduced Model"
+
+df.bm <- rbind(df.fm.melt, df.rm.melt)
+
+library(ggplot2)
+ggplot(df.bm, aes(x = time, y = log(value), color = variable))+
+  geom_line()+
+  facet_wrap(~model, scales = "free")
+
 ```
 
-Now that we have a functioning moel we can try to scan the parameter space of interaction strength as it was done in the mentioned paper.
+Now that we have a functioning model we can try to scan the parameter space of **interaction strength** as it was done in the mentioned paper.
 
-Interaction strength was calculated as $$ \alpha_{i,j} $$
+The per-capita interaction strength of resource i on consumer k was defined as $$ \alpha_{k,i} = \frac{K_k(i)}{i} = \frac{\Omega_{ki}x_ky_k}{\Omega_{ki}i + (1 - \Omega_{ki})j + i_0} $$
+
+The **interaction strength** was defined as the maximum per capita interaction strength of resource of resource species j on the consumer species i.
+Per capita interaction strength is maximum when densities of j and i approach zero.
+Therefore: $$ I_{ki} = \frac{\Omega_{ki}x_ky_k}{i_0} $$
+
+But we will get to that later. Now a sample of what the algorithm can do.
+
+First of all we need to focus mainly on an abstract relations between parameters.
+We are not interested how did parameters change but rather what phenomenons (like a change in interaction strength) can explain, at least qualitatively, the observed patterns in LRR.
+
+Possible hypotheses.
+The advantage of the above model is a possibility of estimation of plausible values of the $x_k$ and $y_k$ parameters as well as working with interaction strengths.
+This allows us to form our hypothesis in terms of interaction strengths:
+
+1. Spiders feed more on herbivores - interaction strength increases in the absence of predators?
+What I would need is a model where the capture rates are explicit. But I can maybe work with what I have here.
+
+What are the parameters and what do I know about them.
+
+```{r}
+# I can manipulate three parameters, lets say
+runModel <- function(r=1,K=1, o_BS=0.7, o_BH=0.3, o_HP=1, o_SH=1, x_H=0.01, x_B=0.0008,
+                     x_S=0.02, y_H=2.009, y_B=4, y_S=3.5, P0=0.16129,H0=0.9, rr = 1, rK = 1,
+                     ro_HP = 1, ro_SH = 1, rx_H=0.01, rx_S=0.02, ry_H  = 2.009, ry_S  = 3.5,
+                     rP0=0.16129,rH0=0.9,
+                     stateFM = c(P = 0.5, H = 0.5, S = 0.5, B = 0.5),
+                     stateRM = c(P = 0.5, H = 0.5, S = 0.5),
+                     times = seq(0,900, by = 0.1)) {
+  params_fm <- c(r=r,
+                 K=K,
+                 o_BS = o_BS,
+                 o_BH = o_BH,
+                 o_HP = o_HP,
+                 o_SH = o_SH,
+                 x_H  = x_H,
+                 x_B  = x_B,
+                 x_S  = x_S,
+                 y_H  = y_H,
+                 y_B  = y_B,
+                 y_S  = y_S,
+                 P0   = P0,
+                 H0   = H0) 
+  params_rm <- c(r    = rr,
+                 K    = rK,
+                 o_HP = ro_HP,
+                 o_SH = ro_SH,
+                 x_H  = rx_H,
+                 x_S  = rx_S,
+                 y_H  = ry_H,
+                 y_S  = ry_S,
+                 P0   = rP0,
+                 H0   = rH0) 
+# Itegrate 
+outFM <- ode(y = stateFM, times = times, func = fullModule,    parms = params_fm)
+outRM <- ode(y = stateRM, times = times, func = reducedModule, parms = params_rm)
+df.fm <- data.frame(time = 1:nrow(outFM), outFM[,-1])
+df.fm.melt <- melt(df.fm, id = "time")
+df.fm.melt$model <- "Full Model"
+df.rm <- data.frame(time = 1:nrow(outRM), outRM[,-1])
+df.rm.melt <- melt(df.rm, id = "time")
+df.rm.melt$model <- "Reduced Model"
+df.bm <- rbind(df.fm.melt, df.rm.melt)
+return(df.bm)
+}
+
+x_H.seq <- seq(from = 0.001, to = 0.2, by = 0.005)
+
+full.data <- data.frame()
+
+for (x_H.val in x_H.seq) {
+  mod.data <- runModel(x_H = x_H.val)
+  mod.data$x_H.par <- x_H.val
+  full.data <- rbind(full.data, mod.data)
+  print(x_H.val)
+}
+
+ggplot(full.data, aes(x = time, y = log(value), color = variable))+
+  geom_line()+
+  facet_grid(model~x_H.par)
+
+```
+```{r}
+
+# Build the model-up starting with just plants add herbivores, spiders, and birds
+
+# Step 1 plants
+
+stateS1 <- c()
+
+```

### 2023-02-15 code/food_chain_model.R

@@ -1,47 +1,94 @@
-# My idea is to investigate using simple mechanistic model 
-# under which conditions we would be able to reproduce the results seen 
-# in the study.
+# # My idea is to investigate using simple mechanistic model 
+# # under which conditions we would be able to reproduce the results seen 
+# # in the study.
+# 
+# # The most important aspects of the model is its structure and the experiment. 
+# # - Parameters will be subjected to simulations to evaluate the parameter space.
+# # - Parameters evaluated again after the top-predators are removed to see 
+# #   where we are in the parameter space 
+# # - Variation in parameters should be included.
+# # - Individual-based model could be implemented as well to model behavioral changes.
+# 
+# # What is going to be the benchmark? 
+# # Biomass ratios 
+# # and variation in biomass
+# 
+# 
+# 
+# # Model for the diamond module is taken from McCann, Hastings and Huxel (1998) Nature.
+# 
+# library(deSolve)
+# 
+# # Parameters and variables
+# # R - resource density
+# # C1 is the density of the first consumer species (herbivores)
+# # C2 density of the second consumer (spiders)
+# # P density of top predator
+# 
+# # K - carrying capacity 
+# # P0, H0 - half saturation densities of the resource and C1
+# # x_i is mass specific metabolic rate of species i measured relative to the
+# #     production-to-biomass ratio of the resource population 
+# # y_i is ingestion rate per unit metabolic rate of species i
+# # O_ij is a fraction indicating the preference of species i for consuming 
+# #     resource j
+# # r - growth rate of plants (equals to 1)
+# 
+# # I will model functional responses explicitly
+# # Herbivores are only ones eating the plants
+# # Herbivores are consumed by both top-predators (B for birds) and spiders (S)
+# # Birds eat both herbivores and spiders.
+# 
+# fullModule <- function(t, state, parameters) {
+#   with(as.list(c(state, parameters)),{
+#     dP <-  r*P*(1-P/K)-(o_HP*x_H*y_H*P*H)/(P+P0) # C1 is the only consumer of the Resource
[…]

### 2023-02-15 code/food_chain_model_RM.R

@@ -1,0 +1,81 @@
+library(deSolve)
+library(spatialEco)
+
+
+# Three species RM model
+reducedRMM <- function(t, state, parameters) {
+  # This models the a simple food chain without the top predators
+  with(as.list(c(state, parameters)),{
+    dP <-  r * P * (1 - P/K) - (a_H * P * H)/(P + P0) # Plant
+    dH <- (e_H * a_H * P * H)/(P + P0) - (m_H * H) - (a_S * S * H)/(H + H0)
+    dS <- (e_S * a_S * S * H)/(H + H0) - (m_S * S)
+    list(c(dP,dH,dS))
+  })
+}
+
+
+# Parameters
+params <- c(r = 1,
+            K = 2,
+            P0 = 0.5,
+            
+            a_H = 1.8,
+            m_H = 0.5,
+            e_H = 0.5,
+            H0 = 0.3,
+            
+            a_S = 1.15,
+            m_S = 0.3,
+            e_S = 0.5 
+            )
+
+times <- seq(0,1000, by = 0.01)
+
+states <- c(P = 0.5,H = 0.5, S = 0.5)
+
+out <- ode(y = states, 
+           times = times, 
+           func = reducedRMM, 
+           parms = params)
+
+matplot(log(out[,-1]), type = "l")
+
+# Analyse the model
+getPoints <- function(out, col = "P"){
+  localmm <- local.min.max(out[(dim(out)[1]/2):(dim(out)[1]-dim(out)[1]*0.1), col], plot = F)[c(1,2)]
+  lapply(localmm, function(x){x[10:length(x)]})
+}
+
+# For the a_S plot herbivore bifurcations
+plotBifurcations <- function (aS.range = seq(0.5,3, by = 0.01)) {
+  plot(x=aS.range, y = seq(0,1, length = length(aS.range)),  type = "n")
+  point.list <- list()
+  for(aS in aS.range){
+    params["a_S"] <<- aS
+    print(paste("Now taking a_S = ", aS))
+    curr.ode <- ode(y = states, 
+               times = times, 
+               func = reducedRMM, 
+               parms = params)
+    pointsToPlot <- unlist(getPoints(curr.ode, col = "P"))
+    points(x = rep(aS, length(pointsToPlot)), y = pointsToPlot, pch = 19)
[…]

### 2023-03-15 ms1_data_analysis.Rmd
Simple model for SLA added to the summary.

--- 2023-01-18 ms1_data_analysis.Rmd
@@
 
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, results='hide',warning=FALSE, message=FALSE)
 
@@
 source('code/log_ratio_extraction.R')
-
-# Over-dispersion test for the density in the fungicide treatment
-# lrdata <- DENSITY.LR
-# lvT <- leveneTest(lratio ~ site, data = lrdata %>% filter(treatment == "f"))
-# lvtpval <- lvT$`Pr(>F)`[1]
 ```
@@
 
-library(ggplot2)
-
 # bio, dens, div, rich
@@
 ```
+
 ## LRR correlations
@@
 ```
+
 From the figure we see that there is usually no correlation in Wanang and Numba but it is always present in Yawan.
@@
 ```
+
+### COMMUNITY COMPOSITION
+
+At each elevation insect exclusion significantly affected the community composition [but only when all treatments were included, in pairwise comparisons, i.e., C vs I the composition was not changed significantly].
+Fungicide treatment significantly affected the full plant community composition only at the highest elevation (Fig, 3; Appendix I, Tab.
+S1).
+
+```{r}
+
+library(vegan)
+library(cowplot)
+
+# 1. Load the data
+COMMDATA <- t(COMMDATA)
+COMMVARS
+
+# Select only i,c,p,f
+COMM.SEL <- COMMDATA[grepl("i|c$|p|f|_h", rownames(COMMDATA)), ]
+VARS.SEL <- COMMVARS[grepl("i|c$|p|f|_h", COMMVARS$plotid), ]
+
+treats.to.consider <- c("i","c","h","f","p")
+
+# Dummy-coding of the treatments
+COMMVARS <- COMMVARS %>%
+  mutate(f = ifelse(treatment == "f", 1, 0),
+         h = ifelse(treatment == "h", 1, 0),
+         p = ifelse(treatment == "p", 1, 0),
+         c = ifelse(treatment == "c", 1, 0),
+         i = ifelse(treatment == "i", 1, 0))
+
+# RDA at each elevation
+
+## 3.1 Analyses for each elevation
+
+plot.list <- list()
+
+for(site.id in substr(unique(COMMVARS$site),1,1)){
+
+  print(site.id)
+
+  SUB.VARS <- COMMVARS %>%
+    filter(grepl(paste('^', site.id, sep = ""), site) & treatment %in% treats.to.consider)
+
+  SUB.DATA <- COMMDATA[SUB.VARS$plotid, ]
+  SUB.DATA <- decostand(SUB.DATA, method = 'hellinger')
+  # Perform RDA and ANOVA and print the results
+  rdamod <- rda(as.data.frame(SUB.DATA)~f+h+p+i+Condition(garden), 
+                data = SUB.VARS)
+
+  h <- how(blocks = SUB.VARS$garden)
+
+  print(anova(rdamod, permutations = h, by = "terms"))
+
+  # Goodness
+  gdnsof <- goodness(rdamod, model = "CCA", proportional = T)
+  gof2ax <- gdnsof[,1] + gdnsof[,2]
+  cgof <- gof2ax[gof2ax > 0.2]
+  cgof <- sort(cgof, decreasing = T)
+  sign.names <- names(cgof[!is.na(cgof) & !is.infinite(cgof)])
+
+  colors <- as.numeric(as.factor(SUB.VARS$treatment))
+
+  # Alternative way to obtain these
+  rda.dat <- list(sites = rdamod$CCA$wa[, c(1,2)],
+                  species = rdamod$CCA$v[, c(1,2)],
+                  biplot = rdamod$CCA$biplot[, c(1,2)])
+
+  sites.dat <- as.data.frame(rda.dat$sites) %>%
+    mutate(codes = rownames(rda.dat$sites),
+           treats = ifelse(grepl("c", codes),"c",
+                           ifelse(grepl("f", codes),"f", 
+                                  ifelse(grepl("h", codes), "h", 
+                                         ifelse(grepl("i", codes), "i","p")))),
+           shapes = as.numeric(as.factor(treats)))
+
+  species.dat <- as.data.frame(rda.dat$species) %>%
+    mutate(names = rownames(as.data.frame(rda.dat$species)))
+
+  arrow.dat <- as.data.frame(rda.dat$biplot) %>%
+    mutate(treats = rownames(as.data.frame(rda.dat$biplot)))
+
+  if(site.id == "y"){
+    arr.col <- c("red","black","black","red")
+    arr.lwd <- c(0.5,0.1,0.1,0.5)
+    arr.lty <- c(1,2,2,1)
+  } else {arr.col <- c("black","black","black","red")
+  arr.lwd <- c(0.1,0.1,0.1,0.5)
+  arr.lty <- c(2,2,2,1)
+  }
+
+  p1 <- ggplot(sites.dat,aes(x = RDA1, y = RDA2))+
+    geom_point(aes(colour = treats), pch = sites.dat$shapes+14) + 
+    geom_hline(yintercept = 0, lty = 2, alpha = 0.5) + 
+    geom_vline(xintercept = 0, lty = 2, alpha = 0.5) + 
+    geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
+                                       xend = RDA1, yend = RDA2),
+                 arrow = arrow(length=unit(0.30,"cm"), 
+                               ends="last", 
+                               type = "closed"),
+                 alpha = 0.8, 
+                 col = arr.col,
+                 lwd = arr.lwd,
+                 lty = arr.lty)+
+    geom_text(data = arrow.dat, aes(x = RDA1*1.1, y = RDA2*1.1, 
+                                    label = treats),
+              cex = 5)+
+    theme_minimal()
+
+  # scalar = 2.5
+
+  p2 <- ggplot(species.dat, aes(x = RDA1*1.1, y = RDA2*1.1)) + 
+    geom_text(data = species.dat %>%
+                filter(names %in% sign.names), aes(label = names))+
+    geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
+                                       xend = RDA1, yend = RDA2),
+                 arrow = arrow(length=unit(0.30,"cm"), 
+                               ends="last", 
+                               type = "closed"),
+                 alpha = 0.8, 
+                 col = arr.col,
+                 lwd = arr.lwd,
+                 lty = arr.lty)+
+    geom_text(data = arrow.dat, aes(x = RDA1*1.1, y = RDA2*1.1, 
+                                    label = treats))+
+    theme_minimal()
+
+  plot.list[[site.id]] <- list(sites.ord = as_grob(p1),
+                               species.ord = as_grob(p2))
+
+}
+
+## 3.2 Plot the ordination ----
+ggpubr::ggarrange(plot.list[["w"]]$sites.ord,
+                  plot.list[["w"]]$species.ord,
+                  plot.list[["n"]]$sites.ord,
+                  plot.list[["n"]]$species.ord,
+                  plot.list[["y"]]$sites.ord,
+                  plot.list[["y"]]$species.ord, ncol = 2, nrow = 3)
+
+```
+
+### Bray-Curtis dissimilarity
+
+BC dissimilarity at the control plots was higher in Wanang than in Numba.
+Addition of herbivores created a monotonic decrease in dissimilarity with elevation (Fig.
+4, reduction of dissimilarity means higher predictability of the plant community).
+In Wanang species composition changed mainly through turnover of species, in contrast to two other sites, where community changed through changes in the abundance structure (gradient component).
+1.
+Baseline (?) species turnover is highest in Wanang.
+Gradient changes are smaller in wanang than in Numba.
+2.
+Fungicide did not change the character of the community structure and assembly (patterns are similar in the case of both control and fungicide treatment).
+3.
+Herbivore addition monotonically reduced dissimilarity (plots within the treatment were more similar to each other, reduced unpredictability of succession) by reducing it substantially in Yawan.
+This was caused by an increase in balanced component in numba and its reduction in Yawan.
+This means that species turnover between sites was reduced by additional herbivorous insects at the highest elevation.
+However, in both of these sites balanced component was similar.
+It was high in W and did not change there.
+4.
+Insect removal monotonically reduces plant species turnover.
+Does not change the pattern of dissimilarity between elevations but substantially increases gradient component in Yawan (it increased slightly gradient component in Wanang and reduced significantly in Numba).
+Insect removal significantly reduced the species turnover (balanced component) in Yawan, and increased it in Numba.
+This means that the presence of insects (opposite of what the treatment does) increases species turnover in Y and reduces species turnover in N.
+Simultaneously it increase gradient component in Wanang and reduces it in Numba.
+Interestingly in Yawan further increase of herbivore pressure has the opposite effect again it brings it back to the previous level.
+Having additional herbivores has the same effect as not having insects at all in case of the balanced component.
+Predator treatmen is the same.
+In case of the gradient component only removal of insects was able to change it as indicated above.
+And only in wanang and numba.
+5.
+Top-predators increase balanced and gradient component similar to insecticide treatement in Numba.
+Similarly to H and I it reduces balanced component in Yawan.
+But has no effect on BC dissimilarity.
+In Numba also reduce gradient component but does not do that in Yawan.
+Between sites: Balanced component is lowest in Yawan general BC dissimilarity is lowest in Yawan and gradient in Yawan is higher in Yawan than in Numba but similar to that in Wanang.
+
+```{r}
+
+# NEEDS TO BE FIXED
+
+# library(betapart)
+# 
+# csites <- grepl("_c\\b", COMMVARS$plotid)
+# fsites <- grepl("_f", COMMVARS$plotid)
+# isites <- grepl("_i", COMMVARS$plotid)
+# hsites <- grepl("_h", COMMVARS$plotid)
+# psites <- grepl("_p", COMMVARS$plotid)
+# 
+# wsites <- COMMVARS$site == "wanang"
+# nsites <- COMMVARS$site == "numba"
+# ysites <- COMMVARS$site == "yawan"
+# 
+# # B-C dissimilarities in C plots between sites
+# cdatc <- COMMDATA[grepl("c$", rownames(COMMDATA)), ]
+# cvars <- COMMVARS[rownames(cdatc), ]
+# 
+# # Make a matrix
+# for(i in 1:dim(cdatc)[1]){
+#   
+#   # Compare all other sites to the current i
+#   for (j in (1:dim(cdatc)[1])[-i]) {
+#     val <- beta.pair.abund(cdatc[c(i,j), ])
+#   }
+#   
+# }
+# 
+# # Comparisons between
+# wc.beta <- beta.pair.abund(COMMDATA[COMMVARS[wsites & csites,]$plotid,  ])
+# 
+# beta.df <- data.frame()
+# 
+# for(siteid in c("w","y","n")){
+#   for(treat in c("_c\\b","_f","_h","_i","_p")){
+#     
+#     # Subset the data
+#     nms <- rownames(COMMDATA)
+#     sel.nms <- nms[grepl(siteid, nms) & grepl(treat, nms)]
+#     site.treat.comm <- COMMDATA[sel.nms,]
+#     
+#     print(rownames(site.treat.comm))
+#     
+#     # Calculate beta diversity
+#     subbeta <- beta.pair.abund(site.treat.comm)
+#     
+#     # Add to a data frame
+#     balanced.df <- data.frame(site = siteid,
+#                               treat = treat,
+#                               vals = as.numeric(subbeta$beta.bray.bal),
+#                               component = "balanced")
+#     
+#     gradient.df <- data.frame(site = siteid,
+#                               treat = treat,
+#                               vals = as.numeric(subbeta$beta.bray.gra),
+#                               component = "gradient")
+#     
+#     bray.df <- data.frame(site = siteid,
+#                               treat = treat,
+#                               vals = as.numeric(subbeta$beta.bray),
+#                               component = "bray")
+#     
+#     sub.df <- rbind(balanced.df, gradient.df, bray.df)
+#       
+#     # Merge the datset
+#     beta.df <- rbind(beta.df,sub.df)
+#     
+#   }
+# }
+# beta.df
+# 
+# # Fix the site order
+# beta.df$fsite <- factor(beta.df$site, levels = c("w","n","y"))
+# 
+# # Transform 0s to 0.001 and 1s to 0.999
+# beta.df$trans.vals = ifelse(beta.df$vals == 1, 0.999999, 
+#                             ifelse(beta.df$vals == 0, 0.000001, beta.df$vals))
+# 
+# # Tests for individual facets
+# 
+# library(lsmeans)
+# library(emmeans)
+# library(multcomp)
+# library(multcompView)
+# 
+# cld.dat = data.frame()
+# prws.comp <- list()
+# 
+# # CLD data for comparisons between elevations
+# for (comp in unique(beta.df$component)){
+#   print(comp)
+#   for (trt in unique(beta.df$treat)){
+#     print(trt)
+#     
+#     test.df <- beta.df %>%
+#       filter(treat == trt,
+#              component == comp)
+#     
+#     # print(test.df)
+#     
+#     mod1 <- betareg::betareg(trans.vals ~ fsite, data = test.df)
+#     
+#     lsmod1 <- lsmeans(mod1,
+#                       pairwise ~ fsite,
+#                       adjust="tukey")
+#     
+#     groups <- cld(lsmod1, Letters = c("abcdefg"))
+#     
+#     y_loc = test.df %>%
+#       group_by(fsite) %>%
+#       summarise(y_loc = mean(trans.vals))
+#     
+#     sub.cld <- data.frame(component = comp,
+#                treat = trt,
+#                fsite = groups$fsite,
+#                gr.label = groups$.group,
+#                y_loc = y_loc$y_loc[order(groups$fsite)])
+#     
+#     print(y_loc)
+#     print(sub.cld)
+#     
+#     prws.comp[[paste(comp,trt)]] <- lsmod1
+#     
+#     cld.dat = rbind(cld.dat, sub.cld)
+#     
+#   }
+# }
+# 
+# # cld.dat = data.frame()
+# # prws.comp <- list()
+# 
+# # For data for comparisons between treatments we need pairwise comps 
+# # (control vs treat) - this can be obtained from summary of the betareg
+# 
+# # library(glmmTMB)
+# 
+# pairwise.comp.res <- data.frame()
+# for (comp in unique(beta.df$component)){
+#   
+#   for (ste in unique(beta.df$fsite)){
+#     print(comp)
+#     print(ste)
+#     
+#     test.df <- beta.df %>%
+#       filter(fsite == ste,
+#              component == comp)
+#     
+#     # print(test.df)
+#     
+#     mod1 <- betareg::betareg(trans.vals ~ treat, data = test.df)
+#     
+#     breg.table <- summary(mod1)$coefficients$mean
+#     breg.pvals <- breg.table[,4]
+#     
+#     y_loc = test.df %>%
+#       group_by(treat) %>%
+#       summarise(y_loc = mean(trans.vals))
+#     
+#     pcr.row <- data.frame(comp = comp,
+#                           fsite = ste,
+#                           component = comp,
+#                           treat = names(breg.pvals),
+#                           p.vals = breg.pvals,
+#                           y_loc = y_loc$y_loc,
+#                           y_loc_trt = y_loc$treat)
+#     
+#     pcr.row <- pcr.row %>%
+#       mutate(label = ifelse(p.vals <= 0.05 & treat != "(Intercept)", "*",""))
+#     
+#     pairwise.comp.res <- rbind(pairwise.comp.res,pcr.row)
+#     
+#   }
+# }
+# 
+# pairwise.comp.res
+# 
+# # Plot
+# # tiff(filename='figures/beta_part_panel.tif',
+# #      height=5600,
+# #      width=5600,
+# #      units='px',
+# #      res=800,compression='lzw')
+# beta.part.plot <- ggplot(beta.df, aes(x = fsite, y = vals))+
+#   geom_jitter(width = 0.1, alpha = 0.2, shape = 16, size = 1.8)+
+#   stat_summary(fun.y = mean, na.rm = T,
+#                geom = "point", size = 3, col = "gray30")+
+#   stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
+#   stat_summary(fun.y = mean, aes(group = treat),
+#                geom = "line", alpha = 0.8, col = "gray30", lty = 2)+
+#   geom_text(data = cld.dat,
+#             aes(x = fsite, y = y_loc + 0.5,
+#                 label = gr.label),
+#             size = 3)+
+#   facet_grid(cols = vars(component), rows = vars(treat))
+# beta.part.plot + theme_bw()
+# 
+# # Plot for the component comparisons
+# pairwise.comp.res$fsite <- factor(pairwise.comp.res$fsite, levels = c("w","n","y"))
+# 
+# # tiff(filename='figures/beta_part_panel_treats.tif',
+# #      height=5600,
+# #      width=5600,
+# #      units='px',
+# #      res=800,compression='lzw')
+# 
+# # Alternative boxplot
+# beta.part.plot <- ggplot(beta.df, aes(x = treat, y = vals))+
+#   geom_boxplot()+
+#   # stat_summary(fun.y = mean, na.rm = T,
+#   #              geom = "point", size = 3, col = "gray30")+
+#   # stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
+#   geom_text(data = pairwise.comp.res,
+#             aes(x = y_loc_trt, y = y_loc + 0.5,
+#                 label = label), cex = 6,  col = "red", lwd = 2)+
+#   facet_grid(component ~ fsite)
+# 
+# beta.part.plot + theme_bw()
+
+```
+
+## Are changes random or deterministic for different treatments?
+
+At the highest studied elevation (Yawan) all treatment caused a shift from composition being deterministic in their divergence (predictably unpredictable to random).
+Similar shift, but weaker in cases of insecticide and fungicide was observed for Wanang.
+Strikingly different assembly was observed for Numba, where communities showed rather random assembly and only addition of herbivores increased divergence of the community composition.
+
+```{r}
+
+all.pairs_max_accuracy <- read.csv("data/raup_crick/all.pairs_max_accuracy.csv")
+all.pairs_max_accuracy$fsite <- factor(all.pairs_max_accuracy$fsite, 
+                                       levels = c("w","n","y"), labels = c("Wanang",
+                                                                           "Numba",
+                                                                           "Yawan"))
+
+test.df <- data.frame()
+for (st in unique(all.pairs_max_accuracy$fsite)){
+  for (cp in unique(all.pairs_max_accuracy$comp)){
+
+    print(paste(st, cp))
+
+    subdat <- all.pairs_max_accuracy %>%
+      filter(fsite == st & comp == cp)
+
+    mod <- nlme::lme(RC ~ treatment,
+                     random = ~1|comp.id,
+                     data = subdat)
+
+    # print(t.test(RC ~ treatment, data = subdat, paired = TRUE))
+
+    test.df <- rbind(test.df, data.frame(fsite = st,
+                                         comp = cp,
+                                         group1 = tolower(substr(cp,1,1)),
+                                         group2 = tolower(substr(cp,6,6)),
+                                         test.res = summary(mod)$tTable[2,5]))
+
+    print(summary(mod))
+
+  }
+}
+
+test.df <- test.df %>%
+  mutate(label = ifelse(test.res < 0.05, "*"," "))
+
+ggplot(all.pairs_max_accuracy, aes(x = treatment, y = RC))+
+  # geom_boxplot()+
+  ylim(c(-1.25,1.5))+
+  stat_summary(fun.y = mean, na.rm = T,
+               geom = "point", size = 4)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  theme_bw()+
+  ggtitle(paste("Accuracy = 0.001"))+
+  geom_hline(yintercept = 0, lty = 2)+
+  geom_hline(yintercept = 1, lty = 3)+
+  geom_hline(yintercept = -1, lty = 3)+
+  facet_grid(vars(fsite), vars(comp), scales = "free")+
+  ggpubr::stat_pvalue_manual(data = test.df[test.df$label == "*",], label = "label",
+                     y.position = 1.25, 
+                     label.size = 6)
+
+```
+
+Making sense out of the above:
+
+* Insects shape community composition at all elevations.
+Fungi affected community composition only in Yawan.
+Fungi increase the species richness, without affecting any other characteristics of the community (beta-diversity components).
+Whereas in Wanang despite having effect on richness and density the effects on the community composition was not visible possibly because of high baseline dissimilarity (variability in community composition) between sites.
+NUmba seems to be specific.
+Despite the effect being present they are being balanced in a way that left the community characteristics unchanged.
+
+* Insect herbivores (P, I, H combined) can suppress plants in communities with low variability in species composition, which allows for new species to establish.
+Moreover, **insects reduce dissimilarity for highly variable communities** (is this true?).
+Exclusion of top predators has the same effect as insect exclusion on species turnover in Numba and Yawan.
+This is a probable result for a strong top-down control of the tri trophic interactions.
+Addition of generalist herbivores also correlated with the predator effects.
+If the effects of insects are mainly caused by specialists, then generalists interfering with their feeding could have the same results.
+But the effects can be more subtle than those of herbivore removal.
+
+### Trait based analyses
+
+In this section we focus on plant species traits and how these can help us predict the performance of individual species at a given plot.
+First we attempt to predict simply whether a species, considering it's traits, will be lost, gained or will remain in various treatment plots.
+We used random forest to investigate the importance of traits in predicting the species faith.
+Prediction accuracy and classifier_RF will tell us how good well the set of traits is (and which traits are most important in predicting faith) for the faith of individuals.
+
+First we analyze only traits ignoring location. Later we can mak an attempt to see wether importance of
+
+```{r}
+
+source("code/pgls_data_prep.R")
+
+subsetVars <- c("c", "i")
+
+# 1. Prepare data ----
+# Add richness
+richdat <- data %>%
+  group_by(site, garden, treatment) %>%
+  summarise(richness = n()) %>%
+  mutate(sgt = paste(site,garden,treatment, sep = "_"))
+
+data.woody.brut <- data.woody.brut %>%
+  mutate(sgt = paste(site,garden,treatment, sep = "_"))
+
+dwbRich <- merge(data.woody.brut, richdat, by = "sgt")
+
+# Filter and add estimated leaf area
+dwbRichF <- dwbRich %>%
+  mutate(est.leaf.area = num_leaf_weight * sla.w.m2kg1,
+         lsla.d = log(sla.d.m2kg1),
+         ltb = log(as.numeric(tot_bio)),
+         labu = log(as.numeric(no_stems+1))) %>%
+  dplyr::select(site.x, garden.x, treatment.x, 
+         unified_names, tot_bio, no_stems, 
+         perc.herb, water.cont, est.leaf.area, sla.d.m2kg1,
+         richness, ltb, lsla.d, labu) %>%
+  filter(water.cont > 0)
+
+names(dwbRichF) <- c("site", "garden", "treatment", "unified_names","tot_bio",
+                     "no_stems", "perc.herb", "water.cont", "est.leaf.area",
+                     "sla.d.m2kg1", "richness", "ltb","lsla.d","labu")
+
+# Subset a data-set to only control and treatment plots
+subDat <- dwbRichF %>%
+  filter(treatment %in% subsetVars)
+
+lgsData <- data.frame()
+
+for (st in unique(subDat$site)){
+  for (gard in unique(subDat$garden)){
+
+    stGardDat <- subDat %>%
+            filter(site == st,
+                   garden == gard)
+
+    # Some quality control
+    if(dim(stGardDat)[1] == 0){
+      next
+      }
+    if(length(unique(stGardDat$treatment)) < 2){
+      next
+      }
+
+    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
+    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
+
+    spLost <- specInCtr[!(specInCtr %in% specInTrt)]
+    spGained <- specInTrt[!(specInTrt %in% specInCtr)]
+
+    stGardDat$pg.cat <- ifelse(stGardDat$unified_names %in% spLost, "lost",
+                               ifelse(stGardDat$unified_names %in% spGained, "gained", "stayed"))
+
+    lgsData <- rbind(lgsData, stGardDat)
+
+  }
+}
+```
+
+I perform the sane analyses for each treatment.
+
+```{r}
+
+# Function that performs RF analysis for a given treatment
+
+# It queries the lgsData that needs to be preapred using separate code (as above)
+
+getLgsData <- function (comp.trt = "i", data) {
+
+  # This function queries the data to compute log
+
+}
+
+# Returns:
+# 1) Classifier RF table
+# 2) Confusion matrix
+# 3) Regression tree example graph
+
+runRF <- function (lgsData) {
+
+}
+
+```
+
+```{r}
+# Random Forest
+# Loading package
+library(caTools)
+library(randomForest)
+
+rfData <- lgsData %>%
+  mutate(pg.cat = as.factor(pg.cat),
+         lest.leaf.area = log(est.leaf.area),
+         lstem = log(no_stems)) %>%
+  dplyr::select(pg.cat, lstem, perc.herb, water.cont,
+                lest.leaf.area, richness, ltb, lsla.d, treatment)
+
+# Try standardized, non-correlated variables
+rfDataRF <- rfData %>%
+  filter(!is.infinite(lest.leaf.area)) %>%
+  dplyr::select(pg.cat, perc.herb, water.cont,
+                richness, ltb, lsla.d)
+
+# Standardized
+rfDataRF[,-1] <- scale(rfDataRF[,-1])
+
+# Splitting data in train and test data
+split <- sample.split(rfDataRF, SplitRatio = 0.7)
+train <- subset(rfDataRF, split == "TRUE")
+test <- subset(rfDataRF, split == "FALSE")
+
+classifier_RF = randomForest(x = train[-1],
+                             y = train$pg.cat,
+                             ntree = 500)
+
+classifier_RF
+
+# Predicting the Test set results
+y_pred = predict(classifier_RF, newdata = test[-1])
+
+# Variable importance plot
+varImpPlot(classifier_RF)
+```
+
+Below is a confusion matrix for the RF
+
+```{r, results='asis'}
+# Confusion Matrix
+confusion_mtx = table(test[, 1], y_pred)
+print(knitr::kable(confusion_mtx))
+```
+
+Using collected variables we cannot predict the faith of an individual species with high certainty.
+We can only predict which plants can be observed (stayed) at both control and insecticide treated plots.
+Error of that prediction is relatively low (0.16).
+For this prediction total biomass is the most important.
+SLA is least important, together with the number of stems.
+To visualize the division we run a single regression tree.
+
+```{r, fig.width = 25, fig.height= 10}
+
+library(rpart)
+library(rpart.plot)
+
+# Regression C50
+library(C50)
+library(printr)
+
+C50Fit <- C5.0(pg.cat ~., data=train,
+               control = C5.0Control(minCases = 5))
+
+C50pred <- predict(object=C50Fit, 
+                   newdata=test, 
+                   type="class")
+
+plot(C50Fit, cex = 0.5)
+
+```
+
+Lets see how good of a job the regression tree did.
+
+```{r, results='asis'}
+
+print(knitr::kable(table(C50pred, test$pg.cat)))
+
+```
+
+The above tree is just a one example of hundreds of averaged trees used in the random forest.
+A few key conclusions can be drawn from the above analyses is that we cannot in certain predict the faith of a tree species.
+But its biomass is the best predictor together with water content.
+SLA seems to be the least important.
+
+Let us now move to test SLA ability to predict the biomass of an individual.
+This is a bit tricky because we will use SLA estimated at the end of the experiment.
+From the leaf frame analyses that spanned all plant species in all treatment plots we saw high plasticity of that trait.
+
+## Trait-based predictions of a species success
+
+### At the level of the whole community
+
+```{r}
+# CWM vs richness
+```
+
+Community-weighted mean SLA (CWM SLA) was higher at low diversity plots.
+This means that species with RELATIVELY higher SLA were the ones that dominated the communities.
+In Numba baseline SLA was reduced even further.
+In Numba also fungicide treatment had lower baseline SLA.
+
+In both cases slope of the relationship was increased (less steep, it was flatter, no relationship).
+
+```{r, fig.cap='Community weighted mean vs number of species at a site for the studied treatments at three elevations. Size of a point indicates total plant biomass of that plot.'}
+
+# SLA values of individual plant species for each treatment.
+
+# I am using the data.woody.brut data-set from the pgls_data_prep script. It has
+# filtered woody palnt names, cleaned sla.d values and limited sla range.
+
+# Define CWM function
+cwmFun <- function(sla.d.m2kg1, tot_bio){
+  # summarize function: CWM for sites
+  return(mean(sla.d.m2kg1*(tot_bio/sum(tot_bio, na.rm = T)), na.rm = T))
+}
+
+# siteSLA <- data.woody.brut %>%
+#   filter(unified_names %in% trait.dat[trait.dat$Life.form == "woody", ]$dsName) %>%
+#   mutate(tot_bio = ifelse(is.infinite(tot_bio), NA, tot_bio),
+#          sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1)) %>%
+#   group_by(site, garden, treatment) %>%
+#   summarise(cwm.sla = cwmFun(sla.d.m2kg1, n.tot_bio),
+#             sp.rich = n())
+
+siteSLA <- data.woody.brut %>%
+  filter(treatment %in% c('i','c','p','h', 'f')) %>%
+  mutate(tot_bio = ifelse(is.infinite(tot_bio), NA, tot_bio),
+         sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1)) %>%
+  group_by(site, garden, treatment) %>%
+  summarise(cwm.sla = cwmFun(sla.d.m2kg1, n.tot_bio),
+            sp.rich = n(),
+            total.bio = sum(n.tot_bio))
+
+siteSLAdf <- as.data.frame(siteSLA)
+
+rownames(siteSLAdf) <- paste(substr(siteSLA$site,1,1), 
+                             siteSLA$garden, 
+                             "_", 
+                             siteSLA$treatment, sep = "")
+
+# Plot of CWM vs species richness
+require(scales)
+p1 <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.sla, color = treatment))+
+  geom_point(aes(size = total.bio)) + 
+  scale_x_continuous(trans = 'log2',
+    breaks = trans_breaks("log2", function(x) 2^x),
+    labels = trans_format("log2", math_format(2^.x))) +
+  scale_y_continuous(trans = 'log2',
+    breaks = trans_breaks("log2", function(x) 2^x),
+    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()
+
+p2 <- ggplot(siteSLAdf, aes(x = treatment, y = cwm.sla, fill = site))+
+  geom_boxplot()+
+  scale_y_continuous(trans = 'log2',
+    breaks = trans_breaks("log2", function(x) 2^x),
+    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()
+
+p3 <- ggplot(siteSLAdf, aes(x = site, y = cwm.sla, fill = treatment))+
+  geom_boxplot()+
+  scale_y_continuous(trans = 'log2',
+    breaks = trans_breaks("log2", function(x) 2^x),
+    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()
+
+ggpubr::ggarrange(p1,p2, p3)
+
+# What is that value for the Y, g8, p?
+# This seem to be an outlier - very low total biomass
+# data.woody.brut %>% filter(site == 'yawan',
+#                            garden == 'g8',
+#                            treatment == 'p')
+
+```
+```{r}
+# Test for the relationship and the differences
+require(sjPlot)
+cwm.lm <- lm(log(cwm.sla)~log(sp.rich)*treatment + site, data = siteSLAdf)
+cwm.lm.nosite <- lm(log(cwm.sla)~log(sp.rich)*treatment, data = siteSLAdf)
+cwm.lm.nosite.noint <- lm(log(cwm.sla)~log(sp.rich)+treatment, data = siteSLAdf)
+cwm.lm.onlysp <- lm(log(cwm.sla)~log(sp.rich), data = siteSLAdf)
+
+# cwm.glm <- glm(cwm.sla~sp.rich*treatment + site, data = siteSLAdf,
+#              family = gaussian(link = 'log'))
+# cwm.glm.nosite <- glm(cwm.sla~sp.rich*treatment, data = siteSLAdf,
+#                     family = gaussian(link = 'log'))
+# cwm.glm.nosite.noint <- glm(cwm.sla~sp.rich+treatment, data = siteSLAdf, 
+#                           family = gaussian(link = 'log'))
+# cwm.glm.onlysp <- glm(cwm.sla~sp.rich, data = siteSLAdf, family = gaussian(link = 'log'))
+
+# tab_model(cwm.lm.nosite.noint)
+
+# anova(cwm.lm.nosite, cwm.lm.onlysp)
+knitr::kable(AIC(cwm.lm, cwm.lm.nosite, cwm.lm.nosite.noint, cwm.lm.onlysp))
+# knitr::kable(AIC(cwm.glm, cwm.glm.nosite, cwm.glm.nosite.noint, cwm.glm.onlysp,
+                 # cwm.lm, cwm.lm.nosite, cwm.lm.nosite.noint, cwm.lm.onlysp))
+
+# tab_model(cwm.glm)
+tab_model(cwm.lm.onlysp)
+
+# plot(cwm.glm)
+# plot(cwm.lm)
+
+# there are no differences
+
+```
+
+The CWM and species number relation is strong and does not change between the treatments and sites.
+
+### Individual species performance
+
+Our original idea was that SLA (approximation of a growth rate) should become a better predictor of plant success (fitness, measured as total biomass in a plot, or positive change in biomass - deltaBio) with elevation.
+
+## SLA
+
+### Can SLA predict change in composition?
+
+SLA doesn’t predict RDA position along C->I, C->H, C->P, C -> F axes, at any elevation.
+
+```{r}
+
+# Figure
+
+```
+
+Community composition change under a treatment may be independent of SLA.
+[key indicator of functionally important aspects of photosynthetic capacity, and is positively correlated with ecosystem productivity (White et al., 2000, Madani et al., 2017, Anderson et al., 2020).
+Reflects the abilities of light capture and adaption to environments of plants (Gao, Wang, and Zhang 2022)]
+
+### Can SLA predict 'success' of a plant species?
+
+In control plots with increasing elevation we hypothesize that SLA (used here as a proxy for the growth rate) should predict species' success (fitness, dominance, change in biomass in response to the treatment) better at higher elevation.
+Specifically, we expect that plants with higher SLA should be characterized by greater ability to dominate the community (higher total biomass - or should it be relative biomass?
+Or should it be biomass change?) in treatments where natural enemies were removed.
+We assume that in these cases inter-specific plant competition should be driven mainly by growth rates.
+
+However there are a few contingent effects that make interpretation of this relationship problematic:
+
+1.
+SLA of a species may be *determined by the community in which it resides*.
+CWM and richness dependency points to that.
+However, at the *individual level we do not see that species generally change their SLA in response to richness of the plot* on which they reside.
+This suggests that increased CWM SLA at low richness plots is driven by dominance of high SLA plants rather than changes in SLA by the same dominants, as would be suggested by the lack of differences in the community composition.
+This lack of significant effect of treatments in the RDA analysis may in turn be caused by a large variation between gardens (high variability in the control sites).
+
+```{r}
+
+# SLA of selected species at different treatments.
+
+# Sla of selected species vs species richness
+
+```
+
+2. From our data we see that at some treatments same species may have different average SLA
+
+3.
+Change in biomass (delta biomass) may be a better measurement of a species' response to a treatment.
+This excludes also species that were lost or gained as a response to a treatment.
+However, it was shown in the previous analysis that these plants were rather small and either in poor condition or just got established at a plot.
+
+4. Change in biomass may be correlated to change in SLA (LRR like plots: what are the results of this?)
+
+```{r}
+
+# Change in SLA vs change in Biomass
+
+```
+
+5. Biomass/$\Delta$Biomass may be smaller if there are many individuals of the same species around. Therefore I think we should **include also the effect of the co-specifics’ abundance**.
+
+6.
+Additionally total biomass may influence the biomass change, i.e, change in biomass may be smaller for smaller plants and larger for larger plants.
+What does this measure mean?
+Delta BIO measures how dominance has changed.
+If there was a dominant plant in the control and its biomass decreased a lot in the treatment what that would mean?
+
+Taking the above considerations into account we formulated the final statistical model.
+The response variable was the $\Delta$Biomass.
+We tested whether it is dependent on SLA but also include the corresponding *site richness* and the *abundance of co-specifics* as co-variate.
+To test treatment influence on $\Delta$Biomass vs SLA relationship we included the interaction between SLA and treatment variable.
+We repeated model for each site and evaluated the model performance to compare power of SLA to predict the biomass changes at each elevation.
+We added a quadratic term for the Phylogenetic correlations are preserved:
+
+#### Complex model model
+$\Delta$Biomass ~ 
+SLA $\times$ site + SLA^2 $\times$ site + 
+water.content $\times$ site + water.content^2 $\times$ site + 
+perc.herb $\times$ site  + perc.herb^2 $\times$ site + 
+richness + abundance + $\Delta$SLA + 
+(Phylogenetic structure).
+
+Random effect is not necessary because we have paired comparisons of control vs treatment within the same garden.
+**Another option** is to include site as a random factor and get rid of all the interaction terms.
+
+This model needs to be applied separately for each treatment comparison.
+
+First we analyse the focused model, where we use only SLA as a predictor of species' success.
+Further we compare the models' performance between sites to explore the idea that SLA is an increasingly better predictor of a species responses with elevations.
+
+#### 1.1 Insecticide
+
+```{r}
+# Create data-set for a given treatment
+deltaData <- data.frame()
+
+# Change total bio variable into numeric
+lgsData$tot_bio <- as.numeric(lgsData$tot_bio)
+
+# Loop through each species in each garden and site and calculate their deltaBio
+for (st in unique(lgsData$site)){
+
+  for (gard in unique(lgsData$garden)){
+
+    stGardDat <- lgsData %>%
+      filter(site == st,
+             garden == gard)
+
+    if (dim(stGardDat)[1] == 0){
+      # If subset is empty skip to another one
+      next
+    }
+
+    if (length(unique(stGardDat$treatment)) < 2){
+      # If there is species in only one treatment skip to another one
+      next
+    }
+
+    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
+    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
+    specInBoth <- specInCtr[specInCtr %in% specInTrt]
+
+    cDat <- stGardDat %>%
+      filter(unified_names %in% specInBoth,
+             treatment == subsetVars[1]) %>%
+      arrange(unified_names)
+    tDat <- stGardDat %>%
+      filter(unified_names %in% specInBoth,
+             treatment == subsetVars[2]) %>%
+      arrange(unified_names)
+
+    # Calculate LRR for biomass and SLA change
+
+    cDat$comparison <- subsetVars[2]
+
+    if (subsetVars[2] == "h"){
+      cDat$lrrtb <- log(tDat$tot_bio/cDat$tot_bio)
+      cDat$lrrsla <- log(tDat$sla.d.m2kg1/cDat$sla.d.m2kg1)
+    } else{
+      cDat$lrrtb <- log(cDat$tot_bio/tDat$tot_bio)
+      cDat$lrrsla <- log(cDat$sla.d.m2kg1/tDat$sla.d.m2kg1)
+    }
+
+    deltaData <- rbind(deltaData, cDat)
+
+  }
+}
+
+# Log transform estimated leaf area
+deltaData <- deltaData %>%
+          mutate(lest.leaf.area = log(est.leaf.area))%>%
+  filter(!is.infinite(lest.leaf.area))
+
+# The Focused model
+# Only the predictive power of SLA, water content and percentage herbivory on deltaBio
+
+# Wanang ----
+wdata <- deltaData %>% 
+  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
+  filter(site == "wanang")
+
+dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d,
+                     data = wdata,
+                     method = "ML",
+                     correlation = corMartins(2.718282, 
+                                              phy = myTree.rand.poly, 
+                                              form = ~unified_names,
+                                              fixed = T),
+                     na.action = na.omit)
+
+# Model results
+
+# Predict values
+pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.w, wdata, 
+                                       se.fit = T)
+wdata <- wdata %>%
+  mutate(preds = pred.vals$fit,
+         se.preds = pred.vals$se.fit,
+         se.upr = preds + se.preds,
+         se.lwr = preds - se.preds)
+
+# print the MSE
+mse_focused_df <- data.frame(mse = with(wdata, mean((lrrtb - preds))^2),
+                             treatment = "Insecticide", site = "Wanang")
+
+# Numba ----
+ndata <- deltaData %>% 
+  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
+  filter(site == "numba")
+
+dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d,
+                     data = ndata,
+                     method = "ML",
+                     correlation = corMartins(2.718282, 
+                                              phy = myTree.rand.poly, 
+                                              form = ~unified_names,
+                                              fixed = T),
+                     na.action = na.omit)
+
+# Predict values
+pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.n, ndata, 
+                                       se.fit = T)
+ndata <- ndata %>%
+  mutate(preds = pred.vals$fit,
+         se.preds = pred.vals$se.fit,
+         se.upr = preds + se.preds,
+         se.lwr = preds - se.preds)
+
+# print the MSE
+mse_focused_df <- rbind(mse_focused_df,  
+                        data.frame(mse = with(ndata, mean((lrrtb - preds))^2),
+                                   treatment = "Insecticide", site = "Numba"))
+
+# Yawan ----
+ydata <- deltaData %>% 
+  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
+  filter(site == "yawan")
+
+# The model
+dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d,
+                     data = ydata,
+                     method = "ML",
+                     correlation = corMartins(2.718282, 
+                                              phy = myTree.rand.poly, 
+                                              form = ~unified_names,
+                                              fixed = T),
+                     na.action = na.omit)
+
+# Predict values
+pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.y, ydata, 
+                                       se.fit = T)
+ydata <- ydata %>%
+  mutate(preds = pred.vals$fit,
+         se.preds = pred.vals$se.fit,
+         se.upr = preds + se.preds,
+         se.lwr = preds - se.preds)
+
+# print the MSE
+mse_focused_df <- rbind(mse_focused_df,  
+                        data.frame(mse = with(ydata, mean((lrrtb - preds))^2),
+                                   treatment = "Insecticide", site = "Yawan"))
+```
+
+```{r, results='asis'}
+library(sjPlot)
+
+tab_model(dd.gls.focsed.w, dd.gls.focsed.n, dd.gls.focsed.y)
+
+```
+
+```{r, fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}
+
+# Plot the relationship and add predicted values
+wpredPlot <- ggplot(wdata, aes(y = lrrtb, x = lsla.d))+
+  geom_point() +
+  geom_line(data = wdata, aes(x = lsla.d, y = preds), lwd = 1,
+            lty = 2)+
+  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) +
+  ggtitle("Wanang 200 m a.s.l.")
+
+npredPlot <- ggplot(ndata, aes(y = lrrtb, x = lsla.d))+
+  geom_point() +
+  geom_line(data = ndata, aes(x = lsla.d, y = preds), 
+            lwd = 1, lty = 2)+
+  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
+  ggtitle("Numba 750 m a.s.l.")
+
+ypredPlot <- ggplot(ydata, aes(y = lrrtb, x = lsla.d))+
+  geom_point() +
+  geom_line(data = ydata, aes(x = lsla.d, y = preds), lwd = 1)+
+  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
+  ggtitle("Yawan 1900 m a.s.l.")
+
+ggpubr::ggarrange(wpredPlot, npredPlot, ypredPlot,
+                  nrow = 1)
+
+```
+
+```{r}
+# Plot the relationship and add predicted values
+ggplot(rbind(ydata, ndata, wdata), aes(y = lrrtb, x = lsla.d, color = site))+
+  geom_point() + 
+  geom_line(data = rbind(ydata, ndata, wdata), 
+            aes(x = lsla.d, y = preds), lwd = 1)+
+  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15)
+
+sjPlot::tab_df(mse_focused_df)
+
+```
+
+The above shows that focused model, where we used SLA as the only predictor of the plant change in biomass has a declining ability to predict change in biomass at the insecticide plot.

### 2023-03-15 code/git_setup.R

@@ -1,5 +1,7 @@
 # PAT 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
-# usethis::create_github_token()
-# gitcreds::gitcreds_set()
+usethis::create_github_token()
+gitcreds::gitcreds_set()
+
+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 
 
 # https://happygitwithr.com/https-pat.html

### 2023-03-15 code/pgls_data_prep.R

@@ -1,0 +1,157 @@
+# 0. Data preparation (move to a separate script) ----
+library(dplyr)
+library(lme4)
+library(lmerTest)
+library(MASS)
+library(ggplot2)
+library(EnvStats)
+library(ape)
+
+## 0.1 Read the data with traits ----
+
+### Biomass data
+data <- read.csv("data/main.data.w.traits.csv")     
+### Type data (woody/non-woody)
+trait.dat <- read.csv("data/suppl_plant_traits.csv")
+### Phylogenetic tree
+myTree <- read.tree('data/phylogeny/V_Phylomaker_tree.txt')
+
+### Convert labels of the phylogenetic tree to lower case
+myTree$tip.label <- tolower(myTree$tip.label)
+
+# Randomly resolve polytomies
+myTree.rand.poly <- multi2di(myTree)
+
+### Convert biomass data into numerical values (number format from European to US)
+data$tot_bio <- gsub(",", ".", data$tot_bio) # substitute commas to dots
+data$n.tot_bio <- as.numeric(data$tot_bio) # woody-non woody data
+
+## 0.2 Filter only woody plants and try to clean up the SLA values ----
+woody.names <- trait.dat[trait.dat$Life.form == "woody",]$dsName
+data.woody <- data %>%
+  filter(unified_names %in% woody.names)
+
+### Check the area lost histograms
+# ggplot(data.woody, aes(log(sla.d.m2kg1), colour = site))+
+#   geom_histogram()
+
+## 0.3 Change inf values to NAs ----
+data.woody[is.infinite(data.woody$sla.d.m2kg1), ]$sla.d.m2kg1 <- NA
+
+
+replaceWithMedian <- function(data.woody, name = "piper_recessum", threshold = 100){
+  
+  # This function replaces missing values for a given species with
+  # a SLA median for that species.
+  # It returns data frame with the changes.
+  
+  # Subset the data
+  piper.dat <- data.woody %>% filter(unified_names == name)
+  print("step 1")
+  
+  # Change extremes and NA's into a meadian value
+  new.sla.val <- median(piper.dat$sla.d.m2kg1, na.rm = T)
+  
+  if(length(piper.dat$sla.d.m2kg1) == 1){
[…]

### 2023-03-24 ms1_data_analysis.Rmd
Before introducing FUNGICIDE treatment analyses.

@@
 
-#### Complex model model
-$\Delta$Biomass ~ 
-SLA $\times$ site + SLA^2 $\times$ site + 
-water.content $\times$ site + water.content^2 $\times$ site + 
-perc.herb $\times$ site  + perc.herb^2 $\times$ site + 
-richness + abundance + $\Delta$SLA + 
-(Phylogenetic structure).
-
-Random effect is not necessary because we have paired comparisons of control vs treatment within the same garden.
-**Another option** is to include site as a random factor and get rid of all the interaction terms.
-
-This model needs to be applied separately for each treatment comparison.
-
-First we analyse the focused model, where we use only SLA as a predictor of species' success.
-Further we compare the models' performance between sites to explore the idea that SLA is an increasingly better predictor of a species responses with elevations.
-
-#### 1.1 Insecticide
+Explore that visually
+
+#### Insecticide
+
+##### Change in Biomass: log-respose ratio of a species' biomass
 
 ```{r}
+
+# Get the delta data ready for th e given subset vars (comparison)
+
+subsetVars <- c("c", "i")
+
 # Create data-set for a given treatment
@@
 
+```
+
+Below I explore nonlinear relationships of the biomass log-ratios with species traits. Nonlinear relationships are introduced to allow for optimum trait values.
+
+```{r}
+
+# Log-response ratios of total biomass vs of sla
+
+# DeltaBiomass vs SLA at each site
+slaPlot <- ggplot(deltaData, aes(x = lsla.d, y = lrrtb, color = site)) +
+  geom_point()+
+  geom_smooth(method = "lm", formula = y ~ x + I(x^2))
+
+# There might be some effect of site.
+# DeltaBiomass vs SLA at each site
+wcPlot <- ggplot(deltaData, aes(x = water.cont, y = lrrtb, color = site)) +
+  geom_point()+
+  geom_smooth(method = "lm", formula = y ~ x + I(x^2))
+# Seems like the effect is not very strong either
+
+herbPlot <- ggplot(deltaData, aes(x = perc.herb, y = lrrtb, color = site)) +
+  geom_point()+
+  geom_smooth(method = "lm", formula = y ~ x + I(x^2)) + 
+  scale_x_continuous(trans = 'log2')
+
+ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+It seems that in some cases the nonlinear relationship may be important.
+Generally water content looks linear for each elevation.
+We will include this and build also a model that models log-ratio biomass vs water content as a linear relationship.
+We arrive to the final definition of the full model.
+
+$\Delta$Biomass ~ SLA $\times$ site + SLA^2 $\times$ site + 
+water.content $\times$ site + water.content^2 $\times$ site + 
+perc.herb $\times$ site  + perc.herb^2 $\times$ site + 
+richness + abundance + $\Delta$SLA + 
+(Phylogenetic structure).
+
+```{r}
+simple.sla <- gls(
+  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + richness + labu + lrrsla,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+just.sla <- gls(
+  lrrtb ~ lsla.d*site + I(lsla.d^2)*site,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+AIC(just.sla, simple.sla)
+anova(just.sla, simple.sla) # simple SLA is much bette
+tab_model(simple.sla) # But only the lrr SLA is significant.
+
+# Run the full model.
+full.sla <- gls(
+  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + water.cont*site + I(water.cont^2)*site + perc.herb*site + I(perc.herb^2)*site + richness + labu + lrrsla,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+# Assuming linear relationships
+full.sla.linear <- gls(
+  lrrtb ~ lsla.d*site + water.cont*site + perc.herb*site + richness + labu + lrrsla,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+anova(simple.sla, full.sla.linear) # full SLA linear is much better that the one using SLA
+tab_model(full.sla.linear)
+
+full.sla.linear.noint <- gls(
+  lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla + site,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+anova(full.sla.linear, full.sla.linear.noint)
+# Removing interactions does not improve the model significantly
+tab_model(full.sla.linear.noint)
+
+# Add interaction only to the water content and sla
+full.sla.linear.intslawc <- gls(
+  lrrtb ~ lsla.d*site + water.cont*site + perc.herb + richness + labu + lrrsla + site,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+anova(full.sla.linear.noint, full.sla.linear.intslawc)
+# No effect
+tab_model(full.sla.linear.noint)
+
+full.sla.linear.intsla <- gls(
+  lrrtb ~ lsla.d*site + water.cont + perc.herb + richness + labu + lrrsla + site,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+anova(full.sla.linear.noint, full.sla.linear.intsla)
+# Any for of iteraction is not improving the model
+tab_model(full.sla.linear.noint) # Thus the simplest model seems to be best.
+
+# I further compare to the full.sla.linear.noint model.
+# Only water content is modeled as linear
+full.sla.custom <-  gls(
+  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + water.cont*site + perc.herb*site + I(perc.herb^2)*site + richness + labu + lrrsla,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+anova(full.sla.linear.noint, full.sla.custom)
+
+full.sla.custom.noint <-  gls(
+  lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + perc.herb + I(perc.herb^2) + richness + labu + lrrsla,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+full.sla.linear.noint.nosite <- gls(
+  lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
+                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+  method = "ML"
+  )
+
+# Withouth a site it is even better
+anova(full.sla.linear.noint, full.sla.linear.noint.nosite)
+
+# The linear model remains the best one.
+plot(full.sla.linear.noint.nosite) # This plot looks best
+```
+
+Table. Results of the selected model for the deltaBio.
+
+```{r}
+sjPlot::tab_model(full.sla.linear.noint.nosite)
+```
+
+Nevertheless it is not a strong correlation and does not explain a lot of variation in the biomass change.
+
+```{r, fig.cap="Observed vs predicted values for the lsls.d and "}
+preddata <- data.frame(lsla.d = seq(from = min(deltaData$lsla.d, na.rm = T), 
+                                  to = max(deltaData$lsla.d, na.rm = T), 
+                                  length = 100),
+                       water.cont = mean(deltaData$water.cont, na.rm = T), 
+                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
+                       richness= mean(deltaData$richness, na.rm = T), 
+                       labu= mean(deltaData$labu, na.rm = T),
+                       lrrsla= mean(deltaData$lrrsla, na.rm = T))
+
+preddatalrrsla <- data.frame(lsla.d = mean(deltaData$lsla.d, na.rm = T),
+                       water.cont = mean(deltaData$water.cont, na.rm = T), 
+                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
+                       richness= mean(deltaData$richness, na.rm = T), 
+                       labu= mean(deltaData$labu, na.rm = T),
+                       lrrsla=  seq(from = min(deltaData$lrrsla, na.rm = T), 
+                                  to = max(deltaData$lrrsla, na.rm = T), 
+                                  length = 100))
+
+visData <- deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label)
+
+predvals <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddata,
+                    newdtype = 'response')
+
+preddata$predvals <- predvals$fit
+preddata$predse <- predvals$se.fit
+
+predvalslrrsla <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddatalrrsla,
+                    newdtype = 'response')
+
+preddatalrrsla$predvals <- predvalslrrsla$fit
+preddatalrrsla$predse <- predvalslrrsla$se.fit
+
+# For lsla.d
+p1 <- ggplot(preddata, aes(x = lsla.d, y = predvals)) + 
+  geom_line()+
+  geom_ribbon(aes(ymin = predvals - predse,
+                  ymax = predvals + predse), 
+              alpha = 0.1) + 
+  geom_point(data = visData, aes(x = lsla.d, y = lrrtb)) + 
+  ggtitle("Log(SLA) vs Log-response ratio of a species' biomass")
+
+# For lrrsla
+p2 <- ggplot(preddatalrrsla, aes(x = lrrsla, y = predvals)) + 
+  geom_line()+
+  geom_ribbon(aes(ymin = predvals - predse,
+                  ymax = predvals + predse), 
+              alpha = 0.1) + 
+  geom_point(data = visData, aes(x = lrrsla, y = lrrtb)) +
+  ggtitle("Delta(SLA) vs Log-response ratio of a species' biomass")
+
+# ggplot(visData, aes(x = lsla.d, y = lrrtb)) + 
+#   geom_point(aes(size = tot_bio, alpha = 0.05), pch = 19)+
+#   geom_point(aes(x = lsla.d, y = predvals, col = "red")) + 
+#   geom_smooth(aes(x = lsla.d, y = predvals), method = "lm")
+
+ggpubr::ggarrange(p1,p2)
+
+```
+
+In general there is some evidence for the change in total biomass to be driven by the sla.
+Water content and log-response ratios of SLA are significantly affecting the change in total biomass.
+The change however might be determined by a strong correlation at one site only: Wanang.
+Higher water content results in positive change in biomass of a species, while if we have large increase in SLA of a species that was correlated with reduction of a species' biomass.
+**None of the interactions with site were significant.**
+
+Random effect is not necessary because we have paired comparisons of control vs treatment within the same garden.
+**Another option** is to include site as a random factor and get rid of all the interaction terms.
+
+This model needs to be applied separately for each treatment comparison.
+
+First we analyse the focused model, where we use only SLA as a predictor of species' success.
+Further we compare the models' performance between sites to explore the idea that SLA is an increasingly better predictor of a species responses with elevations.
+
+The model above has the interactive effect of site, but later I perform analyses at each site separately. **Which approach should I choose?**
+
+However, when we analyse the data separately for each elevation we are getting a different results **if we only use the lsla.d variable**.
+The more realistic models could change this relationship, i.e., when we include also the effects of species richness and other traits as we did above.
+
+```{r, include=FALSE}
 # The Focused model
@@
 
-dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d,
+# dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d,
+#                      data = wdata,
+#                      method = "ML",
+#                      correlation = corMartins(2.718282, 
+#                                               phy = myTree.rand.poly, 
+#                                               form = ~unified_names,
+#                                               fixed = T),
+#                      na.action = na.omit)
+dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                      data = wdata,
@@
 
-dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d,
+# dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d,
+#                      data = ndata,
+#                      method = "ML",
+#                      correlation = corMartins(2.718282, 
+#                                               phy = myTree.rand.poly, 
+#                                               form = ~unified_names,
+#                                               fixed = T),
+#                      na.action = na.omit)
+dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                      data = ndata,
@@
 # The model
-dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d,
+# dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d,
+#                      data = ydata,
+#                      method = "ML",
+#                      correlation = corMartins(2.718282, 
+#                                               phy = myTree.rand.poly, 
+#                                               form = ~unified_names,
+#                                               fixed = T),
+#                      na.action = na.omit)
+
+dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
                      data = ydata,
@@
                                    treatment = "Insecticide", site = "Yawan"))
-```
-
-```{r, results='asis'}
+
 library(sjPlot)
@@
 
-```
-
-```{r, fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}
+# fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}
 
@@
                   nrow = 1)
-
-```
-
-```{r}
 # Plot the relationship and add predicted values
@@
 The above shows that focused model, where we used SLA as the only predictor of the plant change in biomass has a declining ability to predict change in biomass at the insecticide plot.
+
+# Fungicide

### 2023-03-28 ms1_data_analysis.Rmd
SLA models for fungicide and herbivore addition.

@@
 # Get the delta data ready for th e given subset vars (comparison)
-
 subsetVars <- c("c", "i")
@@
 ```{r}
+# Model explorations
 simple.sla <- gls(
@@
 
-AIC(just.sla, simple.sla)
-anova(just.sla, simple.sla) # simple SLA is much bette
-tab_model(simple.sla) # But only the lrr SLA is significant.
+# AIC(just.sla, simple.sla)
+# anova(just.sla, simple.sla) # simple SLA is much bette
+# tab_model(simple.sla) # But only the lrr SLA is significant.
 
@@
 
-anova(simple.sla, full.sla.linear) # full SLA linear is much better that the one using SLA
-tab_model(full.sla.linear)
+# anova(simple.sla, full.sla.linear) # full SLA linear is much better that the one using SLA
+# tab_model(full.sla.linear)
 
@@
 
-anova(full.sla.linear, full.sla.linear.noint)
+# anova(full.sla.linear, full.sla.linear.noint)
 # Removing interactions does not improve the model significantly
-tab_model(full.sla.linear.noint)
+# tab_model(full.sla.linear.noint)
 
@@
 
-anova(full.sla.linear.noint, full.sla.linear.intslawc)
+# anova(full.sla.linear.noint, full.sla.linear.intslawc)
 # No effect
-tab_model(full.sla.linear.noint)
+# tab_model(full.sla.linear.noint)
 
@@
 
-anova(full.sla.linear.noint, full.sla.linear.intsla)
+# anova(full.sla.linear.noint, full.sla.linear.intsla)
 # Any for of iteraction is not improving the model
@@
 
-anova(full.sla.linear.noint, full.sla.custom)
+# anova(full.sla.linear.noint, full.sla.custom)
 
@@
 # Withouth a site it is even better
-anova(full.sla.linear.noint, full.sla.linear.noint.nosite)
+# anova(full.sla.linear.noint, full.sla.linear.noint.nosite)
 
 # The linear model remains the best one.
-plot(full.sla.linear.noint.nosite) # This plot looks best
+# plot(full.sla.linear.noint.nosite) # This plot looks best
+# tab_model(full.sla.linear.noint.nosite)
+
 ```
@@
 # Fungicide
+
+```{r}
+subsetVars <- c("c", "f")
+
+source('code/get_data_ready_for_the_sla_analyses.R')
+
+ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+
+```{r}
+
+source('code/model_definitions.R')
+
+# Are traits improving the null model
+
+# Model selection.
+# 1. Full model (quadratic, and site interactions) vs no traits
+# anova(no.traits, full.model)
+# anova(no.traits, full.model.noint)
+# anova(no.traits, all.linear)
+# anova(no.traits, all.linear.noint)
+
+# 2. Individual traits
+
+# SLA
+# Check quadratic
+# anova(no.traits, just.sla.quadratic.noint)
+# Check linear
+# anova(no.traits, just.sla.linear.noint)
+# Check quadratic interation
+# anova(no.traits, just.sla.quadratic)
+# Check linear interaction
+# anova(no.traits, just.sla.linear)
+
+# WATER
+# Check quadratic
+# anova(no.traits, just.water.quadratic.noint)
+# Check linear
+# anova(no.traits, just.water.linear.noint)
+# Check quadratic interation
+# anova(no.traits, just.water.quadratic)
+# Check linear interaction
+# anova(no.traits, just.water.linear)
+
+# HERBIVORY
+# Check quadratic
+# anova(no.traits, just.herb.quadratic.noint)
+# Check linear
+# anova(no.traits, just.herb.linear.noint)
+# Check quadratic interation
+# anova(no.traits, just.herb.quadratic)
+# Check linear interaction
+# anova(no.traits, just.herb.linear)
+
+# 2. Model with all traits linear/quadratic, what combination?
+
+# quadratic SLA, HERB
+# anova(no.traits, q.sla.q.herb)
+# anova(no.traits, q.sla.q.herb.noint)
+# quadratic SLA, linear HERB
+# anova(no.traits, q.sla.l.herb)
+# anova(no.traits, q.sla.l.herb.noint)
+# linear SLA, quadratic HERB
+# anova(no.traits, l.sla.q.herb)
+# anova(no.traits, l.sla.q.herb.noint)
+# linear SLA, linear HERB
+# anova(no.traits, l.sla.l.herb)
+# anova(no.traits, l.sla.l.herb.noint)
+
+# quadratic HERB, WATER
+# anova(no.traits, q.herb.q.water)
+# anova(no.traits, q.herb.q.water.noint)
+# quadratic HERB, linear WATER
+# anova(no.traits, q.herb.l.water)
+# anova(no.traits, q.herb.l.water.noint)
+# linear HERB, quadratic WATER
+# anova(no.traits, l.herb.q.water)
+# anova(no.traits, l.herb.q.water.noint)
+# linear HERB, linear WATER
+# anova(no.traits, l.herb.l.water)
+# anova(no.traits, l.herb.l.water.noint)
+
+# quadratic SLA, WATER
+# anova(no.traits, q.sla.q.water)
+# anova(no.traits, q.sla.q.water.noint)
+# quadratic SLA, linear WATER
+# anova(no.traits, q.sla.l.water)
+# anova(no.traits, q.sla.l.water.noint)
+# linear SLA, quadratic WATER
+# anova(no.traits, l.sla.q.water)
+# anova(no.traits, l.sla.q.water.noint)
+# linear SLA, linear WATER
+# anova(no.traits, l.sla.l.water)
+# anova(no.traits, l.sla.l.water.noint)
+
+# So the best model is the one without the traits!
+tab_model(no.traits)
+
+```
+
+# Additional herbivores
+
+```{r}
+subsetVars <- c("c", "h")
+
+source('code/get_data_ready_for_the_sla_analyses.R')
+
+ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+
+```{r}
+
+source('code/model_definitions.R')
+
+# Are traits improving the null model
+
+# Model selection.
+# 1. Full model (quadratic, and site interactions) vs no traits
+anova(no.traits, full.model) # full model is much better
+anova(full.model, full.model.noint)
+
+# interaction model is as good as the no interaction. Keep the model without the interaction
+
+# 2. Is the quadratic relationship significant
+# anova(full.model.noint, all.linear)            # AIC is lower for the full model noint
+# anova(full.model.noint, all.linear.noint)      # AIC is lower for the full model noint
+
+# Traits have nonlinear relationships with the delta biomass.
+
+# 3. REMOVE SLA
+# 3.1 anova(full.model.noint, q.herb.q.water.noint) # withouth sla is slightly better (AIC)
+# 3.2 anova(full.model.noint, q.herb.q.water)
+# 3.3 anova(full.model.noint, q.herb.l.water.noint) # Full model is than this !!!
+# 3.4 anova(full.model.noint, q.herb.l.water)
+# 3.5 anova(full.model.noint, l.herb.l.water.noint) # Full model still better
+# 3.6 anova(full.model.noint, l.herb.l.water) #
+
+# Including SLA did not mprove the model significantly
+# tab_model(q.herb.q.water.noint)
+
+# But if we were to start first by remving perc.herbivory?
+# anova(full.model.noint, q.sla.q.water.noint) # Model without herbivory is better
+# anova(full.model.noint, q.sla.q.water)       # Full model is better
+# anova(q.sla.q.water, q.sla.q.water.noint)    # No interaction is better
+# anova(full.model.noint, q.sla.l.water.noint)   # Full model is still better
+# anova(full.model.noint, q.sla.l.water)         # Full model wins again
+# anova(full.model.noint, l.sla.l.water.noint)   # And again by AIC criterium
+# anova(full.model.noint, l.sla.l.water)         # And again
+
+# But if we were to start first by remving water.content?
+# anova(full.model.noint, q.sla.q.herb.noint) # Model with herbivory is much better
+# anova(full.model.noint, q.sla.q.herb)       # Full model is better
+# anova(q.sla.q.herb, q.sla.q.herb.noint)    # No interaction is better
+# anova(full.model.noint, q.sla.l.herb.noint)   # Full model is still better
+# anova(full.model.noint, q.sla.l.herb)         # Full model wins again
+# anova(full.model.noint, l.sla.l.herb.noint)   # And again by AIC criterium
+# anova(full.model.noint, l.sla.l.herb)         # And again
+
+# The above suggests that models without hebivory and sla do better job. Water content is the key trait that helps predict the change in biomass in response to the treatment.
+
+# SLA
+# Check quadratic
+# anova(full.model.noint, just.sla.quadratic.noint)
+# Check linear
+# anova(full.model.noint, just.sla.linear.noint)
+# Check quadratic interation
+# anova(full.model.noint, just.sla.quadratic)
+# Check linear interaction
+# anova(full.model.noint, just.sla.linear)
+
+# WATER
+# Check quadratic
+anova(full.model.noint, just.water.quadratic.noint) # Model with just water content
+
+# Is the quadratic form necessary
+anova(just.water.quadratic.noint, just.water.linear.noint) # no difference but quadratic has lower AIC
+
+# So the best model is the one without the traits!
+tab_model(just.water.quadratic.noint)
+
+```

### 2023-03-28 code/get_data_ready_for_the_sla_analyses.R

@@ -1,0 +1,119 @@
+#### ------------------------------------------------------
+# 1. Subset a data-set to only control and treatment plots
+subDat <- dwbRichF %>%
+  filter(treatment %in% subsetVars)
+
+lgsData <- data.frame()
+
+for (st in unique(subDat$site)){
+  for (gard in unique(subDat$garden)){
+    
+    stGardDat <- subDat %>%
+      filter(site == st,
+             garden == gard)
+    
+    # Some quality control
+    if(dim(stGardDat)[1] == 0){
+      next
+    }
+    if(length(unique(stGardDat$treatment)) < 2){
+      next
+    }
+    
+    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
+    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
+    
+    spLost <- specInCtr[!(specInCtr %in% specInTrt)]
+    spGained <- specInTrt[!(specInTrt %in% specInCtr)]
+    
+    stGardDat$pg.cat <- ifelse(stGardDat$unified_names %in% spLost, "lost",
+                               ifelse(stGardDat$unified_names %in% spGained, "gained", "stayed"))
+    
+    
+    lgsData <- rbind(lgsData, stGardDat)
+    
+    
+  }
+}
+
+### --------------------------------------------------------
+# 2. Get the delta data ready for th e given subset vars (comparison)
+
+# Create data-set for a given treatment
+deltaData <- data.frame()
+
+# Change total bio variable into numeric
+lgsData$tot_bio <- as.numeric(lgsData$tot_bio)
+
+# Loop through each species in each garden and site and calculate their deltaBio
+for (st in unique(lgsData$site)){
+  
+  for (gard in unique(lgsData$garden)){
+    
+    stGardDat <- lgsData %>%
+      filter(site == st,
+             garden == gard)
+    
+    
+    if (dim(stGardDat)[1] == 0){
+      # If subset is empty skip to another one
+      next
+    }
+    
+    if (length(unique(stGardDat$treatment)) < 2){
+      # If there is species in only one treatment skip to another one
+      next
+    }
+    
[…]

### 2023-03-28 code/model_definitions.R

@@ -1,0 +1,375 @@
+modelDefinitions <-  function () {
+  
+  # Runs all of model definitions
+  
+  # NO TRAITS
+  no.traits <- gls(
+    lrrtb ~ richness + labu + lrrsla + site,
+    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    method = "ML"
+  )
+  
+  # FULL MODEL
+  full.model <- gls(
+    lrrtb ~ lsla.d*site+ I(lsla.d^2)*site+water.cont*site+I(water.cont^2)*site+perc.herb*site+I(perc.herb^2)*site+richness+labu+lrrsla+site,
+    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    method = "ML"
+  )
+  
+  # FULL MODEL
+  full.model.noint <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) + perc.herb + I(perc.herb^2) + richness + labu + lrrsla + site,
+    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    method = "ML"
+  )
+  
+  all.linear <- gls(
+    lrrtb ~ lsla.d*site + water.cont*site + perc.herb*site + richness + labu + lrrsla + site,
+    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    method = "ML"
+  )
+  
+  all.linear.noint <- gls(
+    lrrtb ~ lsla.d+water.cont+perc.herb+richness+labu+lrrsla+site,
+    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    method = "ML"
+  )
+  
+  # Individual traits-------------------------------------------------
+  
+  ## SLA
+  just.sla.linear.noint <- gls(
+    lrrtb ~ lsla.d + richness + labu + lrrsla + site,
+    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    method = "ML"
+  )
+  
+  just.sla.linear <- gls(
+    lrrtb ~ lsla.d*site + richness + labu + lrrsla + site,
+    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    method = "ML"
+  )
+  
+  just.sla.quadratic.noint <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + richness + labu + lrrsla + site,
[…]

### 2023-03-29 ms1_data_analysis.Rmd
Before introducing the phylogenetic correlation to the deltaBiomass models.

@@
 
-```{r}
+```{r, results='asis'}
 # Model explorations
@@
 
-```{r}
+```{r, results='asis'}
 
@@
 # 1. Full model (quadratic, and site interactions) vs no traits
-anova(no.traits, full.model) # full model is much better
-anova(full.model, full.model.noint)
+# anova(no.traits, full.model) # full model is much better
+# anova(full.model, full.model.noint)
 
@@
 # Check quadratic
-anova(full.model.noint, just.water.quadratic.noint) # Model with just water content
+# anova(full.model.noint, just.water.quadratic.noint) # Model with just water content
 
 # Is the quadratic form necessary
-anova(just.water.quadratic.noint, just.water.linear.noint) # no difference but quadratic has lower AIC
+# anova(just.water.quadratic.noint, just.water.linear.noint) # no difference but quadratic has lower AIC
 
@@
 ```
+
+# Predators
+
+```{r}
+subsetVars <- c("c", "p")
+
+source('code/get_data_ready_for_the_sla_analyses.R')
+
+ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+
+```{r, results='asis'}
+
+source('code/model_definitions.R')
+
+# Are traits improving the null model
+
+# Model selection.
+# 1. Full model (quadratic, and site interactions) vs no traits
+# anova(no.traits, full.model)            # full model is better
+# anova(full.model, full.model.noint)     # No difference, so we can drop the interaction
+
+# interaction model is as good as the no interaction. Keep the model without the interaction
+
+# 2. Is the quadratic relationship significant
+# anova(full.model.noint, all.linear)            # Quadratic is a bit better
+# anova(full.model.noint, all.linear.noint)      # All linear no int is more efficient
+
+# Traits have nonlinear relationships with the delta biomass.
+
+# 3. REMOVE SLA
+
+# anova(full.model.noint, q.herb.q.water.noint) # withouth sla is slightly better (AIC)
+# anova(full.model.noint, q.herb.q.water)
+# anova(full.model.noint, q.herb.l.water.noint) # Linear water and quadratic herbivory works even better
+# anova(full.model.noint, q.herb.l.water) # interactions make it worse
+# anova(full.model.noint, l.herb.l.water.noint) # Full model still better
+# anova(q.herb.l.water.noint, l.herb.l.water.noint) # All linear? Slightly better
+# anova(full.model.noint, l.herb.l.water)
+
+# All linear with interaction seems to work best, but the interaction is not improving the model so I will keep it out.
+
+# Including SLA did not mprove the model significantly
+# tab_model(l.herb.l.water.noint)
+# anova(no.traits, l.herb.l.water.noint)
+
+# But if we were to start first by remving perc.herbivory?
+# anova(full.model.noint, q.sla.q.water.noint) # Model without herbivory is not better
+# anova(full.model.noint, q.sla.q.water)       # Interaction makes it worse
+
+# anova(full.model.noint, q.sla.l.water.noint)   # Full model is still better
+# anova(full.model.noint, q.sla.l.water)         # Full model wins again
+# anova(full.model.noint, l.sla.l.water.noint)   # And again by AIC criterium
+# anova(full.model.noint, l.sla.l.water)         # And again
+
+# But if we were to start first by remving water.content?
+# anova(full.model.noint, q.sla.q.herb.noint) # Model with herbivory is much better
+# anova(full.model.noint, q.sla.q.herb)       # Full model is better
+# anova(q.sla.q.herb, q.sla.q.herb.noint)    # No interaction is better
+# anova(full.model.noint, q.sla.l.herb.noint)   # Full model is still better
+# anova(full.model.noint, q.sla.l.herb)         # Full model wins again
+# anova(full.model.noint, l.sla.l.herb.noint)   # And again by AIC criterium
+# anova(full.model.noint, l.sla.l.herb)         # And again
+
+# The above suggests that models without hebivory and sla do better job. 
+# Water content is again the key trait that helps predict the change in biomass in response to the treatment.
+
+# WATER
+# Check quadratic
+# anova(l.herb.l.water.noint, just.water.quadratic.noint) 
+# Model with just water content is worse than that without herbivores
+
+# Is the quadratic form necessary
+# anova(l.herb.l.water.noint, just.water.linear.noint) # Herbivory is an important trait 
+# anova(l.herb.l.water.noint, just.herb.linear.noint) # Water is an important trait, both should stay in the model
+
+# So the best model is the one without the traits!
+tab_model(l.herb.l.water.noint)
+
+```

### 2023-03-29 code/model_definitions.R

@@ -6,7 +6,12 @@ modelDefinitions <-  function () {
   no.traits <- gls(
     lrrtb ~ richness + labu + lrrsla + site,
     data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-    method = "ML"
+    method = "ML",
+    correlation = corMartins(2.718282, 
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
   )
   
   # FULL MODEL

### 2023-03-30 ms1_data_analysis.Rmd
Before removing the focused model

--- 2023-03-28 ms1_data_analysis.Rmd
@@
 library(dplyr)
+library(sjPlot)
 
@@
 
-##### Change in Biomass: log-respose ratio of a species' biomass
-
 ```{r}
@@
 
-# Create data-set for a given treatment
-deltaData <- data.frame()
-
-# Change total bio variable into numeric
-lgsData$tot_bio <- as.numeric(lgsData$tot_bio)
-
-# Loop through each species in each garden and site and calculate their deltaBio
-for (st in unique(lgsData$site)){
-
-  for (gard in unique(lgsData$garden)){
-
-    stGardDat <- lgsData %>%
-      filter(site == st,
-             garden == gard)
-
-    if (dim(stGardDat)[1] == 0){
-      # If subset is empty skip to another one
-      next
-    }
-
-    if (length(unique(stGardDat$treatment)) < 2){
-      # If there is species in only one treatment skip to another one
-      next
-    }
-
-    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
-    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
-    specInBoth <- specInCtr[specInCtr %in% specInTrt]
-
-    cDat <- stGardDat %>%
-      filter(unified_names %in% specInBoth,
-             treatment == subsetVars[1]) %>%
-      arrange(unified_names)
-    tDat <- stGardDat %>%
-      filter(unified_names %in% specInBoth,
-             treatment == subsetVars[2]) %>%
-      arrange(unified_names)
-
-    # Calculate LRR for biomass and SLA change
-
-    cDat$comparison <- subsetVars[2]
-
-    if (subsetVars[2] == "h"){
-      cDat$lrrtb <- log(tDat$tot_bio/cDat$tot_bio)
-      cDat$lrrsla <- log(tDat$sla.d.m2kg1/cDat$sla.d.m2kg1)
-    } else{
-      cDat$lrrtb <- log(cDat$tot_bio/tDat$tot_bio)
-      cDat$lrrsla <- log(cDat$sla.d.m2kg1/tDat$sla.d.m2kg1)
-    }
-
-    deltaData <- rbind(deltaData, cDat)
-
-  }
-}
-
-# Log transform estimated leaf area
-deltaData <- deltaData %>%
-          mutate(lest.leaf.area = log(est.leaf.area))%>%
-  filter(!is.infinite(lest.leaf.area))
+source('code/get_data_ready_for_the_sla_analyses.R')
+
+ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+# # Create data-set for a given treatment
+# deltaData <- data.frame()
+# 
+# # Change total bio variable into numeric
+# lgsData$tot_bio <- as.numeric(lgsData$tot_bio)
+# 
+# # Loop through each species in each garden and site and calculate their deltaBio
+# for (st in unique(lgsData$site)){
+#   
+#   for (gard in unique(lgsData$garden)){
+#     
+#     stGardDat <- lgsData %>%
+#       filter(site == st,
+#              garden == gard)
+#     
+#     
+#     if (dim(stGardDat)[1] == 0){
+#       # If subset is empty skip to another one
+#       next
+#     }
+#     
+#     if (length(unique(stGardDat$treatment)) < 2){
+#       # If there is species in only one treatment skip to another one
+#       next
+#     }
+#     
+#     specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
+#     specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
+#     specInBoth <- specInCtr[specInCtr %in% specInTrt]
+#     
+#     cDat <- stGardDat %>%
+#       filter(unified_names %in% specInBoth,
+#              treatment == subsetVars[1]) %>%
+#       arrange(unified_names)
+#     tDat <- stGardDat %>%
+#       filter(unified_names %in% specInBoth,
+#              treatment == subsetVars[2]) %>%
+#       arrange(unified_names)
+#     
+#     # Calculate LRR for biomass and SLA change
+#     
+#     cDat$comparison <- subsetVars[2]
+#     
+#     if (subsetVars[2] == "h"){
+#       cDat$lrrtb <- log(tDat$tot_bio/cDat$tot_bio)
+#       cDat$lrrsla <- log(tDat$sla.d.m2kg1/cDat$sla.d.m2kg1)
+#     } else{
+#       cDat$lrrtb <- log(cDat$tot_bio/tDat$tot_bio)
+#       cDat$lrrsla <- log(cDat$sla.d.m2kg1/tDat$sla.d.m2kg1)
+#     }
+#     
+#     deltaData <- rbind(deltaData, cDat)
+#     
+#   }
+# }
+# 
+# 
+# # Log transform estimated leaf area
+# deltaData <- deltaData %>%
+#           mutate(lest.leaf.area = log(est.leaf.area))%>%
+#   filter(!is.infinite(lest.leaf.area))
 
@@
 
-```{r}
-
-# Log-response ratios of total biomass vs of sla
-
-# DeltaBiomass vs SLA at each site
-slaPlot <- ggplot(deltaData, aes(x = lsla.d, y = lrrtb, color = site)) +
-  geom_point()+
-  geom_smooth(method = "lm", formula = y ~ x + I(x^2))
-
-# There might be some effect of site.
-# DeltaBiomass vs SLA at each site
-wcPlot <- ggplot(deltaData, aes(x = water.cont, y = lrrtb, color = site)) +
-  geom_point()+
-  geom_smooth(method = "lm", formula = y ~ x + I(x^2))
-# Seems like the effect is not very strong either
-
-herbPlot <- ggplot(deltaData, aes(x = perc.herb, y = lrrtb, color = site)) +
-  geom_point()+
-  geom_smooth(method = "lm", formula = y ~ x + I(x^2)) + 
-  scale_x_continuous(trans = 'log2')
-
-ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
-
-```
 It seems that in some cases the nonlinear relationship may be important.
@@
 
-```{r}
+```{r, results='asis'}
+
 # Model explorations
@@
 Table. Results of the selected model for the deltaBio.
-
 ```{r}
@@
 # Wanang ----
-wdata <- deltaData %>% 
-  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
-  filter(site == "wanang")
+# wdata <- deltaData %>% 
+#   filter(unified_names %in% myTree.rand.poly$tip.label) %>%
+#   filter(site == "wanang")
 
@@
 #                      na.action = na.omit)
-dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
-                     data = wdata,
-                     method = "ML",
-                     correlation = corMartins(2.718282, 
-                                              phy = myTree.rand.poly, 
-                                              form = ~unified_names,
-                                              fixed = T),
-                     na.action = na.omit)
+# dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
+#                      data = wdata,
+#                      method = "ML",
+#                      correlation = corMartins(2.718282, 
+#                                               phy = myTree.rand.poly, 
+#                                               form = ~unified_names,
+#                                               fixed = T),
+#                      na.action = na.omit)
 
@@
 # Predict values
-pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.w, wdata, 
-                                       se.fit = T)
-wdata <- wdata %>%
-  mutate(preds = pred.vals$fit,
-         se.preds = pred.vals$se.fit,
-         se.upr = preds + se.preds,
-         se.lwr = preds - se.preds)
-
-# print the MSE
-mse_focused_df <- data.frame(mse = with(wdata, mean((lrrtb - preds))^2),
-                             treatment = "Insecticide", site = "Wanang")
+# pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.w, wdata, 
+#                                        se.fit = T)
+# wdata <- wdata %>%
+#   mutate(preds = pred.vals$fit,
+#          se.preds = pred.vals$se.fit,
+#          se.upr = preds + se.preds,
+#          se.lwr = preds - se.preds)
+# 
+# # print the MSE
+# mse_focused_df <- data.frame(mse = with(wdata, mean((lrrtb - preds))^2),
+#                              treatment = "Insecticide", site = "Wanang")
 
 # Numba ----
-ndata <- deltaData %>% 
-  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
-  filter(site == "numba")
+# ndata <- deltaData %>% 
+#   filter(unified_names %in% myTree.rand.poly$tip.label) %>%
+#   filter(site == "numba")
 
@@
 #                      na.action = na.omit)
-dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
-                     data = ndata,
-                     method = "ML",
-                     correlation = corMartins(2.718282, 
-                                              phy = myTree.rand.poly, 
-                                              form = ~unified_names,
-                                              fixed = T),
-                     na.action = na.omit)
+# dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
+#                      data = ndata,
+#                      method = "ML",
+#                      correlation = corMartins(2.718282, 
+#                                               phy = myTree.rand.poly, 
+#                                               form = ~unified_names,
+#                                               fixed = T),
+#                      na.action = na.omit)
 
 # Predict values
-pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.n, ndata, 
-                                       se.fit = T)
-ndata <- ndata %>%
-  mutate(preds = pred.vals$fit,
-         se.preds = pred.vals$se.fit,
-         se.upr = preds + se.preds,
-         se.lwr = preds - se.preds)
-
-# print the MSE
-mse_focused_df <- rbind(mse_focused_df,  
-                        data.frame(mse = with(ndata, mean((lrrtb - preds))^2),
-                                   treatment = "Insecticide", site = "Numba"))
+# pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.n, ndata, 
+#                                        se.fit = T)
+# ndata <- ndata %>%
+#   mutate(preds = pred.vals$fit,
+#          se.preds = pred.vals$se.fit,
+#          se.upr = preds + se.preds,
+#          se.lwr = preds - se.preds)
+# 
+# # print the MSE
+# mse_focused_df <- rbind(mse_focused_df,  
+#                         data.frame(mse = with(ndata, mean((lrrtb - preds))^2),
+#                                    treatment = "Insecticide", site = "Numba"))
 
 # Yawan ----
-ydata <- deltaData %>% 
-  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
-  filter(site == "yawan")
+# ydata <- deltaData %>% 
+#   filter(unified_names %in% myTree.rand.poly$tip.label) %>%
+#   filter(site == "yawan")
 
@@
 
-dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
-                     data = ydata,
-                     method = "ML",
-                     correlation = corMartins(2.718282,
-                                              phy = myTree.rand.poly,
-                                              form = ~unified_names,
-                                              fixed = T),
-                     na.action = na.omit)
+# dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
+#                      data = ydata,
+#                      method = "ML",
+#                      correlation = corMartins(2.718282,
+#                                               phy = myTree.rand.poly,
+#                                               form = ~unified_names,
+#                                               fixed = T),
+#                      na.action = na.omit)
 
 # Predict values
-pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.y, ydata, 
-                                       se.fit = T)
-ydata <- ydata %>%
-  mutate(preds = pred.vals$fit,
-         se.preds = pred.vals$se.fit,
-         se.upr = preds + se.preds,
-         se.lwr = preds - se.preds)
-
-# print the MSE
-mse_focused_df <- rbind(mse_focused_df,  
-                        data.frame(mse = with(ydata, mean((lrrtb - preds))^2),
-                                   treatment = "Insecticide", site = "Yawan"))
-
-library(sjPlot)
-
-tab_model(dd.gls.focsed.w, dd.gls.focsed.n, dd.gls.focsed.y)
-
-# fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}
-
-# Plot the relationship and add predicted values
-wpredPlot <- ggplot(wdata, aes(y = lrrtb, x = lsla.d))+
-  geom_point() +
-  geom_line(data = wdata, aes(x = lsla.d, y = preds), lwd = 1,
-            lty = 2)+
+# pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.y, ydata, 
+#                                        se.fit = T)
+# ydata <- ydata %>%
+#   mutate(preds = pred.vals$fit,
+#          se.preds = pred.vals$se.fit,
+#          se.upr = preds + se.preds,
+#          se.lwr = preds - se.preds)
+# 
+# # print the MSE
+# mse_focused_df <- rbind(mse_focused_df,  
+#                         data.frame(mse = with(ydata, mean((lrrtb - preds))^2),
+#                                    treatment = "Insecticide", site = "Yawan"))
+
+# tab_model(dd.gls.focsed.w, dd.gls.focsed.n, dd.gls.focsed.y)
+# 
+# 
+# # fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}
+# 
+# # Plot the relationship and add predicted values
+# wpredPlot <- ggplot(wdata, aes(y = lrrtb, x = lsla.d))+
+#   geom_point() +
+#   geom_line(data = wdata, aes(x = lsla.d, y = preds), lwd = 1,
+#             lty = 2)+
+#   # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) +
+#   ggtitle("Wanang 200 m a.s.l.")
+# 
+# npredPlot <- ggplot(ndata, aes(y = lrrtb, x = lsla.d))+
+#   geom_point() +
+#   geom_line(data = ndata, aes(x = lsla.d, y = preds), 
+#             lwd = 1, lty = 2)+
+#   # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
+#   ggtitle("Numba 750 m a.s.l.")
+# 
+# ypredPlot <- ggplot(ydata, aes(y = lrrtb, x = lsla.d))+
+#   geom_point() +
+#   geom_line(data = ydata, aes(x = lsla.d, y = preds), lwd = 1)+
 #   geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
-  ggtitle("Wanang 200 m a.s.l.")
-
-npredPlot <- ggplot(ndata, aes(y = lrrtb, x = lsla.d))+
-  geom_point() +
-  geom_line(data = ndata, aes(x = lsla.d, y = preds), 
-            lwd = 1, lty = 2)+
-  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
-  ggtitle("Numba 750 m a.s.l.")
-
-ypredPlot <- ggplot(ydata, aes(y = lrrtb, x = lsla.d))+
-  geom_point() +
-  geom_line(data = ydata, aes(x = lsla.d, y = preds), lwd = 1)+
-  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
-  ggtitle("Yawan 1900 m a.s.l.")
-
-ggpubr::ggarrange(wpredPlot, npredPlot, ypredPlot,
-                  nrow = 1)
-# Plot the relationship and add predicted values
-ggplot(rbind(ydata, ndata, wdata), aes(y = lrrtb, x = lsla.d, color = site))+
-  geom_point() + 
-  geom_line(data = rbind(ydata, ndata, wdata), 
-            aes(x = lsla.d, y = preds), lwd = 1)+
-  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15)
-
-sjPlot::tab_df(mse_focused_df)
+#   ggtitle("Yawan 1900 m a.s.l.")
+# 
+# ggpubr::ggarrange(wpredPlot, npredPlot, ypredPlot,
+#                   nrow = 1)
+# # Plot the relationship and add predicted values
+# ggplot(rbind(ydata, ndata, wdata), aes(y = lrrtb, x = lsla.d, color = site))+
+#   geom_point() + 
+#   geom_line(data = rbind(ydata, ndata, wdata), 
+#             aes(x = lsla.d, y = preds), lwd = 1)+
+#   geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15)
+# 
+# sjPlot::tab_df(mse_focused_df)
 
@@
 
-# Fungicide
+#### Fungicide
 
@@
 
-# Additional herbivores
+#### Additional herbivores
 
@@
 ```
+
+#### Predators
+
+```{r}
+subsetVars <- c("c", "p")
+
+source('code/get_data_ready_for_the_sla_analyses.R')
+
+ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+
+```{r, results='asis'}
+
+source('code/model_definitions_pgls.R')
+
+# Are traits improving the null model
+
+# Model selection.
+
+# 1. Full model (quadratic, and site interactions) vs no traits
+anova(no.traits, full.model)            # full model is better
+
+# 2. Removal of individual traits
+
+# 2.1 Remove SLA
+anova(full.model, q.herb.q.water) # withouth sla is slightly better (AIC)
+
+# 2.2 Remove herbivory
+anova(full.model, q.sla.q.water.noint) # Model without herbivory is not better
+
+# 2.3 Remove water content?
+anova(full.model, q.sla.q.herb) # Model with herbivory is much better
+
+# 3. Reduce model even further - 1 trait models.
+anova(full.model, just.sla.quadratic)
+anova(full.model, just.herb.quadratic)
+anova(full.model, just.water.quadratic)
+
+# Seems like they all stay!
+
+# 4. Quadratic vs linear
+anova(full.model, all.linear) # Full model still better
+
+# 5. Interaction reduction
+anova(full.model, full.model.noint)
+
+# So the best model is the one without the traits!
+tab_model(full.model)
+
+# Predictions and vizualizations!
+traitnm <- "lsla.d"
+predictAndVisualize <- function (deltaData, traitnm, n=100) {
+  other_traits <- names(deltaData)[!(names(deltaData) %in% c(traitnm, 
+                                                             "site", 
+                                                             "garden", 
+                                                             "treatment", 
+                                                             "unified_names",
+                                                             "pg.cat",
+                                                             "comparison"))]
+
+  focal.column = seq(from = min(deltaData[, traitnm], na.rm = T),
+                     to = max(deltaData[, traitnm], na.rm = T), 
+                     length = n)
+  other.vals <- t(apply(deltaData[, other_traits], 2, mean, na.rm = T))
+  preddata <- cbind(focal.column,
+        do.call("rbind", replicate(n, other.vals, simplify = FALSE)))
+  colnames(preddata)[1] <- traitnm
+  preddata <- rbind(preddata, preddata, preddata)
+  preddata$site <- rep(c("wanang", "numba", "yawan"), each = n)
+
+  # predvals <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddata,
+  #                   newdtype = 'response')
+  # preddata$predvals <- predvals$fit
+  # preddata$predse <- predvals$se.fit
+
+}
+
+preddata <- data.frame(lsla.d = seq(from = min(deltaData$lsla.d, na.rm = T), 
+                                  to = max(deltaData$lsla.d, na.rm = T), 
+                                  length = 100),
+                       water.cont = mean(deltaData$water.cont, na.rm = T), 
+                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
+                       richness= mean(deltaData$richness, na.rm = T), 
+                       labu= mean(deltaData$labu, na.rm = T),
+                       lrrsla= mean(deltaData$lrrsla, na.rm = T))
+
+preddatalrrsla <- data.frame(lsla.d = mean(deltaData$lsla.d, na.rm = T),
+                       water.cont = mean(deltaData$water.cont, na.rm = T), 
+                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
+                       richness= mean(deltaData$richness, na.rm = T), 
+                       labu= mean(deltaData$labu, na.rm = T),
+                       lrrsla=  seq(from = min(deltaData$lrrsla, na.rm = T), 
+                                  to = max(deltaData$lrrsla, na.rm = T), 
+                                  length = 100))
+
+visData <- deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label)
+
+predvals <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddata,
+                    newdtype = 'response')
+preddata$predvals <- predvals$fit
+preddata$predse <- predvals$se.fit
+
+predvalslrrsla <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddatalrrsla,
+                    newdtype = 'response')
+
+preddatalrrsla$predvals <- predvalslrrsla$fit
+preddatalrrsla$predse <- predvalslrrsla$se.fit
+
+```

### 2023-03-30 code/model_definitions.R

@@ -1,3 +1,5 @@
+moddata <- deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label)
+
 modelDefinitions <-  function () {
   
   # Runs all of model definitions
@@ -5,38 +7,38 @@ modelDefinitions <-  function () {
   # NO TRAITS
   no.traits <- gls(
     lrrtb ~ richness + labu + lrrsla + site,
-    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    data = moddata,
     method = "ML",
-    correlation = corMartins(2.718282, 
-                             phy = myTree.rand.poly,
-                             form = ~unified_names,
-                             fixed = T),
-    na.action = na.omit
+    # correlation = corMartins(2.718282, 
+    #                          phy = myTree.rand.poly,
+    #                          form = ~unified_names,
+    #                          fixed = T),
+    # na.action = na.omit
   )
   
   # FULL MODEL
   full.model <- gls(
     lrrtb ~ lsla.d*site+ I(lsla.d^2)*site+water.cont*site+I(water.cont^2)*site+perc.herb*site+I(perc.herb^2)*site+richness+labu+lrrsla+site,
-    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    data = ),
     method = "ML"
   )
   
   # FULL MODEL
   full.model.noint <- gls(
     lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) + perc.herb + I(perc.herb^2) + richness + labu + lrrsla + site,
-    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    data = moddata,
     method = "ML"
   )
   
   all.linear <- gls(
     lrrtb ~ lsla.d*site + water.cont*site + perc.herb*site + richness + labu + lrrsla + site,
-    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
+    data = moddata,
     method = "ML"
   )
   
   all.linear.noint <- gls(
     lrrtb ~ lsla.d+water.cont+perc.herb+richness+labu+lrrsla+site,
-    data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
[…]

### 2023-03-30 code/model_definitions_pgls.R

@@ -1,0 +1,518 @@
+moddata <- deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label)
+
+modelDefinitions <-  function () {
+  
+  # Runs all of model definitions
+  
+  # NO TRAITS
+  no.traits <- gls(
+    lrrtb ~ richness + labu + lrrsla + site,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  # FULL MODEL
+  full.model <- gls(
+    lrrtb ~ lsla.d*site+ I(lsla.d^2)*site+water.cont*site+I(water.cont^2)*site+perc.herb*site+I(perc.herb^2)*site+richness+labu+lrrsla+site,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  # FULL MODEL
+  full.model.noint <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) + perc.herb + I(perc.herb^2) + richness + labu + lrrsla + site,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  all.linear <- gls(
+    lrrtb ~ lsla.d*site + water.cont*site + perc.herb*site + richness + labu + lrrsla + site,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  all.linear.noint <- gls(
+    lrrtb ~ lsla.d+water.cont+perc.herb+richness+labu+lrrsla+site,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  # Individual traits-------------------------------------------------
+  
+  ## SLA
[…]

### 2023-04-11 ms1_data_analysis.Rmd
Prediction plots for I, P, H

--- 2023-03-28 ms1_data_analysis.Rmd
@@
 source('code/log_ratio_extraction.R')
+
+library(ggplot2)
+library(dplyr)
+library(sjPlot)
+library(AICcmodavg)
+
 ```
@@
 
-Figure 1. Median, 1st, second quadrilles, 1.5 IRQ distance and outliers in biomass, richness diversity and density at each location/elevation, but some Yawan sites were extremely productive.
-
+Figure 1. Median, 1st, second quadrilles, 1.5 IRQ distance, and outliers (labeled) for logarithm of biomass, richness diversity, and density for the three studied location/elevation.
 ```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure S1 Characteistics"}
-
-library(ggplot2)
-library(dplyr)
 
@@
 
-summary(trtlm)
-summary(trtgls)
+# summary(trtlm)
+# summary(trtgls)
 
@@
 Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation.
-
 ```{r}
@@
       # Test for noramlity
-      print(shapiro.test(lrrtd$x))
-      print(shapiro.test(lrrtd$y))
+      # print(shapiro.test(lrrtd$x))
+      # print(shapiro.test(lrrtd$y))
 
@@
 
-      print(lm.test.p)
+      # print(lm.test.p)
 
@@
 
-ggplot(cpdata, aes(x = x, y = y)) +
-  geom_point(size = 2, alpha = 0.4)+
-  stat_smooth(data = cpdata %>%
-                filter(sign == "signif"),
-              method ="lm")+
-  geom_hline(yintercept = 0)+
-  geom_vline(xintercept = 0)+
-  ylab("Effect size of insects")+
-  xlab("Effect size")+
-  theme_bw()+
-  facet_grid(response*descriptor~site)
+# ggplot(cpdata, aes(x = x, y = y)) +
+#   geom_point(size = 2, alpha = 0.4)+
+#   stat_smooth(data = cpdata %>%
+#                 filter(sign == "signif"),
+#               method ="lm")+
+#   geom_hline(yintercept = 0)+
+#   geom_vline(xintercept = 0)+
+#   ylab("Effect size of insects")+
+#   xlab("Effect size")+
+#   theme_bw()+
+#   facet_grid(response*descriptor~site)
 
@@
 
-ggplot(fidata, aes(x = x, y = y)) +
-  geom_point(size = 2, alpha = 0.4)+
-  stat_smooth(data = fidata %>%
-                filter(sign == "signif"),
-              method ="lm")+
-  geom_hline(yintercept = 0)+
-  geom_vline(xintercept = 0)+
-  xlab("Predator removal")+
-  ylab("Predator removal and insect addition")+
-  theme_bw()+
-  facet_grid(descriptor~site)
+# ggplot(fidata, aes(x = x, y = y)) +
+#   geom_point(size = 2, alpha = 0.4)+
+#   stat_smooth(data = fidata %>%
+#                 filter(sign == "signif"),
+#               method ="lm")+
+#   geom_hline(yintercept = 0)+
+#   geom_vline(xintercept = 0)+
+#   xlab("Predator removal")+
+#   ylab("Predator removal and insect addition")+
+#   theme_bw()+
+#   facet_grid(descriptor~site)
 
@@
       # Test for normality
-      print(shapiro.test(lrrtd$x))
-      print(shapiro.test(lrrtd$y))
+      # print(shapiro.test(lrrtd$x))
+      # print(shapiro.test(lrrtd$y))
 
@@
 
-ggplot(fidata, aes(x = x, y = y)) +
-  geom_point(size = 2, alpha = 0.4)+
-  stat_smooth(data = fidata %>%
-                filter(sign == "signif"),
-              method ="lm")+
-  geom_hline(yintercept = 0)+
-  geom_vline(xintercept = 0)+
-  ylab("Effect size of insects")+
-  xlab("Effect size of fungi")+
-  theme_bw()+
-  facet_grid(descriptor~site)
+# ggplot(fidata, aes(x = x, y = y)) +
+#   geom_point(size = 2, alpha = 0.4)+
+#   stat_smooth(data = fidata %>%
+#                 filter(sign == "signif"),
+#               method ="lm")+
+#   geom_hline(yintercept = 0)+
+#   geom_vline(xintercept = 0)+
+#   ylab("Effect size of insects")+
+#   xlab("Effect size of fungi")+
+#   theme_bw()+
+#   facet_grid(descriptor~site)
 
@@
 
-    print(summary(mod))
+    # print(summary(mod))
 
@@
 
-classifier_RF
+# classifier_RF
 
@@
 # Variable importance plot
-varImpPlot(classifier_RF)
+# varImpPlot(classifier_RF)
 ```
@@
 
-Explore that visually
-
-#### Insecticide
-
-##### Change in Biomass: log-respose ratio of a species' biomass
-
-```{r}
-
-# Get the delta data ready for th e given subset vars (comparison)
-subsetVars <- c("c", "i")
-
-# Create data-set for a given treatment
-deltaData <- data.frame()
-
-# Change total bio variable into numeric
-lgsData$tot_bio <- as.numeric(lgsData$tot_bio)
-
-# Loop through each species in each garden and site and calculate their deltaBio
-for (st in unique(lgsData$site)){
-
-  for (gard in unique(lgsData$garden)){
-
-    stGardDat <- lgsData %>%
-      filter(site == st,
-             garden == gard)
-
-    if (dim(stGardDat)[1] == 0){
-      # If subset is empty skip to another one
-      next
-    }
-
-    if (length(unique(stGardDat$treatment)) < 2){
-      # If there is species in only one treatment skip to another one
-      next
-    }
-
-    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
-    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
-    specInBoth <- specInCtr[specInCtr %in% specInTrt]
-
-    cDat <- stGardDat %>%
-      filter(unified_names %in% specInBoth,
-             treatment == subsetVars[1]) %>%
-      arrange(unified_names)
-    tDat <- stGardDat %>%
-      filter(unified_names %in% specInBoth,
-             treatment == subsetVars[2]) %>%
-      arrange(unified_names)
-
-    # Calculate LRR for biomass and SLA change
-
-    cDat$comparison <- subsetVars[2]
-
-    if (subsetVars[2] == "h"){
-      cDat$lrrtb <- log(tDat$tot_bio/cDat$tot_bio)
-      cDat$lrrsla <- log(tDat$sla.d.m2kg1/cDat$sla.d.m2kg1)
-    } else{
-      cDat$lrrtb <- log(cDat$tot_bio/tDat$tot_bio)
-      cDat$lrrsla <- log(cDat$sla.d.m2kg1/tDat$sla.d.m2kg1)
-    }
-
-    deltaData <- rbind(deltaData, cDat)
-
-  }
-}
-
-# Log transform estimated leaf area
-deltaData <- deltaData %>%
-          mutate(lest.leaf.area = log(est.leaf.area))%>%
-  filter(!is.infinite(lest.leaf.area))
-
-```
-
 Below I explore nonlinear relationships of the biomass log-ratios with species traits. Nonlinear relationships are introduced to allow for optimum trait values.
 
-```{r}
-
-# Log-response ratios of total biomass vs of sla
-
-# DeltaBiomass vs SLA at each site
-slaPlot <- ggplot(deltaData, aes(x = lsla.d, y = lrrtb, color = site)) +
-  geom_point()+
-  geom_smooth(method = "lm", formula = y ~ x + I(x^2))
-
-# There might be some effect of site.
-# DeltaBiomass vs SLA at each site
-wcPlot <- ggplot(deltaData, aes(x = water.cont, y = lrrtb, color = site)) +
-  geom_point()+
-  geom_smooth(method = "lm", formula = y ~ x + I(x^2))
-# Seems like the effect is not very strong either
-
-herbPlot <- ggplot(deltaData, aes(x = perc.herb, y = lrrtb, color = site)) +
-  geom_point()+
-  geom_smooth(method = "lm", formula = y ~ x + I(x^2)) + 
-  scale_x_continuous(trans = 'log2')
-
-ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
-
-```
 It seems that in some cases the nonlinear relationship may be important.
@@
 
+#### Insecticide
+
 ```{r}
-# Model explorations
-simple.sla <- gls(
-  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + richness + labu + lrrsla,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-just.sla <- gls(
-  lrrtb ~ lsla.d*site + I(lsla.d^2)*site,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-# AIC(just.sla, simple.sla)
-# anova(just.sla, simple.sla) # simple SLA is much bette
-# tab_model(simple.sla) # But only the lrr SLA is significant.
-
-# Run the full model.
-full.sla <- gls(
-  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + water.cont*site + I(water.cont^2)*site + perc.herb*site + I(perc.herb^2)*site + richness + labu + lrrsla,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-# Assuming linear relationships
-full.sla.linear <- gls(
-  lrrtb ~ lsla.d*site + water.cont*site + perc.herb*site + richness + labu + lrrsla,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-# anova(simple.sla, full.sla.linear) # full SLA linear is much better that the one using SLA
-# tab_model(full.sla.linear)
-
-full.sla.linear.noint <- gls(
-  lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla + site,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-# anova(full.sla.linear, full.sla.linear.noint)
-# Removing interactions does not improve the model significantly
-# tab_model(full.sla.linear.noint)
-
-# Add interaction only to the water content and sla
-full.sla.linear.intslawc <- gls(
-  lrrtb ~ lsla.d*site + water.cont*site + perc.herb + richness + labu + lrrsla + site,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-# anova(full.sla.linear.noint, full.sla.linear.intslawc)
-# No effect
-# tab_model(full.sla.linear.noint)
-
-full.sla.linear.intsla <- gls(
-  lrrtb ~ lsla.d*site + water.cont + perc.herb + richness + labu + lrrsla + site,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-# anova(full.sla.linear.noint, full.sla.linear.intsla)
-# Any for of iteraction is not improving the model
-tab_model(full.sla.linear.noint) # Thus the simplest model seems to be best.
-
-# I further compare to the full.sla.linear.noint model.
-# Only water content is modeled as linear
-full.sla.custom <-  gls(
-  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + water.cont*site + perc.herb*site + I(perc.herb^2)*site + richness + labu + lrrsla,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-# anova(full.sla.linear.noint, full.sla.custom)
-
-full.sla.custom.noint <-  gls(
-  lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + perc.herb + I(perc.herb^2) + richness + labu + lrrsla,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-full.sla.linear.noint.nosite <- gls(
-  lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
-                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
-  method = "ML"
-  )
-
-# Withouth a site it is even better
-# anova(full.sla.linear.noint, full.sla.linear.noint.nosite)
-
-# The linear model remains the best one.
-# plot(full.sla.linear.noint.nosite) # This plot looks best
-# tab_model(full.sla.linear.noint.nosite)
-
-```
-
-Table. Results of the selected model for the deltaBio.
-
-```{r}
-sjPlot::tab_model(full.sla.linear.noint.nosite)
+
+# Get the delta data ready for th e given subset vars (comparison)
+subsetVars <- c("c", "i")
+
+source('code/get_data_ready_for_the_sla_analyses.R')
+
+# ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+
+```{r, results='asis'}
+
+source('code/model_definitions_pgls.R')
+
+# Are traits improving the null model
+
+# Model selection.
+
+# 1. Full model (quadratic, and site interactions) vs no traits
+# anova(no.traits, full.model)            # full model is better
+
+# 2. Removal of individual traits
+
+# 2.1 Remove SLA
+# anova(full.model, q.herb.q.water) # no-difference, keep model without the SLA a simpler model?
+# 2.2 Remove herbivory
+# anova(full.model, q.sla.q.water.noint) # Model without herbivory is significantly worse
+# 2.3 Remove water content?
+# anova(full.model, q.sla.q.herb) # Model without water is significantly worse
+
+# Herbivory and water are the most important!!!
+
+# 3. Reduce model even further - 1 trait models.
+# anova(q.herb.q.water, just.sla.quadratic)   # Two trait model better
+# anova(q.herb.q.water, just.herb.quadratic)  # Two trait model better
+# anova(q.herb.q.water, just.water.quadratic) # Two trait model better
+
+# 4. Quadratic vs linear
+# anova(q.herb.q.water, l.herb.l.water) # Quadratic model is still better
+
+# 5. Interaction reduction
+# anova(q.herb.q.water, q.herb.q.water.noint) # Model with interaction is better
+# anova(q.herb.q.water, just.water.quadratic.noint) # Model with interaction is better
+# anova(q.herb.q.water, just.water.linear.noint)
+
+```
+
+```{r, results='asis'}
+# So the best model is the one without the traits!
+# tab_model(q.herb.q.water)
+# tab_model(just.water.quadratic.noint)
+i.mod <- q.herb.q.water
+i.dD <- deltaData
 ```
@@
 Nevertheless it is not a strong correlation and does not explain a lot of variation in the biomass change.
-
-```{r, fig.cap="Observed vs predicted values for the lsls.d and "}
-preddata <- data.frame(lsla.d = seq(from = min(deltaData$lsla.d, na.rm = T), 
-                                  to = max(deltaData$lsla.d, na.rm = T), 
-                                  length = 100),
-                       water.cont = mean(deltaData$water.cont, na.rm = T), 
-                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
-                       richness= mean(deltaData$richness, na.rm = T), 
-                       labu= mean(deltaData$labu, na.rm = T),
-                       lrrsla= mean(deltaData$lrrsla, na.rm = T))
-
-preddatalrrsla <- data.frame(lsla.d = mean(deltaData$lsla.d, na.rm = T),
-                       water.cont = mean(deltaData$water.cont, na.rm = T), 
-                       perc.herb= mean(deltaData$perc.herb, na.rm = T), 
-                       richness= mean(deltaData$richness, na.rm = T), 
-                       labu= mean(deltaData$labu, na.rm = T),
-                       lrrsla=  seq(from = min(deltaData$lrrsla, na.rm = T), 
-                                  to = max(deltaData$lrrsla, na.rm = T), 
-                                  length = 100))
-
-visData <- deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label)
-
-predvals <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddata,
-                    newdtype = 'response')
-
-preddata$predvals <- predvals$fit
-preddata$predse <- predvals$se.fit
-
-predvalslrrsla <- AICcmodavg::predictSE.gls(full.sla.linear.noint.nosite, newdata = preddatalrrsla,
-                    newdtype = 'response')
-
-preddatalrrsla$predvals <- predvalslrrsla$fit
-preddatalrrsla$predse <- predvalslrrsla$se.fit
-
-# For lsla.d
-p1 <- ggplot(preddata, aes(x = lsla.d, y = predvals)) + 
-  geom_line()+
-  geom_ribbon(aes(ymin = predvals - predse,
-                  ymax = predvals + predse), 
-              alpha = 0.1) + 
-  geom_point(data = visData, aes(x = lsla.d, y = lrrtb)) + 
-  ggtitle("Log(SLA) vs Log-response ratio of a species' biomass")
-
-# For lrrsla
-p2 <- ggplot(preddatalrrsla, aes(x = lrrsla, y = predvals)) + 
-  geom_line()+
-  geom_ribbon(aes(ymin = predvals - predse,
-                  ymax = predvals + predse), 
-              alpha = 0.1) + 
-  geom_point(data = visData, aes(x = lrrsla, y = lrrtb)) +
-  ggtitle("Delta(SLA) vs Log-response ratio of a species' biomass")
-
-# ggplot(visData, aes(x = lsla.d, y = lrrtb)) + 
-#   geom_point(aes(size = tot_bio, alpha = 0.05), pch = 19)+
-#   geom_point(aes(x = lsla.d, y = predvals, col = "red")) + 
-#   geom_smooth(aes(x = lsla.d, y = predvals), method = "lm")
-
-ggpubr::ggarrange(p1,p2)
-
-```
 
@@
 
-```{r, include=FALSE}
-# The Focused model
-# Only the predictive power of SLA, water content and percentage herbivory on deltaBio
-
-# Wanang ----
-wdata <- deltaData %>% 
-  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
-  filter(site == "wanang")
-
-# dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d,
-#                      data = wdata,
-#                      method = "ML",
-#                      correlation = corMartins(2.718282, 
-#                                               phy = myTree.rand.poly, 
-#                                               form = ~unified_names,
-#                                               fixed = T),
-#                      na.action = na.omit)
-dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
-                     data = wdata,
-                     method = "ML",
-                     correlation = corMartins(2.718282, 
-                                              phy = myTree.rand.poly, 
-                                              form = ~unified_names,
-                                              fixed = T),
-                     na.action = na.omit)
-
-# Model results
-
-# Predict values
-pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.w, wdata, 
-                                       se.fit = T)
-wdata <- wdata %>%
-  mutate(preds = pred.vals$fit,
-         se.preds = pred.vals$se.fit,
-         se.upr = preds + se.preds,
-         se.lwr = preds - se.preds)
-
-# print the MSE
-mse_focused_df <- data.frame(mse = with(wdata, mean((lrrtb - preds))^2),
-                             treatment = "Insecticide", site = "Wanang")
-
-# Numba ----
-ndata <- deltaData %>% 
-  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
-  filter(site == "numba")
-
-# dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d,
-#                      data = ndata,
-#                      method = "ML",
-#                      correlation = corMartins(2.718282, 
-#                                               phy = myTree.rand.poly, 
-#                                               form = ~unified_names,
-#                                               fixed = T),
-#                      na.action = na.omit)
-dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
-                     data = ndata,
-                     method = "ML",
-                     correlation = corMartins(2.718282, 
-                                              phy = myTree.rand.poly, 
-                                              form = ~unified_names,
-                                              fixed = T),
-                     na.action = na.omit)
-
-# Predict values
-pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.n, ndata, 
-                                       se.fit = T)
-ndata <- ndata %>%
-  mutate(preds = pred.vals$fit,
-         se.preds = pred.vals$se.fit,
-         se.upr = preds + se.preds,
-         se.lwr = preds - se.preds)
-
-# print the MSE
-mse_focused_df <- rbind(mse_focused_df,  
-                        data.frame(mse = with(ndata, mean((lrrtb - preds))^2),
-                                   treatment = "Insecticide", site = "Numba"))
-
-# Yawan ----
-ydata <- deltaData %>% 
-  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
-  filter(site == "yawan")
-
-# The model
-# dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d,
-#                      data = ydata,
-#                      method = "ML",
-#                      correlation = corMartins(2.718282, 
-#                                               phy = myTree.rand.poly, 
-#                                               form = ~unified_names,
-#                                               fixed = T),
-#                      na.action = na.omit)
-
-dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d + water.cont + perc.herb + richness + labu + lrrsla,
-                     data = ydata,
-                     method = "ML",
-                     correlation = corMartins(2.718282,
-                                              phy = myTree.rand.poly,
-                                              form = ~unified_names,
-                                              fixed = T),
-                     na.action = na.omit)
-
-# Predict values
-pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.y, ydata, 
-                                       se.fit = T)
-ydata <- ydata %>%
-  mutate(preds = pred.vals$fit,
-         se.preds = pred.vals$se.fit,
-         se.upr = preds + se.preds,
-         se.lwr = preds - se.preds)
-
-# print the MSE
-mse_focused_df <- rbind(mse_focused_df,  
-                        data.frame(mse = with(ydata, mean((lrrtb - preds))^2),
-                                   treatment = "Insecticide", site = "Yawan"))
-
-library(sjPlot)
-
-tab_model(dd.gls.focsed.w, dd.gls.focsed.n, dd.gls.focsed.y)
-
-# fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}
-
-# Plot the relationship and add predicted values
-wpredPlot <- ggplot(wdata, aes(y = lrrtb, x = lsla.d))+
-  geom_point() +
-  geom_line(data = wdata, aes(x = lsla.d, y = preds), lwd = 1,
-            lty = 2)+
-  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) +
-  ggtitle("Wanang 200 m a.s.l.")
-
-npredPlot <- ggplot(ndata, aes(y = lrrtb, x = lsla.d))+
-  geom_point() +
-  geom_line(data = ndata, aes(x = lsla.d, y = preds), 
-            lwd = 1, lty = 2)+
-  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
-  ggtitle("Numba 750 m a.s.l.")
-
-ypredPlot <- ggplot(ydata, aes(y = lrrtb, x = lsla.d))+
-  geom_point() +
-  geom_line(data = ydata, aes(x = lsla.d, y = preds), lwd = 1)+
-  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
-  ggtitle("Yawan 1900 m a.s.l.")
-
-ggpubr::ggarrange(wpredPlot, npredPlot, ypredPlot,
-                  nrow = 1)
-# Plot the relationship and add predicted values
-ggplot(rbind(ydata, ndata, wdata), aes(y = lrrtb, x = lsla.d, color = site))+
-  geom_point() + 
-  geom_line(data = rbind(ydata, ndata, wdata), 
-            aes(x = lsla.d, y = preds), lwd = 1)+
-  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15)
-
-sjPlot::tab_df(mse_focused_df)
-
-```
-
-The above shows that focused model, where we used SLA as the only predictor of the plant change in biomass has a declining ability to predict change in biomass at the insecticide plot.
-
-# Fungicide
+#### Fungicide
 
 ```{r}
+
 subsetVars <- c("c", "f")
@@
 
-ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
-
-```
-
-```{r}
-
-source('code/model_definitions.R')
+# ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+
+```{r, results='asis'}
+
+source('code/model_definitions_pgls.R')
 
@@
 # Model selection.
+
 # 1. Full model (quadratic, and site interactions) vs no traits
-# anova(no.traits, full.model)
-# anova(no.traits, full.model.noint)
-# anova(no.traits, all.linear)
-# anova(no.traits, all.linear.noint)
-
-# 2. Individual traits
-
-# SLA
-# Check quadratic
-# anova(no.traits, just.sla.quadratic.noint)
-# Check linear
-# anova(no.traits, just.sla.linear.noint)
-# Check quadratic interation
-# anova(no.traits, just.sla.quadratic)
-# Check linear interaction
-# anova(no.traits, just.sla.linear)
-
-# WATER
-# Check quadratic
-# anova(no.traits, just.water.quadratic.noint)
-# Check linear
-# anova(no.traits, just.water.linear.noint)
-# Check quadratic interation
-# anova(no.traits, just.water.quadratic)
-# Check linear interaction
-# anova(no.traits, just.water.linear)
-
-# HERBIVORY
-# Check quadratic
-# anova(no.traits, just.herb.quadratic.noint)
-# Check linear
-# anova(no.traits, just.herb.linear.noint)
-# Check quadratic interation
-# anova(no.traits, just.herb.quadratic)
-# Check linear interaction
-# anova(no.traits, just.herb.linear)
-
-# 2. Model with all traits linear/quadratic, what combination?
-
-# quadratic SLA, HERB
-# anova(no.traits, q.sla.q.herb)
-# anova(no.traits, q.sla.q.herb.noint)
-# quadratic SLA, linear HERB
-# anova(no.traits, q.sla.l.herb)
-# anova(no.traits, q.sla.l.herb.noint)
-# linear SLA, quadratic HERB
-# anova(no.traits, l.sla.q.herb)
-# anova(no.traits, l.sla.q.herb.noint)
-# linear SLA, linear HERB
-# anova(no.traits, l.sla.l.herb)
-# anova(no.traits, l.sla.l.herb.noint)
-
-# quadratic HERB, WATER
-# anova(no.traits, q.herb.q.water)
-# anova(no.traits, q.herb.q.water.noint)
-# quadratic HERB, linear WATER
-# anova(no.traits, q.herb.l.water)
-# anova(no.traits, q.herb.l.water.noint)
-# linear HERB, quadratic WATER
-# anova(no.traits, l.herb.q.water)
-# anova(no.traits, l.herb.q.water.noint)
-# linear HERB, linear WATER
-# anova(no.traits, l.herb.l.water)
-# anova(no.traits, l.herb.l.water.noint)
-
-# quadratic SLA, WATER
-# anova(no.traits, q.sla.q.water)
-# anova(no.traits, q.sla.q.water.noint)
-# quadratic SLA, linear WATER
-# anova(no.traits, q.sla.l.water)
-# anova(no.traits, q.sla.l.water.noint)
-# linear SLA, quadratic WATER
-# anova(no.traits, l.sla.q.water)
-# anova(no.traits, l.sla.q.water.noint)
-# linear SLA, linear WATER
-# anova(no.traits, l.sla.l.water)
-# anova(no.traits, l.sla.l.water.noint)
-
-# So the best model is the one without the traits!
-tab_model(no.traits)
-
-```
-
-# Additional herbivores
+# anova(no.traits, full.model)            # full model is better
+
+# 2. Removal of individual traits
+
+# 2.1 Remove SLA
+# anova(full.model, q.herb.q.water) # without sla is slightly better (AIC)
+
+# 2.2 Remove herbivory
+# anova(full.model, q.sla.q.water) # Model without herbivory is not better
+
+# Both q.herb.q.water and q.sla.q.water are simpler, but which one is better?
+# anova(full.model, q.sla.q.herb) # Model without water is significantly worse
+
+# 2.3 Remove water content?
+# anova(full.model, q.sla.q.herb) # Model with herbivory is much better!
+# Water is important!!!
+
+# anova(q.herb.q.water, q.sla.q.water) # q.sla.q.water is better based on AIC
+
+# 3. Reduce model even further - 1 trait models.
+# anova(q.sla.q.water, just.sla.quadratic)  # q.sla.q.water is still better
+# anova(q.sla.q.water, just.herb.quadratic) # q.sla.q.water is still better
+# anova(q.sla.q.water, just.water.quadratic) # model with just water is as good as q.sla.q.water
+
+# model with just water is not different form the full model, and it has lower AIC
+
+# Seems like they all stay!
+
+# 4. Quadratic vs linear
+# anova(just.water.quadratic, just.water.linear) # Linear is simpler model still better
+
+# 5. Interaction reduction
+# anova(just.water.linear, just.water.linear.noint) # Interaction does not bring anything to the model
+
+```
+
+```{r, results='asis'}
+
+# So the best model is the one with only water!
+# tab_model(just.water.linear.noint)
+f.mod <- just.water.linear.noint
+f.dD <- deltaData
+```
+
+#### Additional herbivores
 
@@
 
-ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
-
-```
+# ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+
+```{r, results='asis'}
+
+source('code/model_definitions_pgls.R')
+
+# Are traits improving the null model
+
+# Model selection.
+
+# 1. Full model (quadratic, and site interactions) vs no traits
+# anova(no.traits, full.model)            # full model is better
+
+# 2. Removal of individual traits
+
+# 2.1 Remove SLA
+# anova(full.model, q.herb.q.water) # without sla is slightly better (AIC)
+
+# 2.2 Remove herbivory
+# anova(full.model, q.sla.q.water) # Model without herbivory is not better
+
+# 2.3 Remove water content?
+# anova(full.model, q.sla.q.herb) # Model with herbivory is much better
+
+# Two candidate models:
+# anova(q.herb.q.water,q.sla.q.water) # q.sla.q.water is a bit better (AIC)
+
+# 3. Reduce model even further - 1 trait models.
+# anova(q.sla.q.water, just.sla.quadratic) # MOdel with water is significantly better than just sla
+# anova(q.sla.q.water, just.herb.quadratic) # It is better than just herbivory
+# anova(q.sla.q.water, just.water.quadratic) # But is not better than model with just water
+
+# Seems like they all stay!
+
+# 4. Quadratic vs linear
+# anova(just.water.quadratic, just.water.linear) # Linear model is not simpler, it has slightly higher AIC
+
+# 5. Interaction reduction
+# anova(just.water.linear, just.water.linear.noint) # Interaction model is significantly better
+```
+
+```{r, results='asis'}
+# So the best model is ... !
+# tab_model(just.water.linear)
+h.mod <- just.water.linear
+h.dD <- deltaData
+```
+
+#### Predators
 
 ```{r}
-
-source('code/model_definitions.R')
+subsetVars <- c("c", "p")
+
+source('code/get_data_ready_for_the_sla_analyses.R')
+
+# ggpubr::ggarrange(slaPlot, wcPlot, herbPlot)
+
+```
+
+```{r, results='asis'}
+
+source('code/model_definitions_pgls.R')
 
@@
 # Model selection.
+
 # 1. Full model (quadratic, and site interactions) vs no traits
-anova(no.traits, full.model) # full model is much better
-anova(full.model, full.model.noint)
-
-# interaction model is as good as the no interaction. Keep the model without the interaction
-
-# 2. Is the quadratic relationship significant
-# anova(full.model.noint, all.linear)            # AIC is lower for the full model noint
-# anova(full.model.noint, all.linear.noint)      # AIC is lower for the full model noint
-
-# Traits have nonlinear relationships with the delta biomass.
-
-# 3. REMOVE SLA
-# 3.1 anova(full.model.noint, q.herb.q.water.noint) # withouth sla is slightly better (AIC)
-# 3.2 anova(full.model.noint, q.herb.q.water)
-# 3.3 anova(full.model.noint, q.herb.l.water.noint) # Full model is than this !!!
-# 3.4 anova(full.model.noint, q.herb.l.water)
-# 3.5 anova(full.model.noint, l.herb.l.water.noint) # Full model still better
-# 3.6 anova(full.model.noint, l.herb.l.water) #
-
-# Including SLA did not mprove the model significantly
-# tab_model(q.herb.q.water.noint)
-
-# But if we were to start first by remving perc.herbivory?
-# anova(full.model.noint, q.sla.q.water.noint) # Model without herbivory is better
-# anova(full.model.noint, q.sla.q.water)       # Full model is better
-# anova(q.sla.q.water, q.sla.q.water.noint)    # No interaction is better
-# anova(full.model.noint, q.sla.l.water.noint)   # Full model is still better
-# anova(full.model.noint, q.sla.l.water)         # Full model wins again
-# anova(full.model.noint, l.sla.l.water.noint)   # And again by AIC criterium
-# anova(full.model.noint, l.sla.l.water)         # And again
-
-# But if we were to start first by remving water.content?
-# anova(full.model.noint, q.sla.q.herb.noint) # Model with herbivory is much better
-# anova(full.model.noint, q.sla.q.herb)       # Full model is better
-# anova(q.sla.q.herb, q.sla.q.herb.noint)    # No interaction is better
-# anova(full.model.noint, q.sla.l.herb.noint)   # Full model is still better
-# anova(full.model.noint, q.sla.l.herb)         # Full model wins again
-# anova(full.model.noint, l.sla.l.herb.noint)   # And again by AIC criterium
-# anova(full.model.noint, l.sla.l.herb)         # And again
-
-# The above suggests that models without hebivory and sla do better job. Water content is the key trait that helps predict the change in biomass in response to the treatment.
-
-# SLA
-# Check quadratic
-# anova(full.model.noint, just.sla.quadratic.noint)
-# Check linear
-# anova(full.model.noint, just.sla.linear.noint)
-# Check quadratic interation
-# anova(full.model.noint, just.sla.quadratic)
-# Check linear interaction
-# anova(full.model.noint, just.sla.linear)
-
-# WATER
-# Check quadratic
-anova(full.model.noint, just.water.quadratic.noint) # Model with just water content
-
-# Is the quadratic form necessary
-anova(just.water.quadratic.noint, just.water.linear.noint) # no difference but quadratic has lower AIC
-
+# anova(no.traits, full.model)            # full model is better
+
+# 2. Removal of individual traits
+
+# 2.1 Remove SLA
+# anova(full.model, q.herb.q.water) # Full model is significanlty better
+
+# 2.2 Remove herbivory
+# anova(full.model, q.sla.q.water.noint) # Full model still better
+
+# 2.3 Remove water content?
+# anova(full.model, q.sla.q.herb) # Full model still better
+
+# 3. Reduce model even further - 1 trait models.
+# anova(full.model, just.sla.quadratic)
+# anova(full.model, just.herb.quadratic)
+# anova(full.model, just.water.quadratic)
+
+# Seems like they all stay!
+
+# 4. Quadratic vs linear
+# anova(full.model, all.linear) # Full model still better
+
+# 5. Interaction reduction
+# anova(full.model, full.model.noint) # Full model still better
+```
+
+```{r, results='asis'}
 # So the best model is the one without the traits!
-tab_model(just.water.quadratic.noint)
-
-```
+# tab_model(full.model)
+p.mod <- full.model
+p.dD <- deltaData
+```
+
+### SLA models for treatments
+
+```{r, results='asis'}
+
+tab_model(p.mod, i.mod, h.mod, f.mod,
+          pred.labels = c("Intercept", 
+                          "Log[SLA]-linear", 
+                          "Numba", 
+                          "Yawan",
+                          "Log[SLA]-quadratic", 
+                          "Water content - linear", 
+                          "Water content - quadratic",
+                          "Herbivory [%] - linear",
+                          "Herbivory [%] - quadratic",
+                          "Species richness",
+                          "Abundance",
+                          "LRR SLA",
+                          "Log[SLA]-linear : Numba",
+                          "Log[SLA]-linear : Yawan",
+                          "Log[SLA]-quadratic : Numba",
+                          "Log[SLA]-quadratic : Yawan",
+                          "Water content-linear : Numba",
+                          "Water content-linear : Yawan",
+                          "Water content-quadratic : Numba",
+                          "Water content-quadratic : Yawan",
+                          "Herbivory-linear : Numba",
+                          "Herbivory-linear : Yawan",
+                          "Herbivory-quadratic : Numba",
+                          "Herbivory-quadratic : Yawan"),
+          dv.labels = c("Predator", "Insecticide", "Extra herbivory", "Fungicide"),
+          string.pred = "Coeffcient",
+          string.ci = "Conf. Int (95%)",
+          string.p = "P-Value")
+
+```
+
+### Predictions and visualizations
+
+```{r}
+source("code/predictAndVisualize_function.R")
+
+# Predator --------------------------
+pslap <- predictAndVisualize(p.mod, 
+                    p.dD,
+                    "lsla.d",
+                    n = 100,
+                    titlestring = "Predator - A",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Log(SLA)")
+
+# Focal traits
+
+# Co-variates
+psitep <- predictAndVisualize(p.mod, 
+                    p.dD,
+                    "fsite",
+                    n = 100,
+                    titlestring = "Predator - B",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Sites")
+
+plabup <- predictAndVisualize(p.mod, 
+                    p.dD,
+                    "labu",
+                    n = 100,
+                    titlestring = "Predator - C",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Log(stem number)")
+
+# Insecticide-----------------------
+ilabup <- predictAndVisualize(i.mod, 
+                    i.dD,
+                    "labu",
+                    n = 100,
+                    titlestring = "Insecticide - C",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Log(stem number)")
+
+# Extra herbivory-------------------
+hsitep <- predictAndVisualize(h.mod, 
+                    h.dD,
+                    "fsite",
+                    n = 100,
+                    titlestring = "Herbivory - A",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Sites")
+
+hwaterp <- predictAndVisualize(h.mod, 
+                    h.dD,
+                    "water.cont",
+                    n = 100,
+                    titlestring = "Herbivory - A",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Water content")
+
+hlabup <- predictAndVisualize(h.mod, 
+                    h.dD,
+                    "labu",
+                    n = 100,
+                    titlestring = "Herbivory - B",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Log(stem number)")
+
+hlrrslap <- predictAndVisualize(h.mod, 
+                    h.dD,
+                    "lrrsla",
+                    n = 100,
+                    titlestring = "Herbivory - B",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "LRR SLA")
+
+# Plot
+ggpubr::ggarrange(pslap, psitep, plabup)
+
+ggpubr::ggarrange(ilabup)
+
+ggpubr::ggarrange(hwaterp, hsitep, hlabup, hlrrslap)
+
+```
+
+```{r}
+# Insecticide
+
+```

### 2023-04-11 code/get_data_ready_for_the_sla_analyses.R

@@ -98,7 +98,10 @@ for (st in unique(lgsData$site)){
 ####--------------------------------------------------
 # 3. Log transform estimated leaf area
 deltaData <- deltaData %>%
-  mutate(lest.leaf.area = log(est.leaf.area))%>%
+  mutate(lest.leaf.area = log(est.leaf.area),
+         fsite = factor(site, 
+                        levels = c("wanang", "numba","yawan")),
+         perc.herb = c(scale(perc.herb)))%>%
   filter(!is.infinite(lest.leaf.area))
 
 # DeltaBiomass vs SLA at each site

### 2023-04-11 code/model_definitions_pgls.R

@@ -6,7 +6,7 @@ modelDefinitions <-  function () {
   
   # NO TRAITS
   no.traits <- gls(
-    lrrtb ~ richness + labu + lrrsla + site,
+    lrrtb ~ richness + labu + lrrsla + fsite,
     data = moddata,
     method = "ML",
     correlation = corMartins(2.718282,
@@ -18,7 +18,7 @@ modelDefinitions <-  function () {
   
   # FULL MODEL
   full.model <- gls(
-    lrrtb ~ lsla.d*site+ I(lsla.d^2)*site+water.cont*site+I(water.cont^2)*site+perc.herb*site+I(perc.herb^2)*site+richness+labu+lrrsla+site,
+    lrrtb ~ lsla.d*fsite+ I(lsla.d^2)*fsite+water.cont*fsite+I(water.cont^2)*fsite+perc.herb*fsite+I(perc.herb^2)*fsite+richness+labu+lrrsla+fsite,
     data = moddata,
     method = "ML",
     correlation = corMartins(2.718282,
@@ -30,7 +30,7 @@ modelDefinitions <-  function () {
   
   # FULL MODEL
   full.model.noint <- gls(
-    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) + perc.herb + I(perc.herb^2) + richness + labu + lrrsla + site,
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) + perc.herb + I(perc.herb^2) + richness + labu + lrrsla + fsite,
     data = moddata,
     method = "ML",
     correlation = corMartins(2.718282,
@@ -41,7 +41,7 @@ modelDefinitions <-  function () {
   )
   
   all.linear <- gls(
-    lrrtb ~ lsla.d*site + water.cont*site + perc.herb*site + richness + labu + lrrsla + site,
+    lrrtb ~ lsla.d*fsite + water.cont*fsite + perc.herb*fsite + richness + labu + lrrsla + fsite,
     data = moddata,
     method = "ML",
     correlation = corMartins(2.718282,
@@ -52,7 +52,7 @@ modelDefinitions <-  function () {
   )
   
   all.linear.noint <- gls(
-    lrrtb ~ lsla.d+water.cont+perc.herb+richness+labu+lrrsla+site,
+    lrrtb ~ lsla.d+water.cont+perc.herb+richness+labu+lrrsla+fsite,
     data = moddata,
[…]

### 2023-04-11 code/predictAndVisualize_function.R

@@ -1,0 +1,191 @@
+# Predictions and visualizations!
+# predictAndVisualize <- function (model,
+#                                  deltaData,
+#                                  traitnm,
+#                                  n=100,
+#                                  titlestring = "",
+#                                  ylabstring = "",
+#                                  xlabstring = ""
+# ) {
+# 
+#   other_traits <- names(deltaData)[!(names(deltaData) %in% c(traitnm,
+#                                                              "fsite",
+#                                                              "site",
+#                                                              "garden",
+#                                                              "treatment",
+#                                                              "unified_names",
+#                                                              "pg.cat",
+#                                                              "comparison"))]
+# 
+#   # Create a value gradient for the focal variable.
+#   # focal.column = seq(from = min(deltaData[, traitnm], na.rm = T),
+#   #                    to = max(deltaData[, traitnm], na.rm = T),
+#   #                    length = n)
+# 
+#   focal.column.w = seq(from = min(deltaData[deltaData$site == "wanang",
+#                                             traitnm], na.rm = T),
+#                        to = max(deltaData[deltaData$site == "wanang",
+#                                           traitnm], na.rm = T),
+#                        length = n)
+#   focal.column.n = seq(from = min(deltaData[deltaData$site == "numba", traitnm], na.rm = T),
+#                        to = max(deltaData[deltaData$site == "numba",
+#                                           traitnm], na.rm = T),
+#                        length = n)
+#   focal.column.y = seq(from = min(deltaData[deltaData$site == "yawan", traitnm], na.rm = T),
+#                        to = max(deltaData[deltaData$site == "yawan",
+#                                           traitnm], na.rm = T),
+#                        length = n)
+# 
+#   # Assume average values for the rest.
+#   other.vals.w <- t(apply(deltaData[deltaData$site == "wanang",
+#                                     other_traits], 2, mean, na.rm = T))
+#   other.vals.n <- t(apply(deltaData[deltaData$site == "numba",
+#                                     other_traits], 2, mean, na.rm = T))
+#   other.vals.y <- t(apply(deltaData[deltaData$site == "yawan",
[…]

### 2023-04-11 code/specific_functions.R

@@ -57,7 +57,7 @@ getLogRatio <- function (dataset,
   
   for (gard in unique(dset$garden)){
     
-    print(gard)
+    # print(gard)
     
     # Dataset only for control plot
     based <- dset %>%
@@ -110,19 +110,19 @@ runTest <- function(lrdata, treatment_code = "f", label){
     testStat <- aovtrt$`F-value`
     testP <- aovtrt$`p-value`
     
-    print("Model runned")
+    # print("Model runned")
     
     # Tukey post-hoc
     
-    print("Stage1")
+    # print("Stage1")
     # emm <- emmeans(object = trtlm, 
     #                         specs = "site",
     #                adjust = "mvt")
     
-    print("Post hoc not runned")
+    # print("Post hoc not runned")
     
     zero.dif <- ifelse(summary(trtlm)$tTable[,4] > 0.05, "","*")
-    print("zero.dif assigned")
+    # print("zero.dif assigned")
     
     # mod_means <- multcomp::cld(object = emm,
     #                            Letters = letters)$.group

### 2023-04-13 ms1_data_analysis.Rmd
Visualizations for the SLA model mostly done. Getting ready to impolement random forest models for the rest of the treatments.

@@
 
+# ENCLOSED IN THE get_ready_for_the_sla_analyses
 # Subset a data-set to only control and treatment plots 
@@
 subsetVars <- c("c", "i")
-
 source('code/get_data_ready_for_the_sla_analyses.R')
@@
                     n = 100,
-                    titlestring = "Predator - A",
+                    titlestring = "Predator - Specific Leaf Area",
                     ylabstring = "Log-response ratio of total biomass.",
@@
                     n = 100,
-                    titlestring = "Predator - B",
+                    titlestring = "Predator - Site",
                     ylabstring = "Log-response ratio of total biomass.",
@@
                     n = 100,
-                    titlestring = "Predator - C",
+                    titlestring = "Predator - Abundance",
                     ylabstring = "Log-response ratio of total biomass.",
                     xlabstring = "Log(stem number)")
+prichp <- predictAndVisualize(p.mod, 
+                    p.dD,
+                    "richness",
+                    n = 100,
+                    titlestring = "Predator - Richness",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Number od species")
+plrrslap <- predictAndVisualize(p.mod, 
+                    p.dD,
+                    "lrrsla",
+                    n = 100,
+                    titlestring = "Predator - LRR SLA",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "LRR SLA")
 
@@
                     n = 100,
-                    titlestring = "Insecticide - C",
+                    titlestring = "Insecticide - Abundance",
                     ylabstring = "Log-response ratio of total biomass.",
@@
                     n = 100,
-                    titlestring = "Herbivory - A",
+                    titlestring = "Herbivory - Site",
                     ylabstring = "Log-response ratio of total biomass.",
@@
                     n = 100,
-                    titlestring = "Herbivory - A",
+                    titlestring = "Herbivory - Water Content",
                     ylabstring = "Log-response ratio of total biomass.",
@@
                     n = 100,
-                    titlestring = "Herbivory - B",
+                    titlestring = "Herbivory - Abundance",
                     ylabstring = "Log-response ratio of total biomass.",
@@
                     n = 100,
-                    titlestring = "Herbivory - B",
+                    titlestring = "Herbivory - LRR SLA",
                     ylabstring = "Log-response ratio of total biomass.",
@@
 
+# Fungicide
+
+fwaterp <- predictAndVisualize(f.mod, 
+                    f.dD,
+                    "water.cont",
+                    n = 100,
+                    titlestring = "Fungicide - Water Content",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Water content")
+
+fsitep <- predictAndVisualize(f.mod, 
+                    f.dD,
+                    "fsite",
+                    n = 100,
+                    titlestring = "Fungicide - Site",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Site")
+
+frichp <- predictAndVisualize(f.mod, 
+                    f.dD,
+                    "richness",
+                    n = 100,
+                    titlestring = "Fungicide - Richness",
+                    ylabstring = "Log-response ratio of total biomass.",
+                    xlabstring = "Number of species per plot")
+
 # Plot
-ggpubr::ggarrange(pslap, psitep, plabup)
-
-ggpubr::ggarrange(ilabup)
-
-ggpubr::ggarrange(hwaterp, hsitep, hlabup, hlrrslap)
+ggpubr::ggarrange(pslap +
+                    scale_y_continuous(limits = c(-5, 10)),
+                  fwaterp + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  hwaterp + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1, legend = F)
+ggpubr::ggarrange(psitep + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  fsitep + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  hsitep + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1,legend = F)
+ggpubr::ggarrange(prichp + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  frichp + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1, legend = F)
+ggpubr::ggarrange(plabup + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  ilabup + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  hlabup + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1, legend = F)
+ggpubr::ggarrange(plrrslap +
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  hlrrslap + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1, common.legend = T)
 

### 2023-04-14 ms1_data_analysis.Rmd
All results added.

@@
 library(AICcmodavg)
-
-```
-
-We collected data on biomass (total of `r round(sum(BIOMASS$tot_bio), 0)` kg), density, and plant traits: area lost to herbivores, water content, leaf dry matter content and specific leaf area for `r length(unique(data$unified_names))` woody plant species from three tropical forest sites on elevations spanning `r 1900 - 200` m from 200 to 1900 m a.s.l.
-Low elevation is Wanang (200 m a.s.l.), Mid-elevation is Numba (700 m a.s.l.), and High elevation is Yawan (1900 m a.s.l.).
+library(caTools)
+library(randomForest)
+library(rpart)
+library(rpart.plot)
+library(C50)
+library(printr)
+library(betapart)
+library(lsmeans)
+library(emmeans)
+library(multcomp)
+library(multcompView)
+
+```
+
+We collected data on biomass (total of `r round(sum(BIOMASS$tot_bio), 0)` kg), density, and plant traits: area lost to herbivores, water content, leaf dry matter content and specific leaf area for `r length(unique(data$unified_names))` woody plant species from three tropical forest sites: low elevation (Wanang, 200 m a.s.l.); mid-elevation (Numba, 700 m a.s.l.); high elevation (Yawan, 1900 m a.s.l.).
 We refer to the sites as Low, Mid- and High elevation throughout the text.
 
-Figure 1. Median, 1st, second quadrilles, 1.5 IRQ distance, and outliers (labeled) for logarithm of biomass, richness diversity, and density for the three studied location/elevation.
-```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure S1 Characteistics"}
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
 
@@
 
+# Tests and labels
+
+labelsS1 <- data.frame()
+
+for (desc in unique(elevdataT$descriptor)){
+  for (trt in unique(elevdataT$treatment)){
+    subdat <- elevdataT %>% filter(descriptor == desc & treatment == trt)
+    subdat %>% 
+      group_by(fsite) %>%
+      summarize(ycoord = mean(valueTrans)) %>%
+      dplyr::select(ycoord)
+    mod1 <- lm(valueTrans ~ fsite, data = subdat)
+    anova(mod1)
+    lbls <- cld(object = emmeans(mod1, specs = "fsite"),Letters = c("a","b","c"))$.group
+
+    print(paste(desc, trt, paste0(lbls, collapse = "")))
+
+    labelsS1 <- rbind(labelsS1, data.frame(descriptor = desc,
+                                           sites = factor(c("wanang", "numba", "yawan"),
+                                                          levels = c("wanang", "numba", "yawan")),
+                                           labels = lbls,
+                                           treatment = trt,
+                                           ycoord = subdat %>% 
+                                             group_by(fsite) %>%
+                                             summarize(ycoord = mean(valueTrans)) %>%
+                                             dplyr::select(ycoord)))
+  }
+}
+
+```
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap=" Figure S1.
+Median, 1st, 2nd quadrilles, 1.5 IRQ distance, and outliers (labeled) for logarithm of biomass, richness diversity, and density at three studied locations/elevations.
+Letters indicate statisitcal differences at the $\alpha$ = 0.05 level from the post hos Tukey test."}
+# Plot
 ggplot(elevdataT, aes(y = valueTrans, x = fsite, label = garden))+
   geom_boxplot()+
-  geom_text(aes(label=outlier), hjust = 1.2, vjust = -0.0005, na.rm=TRUE)+
+  geom_text(aes(label=outlier), 
+            size = 2, 
+            alpha = 0.6,
+            hjust = 1.2, 
+            vjust = -0.0005, 
+            na.rm=TRUE) +
   facet_grid(vars(descriptor), vars(treatment), scales = "free")+
-  theme_bw()
-
-```
-
-Looking at the control plots Numba (the mid-elevation in our study) has the highest abundance and number of species.
-Predictably productivity decreases with elevation.
+  geom_text(data = labelsS1,
+            aes(x = sites, y = ycoord + 2.5, group = trt,
+                label = labels),
+            size = 3) +
+  theme_bw() +
+  theme(strip.text = element_text(size =5.5),
+        axis.text.x = element_text(angle = 45,
+                                   hjust = 1))+
+  xlab("")
+
+```
+
+Elevations were similar in their baseline productivity (total biomass).
+Mid and high elevations had the highest baseline productivity (total biomass).
+Abundance and species richness.
+decreased with elevation.
 Control plots in Numba are more variable than in other two elevations.
@@
 for (file in names(dats)){
-  print(file)
+  # print(file)
   dat <- dats[[file]]
@@
 
-# library(dplyr)
-# library(ggplot2)
-# library(GGally)
-# library(lme4)
-# library(lmerTest)
-# library(ggpmisc)
-# library(tibble)
-
 # LRR correlation plot
-
 ftreats <- unique(panel_data$ftreatment)
@@
 
-      print(paste(resp, desc,loc))
+      # print(paste(resp, desc,loc))
 
@@
 
-      print(paste(desc,loc))
+      # print(paste(desc,loc))
 
@@
 
-### COMMUNITY COMPOSITION
+## COMMUNITY COMPOSITION
 
@@
 
-  print(site.id)
+  # print(site.id)
 
@@
 
-  print(anova(rdamod, permutations = h, by = "terms"))
+  # print(anova(rdamod, permutations = h, by = "terms"))
 
@@
 
-# library(betapart)
-# 
-# csites <- grepl("_c\\b", COMMVARS$plotid)
-# fsites <- grepl("_f", COMMVARS$plotid)
-# isites <- grepl("_i", COMMVARS$plotid)
-# hsites <- grepl("_h", COMMVARS$plotid)
-# psites <- grepl("_p", COMMVARS$plotid)
-# 
-# wsites <- COMMVARS$site == "wanang"
-# nsites <- COMMVARS$site == "numba"
-# ysites <- COMMVARS$site == "yawan"
-# 
-# # B-C dissimilarities in C plots between sites
-# cdatc <- COMMDATA[grepl("c$", rownames(COMMDATA)), ]
-# cvars <- COMMVARS[rownames(cdatc), ]
-# 
-# # Make a matrix
+csites <- grepl("_c\\b", COMMVARS$plotid)
+fsites <- grepl("_f", COMMVARS$plotid)
+isites <- grepl("_i", COMMVARS$plotid)
+hsites <- grepl("_h", COMMVARS$plotid)
+psites <- grepl("_p", COMMVARS$plotid)
+
+wsites <- COMMVARS$site == "wanang"
+nsites <- COMMVARS$site == "numba"
+ysites <- COMMVARS$site == "yawan"
+
+# B-C dissimilarities in C plots between sites
+
+cdatc <- COMMDATA[grepl("c$", rownames(COMMDATA)), ]
+cvars <- COMMVARS[rownames(cdatc), ]
+
+# Make a matrix
 # for(i in 1:dim(cdatc)[1]){
-#   
 #   # Compare all other sites to the current i
@@
 #   }
-#   
 # }
-# 
-# # Comparisons between
-# wc.beta <- beta.pair.abund(COMMDATA[COMMVARS[wsites & csites,]$plotid,  ])
-# 
-# beta.df <- data.frame()
-# 
-# for(siteid in c("w","y","n")){
-#   for(treat in c("_c\\b","_f","_h","_i","_p")){
-#     
-#     # Subset the data
-#     nms <- rownames(COMMDATA)
-#     sel.nms <- nms[grepl(siteid, nms) & grepl(treat, nms)]
-#     site.treat.comm <- COMMDATA[sel.nms,]
-#     
-#     print(rownames(site.treat.comm))
-#     
-#     # Calculate beta diversity
-#     subbeta <- beta.pair.abund(site.treat.comm)
-#     
-#     # Add to a data frame
-#     balanced.df <- data.frame(site = siteid,
-#                               treat = treat,
-#                               vals = as.numeric(subbeta$beta.bray.bal),
-#                               component = "balanced")
-#     
-#     gradient.df <- data.frame(site = siteid,
-#                               treat = treat,
-#                               vals = as.numeric(subbeta$beta.bray.gra),
-#                               component = "gradient")
-#     
-#     bray.df <- data.frame(site = siteid,
-#                               treat = treat,
-#                               vals = as.numeric(subbeta$beta.bray),
-#                               component = "bray")
-#     
-#     sub.df <- rbind(balanced.df, gradient.df, bray.df)
-#       
-#     # Merge the datset
-#     beta.df <- rbind(beta.df,sub.df)
-#     
-#   }
-# }
+
+# Comparisons between
+wc.beta <- beta.pair.abund(COMMDATA[COMMVARS[wsites & csites,]$plotid,  ])
+
+beta.df <- data.frame()
+
+for(siteid in c("w","y","n")){
+  for(treat in c("_c\\b","_f","_h","_i","_p")){
+
+    # Subset the data
+    nms <- rownames(COMMDATA)
+    sel.nms <- nms[grepl(siteid, nms) & grepl(treat, nms)]
+    site.treat.comm <- COMMDATA[sel.nms,]
+
+    print(rownames(site.treat.comm))
+
+    # Calculate beta diversity
+    subbeta <- beta.pair.abund(site.treat.comm)
+
+    # Add to a data frame
+    balanced.df <- data.frame(site = siteid,
+                              treat = treat,
+                              vals = as.numeric(subbeta$beta.bray.bal),
+                              component = "balanced")
+
+    gradient.df <- data.frame(site = siteid,
+                              treat = treat,
+                              vals = as.numeric(subbeta$beta.bray.gra),
+                              component = "gradient")
+
+    bray.df <- data.frame(site = siteid,
+                              treat = treat,
+                              vals = as.numeric(subbeta$beta.bray),
+                              component = "bray")
+
+    sub.df <- rbind(balanced.df, gradient.df, bray.df)
+
+    # Merge the datset
+    beta.df <- rbind(beta.df,sub.df)
+
+  }
+}
+
 # beta.df
-# 
-# # Fix the site order
-# beta.df$fsite <- factor(beta.df$site, levels = c("w","n","y"))
-# 
-# # Transform 0s to 0.001 and 1s to 0.999
-# beta.df$trans.vals = ifelse(beta.df$vals == 1, 0.999999, 
-#                             ifelse(beta.df$vals == 0, 0.000001, beta.df$vals))
-# 
-# # Tests for individual facets
-# 
-# library(lsmeans)
-# library(emmeans)
-# library(multcomp)
-# library(multcompView)
-# 
+
+# Fix the site order
+beta.df$fsite <- factor(beta.df$site, levels = c("w","n","y"))
+
+# Transform 0s to 0.001 and 1s to 0.999
+beta.df$trans.vals = ifelse(beta.df$vals == 1, 0.999999,
+                            ifelse(beta.df$vals == 0, 0.000001, beta.df$vals))
+
+# Tests for individual facets
+cld.dat = data.frame()
+prws.comp <- list()
+
+# CLD data for comparisons between elevations
+for (comp in unique(beta.df$component)){
+  print(comp)
+  for (trt in unique(beta.df$treat)){
+    print(trt)
+
+    test.df <- beta.df %>%
+      filter(treat == trt,
+             component == comp)
+
+    # print(test.df)
+
+    mod1 <- betareg::betareg(trans.vals ~ fsite, data = test.df)
+
+    lsmod1 <- lsmeans(mod1,
+                      pairwise ~ fsite,
+                      adjust="tukey")
+
+    groups <- cld(lsmod1, Letters = c("abcdefg"))
+
+    y_loc = test.df %>%
+      group_by(fsite) %>%
+      summarise(y_loc = mean(trans.vals))
+
+    sub.cld <- data.frame(component = comp,
+               treat = trt,
+               fsite = groups$fsite,
+               gr.label = groups$.group,
+               y_loc = y_loc$y_loc[order(groups$fsite)])
+
+    print(y_loc)
+    print(sub.cld)
+
+    prws.comp[[paste(comp,trt)]] <- lsmod1
+
+    cld.dat = rbind(cld.dat, sub.cld)
+
+  }
+}
+
 # cld.dat = data.frame()
 # prws.comp <- list()
-# 
-# # CLD data for comparisons between elevations
-# for (comp in unique(beta.df$component)){
-#   print(comp)
-#   for (trt in unique(beta.df$treat)){
-#     print(trt)
-#     
-#     test.df <- beta.df %>%
-#       filter(treat == trt,
-#              component == comp)
-#     
-#     # print(test.df)
-#     
-#     mod1 <- betareg::betareg(trans.vals ~ fsite, data = test.df)
-#     
-#     lsmod1 <- lsmeans(mod1,
-#                       pairwise ~ fsite,
-#                       adjust="tukey")
-#     
-#     groups <- cld(lsmod1, Letters = c("abcdefg"))
-#     
-#     y_loc = test.df %>%
-#       group_by(fsite) %>%
-#       summarise(y_loc = mean(trans.vals))
-#     
-#     sub.cld <- data.frame(component = comp,
-#                treat = trt,
-#                fsite = groups$fsite,
-#                gr.label = groups$.group,
-#                y_loc = y_loc$y_loc[order(groups$fsite)])
-#     
-#     print(y_loc)
-#     print(sub.cld)
-#     
-#     prws.comp[[paste(comp,trt)]] <- lsmod1
-#     
-#     cld.dat = rbind(cld.dat, sub.cld)
-#     
-#   }
-# }
-# 
-# # cld.dat = data.frame()
-# # prws.comp <- list()
 # 
@@
 # # library(glmmTMB)
-# 
-# pairwise.comp.res <- data.frame()
-# for (comp in unique(beta.df$component)){
-#   
-#   for (ste in unique(beta.df$fsite)){
-#     print(comp)
-#     print(ste)
-#     
-#     test.df <- beta.df %>%
-#       filter(fsite == ste,
-#              component == comp)
-#     
-#     # print(test.df)
-#     
-#     mod1 <- betareg::betareg(trans.vals ~ treat, data = test.df)
-#     
-#     breg.table <- summary(mod1)$coefficients$mean
-#     breg.pvals <- breg.table[,4]
-#     
-#     y_loc = test.df %>%
-#       group_by(treat) %>%
-#       summarise(y_loc = mean(trans.vals))
-#     
-#     pcr.row <- data.frame(comp = comp,
-#                           fsite = ste,
-#                           component = comp,
-#                           treat = names(breg.pvals),
-#                           p.vals = breg.pvals,
-#                           y_loc = y_loc$y_loc,
-#                           y_loc_trt = y_loc$treat)
-#     
-#     pcr.row <- pcr.row %>%
-#       mutate(label = ifelse(p.vals <= 0.05 & treat != "(Intercept)", "*",""))
-#     
-#     pairwise.comp.res <- rbind(pairwise.comp.res,pcr.row)
-#     
-#   }
-# }
-# 
-# pairwise.comp.res
-# 
+
+pairwise.comp.res <- data.frame()
+for (comp in unique(beta.df$component)){
+
+  for (ste in unique(beta.df$fsite)){
+    print(comp)
+    print(ste)
+
+    test.df <- beta.df %>%
+      filter(fsite == ste,
+             component == comp)
+
+    # print(test.df)
+
+    mod1 <- betareg::betareg(trans.vals ~ treat, data = test.df)
+
+    breg.table <- summary(mod1)$coefficients$mean
+    breg.pvals <- breg.table[,4]
+
+    y_loc = test.df %>%
+      group_by(treat) %>%
+      summarise(y_loc = mean(trans.vals))
+
+    pcr.row <- data.frame(comp = comp,
+                          fsite = ste,
+                          component = comp,
+                          treat = names(breg.pvals),
+                          p.vals = breg.pvals,
+                          y_loc = y_loc$y_loc,
+                          y_loc_trt = y_loc$treat)
+
+    pcr.row <- pcr.row %>%
+      mutate(label = ifelse(p.vals <= 0.05 & treat != "(Intercept)", "*",""))
+
+    pairwise.comp.res <- rbind(pairwise.comp.res,pcr.row)
+
+  }
+}
+
+pairwise.comp.res
+
 # # Plot
@@
 # #      res=800,compression='lzw')
-# beta.part.plot <- ggplot(beta.df, aes(x = fsite, y = vals))+
-#   geom_jitter(width = 0.1, alpha = 0.2, shape = 16, size = 1.8)+
-#   stat_summary(fun.y = mean, na.rm = T,
-#                geom = "point", size = 3, col = "gray30")+
-#   stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
-#   stat_summary(fun.y = mean, aes(group = treat),
-#                geom = "line", alpha = 0.8, col = "gray30", lty = 2)+
-#   geom_text(data = cld.dat,
-#             aes(x = fsite, y = y_loc + 0.5,
-#                 label = gr.label),
-#             size = 3)+
-#   facet_grid(cols = vars(component), rows = vars(treat))
-# beta.part.plot + theme_bw()
-# 
-# # Plot for the component comparisons
-# pairwise.comp.res$fsite <- factor(pairwise.comp.res$fsite, levels = c("w","n","y"))
-# 
+
+beta.part.plot1 <- ggplot(beta.df, aes(x = fsite, y = vals))+
+  geom_jitter(width = 0.1, alpha = 0.2, shape = 16, size = 1.8)+
+  stat_summary(fun.y = mean, na.rm = T,
+               geom = "point", size = 3, col = "gray30")+
+  stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
+  stat_summary(fun.y = mean, aes(group = treat),
+               geom = "line", alpha = 0.8, col = "gray30", lty = 2)+
+  geom_text(data = cld.dat,
+            aes(x = fsite, y = y_loc + 0.5,
+                label = gr.label),
+            size = 3)+
+  facet_grid(cols = vars(component), rows = vars(treat)) + 
+  theme_bw()
+
+# Plot for the component comparisons
+pairwise.comp.res$fsite <- factor(pairwise.comp.res$fsite, levels = c("w","n","y"))
+
 # # tiff(filename='figures/beta_part_panel_treats.tif',
@@
 # #      res=800,compression='lzw')
-# 
-# # Alternative boxplot
-# beta.part.plot <- ggplot(beta.df, aes(x = treat, y = vals))+
-#   geom_boxplot()+
-#   # stat_summary(fun.y = mean, na.rm = T,
-#   #              geom = "point", size = 3, col = "gray30")+
-#   # stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
-#   geom_text(data = pairwise.comp.res,
-#             aes(x = y_loc_trt, y = y_loc + 0.5,
-#                 label = label), cex = 6,  col = "red", lwd = 2)+
-#   facet_grid(component ~ fsite)
-# 
-# beta.part.plot + theme_bw()
-
+
+# Alternative boxplot
+beta.part.plot2 <- ggplot(beta.df, aes(x = treat, y = vals))+
+  geom_boxplot()+
+  # stat_summary(fun.y = mean, na.rm = T,
+  #              geom = "point", size = 3, col = "gray30")+
+  # stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
+  geom_text(data = pairwise.comp.res,
+            aes(x = y_loc_trt, y = y_loc + 0.5,
+                label = label), cex = 6,  col = "red", lwd = 2)+
+  facet_grid(component ~ fsite) + 
+  theme_bw()
+```
+
+```{r, fig.width=10, fig.height=18}
+ggpubr::ggarrange(beta.part.plot1,
+                  beta.part.plot2, nrow = 2)
 ```
@@
 
-### Trait based analyses
+### Traits and species faith
 
@@
 # Subset a data-set to only control and treatment plots 
-subDat <- dwbRichF %>%
-  filter(treatment %in% subsetVars)
-
-lgsData <- data.frame()
-
-for (st in unique(subDat$site)){
-  for (gard in unique(subDat$garden)){
-
-    stGardDat <- subDat %>%
-            filter(site == st,
-                   garden == gard)
-
-    # Some quality control
-    if(dim(stGardDat)[1] == 0){
-      next
-      }
-    if(length(unique(stGardDat$treatment)) < 2){
-      next
-      }
-
-    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
-    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
-
-    spLost <- specInCtr[!(specInCtr %in% specInTrt)]
-    spGained <- specInTrt[!(specInTrt %in% specInCtr)]
-
-    stGardDat$pg.cat <- ifelse(stGardDat$unified_names %in% spLost, "lost",
-                               ifelse(stGardDat$unified_names %in% spGained, "gained", "stayed"))
-
-    lgsData <- rbind(lgsData, stGardDat)
-
-  }
-}
+# subDat <- dwbRichF %>%
+#   filter(treatment %in% subsetVars)
+# 
+# lgsData <- data.frame()
+# 
+# for (st in unique(subDat$site)){
+#   for (gard in unique(subDat$garden)){
+#     
+#     stGardDat <- subDat %>%
+#             filter(site == st,
+#                    garden == gard)
+#     
+#     # Some quality control
+#     if(dim(stGardDat)[1] == 0){
+#       next
+#       }
+#     if(length(unique(stGardDat$treatment)) < 2){
+#       next
+#       }
+#     
+#     specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
+#     specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
+#     
+#     spLost <- specInCtr[!(specInCtr %in% specInTrt)]
+#     spGained <- specInTrt[!(specInTrt %in% specInCtr)]
+#     
+#     stGardDat$pg.cat <- ifelse(stGardDat$unified_names %in% spLost, "lost",
+#                                ifelse(stGardDat$unified_names %in% spGained, "gained", "stayed"))
+#     
+#     
+#     lgsData <- rbind(lgsData, stGardDat)
+#     
+#     
+#   }
+# }
 ```
@@
 
-```{r}
-
-# Function that performs RF analysis for a given treatment
-
-# It queries the lgsData that needs to be preapred using separate code (as above)
-
-getLgsData <- function (comp.trt = "i", data) {
-
-  # This function queries the data to compute log
-
-}
-
-# Returns:
-# 1) Classifier RF table
-# 2) Confusion matrix
-# 3) Regression tree example graph
-
-runRF <- function (lgsData) {
-
-}
-
-```
-
-```{r}
+```{r, results='asis'}
 # Random Forest
 # Loading package
-library(caTools)
-library(randomForest)
+
+rfAnalysis <- function(lgsData) {
 
@@
                 lest.leaf.area, richness, ltb, lsla.d, treatment)
-
   # Try standardized, non-correlated variables
@@
     filter(!is.infinite(lest.leaf.area)) %>%  
-  dplyr::select(pg.cat, perc.herb, water.cont,
-                richness, ltb, lsla.d)
-
+    dplyr::select(pg.cat, perc.herb, water.cont, richness, ltb, lsla.d)
   # Standardized
@@
 
-# classifier_RF
-
   # Predicting the Test set results
@@
 
+  # Confusion Matrix
+  confusion_mtx = table(test[, 1], y_pred)
+
+  # Regression C50
+  C50Fit <- C5.0(pg.cat ~., data=train,
+                 control = C5.0Control(minCases = 5))
+  C50pred <- predict(object=C50Fit, 
+                     newdata=test, 
+                     type="class")
+
+  # return(list(rf_confusion_matrix = confusion_mtx,
+  #             classifier_for_varImpPlot = varImpPlot(classifier_RF),
+  #             reg_tree_table = table(C50pred, test$pg.cat),
+  #             reg_fit_for_plot = plot(C50Fit, cex = 0.5)))
+
+  return(list(rf_confusion_matrix = confusion_mtx,
+              classifier_for_varImpPlot = classifier_RF,
+              reg_tree_table = table(C50pred, test$pg.cat),
+              reg_fit_for_plot = C50Fit))
+
+  }
+
+# rfData <- lgsData %>%
+#   mutate(pg.cat = as.factor(pg.cat),
+#          lest.leaf.area = log(est.leaf.area),
+#          lstem = log(no_stems)) %>%
+#   dplyr::select(pg.cat, lstem, perc.herb, water.cont,
+#                 lest.leaf.area, richness, ltb, lsla.d, treatment)
+# 
+# # Try standardized, non-correlated variables
+# rfDataRF <- rfData %>%
+#   filter(!is.infinite(lest.leaf.area)) %>%
+#   dplyr::select(pg.cat, perc.herb, water.cont,
+#                 richness, ltb, lsla.d)
+# 
+# # Standardized
+# rfDataRF[,-1] <- scale(rfDataRF[,-1])
+# 
+# # Splitting data in train and test data
+# split <- sample.split(rfDataRF, SplitRatio = 0.7)
+# train <- subset(rfDataRF, split == "TRUE")
+# test <- subset(rfDataRF, split == "FALSE")
+#   
+# classifier_RF = randomForest(x = train[-1],
+#                              y = train$pg.cat,
+#                              ntree = 500)
+
+# classifier_RF
+
+# Predicting the Test set results
+# y_pred = predict(classifier_RF, newdata = test[-1])
+
 # Variable importance plot
 # varImpPlot(classifier_RF)
-```
-
-Below is a confusion matrix for the RF
-
-```{r, results='asis'}
-# Confusion Matrix
-confusion_mtx = table(test[, 1], y_pred)
-print(knitr::kable(confusion_mtx))
 ```
@@
 
-library(rpart)
-library(rpart.plot)
-
-# Regression C50
-library(C50)
-library(printr)
-
-C50Fit <- C5.0(pg.cat ~., data=train,
-               control = C5.0Control(minCases = 5))
-
-C50pred <- predict(object=C50Fit, 
-                   newdata=test, 
-                   type="class")
-
-plot(C50Fit, cex = 0.5)
-
-```
-
-Lets see how good of a job the regression tree did.
-
-```{r, results='asis'}
-
-print(knitr::kable(table(C50pred, test$pg.cat)))
+# Run all random forests
+
+subsetVars <- c("c", "p")
+source('code/get_data_ready_for_the_sla_analyses.R')
+prfres <- rfAnalysis(lgsData)
+
+subsetVars <- c("c", "i")
+source('code/get_data_ready_for_the_sla_analyses.R')
+irfres <- rfAnalysis(lgsData)
+
+subsetVars <- c("c", "f")
+source('code/get_data_ready_for_the_sla_analyses.R')
+frfres <- rfAnalysis(lgsData)
+
+subsetVars <- c("c", "h")
+source('code/get_data_ready_for_the_sla_analyses.R')
+hrfres <- rfAnalysis(lgsData)
 
@@
 From the leaf frame analyses that spanned all plant species in all treatment plots we saw high plasticity of that trait.
+
+```{r, figures-side, fig.show="hold", out.width="50%"}
+varImpPlot(prfres$classifier_for_varImpPlot,
+           main = 'Variable Importance for the predator exclusion')
+varImpPlot(irfres$classifier_for_varImpPlot,
+           main = 'Variable Importance for the insect exclusion')
+varImpPlot(frfres$classifier_for_varImpPlot,
+           main = 'Variable Importance for the fungi exclusion')
+varImpPlot(hrfres$classifier_for_varImpPlot,
+           main = 'Variable Importance for the herbivore addition')
+```
 
@@
 ```
-```{r}
+```{r, results='asis'}
 # Test for the relationship and the differences

### 2023-04-21 ms1_data_analysis.Rmd
Adding RDA position vs trait analyses.

@@
 
-## Experimental design
+# Experimental design
 
@@
 
+# Results
+
 We collected data on biomass (total of `r round(sum(BIOMASS$tot_bio), 0)` kg), density, and plant traits: area lost to herbivores, water content, leaf dry matter content and specific leaf area for `r length(unique(data$unified_names))` woody plant species from three tropical forest sites: low elevation (Wanang, 200 m a.s.l.); mid-elevation (Numba, 700 m a.s.l.); high elevation (Yawan, 1900 m a.s.l.).
@@
 
-      print(paste(desc,st,trt))
+      # print(paste(desc,st,trt))
 
@@
 
-    print(paste(desc, trt, paste0(lbls, collapse = "")))
+    # print(paste(desc, trt, paste0(lbls, collapse = "")))
 
@@
 Elevations were similar in their baseline productivity (total biomass).
-Mid and high elevations had the highest baseline productivity (total biomass).
-Abundance and species richness.
-decreased with elevation.
-Control plots in Numba are more variable than in other two elevations.
-Yawan has one site with lowest abundance (garden number seven).However no significant differences in variance were detected (Levene's and Bartellet's tests)
+At the high elevation the plant abundance, richness and diversity was significantly higher than at the low elevation.
+Mid elevation has many outliers.
+However, no significant differences in variance were detected (Table S1)
 
 ## General effects of biotic factors
-
 We calculated log-response ratios for plant biomass, species richness, species diversity and stem density to be able to compare the effect sizes at individual elevations.
@@
 Table 1. Test results for over-dispersion all descriptors at three studied elevations.
-
 ```{r, echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
@@
 
-Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation.
-```{r}
+Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation. General Least Squares assumes over-dispersion in the data
+```{r, results='asis'}
 sjPlot::tab_df(model_comp)
@@
 
-There is generally little evidence for the over-dispersion in the LRR values at individual elevations.
+There is generally no evidence for the over-dispersion in the LRR values at individual elevations.
 Only woody stem density for the predator exclusion showed some evidence of over-dispersion.
@@
 # Create a factor variable for sites
-panel_data$fsite <- factor(panel_data$site, labels = c("Wanang",
-                                      "Numba",
-                                      "Yawan"))
+panel_data$fsite <- factor(panel_data$site, labels = c("Low",
+                                      "Mid",
+                                      "High"))
 
@@
   theme(strip.text = element_text(size =5.5),
-        axis.text.x = element_text(angle = 45,
-                                   hjust = 1))+
+        axis.text.x = element_text(angle = 0,
+                                   hjust = 0.5))+
   geom_hline(yintercept = 0, lty = 2)+
@@
 Insects reduced biomass only at the mid-elevation, increased diversity at high and richness at high and low elevation.
-At high elevation predator removal caused an increase in diversity, richness and density of woody plants.Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.
+At high elevation predator removal caused an increase in diversity, richness and density of woody plants.
+Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.
 
@@
 ## 3.1 Analyses for each elevation
-
+ord.list <- list()
 plot.list <- list()
@@
                 data = SUB.VARS)
+
+  ord.list[[site.id]] <- rdamod
 
@@
     geom_text(data = species.dat %>%
-                filter(names %in% sign.names), aes(label = names))+
+                filter(names %in% sign.names), 
+              aes(label = names, alpha = 0.2))+
     geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
@@
 ```{r}
-
-# NEEDS TO BE FIXED
-
-csites <- grepl("_c\\b", COMMVARS$plotid)
-fsites <- grepl("_f", COMMVARS$plotid)
-isites <- grepl("_i", COMMVARS$plotid)
-hsites <- grepl("_h", COMMVARS$plotid)
-psites <- grepl("_p", COMMVARS$plotid)
-
-wsites <- COMMVARS$site == "wanang"
-nsites <- COMMVARS$site == "numba"
-ysites <- COMMVARS$site == "yawan"
+# csites <- grepl("_c\\b", COMMVARS$plotid)
+# fsites <- grepl("_f", COMMVARS$plotid)
+# isites <- grepl("_i", COMMVARS$plotid)
+# hsites <- grepl("_h", COMMVARS$plotid)
+# psites <- grepl("_p", COMMVARS$plotid)
+
+# wsites <- COMMVARS$site == "wanang"
+# nsites <- COMMVARS$site == "numba"
+# ysites <- COMMVARS$site == "yawan"
 
@@
 
-cdatc <- COMMDATA[grepl("c$", rownames(COMMDATA)), ]
-cvars <- COMMVARS[rownames(cdatc), ]
-
-# Make a matrix
-# for(i in 1:dim(cdatc)[1]){
-#   # Compare all other sites to the current i
-#   for (j in (1:dim(cdatc)[1])[-i]) {
-#     val <- beta.pair.abund(cdatc[c(i,j), ])
-#   }
-# }
+# cdatc <- COMMDATA[grepl("c$", rownames(COMMDATA)), ]
+# cvars <- COMMVARS[rownames(cdatc), ]
 
 # Comparisons between
-wc.beta <- beta.pair.abund(COMMDATA[COMMVARS[wsites & csites,]$plotid,  ])
+# wc.beta <- beta.pair.abund(COMMDATA[COMMVARS[wsites & csites,]$plotid,  ])
 
@@
 
-for(siteid in c("w","y","n")){
+# siteid <- "^w"
+# treat <- "_c\\b"
+
+for(siteid in c("^w","^y","^n")){
   for(treat in c("_c\\b","_f","_h","_i","_p")){
@@
     site.treat.comm <- COMMDATA[sel.nms,]
-
-    print(rownames(site.treat.comm))
 
@@
 
-    # Merge the datset
+    # Merge the data-set
     beta.df <- rbind(beta.df,sub.df)
@@
 # Fix the site order
-beta.df$fsite <- factor(beta.df$site, levels = c("w","n","y"))
+beta.df$fsite <- factor(beta.df$site, levels = c("^w","^n","^y"),
+                        labels = c("w","n","y"))
 
@@
 for (comp in unique(beta.df$component)){
-  print(comp)
   for (trt in unique(beta.df$treat)){
-    print(trt)
 
@@
              component == comp)
-
-    # print(test.df)
 
@@
 
-    print(y_loc)
-    print(sub.cld)
-
     prws.comp[[paste(comp,trt)]] <- lsmod1
@@
 }
-
-# cld.dat = data.frame()
-# prws.comp <- list()
-# 
-# # For data for comparisons between treatments we need pairwise comps 
-# # (control vs treat) - this can be obtained from summary of the betareg
-# 
-# # library(glmmTMB)
 
@@
   for (ste in unique(beta.df$fsite)){
-    print(comp)
-    print(ste)
+    # print(comp)
+    # print(ste)
 
@@
 }
-
-pairwise.comp.res
-
-# # Plot
-# # tiff(filename='figures/beta_part_panel.tif',
-# #      height=5600,
-# #      width=5600,
-# #      units='px',
-# #      res=800,compression='lzw')
 
@@
 # Plot for the component comparisons
-pairwise.comp.res$fsite <- factor(pairwise.comp.res$fsite, levels = c("w","n","y"))
-
-# # tiff(filename='figures/beta_part_panel_treats.tif',
-# #      height=5600,
-# #      width=5600,
-# #      units='px',
-# #      res=800,compression='lzw')
+pairwise.comp.res$fsite <- factor(pairwise.comp.res$fsite,
+                                  levels = c("w","n","y"))
 
@@
             aes(x = y_loc_trt, y = y_loc + 0.5,
-                label = label), cex = 6,  col = "red", lwd = 2)+
+                label = label), cex = 6,  col = "red")+
   facet_grid(component ~ fsite) + 
@@
 
-```{r, fig.width=10, fig.height=18}
+```{r, fig.width=10, fig.height=16}
 ggpubr::ggarrange(beta.part.plot1,
@@
 Exclusion of top predators has the same effect as insect exclusion on species turnover in Numba and Yawan.
-This is a probable result for a strong top-down control of the tri trophic interactions.
+This is a probable result for a strong top-down control of the tri-trophic interactions.
 Addition of generalist herbivores also correlated with the predator effects.
@@
 
-# rfData <- lgsData %>%
-#   mutate(pg.cat = as.factor(pg.cat),
-#          lest.leaf.area = log(est.leaf.area),
-#          lstem = log(no_stems)) %>%
-#   dplyr::select(pg.cat, lstem, perc.herb, water.cont,
-#                 lest.leaf.area, richness, ltb, lsla.d, treatment)
-# 
-# # Try standardized, non-correlated variables
-# rfDataRF <- rfData %>%
-#   filter(!is.infinite(lest.leaf.area)) %>%
-#   dplyr::select(pg.cat, perc.herb, water.cont,
-#                 richness, ltb, lsla.d)
-# 
-# # Standardized
-# rfDataRF[,-1] <- scale(rfDataRF[,-1])
-# 
-# # Splitting data in train and test data
-# split <- sample.split(rfDataRF, SplitRatio = 0.7)
-# train <- subset(rfDataRF, split == "TRUE")
-# test <- subset(rfDataRF, split == "FALSE")
-#   
-# classifier_RF = randomForest(x = train[-1],
-#                              y = train$pg.cat,
-#                              ntree = 500)
-
-# classifier_RF
-
-# Predicting the Test set results
-# y_pred = predict(classifier_RF, newdata = test[-1])
-
-# Variable importance plot
-# varImpPlot(classifier_RF)
 ```
@@
 
-```{r}
-# CWM vs richness
-```
-
 Community-weighted mean SLA (CWM SLA) was higher at low diversity plots.
@@
 require(scales)
-p1 <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.sla, color = treatment))+
-  geom_point(aes(size = total.bio)) + 
+p1 <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.sla))+
+  geom_point(aes(size = total.bio, color = treatment)) + 
+  stat_smooth(method = "lm")+
   scale_x_continuous(trans = 'log2',
@@
 ```
+
 ```{r, results='asis'}
@@
 
-### Can SLA predict change in composition?
-
+### Can traits predict change in composition?
 SLA doesn’t predict RDA position along C->I, C->H, C->P, C -> F axes, at any elevation.
-
-```{r}
-
-# Figure
-
-```
-
 Community composition change under a treatment may be independent of SLA.
@@
 Reflects the abilities of light capture and adaption to environments of plants (Gao, Wang, and Zhang 2022)]
+
+```{r}
+# Calculate position of species and site along a given treatment vector.
+function (oridnation, treatment, type = c('site','species')) {
+  NULL
+}
+
+surfData <- data.woody.brut %>%
+  filter(site == "wanang") %>%
+  group_by(unified_names) %>%
+  summarise(lsla.d = log(mean(sla.d, na.rm=T)),
+            water.cont = mean(water.cont, na.rm = T),
+            herbivory = mean(perc.herb, na.rm = T))
+
+# Add row names to sla data
+surfDataDf <- as.data.frame(surfData)
+tibble::column_to_rownames(surfData, var = 'unified_names')
+rownames(surfDataDf) <- surfDataDf$unified_names
+
+# Get the treatment vector.
+nm <- "f"
+# names(ord.list[["w"]])
+trtVcts <- ord.list[["w"]]$CCA$biplot[nm, c(1,2)]
+
+# Get species vector.
+all.specs <- ord.list[["w"]]$CCA$v[, c(1,2)]
+usednms <- surfData$unified_names[surfData$unified_names %in% rownames(all.specs)]
+specVcts <- all.specs[usednms,]
+
+# Projection of species onto treatment vector.
+VdotU <- ((specVcts %*% trtVcts))
+uNormSq <- (trtVcts %*% trtVcts)
+projUV <- (VdotU/as.numeric(uNormSq)) %*% t(as.matrix(trtVcts))
+
+# Get norm for each projection vector.
+positions <- apply(projUV, 1, function(x) sqrt(x %*% x))
+
+# Add positions to the data frame.
+modDat <- data.frame(value = c(surfDataDf[usednms, ]$lsla.d,
+                               surfDataDf[usednms, ]$water.cont,
+                               surfDataDf[usednms, ]$herbivory),
+                     label = rep(c("SLA", "Water Content", "Herbivory"), 
+                                 each = length(positions)),
+                     pos = positions[usednms])
+
+testDat <- data.frame(sla = surfDataDf[usednms, ]$lsla.d,
+                      water = surfDataDf[usednms, ]$water.cont,
+                      herb = surfDataDf[usednms, ]$herbivory,
+                      pos = positions[usednms])
+
+ggplot(modDat, aes(x = pos, y = value)) +
+  geom_point()+
+  stat_smooth(method = "lm") + 
+  facet_wrap(~label, scales = 'free')
+
+# Tests + phylogeny!!!!
+glsmod1 <- gls(pos ~ sla + water + herb, data = testDat)
+glsmod2 <- gls(pos ~ sla +       + herb, data = testDat)
+glsmod3 <- gls(pos ~ sla + water       , data = testDat)
+glsmod4 <- gls(pos ~               herb, data = testDat)
+glsmod5 <- gls(pos ~       water       , data = testDat)
+glsmod6 <- gls(pos ~ sla               , data = testDat)
+
+summary(glsmod1)
+summary(glsmod2)
+summary(glsmod3)
+summary(glsmod4)
+summary(glsmod5)
+summary(glsmod6)
+
+AIC(glsmod1,glsmod2,glsmod3,glsmod4,glsmod5,glsmod6)
+
+summary(glsmod1)
+
+myTree
+
+# Correlate positions with individual species' with their sla. CWM we can leave out because we know it correlates with the .
+
+```
 

### 2023-04-24 ms1_data_analysis.Rmd
RDA axis predictions with phylogeny.

@@
 
+library(AICcmodavg)
+library(betapart)
+library(caTools)
+library(C50)
+library(dplyr)
+library(emmeans)
+library(effects)
 library(ggplot2)
-library(dplyr)
-library(sjPlot)
-library(AICcmodavg)
-library(caTools)
+library(lsmeans)
+library(multcomp)
+library(multcompView)
+library(printr)
 library(randomForest)
@@
 library(rpart.plot)
-library(C50)
-library(printr)
-library(betapart)
-library(lsmeans)
-library(emmeans)
-library(multcomp)
-library(multcompView)
+library(sjPlot)
 
@@
 }
-
-# ggplot(fidata, aes(x = x, y = y)) +
-#   geom_point(size = 2, alpha = 0.4)+
-#   stat_smooth(data = fidata %>%
-#                 filter(sign == "signif"),
-#               method ="lm")+
-#   geom_hline(yintercept = 0)+
-#   geom_vline(xintercept = 0)+
-#   ylab("Effect size of insects")+
-#   xlab("Effect size of fungi")+
-#   theme_bw()+
-#   facet_grid(descriptor~site)
 
@@
 
-# ENCLOSED IN THE get_ready_for_the_sla_analyses
-# Subset a data-set to only control and treatment plots 
-# subDat <- dwbRichF %>%
-#   filter(treatment %in% subsetVars)
-# 
-# lgsData <- data.frame()
-# 
-# for (st in unique(subDat$site)){
-#   for (gard in unique(subDat$garden)){
-#     
-#     stGardDat <- subDat %>%
-#             filter(site == st,
-#                    garden == gard)
-#     
-#     # Some quality control
-#     if(dim(stGardDat)[1] == 0){
-#       next
-#       }
-#     if(length(unique(stGardDat$treatment)) < 2){
-#       next
-#       }
-#     
-#     specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
-#     specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
-#     
-#     spLost <- specInCtr[!(specInCtr %in% specInTrt)]
-#     spGained <- specInTrt[!(specInTrt %in% specInCtr)]
-#     
-#     stGardDat$pg.cat <- ifelse(stGardDat$unified_names %in% spLost, "lost",
-#                                ifelse(stGardDat$unified_names %in% spGained, "gained", "stayed"))
-#     
-#     
-#     lgsData <- rbind(lgsData, stGardDat)
-#     
-#     
-#   }
-# }
 ```
@@
 
-```{r}
+```{r, results='asis'}
+
 # Calculate position of species and site along a given treatment vector.
-function (oridnation, treatment, type = c('site','species')) {
-  NULL
-}
-
-surfData <- data.woody.brut %>%
-  filter(site == "wanang") %>%
+
+traitPredsCommComp <- function (data = data.woody.brut,
+                                ordination = ord.list,
+                                phylogeny = myTree.rand.poly,
+                                treatment = "f", st = "wanang") {
+
+  surfData <- data %>%
+  filter(site == st) %>%
   group_by(unified_names) %>%
@@
   surfDataDf <- as.data.frame(surfData)
-tibble::column_to_rownames(surfData, var = 'unified_names')
   rownames(surfDataDf) <- surfDataDf$unified_names
@@
   # Get the treatment vector.
-nm <- "f"
-# names(ord.list[["w"]])
-trtVcts <- ord.list[["w"]]$CCA$biplot[nm, c(1,2)]
+  trtVcts <- ordination[[substr(st, 1, 1)]]$CCA$biplot[treatment, c(1,2)]
 
   # Get species vector.
-all.specs <- ord.list[["w"]]$CCA$v[, c(1,2)]
-usednms <- surfData$unified_names[surfData$unified_names %in% rownames(all.specs)]
+  all.specs <- ordination[[substr(st, 1, 1)]]$CCA$v[, c(1,2)]
+  usednms <- surfDataDf$unified_names[surfDataDf$unified_names %in% rownames(all.specs)]
   specVcts <- all.specs[usednms,]
@@
                         herb = surfDataDf[usednms, ]$herbivory,
-                      pos = positions[usednms])
-
-ggplot(modDat, aes(x = pos, y = value)) +
-  geom_point()+
-  stat_smooth(method = "lm") + 
-  facet_wrap(~label, scales = 'free')
+                        pos = positions[usednms],
+                        unified_names = usednms)
+
+  # print(ggplot(modDat, aes(x = pos, y = value)) +
+  #   geom_point()+
+  #   stat_smooth(method = "lm") + 
+  #   facet_wrap(~label, scales = 'free'))
+
+  # Phylogeny ready
+  testDatTree <- testDat %>%
+    filter(unified_names %in% phylogeny$tip.label)
 
   # Tests + phylogeny!!!!
-glsmod1 <- gls(pos ~ sla + water + herb, data = testDat)
-glsmod2 <- gls(pos ~ sla +       + herb, data = testDat)
-glsmod3 <- gls(pos ~ sla + water       , data = testDat)
-glsmod4 <- gls(pos ~               herb, data = testDat)
-glsmod5 <- gls(pos ~       water       , data = testDat)
-glsmod6 <- gls(pos ~ sla               , data = testDat)
-
-summary(glsmod1)
-summary(glsmod2)
-summary(glsmod3)
-summary(glsmod4)
-summary(glsmod5)
-summary(glsmod6)
-
-AIC(glsmod1,glsmod2,glsmod3,glsmod4,glsmod5,glsmod6)
-
-summary(glsmod1)
-
-myTree
-
-# Correlate positions with individual species' with their sla. CWM we can leave out because we know it correlates with the .
+  glsmod1 <- gls(pos ~ sla + water + herb,
+                 data = testDatTree,
+                 method = "ML",
+                 correlation = corMartins(2.718282,
+                       phy = phylogeny,
+                       form = ~unified_names,
+                       fixed = T),
+                 na.action = na.omit)
+
+  # glsmod2 <- gls(pos ~ sla +       + herb, 
+  #                data = testDatTree,
+  #                method = "ML",
+  #                correlation = corMartins(2.718282,
+  #                      phy = myTree.rand.poly,
+  #                      form = ~unified_names,
+  #                      fixed = T),
+  #                na.action = na.omit)
+  # 
+  # glsmod3 <- gls(pos ~ sla + water       ,
+  #                data = testDatTree,
+  #                method = "ML",
+  #                correlation = corMartins(2.718282,
+  #                      phy = myTree.rand.poly,
+  #                      form = ~unified_names,
+  #                      fixed = T),
+  #                na.action = na.omit)
+  # 
+  # glsmod4 <- gls(pos ~               herb,
+  #                data = testDatTree,
+  #                method = "ML",
+  #                correlation = corMartins(2.718282,
+  #                      phy = myTree.rand.poly,
+  #                      form = ~unified_names,
+  #                      fixed = T),
+  #                na.action = na.omit)
+  # 
+  # glsmod5 <- gls(pos ~       water       ,
+  #                data = testDatTree,
+  #                method = "ML",
+  #                correlation = corMartins(2.718282,
+  #                      phy = myTree.rand.poly,
+  #                      form = ~unified_names,
+  #                      fixed = T),
+  #                na.action = na.omit)
+  # glsmod6 <- gls(pos ~ sla               ,
+  #                data = testDatTree,
+  #                method = "ML",
+  #                correlation = corMartins(2.718282,
+  #                      phy = myTree.rand.poly,
+  #                      form = ~unified_names,
+  #                      fixed = T),
+  #                na.action = na.omit)
+
+  # Summary
+
+  tab_model(glsmod1, title = paste(st, treatment))
+
+  # Prediction of the full model
+  # effx1 <- effect("sla", glsmod1, partial.residuals=T)
+  # effx2 <- effect("water", glsmod1, partial.residuals=T)
+  # effx3 <- effect("herb", glsmod1, partial.residuals=T)
+
+  # plot(effx1)
+  # plot(effx2)
+  # plot(effx3)
+
+}
+
+traitPredsCommComp(treatment = "f", st = "wanang")
+traitPredsCommComp(treatment = "f", st = "numba")
+traitPredsCommComp(treatment = "f", st = "yawan")
+
+traitPredsCommComp(treatment = "i", st = "wanang")
+traitPredsCommComp(treatment = "i", st = "numba")
+traitPredsCommComp(treatment = "i", st = "yawan")
 

### 2023-04-24 code/pgls_data_prep.R

@@ -47,7 +47,7 @@ replaceWithMedian <- function(data.woody, name = "piper_recessum", threshold = 1
   
   # Subset the data
   piper.dat <- data.woody %>% filter(unified_names == name)
-  print("step 1")
+  # print("step 1")
   
   # Change extremes and NA's into a meadian value
   new.sla.val <- median(piper.dat$sla.d.m2kg1, na.rm = T)
@@ -60,11 +60,11 @@ replaceWithMedian <- function(data.woody, name = "piper_recessum", threshold = 1
   }
   
   piper.dat.no.na <- piper.dat[!is.na(piper.dat$sla.d.m2kg1),]
-  print("step 2")
+  # print("step 2")
   
   # Replace extremes with the median
   piper.dat.no.na[piper.dat.no.na$sla.d.m2kg1 > threshold, ]$sla.d.m2kg1 <- new.sla.val
-  print("step 3")
+  # print("step 3")
   
   # Add these to the main data
   data.woody[data.woody$X %in% piper.dat.no.na$X, 

### 2023-05-30 ms1_data_analysis.Rmd
Adding the Raup-Crick code. Early stage of the introduction

@@
 
-# Experimental design
+# Materials and methods
 
@@
 
+We calculated log-response ratios for plant biomass, species richness, species diversity and stem density to be able to compare the effect sizes at individual elevations.
+
 # Results
@@
 We refer to the sites as Low, Mid- and High elevation throughout the text.
-
-```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
-
-elev.bind <- rbind(BIOMASS %>% dplyr::select(site, garden, treatment, tot_bio), 
-                   RICHNESS %>% dplyr::select(site, garden, treatment, sp_no), 
-                   RICHNESS %>% dplyr::select(site, garden, treatment, simpson), 
-                   DENSITY %>% dplyr::select(site, garden, treatment, tot.stem.no))
-
-findoutlier <- function(x) {
-  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
-}
-
-elevdata <- elev.bind %>% 
-  mutate(value = ifelse(!is.na(tot_bio), tot_bio, 
-                        ifelse(!is.na(sp_no), sp_no, 
-                               ifelse(!is.na(simpson), simpson, tot.stem.no)))) %>%
-  mutate(descriptor = ifelse(!is.na(tot_bio), "Total biomass [kg]", 
-                        ifelse(!is.na(sp_no), "Number of species", 
-                               ifelse(!is.na(simpson), "Simpson's diversity", "Abundance")))) %>%
-  dplyr::select(site, garden, treatment, value, descriptor) %>%
-  mutate(fsite = factor(site, levels = c("wanang", "numba","yawan"))) %>%
-  arrange(descriptor, fsite, treatment, garden)
-
-elevdataT <- elevdata %>% 
-  mutate(valueTrans = ifelse(descriptor == "Simpson's diversity", 
-                             value, log(value)))
-
-outliers <- c()
-for (desc in unique(elevdataT$descriptor)) {
-  for (st in unique(elevdataT$site)) {
-    for (trt in unique(elevdataT$treatment)) {
-
-      # print(paste(desc,st,trt))
-
-      outliers <- c(outliers,findoutlier(pull(elevdataT %>% filter(descriptor == desc,
-                          site == st,
-                          treatment == trt), valueTrans)))
-
-    }
-
-  }
-}
-elevdataT$outlierBool <- outliers 
-elevdataT <- elevdataT %>% mutate(outlier = ifelse(outlierBool, garden, ""))
-
-# Tests and labels
-
-labelsS1 <- data.frame()
-
-for (desc in unique(elevdataT$descriptor)){
-  for (trt in unique(elevdataT$treatment)){
-    subdat <- elevdataT %>% filter(descriptor == desc & treatment == trt)
-    subdat %>% 
-      group_by(fsite) %>%
-      summarize(ycoord = mean(valueTrans)) %>%
-      dplyr::select(ycoord)
-    mod1 <- lm(valueTrans ~ fsite, data = subdat)
-    anova(mod1)
-    lbls <- cld(object = emmeans(mod1, specs = "fsite"),Letters = c("a","b","c"))$.group
-
-    # print(paste(desc, trt, paste0(lbls, collapse = "")))
-
-    labelsS1 <- rbind(labelsS1, data.frame(descriptor = desc,
-                                           sites = factor(c("wanang", "numba", "yawan"),
-                                                          levels = c("wanang", "numba", "yawan")),
-                                           labels = lbls,
-                                           treatment = trt,
-                                           ycoord = subdat %>% 
-                                             group_by(fsite) %>%
-                                             summarize(ycoord = mean(valueTrans)) %>%
-                                             dplyr::select(ycoord)))
-  }
-}
-
-```
-
-```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap=" Figure S1.
-Median, 1st, 2nd quadrilles, 1.5 IRQ distance, and outliers (labeled) for logarithm of biomass, richness diversity, and density at three studied locations/elevations.
-Letters indicate statisitcal differences at the $\alpha$ = 0.05 level from the post hos Tukey test."}
-# Plot
-ggplot(elevdataT, aes(y = valueTrans, x = fsite, label = garden))+
-  geom_boxplot()+
-  geom_text(aes(label=outlier), 
-            size = 2, 
-            alpha = 0.6,
-            hjust = 1.2, 
-            vjust = -0.0005, 
-            na.rm=TRUE) +
-  facet_grid(vars(descriptor), vars(treatment), scales = "free")+
-  geom_text(data = labelsS1,
-            aes(x = sites, y = ycoord + 2.5, group = trt,
-                label = labels),
-            size = 3) +
-  theme_bw() +
-  theme(strip.text = element_text(size =5.5),
-        axis.text.x = element_text(angle = 45,
-                                   hjust = 1))+
-  xlab("")
-
-```
 
@@
 ## General effects of biotic factors
-We calculated log-response ratios for plant biomass, species richness, species diversity and stem density to be able to compare the effect sizes at individual elevations.
-
-Table 1. Test results for over-dispersion all descriptors at three studied elevations.
-```{r, echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
-sjPlot::tab_df(lrr_test_df)
-```
-
-```{r,  echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
-
-# Results of a GLS model with over-dispersion vs LM model for density LRR in the predator exclusion treatment.
-trtdata <-  trtdata  <- DENSITY.LR %>%
-    filter(treatment == 'p')
-
-trtlm <- lm(lratio ~ 0 + site, data = trtdata)
-
-trtgls <- nlme::gls(lratio ~ 0 + site, data = trtdata,
-                 weights = varIdent(form = ~1|site))
-
-# summary(trtlm)
-# summary(trtgls)
-
-aovtrt <- anova(trtlm)
-
-testStat <- aovtrt$`F-value`
-testP <- aovtrt$`p-value`
-
-# Tukey post-hoc
-emm <- emmeans(object = trtlm,
-               specs = "site",
-               adjust = "mvt")
-
-mod_means <- multcomp::cld(object = emm,
-                           Letters = letters)$.group
-
-zero.dif.lm <- ifelse(summary(trtlm)$coefficients[,4] > 0.05, "","*")
-zero.dif.gls <- ifelse(summary(trtgls)$tTable[,4] > 0.05, "","*")
-
-lmres <- as_tibble(summary(trtlm)$coefficients)
-lmres <- rename(lmres, Estimate = 'Estimate', 
-               `Standard Error` = `Std. Error`, 
-               `T-value` =`t value`,
-               `P-value` = `Pr(>|t|)`)
-
-glsres <- as_tibble(summary(trtgls)$tTable)
-glsres <- rename(glsres,
-                 Estimate = 'Value', 
-                 `Standard Error` = "Std.Error", 
-                 `T-value` = "t-value", 
-                 `P-value` = 'p-value')
-
-model_comp <- rbind(lmres, glsres)
-model_comp$Site <- rep(c("Wanang", "Numba", "Yawan"), 2)
-model_comp$Model <- rep(c("LM", "GLS"), each = 3)
-
-```
-
-Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation. General Least Squares assumes over-dispersion in the data
-```{r, results='asis'}
-sjPlot::tab_df(model_comp)
-```
-
-There is generally no evidence for the over-dispersion in the LRR values at individual elevations.
-Only woody stem density for the predator exclusion showed some evidence of over-dispersion.
-Including this in statistical modle (using 'varIdent(form = ~1|site)' as weights in the GLS function) showed significant difference of LRR in Yawan, that was not observed for LM model.
 
@@
 ```
-
-## Impact of treatments
-
-Positive effect of pathogenic fungi on richness was found at low and high elevations (Fig.
-1) and on density at low elevation.
-Insects reduced biomass only at the mid-elevation, increased diversity at high and richness at high and low elevation.
-At high elevation predator removal caused an increase in diversity, richness and density of woody plants.
-Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.
-
-```{r}
-
-# Check whether the RICHNESS.LR for predators is correct.
-# trtdata <-  trtdata  <- RICHNESS.LR %>%
-#     filter(treatment == 'p')
-
-# trtlm <- lm(lratio ~ 0 + site, data = trtdata)
-
-# summary(trtlm)
-
-# Tukey post-hoc
-# emm <- emmeans(object = trtlm,
-#                specs = "site",
-#                adjust = "mvt")
-
-# mod_means <- multcomp::cld(object = emm,
-#                            Letters = letters)$.group
-
-```
+Positive effects of pathogenic fungi on richness were found at low and high elevations (Fig.
+1) and on density at low elevations.
+Insects decreased biomass only at mid-elevations, increased diversity at high elevations, and increased richness at high and low elevations.
+At high elevations, predator removal increased woody plant diversity, richness, and density.
+Increases in the number of generalist herbivores reduced woody plant biomass and density at low elevations and richness at low and high elevations.
 
 ## LRR correlations
-
-Differences of LRR averages from zero may suggest that the effects are random, generally weak and noisy.
-We now look at the LRR correlation for different treatments to probe some possible mechanisms of the biotic factors' action.
 
@@
 
-# ggplot(cpdata, aes(x = x, y = y)) +
-#   geom_point(size = 2, alpha = 0.4)+
-#   stat_smooth(data = cpdata %>%
-#                 filter(sign == "signif"),
-#               method ="lm")+
-#   geom_hline(yintercept = 0)+
-#   geom_vline(xintercept = 0)+
-#   ylab("Effect size of insects")+
-#   xlab("Effect size")+
-#   theme_bw()+
-#   facet_grid(response*descriptor~site)
-
-```
-
-#### OBSERVATIONS
-
-Correlation between LRR of insect effects were not very common and were observed in `r 9/(3*8) * 100` % of cases.
-Correlations were more often found in Yawan (five out of eight studied correlations) than in Numba (four out of eight) and no significant correlation was found in Wanang.
-If we also include correlations between predator exclusion (P) and P+A this ratio gets significantly higher for Yawan (9 out of 12 correlations).
-
-All observed correlations of the top-predators effects were positively correlated with insect effects, whereas all effects of herbivore addition were negatively correlated with the insect effects.
-
-#### EXPLANATION
-
-If a plant community has a potential to be strongly negatively affected by insects (negative LRR of insects, strong top-down control), then the P+A would result in weaker or positive effects of additional generalist herbivores.
-Top down control seem to be stronger in the highlands and we see more frequent correlation of the LRR's.
-(Model where Intra-guild predators are being strongly limited by top predators.)
-
-What exactly happened after addition can however only be speculated here.
-In case of a  strong bottom-up control should not result in any sort of correlation like in Wanang.
-What I think happened was that spiders could be more active in the exclosures which resulted in their effects being similar to that of a simple exclusion.
-
-A few aspects of insect effects on plants can be important here: 
-If community is saturated with herbivores (Konno) then addition may have negative results.
-Herbivore resource limitation type (relative or absolute). Feeding mode change, adaptive trade off, risk avoidance. All these can only be speculated.
-
-Generalists are worse in consuming any individual plant biomass, therefore they may have positive effects on plants, when natural insect communities have negative effects.
-It is possible that addition of herbivores cause som disruption in the functioning of the insect community effect on vegetation.
-There were no correlation between factors’ effects on vegetation at the low elevation. Herbivore addition and removal were not correlated as expected by the strong bottom-up control.
-
-Simulation showed that it cannot be simply an effect of the effect-abundance relationship between abundance of insects and their effect (quadratic and linear function result in positive relationship).
-
-Top-down effects of predators and their positive correlation with insect effects can be explained well by the bottom-up control of insects by plants.
-It was also assumed in the above simulations.
-Empirically we do not observe any correlation between insect abundance and strength of the effect (lrr_simualtions.R at the end).
-There is also no correlation between richness of a community and its LRR for any of the treatments (lrr_simulations.R).
-
-#### SOME DISCLAIMERS
-
-It is noteworthy that plots where herbivores were added were also covered with nets to prevent increased predation, wich is known to be correlated with increased leaf damage (Katka).
-Therefore it cannot be interpreted directly as a simple increase in herbivory pressure even though this is what should be happening assuming simple chain model and top-down control by top-predators.
-We do not know the faith of the added predators. It is most likely that they did not cause a permanent increase in the species abundance (graph here). However, herbivore damage in general was...
-
-### Predator exclusion and predator exclusion with insect addition
-
-The hypothesis that I have stated above, about increased activity of spiders in exclosures is here being investigated further.
-I want to see whether the addition of herbivores has a positive correlation with the regular exclosures.
-Positive correlation would suggest that the herbivore addition had little to no effect, and that spiders responding to a lack of predators would control the herbivore population even with boosted abundances.
+ggplot(cpdata, aes(x = x, y = y)) +
+  geom_point(size = 2, alpha = 0.4)+
+  stat_smooth(data = cpdata %>%
+                filter(sign == "signif"),
+              method ="lm")+
+  geom_hline(yintercept = 0)+
+  geom_vline(xintercept = 0)+
+  ylab("Effect size of insects")+
+  xlab("Effect size")+
+  theme_bw()+
+  facet_grid(response*descriptor~site)
+
+```
+
+Correlations between LRR of insect effects were not very frequent and were observed in 50% of cases.
+Correlations were found more frequently in Yawan (five out of eight correlations examined) than in Numba (four out of eight) and no significant correlation was found in Wanang.
+When we also include correlations between predator exclusion (P) and P+A, this ratio becomes significantly higher for Yawan (9 out of 12 correlations).
+All observed correlations of top predator effects were positively correlated with insect effects, whereas all effects of herbivore admixture were negatively correlated with insect effects.
 
@@
 
-# ggplot(fidata, aes(x = x, y = y)) +
-#   geom_point(size = 2, alpha = 0.4)+
-#   stat_smooth(data = fidata %>%
-#                 filter(sign == "signif"),
-#               method ="lm")+
-#   geom_hline(yintercept = 0)+
-#   geom_vline(xintercept = 0)+
-#   xlab("Predator removal")+
-#   ylab("Predator removal and insect addition")+
-#   theme_bw()+
-#   facet_grid(descriptor~site)
+ggplot(fidata, aes(x = x, y = y)) +
+  geom_point(size = 2, alpha = 0.4)+
+  stat_smooth(data = fidata %>%
+                filter(sign == "signif"),
+              method ="lm")+
+  geom_hline(yintercept = 0)+
+  geom_vline(xintercept = 0)+
+  xlab("Predator removal")+
+  ylab("Predator removal and insect addition")+
+  theme_bw()+
+  facet_grid(descriptor~site)
 
@@
 }
+
+ggplot(fidata, aes(x = x, y = y)) +
+  geom_point(size = 2, alpha = 0.4)+
+  stat_smooth(data = fidata %>%
+                filter(sign == "signif"),
+              method ="lm")+
+  geom_hline(yintercept = 0)+
+  geom_vline(xintercept = 0)+
+  xlab("Insects")+
+  ylab("Fungicide")+
+  theme_bw()+
+  facet_grid(descriptor~site)
 
@@
 ```
+
+# Discussion
+
+## LRR correlations
+
+  If a plant community has a potential to be strongly negatively affected by insects (negative LRR of insects, strong top-down control), then the P+A results in weaker or positive effects of additional generalist herbivores.
+Top down control seem to be stronger in the highlands and we see more frequent correlation of the LRR's.
+(Model where Intra-guild predators are being strongly limited by top predators.)
+  What exactly happened after addition can however only be speculated here.
+In case of a  strong bottom-up control should not result in any sort of correlation like in Wanang.
+What I think happened was that spiders could be more active in the exclosures which resulted in their effects being similar to that of a simple exclusion.
+  A few aspects of insect effects on plants can be important here: 
+If community is saturated with herbivores (Konno) then addition may have negative results.
+Herbivore resource limitation type (relative or absolute). Feeding mode change, adaptive trade off, risk avoidance. All these can only be speculated.
+  Generalists are worse in consuming any individual plant biomass, therefore they may have positive effects on plants, when natural insect communities have negative effects.
+It is possible that addition of herbivores cause som disruption in the functioning of the insect community effect on vegetation.
+There were no correlation between factors’ effects on vegetation at the low elevation. Herbivore addition and removal were not correlated as expected by the strong bottom-up control.
+  Simulation showed that it cannot be simply an effect of the effect-abundance relationship between abundance of insects and their effect (quadratic and linear function result in positive relationship).
+  Top-down effects of predators and their positive correlation with insect effects can be explained well by the bottom-up control of insects by plants.
+It was also assumed in the above simulations.
+Empirically we do not observe any correlation between insect abundance and strength of the effect (lrr_simualtions.R at the end).
+There is also no correlation between richness of a community and its LRR for any of the treatments (lrr_simulations.R).
+  It is noteworthy that plots where herbivores were added were also covered with nets to prevent increased predation, wich is known to be correlated with increased leaf damage (Katka).
+Therefore it cannot be interpreted directly as a simple increase in herbivory pressure even though this is what should be happening assuming simple chain model and top-down control by top-predators.
+We do not know the faith of the added predators. It is most likely that they did not cause a permanent increase in the species abundance (graph here). However, herbivore damage in general was...
+  The hypothesis that I have stated above, about increased activity of spiders in exclosures is here being investigated further.
+I want to see whether the addition of herbivores has a positive correlation with the regular exclosures.
+Positive correlation would suggest that the herbivore addition had little to no effect, and that spiders responding to a lack of predators would control the herbivore population even with boosted abundances.
+
+If a plant community has the potential to be strongly negatively impacted by insects (negative LRR from insects, strong top-down control), then P+A would result in weaker or positive impacts from additional generalist herbivores.
+Top-down control appears to be stronger in the highlands and we see more frequent correlation of LRRs.
+(Model in which intra-guild predators are strongly constrained by top predators).
+
+However, what exactly happened after the addition can only be speculated here.
+With strong bottom-up control, there should not be a correlation like in Wanang.
+I suspect that the spiders were more active in the exclusion areas, causing their effects to be similar to simple exclusion.
+A few aspects of insect effects on plants may be important here:
+If the community is saturated with herbivores, the addition may have negative effects.
+Nature of resource limitation by herbivores (relative or absolute).
+Change in diet, adaptive trade-off, risk avoidance.
+All of these can only be speculated about.
+Generalists are worse at consuming each plant biomass, so they may have positive effects on plants, while natural insect communities have negative effects.
+It is possible that the addition of herbivores causes a disruption in the functioning of the insect community that affects vegetation.
+There was no correlation between the effects of the factors on vegetation at the low elevations.
+The addition and removal of herbivores were not correlated, as expected from the strong bottom-up control.
+
+The simulations showed that it could not simply be an effect of the relationship between insect abundance and its effect (quadratic and linear functions yield a positive relationship).
+ The top-down effects of predators and their positive correlation with insect effects can be well explained by the bottom-up control of insects by plants.
+This was also assumed in the simulations above.
+Empirically, we observe no correlation between insect abundance and effect strength (lrr_simualtions.R at the end).
+There is also no correlation between the abundance of a community and its LRR for any of the treatments (lrr_simulations.R).
+It is worth noting that the plots where herbivores were used were also covered with nets to prevent increased predation, which is known to correlate with increased leaf damage (Katka).
+Therefore, this cannot be directly interpreted as a simple increase in herbivore pressure, although it should be, assuming a simple chain model and top-down control by top predators.
+We do not know the beliefs of the additional predators.
+Most likely, they did not lead to a permanent increase in species diversity (graph here).
+However, the damage from herbivores in general has been...
+The hypothesis I made above, about increased spider activity in the enclosures, is being further explored here.
+I want to see if the addition of herbivores has a positive correlation with the regular exclosures.
+A positive correlation would suggest that the addition of herbivores has little to no effect and that spiders responding to a lack of predators control the herbivore population even with increased abundance.
+
+# Supplement
+
+There is generally no evidence for the over-dispersion in the LRR values at individual elevations.
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
+
+elev.bind <- rbind(BIOMASS %>% dplyr::select(site, garden, treatment, tot_bio), 
+                   RICHNESS %>% dplyr::select(site, garden, treatment, sp_no), 
+                   RICHNESS %>% dplyr::select(site, garden, treatment, simpson), 
+                   DENSITY %>% dplyr::select(site, garden, treatment, tot.stem.no))
+
+findoutlier <- function(x) {
+  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
+}
+
+elevdata <- elev.bind %>% 
+  mutate(value = ifelse(!is.na(tot_bio), tot_bio, 
+                        ifelse(!is.na(sp_no), sp_no, 
+                               ifelse(!is.na(simpson), simpson, tot.stem.no)))) %>%
+  mutate(descriptor = ifelse(!is.na(tot_bio), "Total biomass [kg]", 
+                        ifelse(!is.na(sp_no), "Number of species", 
+                               ifelse(!is.na(simpson), "Simpson's diversity", "Abundance")))) %>%
+  dplyr::select(site, garden, treatment, value, descriptor) %>%
+  mutate(fsite = factor(site, levels = c("wanang", "numba","yawan"))) %>%
+  arrange(descriptor, fsite, treatment, garden)
+
+elevdataT <- elevdata %>% 
+  mutate(valueTrans = ifelse(descriptor == "Simpson's diversity", 
+                             value, log(value)))
+
+outliers <- c()
+for (desc in unique(elevdataT$descriptor)) {
+  for (st in unique(elevdataT$site)) {
+    for (trt in unique(elevdataT$treatment)) {
+
+      # print(paste(desc,st,trt))
+
+      outliers <- c(outliers,findoutlier(pull(elevdataT %>% filter(descriptor == desc,
+                          site == st,
+                          treatment == trt), valueTrans)))
+
+    }
+
+  }
+}
+elevdataT$outlierBool <- outliers 
+elevdataT <- elevdataT %>% mutate(outlier = ifelse(outlierBool, garden, ""))
+
+# Tests and labels
+
+labelsS1 <- data.frame()
+
+for (desc in unique(elevdataT$descriptor)){
+  for (trt in unique(elevdataT$treatment)){
+    subdat <- elevdataT %>% filter(descriptor == desc & treatment == trt)
+    subdat %>% 
+      group_by(fsite) %>%
+      summarize(ycoord = mean(valueTrans)) %>%
+      dplyr::select(ycoord)
+    mod1 <- lm(valueTrans ~ fsite, data = subdat)
+    anova(mod1)
+    lbls <- cld(object = emmeans(mod1, specs = "fsite"),Letters = c("a","b","c"))$.group
+
+    # print(paste(desc, trt, paste0(lbls, collapse = "")))
+
+    labelsS1 <- rbind(labelsS1, data.frame(descriptor = desc,
+                                           sites = factor(c("wanang", "numba", "yawan"),
+                                                          levels = c("wanang", "numba", "yawan")),
+                                           labels = lbls,
+                                           treatment = trt,
+                                           ycoord = subdat %>% 
+                                             group_by(fsite) %>%
+                                             summarize(ycoord = mean(valueTrans)) %>%
+                                             dplyr::select(ycoord)))
+  }
+}
+
+```
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap=" Figure S1.
+Median, 1st, 2nd quadrilles, 1.5 IRQ distance, and outliers (labeled) for logarithm of biomass, richness diversity, and density at three studied locations/elevations.
+Letters indicate statisitcal differences at the $\alpha$ = 0.05 level from the post hos Tukey test."}
+# Plot
+ggplot(elevdataT, aes(y = valueTrans, x = fsite, label = garden))+
+  geom_boxplot()+
+  geom_text(aes(label=outlier), 
+            size = 2, 
+            alpha = 0.6,
+            hjust = 1.2, 
+            vjust = -0.0005, 
+            na.rm=TRUE) +
+  facet_grid(vars(descriptor), vars(treatment), scales = "free")+
+  geom_text(data = labelsS1,
+            aes(x = sites, y = ycoord + 2.5, group = trt,
+                label = labels),
+            size = 3) +
+  theme_bw() +
+  theme(strip.text = element_text(size =5.5),
+        axis.text.x = element_text(angle = 45,
+                                   hjust = 1))+
+  xlab("")
+
+```
+
+**Table S1**. Test results for over-dispersion all descriptors at three studied elevations.
+```{r, echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
+sjPlot::tab_df(lrr_test_df)
+```
+
+Only woody stem density for the predator exclusion showed some evidence of over-dispersion.
+Including this in statistical model (using *varIdent(form = ~1|site)* as weights in the GLS function) showed significant difference of LRR in Yawan, that was not observed for LM model.
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
+
+elev.bind <- rbind(BIOMASS %>% dplyr::select(site, garden, treatment, tot_bio), 
+                   RICHNESS %>% dplyr::select(site, garden, treatment, sp_no), 
+                   RICHNESS %>% dplyr::select(site, garden, treatment, simpson), 
+                   DENSITY %>% dplyr::select(site, garden, treatment, tot.stem.no))
+
+findoutlier <- function(x) {
+  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
+}
+
+elevdata <- elev.bind %>% 
+  mutate(value = ifelse(!is.na(tot_bio), tot_bio, 
+                        ifelse(!is.na(sp_no), sp_no, 
+                               ifelse(!is.na(simpson), simpson, tot.stem.no)))) %>%
+  mutate(descriptor = ifelse(!is.na(tot_bio), "Total biomass [kg]", 
+                        ifelse(!is.na(sp_no), "Number of species", 
+                               ifelse(!is.na(simpson), "Simpson's diversity", "Abundance")))) %>%
+  dplyr::select(site, garden, treatment, value, descriptor) %>%
+  mutate(fsite = factor(site, levels = c("wanang", "numba","yawan"))) %>%
+  arrange(descriptor, fsite, treatment, garden)
+
+elevdataT <- elevdata %>% 
+  mutate(valueTrans = ifelse(descriptor == "Simpson's diversity", 
+                             value, log(value)))
+
+outliers <- c()
+for (desc in unique(elevdataT$descriptor)) {
+  for (st in unique(elevdataT$site)) {
+    for (trt in unique(elevdataT$treatment)) {
+
+      # print(paste(desc,st,trt))
+
+      outliers <- c(outliers,findoutlier(pull(elevdataT %>% filter(descriptor == desc,
+                          site == st,
+                          treatment == trt), valueTrans)))
+
+    }
+
+  }
+}
+elevdataT$outlierBool <- outliers 
+elevdataT <- elevdataT %>% mutate(outlier = ifelse(outlierBool, garden, ""))
+
+# Tests and labels
+
+labelsS1 <- data.frame()
+
+for (desc in unique(elevdataT$descriptor)){
+  for (trt in unique(elevdataT$treatment)){
+    subdat <- elevdataT %>% filter(descriptor == desc & treatment == trt)
+    subdat %>% 
+      group_by(fsite) %>%
+      summarize(ycoord = mean(valueTrans)) %>%
+      dplyr::select(ycoord)
+    mod1 <- lm(valueTrans ~ fsite, data = subdat)
+    anova(mod1)
+    lbls <- cld(object = emmeans(mod1, specs = "fsite"),Letters = c("a","b","c"))$.group
+
+    # print(paste(desc, trt, paste0(lbls, collapse = "")))
+
+    labelsS1 <- rbind(labelsS1, data.frame(descriptor = desc,
+                                           sites = factor(c("wanang", "numba", "yawan"),
+                                                          levels = c("wanang", "numba", "yawan")),
+                                           labels = lbls,
+                                           treatment = trt,
+                                           ycoord = subdat %>% 
+                                             group_by(fsite) %>%
+                                             summarize(ycoord = mean(valueTrans)) %>%
+                                             dplyr::select(ycoord)))
+  }
+}
+
+```
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap=" Figure S1.
+Median, 1st, 2nd quadrilles, 1.5 IRQ distance, and outliers (labeled) for logarithm of biomass, richness diversity, and density at three studied locations/elevations.
+Letters indicate statisitcal differences at the $\alpha$ = 0.05 level from the post hos Tukey test."}
+# Plot
+ggplot(elevdataT, aes(y = valueTrans, x = fsite, label = garden))+
+  geom_boxplot()+
+  geom_text(aes(label=outlier), 
+            size = 2, 
+            alpha = 0.6,
+            hjust = 1.2, 
+            vjust = -0.0005, 
+            na.rm=TRUE) +
+  facet_grid(vars(descriptor), vars(treatment), scales = "free")+
+  geom_text(data = labelsS1,
+            aes(x = sites, y = ycoord + 2.5, group = trt,
+                label = labels),
+            size = 3) +
+  theme_bw() +
+  theme(strip.text = element_text(size =5.5),
+        axis.text.x = element_text(angle = 45,
+                                   hjust = 1))+
+  xlab("")
+
+```
+
+```{r,  echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
+
+# Results of a GLS model with over-dispersion vs LM model for density LRR in the predator exclusion treatment.
+trtdata <-  trtdata  <- DENSITY.LR %>%
+    filter(treatment == 'p')
+
+trtlm <- lm(lratio ~ 0 + site, data = trtdata)
+
+trtgls <- nlme::gls(lratio ~ 0 + site, data = trtdata,
+                 weights = varIdent(form = ~1|site))
+
+# summary(trtlm)
+# summary(trtgls)
+
+aovtrt <- anova(trtlm)
+
+testStat <- aovtrt$`F-value`
+testP <- aovtrt$`p-value`
+
+# Tukey post-hoc
+emm <- emmeans(object = trtlm,
+               specs = "site",
+               adjust = "mvt")
+
+mod_means <- multcomp::cld(object = emm,
+                           Letters = letters)$.group
+
+zero.dif.lm <- ifelse(summary(trtlm)$coefficients[,4] > 0.05, "","*")
+zero.dif.gls <- ifelse(summary(trtgls)$tTable[,4] > 0.05, "","*")
+
+lmres <- as_tibble(summary(trtlm)$coefficients)
+lmres <- rename(lmres, Estimate = 'Estimate', 
+               `Standard Error` = `Std. Error`, 
+               `T-value` =`t value`,
+               `P-value` = `Pr(>|t|)`)
+
+glsres <- as_tibble(summary(trtgls)$tTable)
+glsres <- rename(glsres,
+                 Estimate = 'Value', 
+                 `Standard Error` = "Std.Error", 
+                 `T-value` = "t-value", 
+                 `P-value` = 'p-value')
+
+model_comp <- rbind(lmres, glsres)
+model_comp$Site <- rep(c("Wanang", "Numba", "Yawan"), 2)
+model_comp$Model <- rep(c("LM", "GLS"), each = 3)
+
+```
+
+**Table S2**. Two models testing difference of LRR of predator exclosures from zero at each elevation. General Least Squares assumes over-dispersion in the data
+```{r, results='asis'}
+sjPlot::tab_df(model_comp)
+```

### 2023-05-30 code/data_processing.R

@@ -1,0 +1,195 @@
+# rm(list = ls())
+library(dplyr)
+library(ggplot2)
+library(vegan)
+
+# 0. Read data ----
+
+mainData <- read.csv("data/main_data_v2.csv")
+trait.dat <- read.csv("data/suppl_plant_traits.csv")
+mainDataT <- read.csv("data/main.data.w.traits.csv")
+
+# Wanang has only total stem abundance, no data on stems (no dbh values),
+# Since I am not calculating basal area for the rest of the sites but only focus on abundance I can move tot abu to the dbh column
+# to build a data summary
+mainData[mainData$site == "wanang", ]$dbh <- mainData[mainData$site == "wanang", ]$tot_abu
+
+# Change treatment labels
+mainData[mainData$treatment == "FUNGICIDE", ]$treatment <- "f"
+mainData[mainData$treatment == "I", ]$treatment <- "i"
+mainData[mainData$treatment == "WEEVIL125", ]$treatment <- "h"
+mainData[mainData$treatment == "WEEVIL25", ]$treatment <- "lh"
+mainData[mainData$treatment == "CONTROL", ]$treatment <- "c"
+mainData[mainData$treatment == "PREDATOR", ]$treatment <- "p"
+mainData[mainData$treatment == "INSECTICIDE", ]$treatment <- "i"
+
+# Convert tot_bio into numeric values (when decimal points are being separated with commas)
+mainData$tot_bio <- gsub(",",".",mainData$tot_bio)
+mainData$tot_bio <- as.numeric(mainData$tot_bio)
+
+mainDataT$tot_bio <- gsub(",",".",mainDataT$tot_bio)
+mainDataT$tot_bio <- as.numeric(mainDataT$tot_bio)
+
+# Dbh is sometimes written as "no stems", zero.
+mainData[mainData$dbh == "no stems",] # NA is dedicated for plants that are not woody.
+mainData$dbh <- as.numeric(mainData$dbh)
+
+
+# 1. Total biomass data ----
+getBioData <- function (mainData){
+  
+  # produces data-set for with raw total biomass for the total plant biomass 
+  # for individual plots and treatments along elevation gradients.
+  
+  # Creates a lookup character vector (primary key).
+  lookup <- paste(mainData$plot,
+                  mainData$treatment,
+                  mainData$garden,
+                  mainData$unified_names,
[…]

### 2023-05-30 code/raup_crick_analysis_v2.R

@@ -1,0 +1,313 @@
+rm(list = ls())
+
+# Start with Wanang C and I
+source("code/data_processing.R")
+COMMDATA <- t(COMMDATA)
+
+# This script performs RC analysis on the experimental elevation plots.
+# For each plot comparison within the same treatment I constrained the
+# plant data to species and biomass observed within the SITE (but not garden).
+# This included all treatments to obtain the most reliable species list for
+# random sampling.
+
+# Declare functions ----
+### Function to subset the main plant data-set
+getPlantMat <- function (COMMDATA, site) {
+  # Takes full dataset and creates a subset for a given garden and site
+  c1 <- grepl(site, rownames(COMMDATA)) # site constraint
+  # c2 <- grepl(garden, rownames(COMMDATA)) # garden constraint
+  return(COMMDATA[c1, ])
+}
+
+### Assemble function
+assemble <- function (siteCode,
+                      plant_mat, # subset (either within garden/or within)
+                      accuracy = 0.001){
+  
+  # Samples random community with constraints defined by the given site code:
+  # number of species an dbiomass randomized is equal to the observed values.
+  
+  # Calculate characteristics of that site
+  empiricalSpNo <- sum(plant_mat[siteCode, ]>0)
+  empiricalBio  <- round(sum(plant_mat[siteCode, ]), 
+                         digits = -log(accuracy)/log(10))
+  spec_patterns <- colSums(plant_mat > 0)
+  bio_patterns <- round(colSums(plant_mat), 
+                        digits = -log(accuracy)/log(10))
+  bio_patterns[bio_patterns == 0] <- accuracy 
+  
+  # How to add variability to these probabilities?
+  spec_prob <- spec_patterns/sum(spec_patterns)
+  
+  # Sample given number of species
+  S1 <- sample(names(spec_prob),
+               empiricalSpNo,
+               replace=FALSE,
+               prob = spec_prob)
+  
+  # Probabilities for the biomass conditioned on the species sampled
+  bio_prob <- bio_patterns[S1]/sum(bio_patterns[S1])
+    
+  
+  # Initiate community with minimal biomass for the S1 sampled species
+  # This assures that all sampled species get to the community because their 
+  # probability of increment might be so small that they never get there.
[…]

### 2023-07-19 ms1_data_analysis.Rmd
Models for LRR

@@
 test.df <- data.frame()
+
+all.pairs_max_accuracy$fsite <- factor(all.pairs_max_accuracy$fsite, levels = c("Wanang",
+                                                  "Numba","Yawan"))
+
 for (st in unique(all.pairs_max_accuracy$fsite)){
@@
   mutate(label = ifelse(test.res < 0.05, "*"," "))
+test.df$fsite <- factor(test.df$fsite, levels = c("Wanang",
+                                                  "Numba",
+                                                  "Yawan"))
 
@@
   mutate(tot_bio = ifelse(is.infinite(tot_bio), NA, tot_bio),
-         sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1)) %>%
+         sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1),
+         water.cont = ifelse(is.infinite(water.cont), NA, water.cont)) %>%
   group_by(site, garden, treatment) %>%
   summarise(cwm.sla = cwmFun(sla.d.m2kg1, n.tot_bio),
+            cwm.wc = cwmFun(water.cont, n.tot_bio),
             sp.rich = n(),
@@
 require(scales)
+
+p1.wc <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.wc))+
+  geom_point(aes(size = total.bio, color = treatment)) + 
+  stat_smooth(method = "lm")+
+  scale_x_continuous(trans = 'log2',
+    breaks = trans_breaks("log2", function(x) 2^x),
+    labels = trans_format("log2", math_format(2^.x))) +
+  scale_y_continuous(trans = 'log2',
+    breaks = trans_breaks("log2", function(x) 2^x),
+    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()
+
 p1 <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.sla))+
@@
 
-ggpubr::ggarrange(p1,p2, p3)
+# ggpubr::ggarrange(p1,p2, p3)
+ggpubr::ggarrange(p1,p1.wc)
 

### 2023-07-19 ms1_supplement.Rmd
Models for LRR

---
title: "Supplement X. Bottom-up drivers of the LRR effects"
author: "Piotr Szefer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
source('code/lrr_and_community_chars_correlations.R')
```

```{r}
tab_model(biolrr.mod0)
```

```{r pressure, echo=FALSE, fig.width=10, fig.height=10}
ggpubr::ggarrange(p1, p1.1,
                  nrow = 2)
```

### 2023-07-19 code/lrr_and_community_chars_correlations.R

@@ -34,22 +34,429 @@ LRR_DEP_TEST_DATA$DivLRR = DivLRR
 LRR_DEP_TEST_DATA$ConDivVal = ConDivVal
 LRR_DEP_TEST_DATA$DenLRR = DenLRR
 LRR_DEP_TEST_DATA$ConDenVal = ConDenVal
+LRR_DEP_TEST_DATA = LRR_DEP_TEST_DATA %>%
+  mutate(SiteTreatment = paste(Site, Treatment, sep = " "))
 
 library(ggplot2)
-ggplot(LRR_DEP_TEST_DATA %>%
-         filter(Treatment == "i"), aes(BioLRR, ConRicVal)) + 
-  geom_point()
 
-ggplot(LRR_DEP_TEST_DATA %>%
-         filter(Treatment == "i"), aes(BioLRR, ConDivVal)) + 
-  geom_point()
+responses <- c("BioLRR","RichLRR","DivLRR","DenLRR")
+predictors <- c("ConBioVal","ConRicVal","ConDivVal","ConDenVal")
 
-ggplot(LRR_DEP_TEST_DATA %>%
-         filter(Treatment == "i"), aes(BioLRR, ConDenVal)) + 
-  geom_point()
+# responses <- c("BioLRR")
+# predictors <- c("ConBioVal","ConRicVal")
 
-ggplot(LRR_DEP_TEST_DATA %>%
-         filter(Treatment == "i"), aes(BioLRR, log(ConBioVal))) + 
-  geom_point()
+plot_list <- list()
+test_list <- list()
+for (resp in responses){
+  for (pred in predictors) {
+    
+    print(paste(resp, pred))
+  
+    pp <- ggplot(LRR_DEP_TEST_DATA %>%
+                   filter(Treatment == "i"), aes_string(x=pred, y=resp, 
+                                        color = "Site")) + 
+      geom_point() + 
+      geom_smooth(method = "lm")+
+      coord_trans(x = 'log')
+      
+    plot_list[[paste(resp, pred)]] <- pp
+    
+    # Test each 
+    
+    
+    
+  }
+}
+
+
+# Herbivory
+hplot_list <- list()
+
+for (resp in responses){
+  for (pred in predictors) {
+    
+    print(paste(resp, pred))
+    
+    pp <- ggplot(LRR_DEP_TEST_DATA %>%
+                   filter(Treatment == "h"), aes_string(x=pred, y=resp, 
[…]

### 2023-07-21 ms1_supplement.Rmd
Experimenting with explanatory power of baseline biomass and richness. Models and plots.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results='hide')
 ```
 
-```{r cars}
+```{r, echo = FALSE, warning=FALSE, message=FALSE, results='hide'}
 source('code/lrr_and_community_chars_correlations.R')
+source('code/elevation_descriptors.R')
 ```
 
-```{r}
-tab_model(biolrr.mod0)
+# Introduction
+
+Aim of our study was to evaluate to what extent pathogenic fungi, insect herbivores, and their predators shape community composition of pioneer stages of tropical forest regeneration.
+Further we were interested in the effect of natural elevation gradient on the direction and strength of these effects.
+
+Looking at biomass, diversity (number of species and Simpson's index values), and density of woody plant species.
+Biomass did not change much with elevation, but diversity and density peaked at the mid elevation.
+
+```{r, fig.cap="Figure 1. Characteristics of three studied sites: biomass, richness, diversity (Simpson's index) and density of woody plants at the control sites."}
+
+ggpubr::ggarrange(bio.plot, rich.plot, div.plot, dens.plot)
+
+```
+
+Now lets look at the effects that the biotic factors had on the vegetation characteristics.
+
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 1.
+Magnitude and direction of fungi, insects and their predators on biomass, richness, diversity and density of woody plants along elevation gradient: Wanang (200 m a.s.l.); Numba (750 m a.s.l.); Yawan (1900 m a.s.l.).
+Log response ratio is a log-ratio of a given descriptor value from plots where a biotic factor was present to where it was absent.
+Grey points represent empirical values.
+Letters indicate statistically significant ($\\alpha$ = 0.05) differences for pairwise comparisons between elevations within a treatment-descriptor combination with Tukey correction for multiplicity.
+For better readability letters are given only in cases where any of the mean were either different form each other or different from zero.
+Zero effect are indicated with a dashed line.
+Stars in the upper index indicate significant ($\\alpha$ = 0.05) differences of the effect mean from zero (significance of the effect at a given site)", fig.width= 8, fig.height=8}
+
+# bio, dens, div, rich
+panel_data <- data.frame()
+
+dats <- list(biomass = BIOMASS.LR, 
+             denssity = DENSITY.LR, 
+             diversity = DIVERSITY.LR, 
+             richness = RICHNESS.LR)
+
+# Stack the data and rename descriptor names
+for (file in names(dats)){
+  # print(file)
+  dat <- dats[[file]]
+  dat$descriptor <- ifelse(grepl("bio", file), "Biomass",
+                           ifelse(grepl("dens", file), "Woody plant density",
+                                  ifelse(grepl("div", file),"Diversity", "Richness")))
+  dat <- dat[, c("site", "garden", "treatment", "lratio","descriptor")]
+  panel_data <- rbind(dat, panel_data)
+}
+
+# Make a renamed factor variable for the treatments
+
+panel_data$ftreatment <- panel_data$treatment
+
+# Fix treatment names
+panel_data$ftreatment[panel_data$ftreatment == "i"] <- "Insects"
+panel_data$ftreatment[panel_data$ftreatment == "p"] <- "Birds, bats and ants"
+panel_data$ftreatment[panel_data$ftreatment == "f"] <- "Fungi"
+panel_data$ftreatment[panel_data$ftreatment == "h"] <- "Increased herbivory"
+panel_data$ftreatment <- factor(panel_data$ftreatment,
+                           levels = c("Fungi",
+                                      "Insects",
+                                      "Birds, bats and ants",
+                                      "Increased herbivory"
+                                      ))
+
+# Set the transparency value
+aph <- 0.5
+
+# Create a factor variable for sites
+panel_data$fsite <- factor(panel_data$site, labels = c("Low",
+                                      "Mid",
+                                      "High"))
+
+# Display panel plot
+ggplot(panel_data, aes(x = fsite, y = lratio))+
+  geom_jitter(width = 0.1, alpha = 0.1, size = 2)+
+
+  # Group means
+  stat_summary(fun = mean, na.rm = T,
+               geom = "point", size = 2,
+               alpha = aph)+
+
+  stat_summary(fun.data = "mean_cl_boot",
+               alpha = aph) +
+
+  # Annotations
+  geom_text(data = panel_data %>%
+              group_by(ftreatment, fsite) %>%
+              summarise(mlr = mean(lratio)),
+            aes(x = fsite, y = mlr + 1.5, group = ftreatment,
+                label = lrr_panel_labs),
+            size = 3)+
+
+  theme_bw()+
+  theme(strip.text = element_text(size =5.5),
+        axis.text.x = element_text(angle = 0,
+                                   hjust = 0.5))+
+  geom_hline(yintercept = 0, lty = 2)+
+  ylab("Log response ratio")+xlab("")+
+  facet_grid(descriptor ~ ftreatment)
+
+```
+
+If we look at the biomass plots and a the effect plots we might be struck by a close relationship between some of these patterns.
+For example, the effect of fungi on biomass is lacking and seems exactly as the biomass patterns at each elevation.
+Effects of fungi, insects and predators on diversity measures seem to be inversely proportional to the patterns in diversity and density for three elevations.
+
+This may suggest a strong bottom-up control of the effect strength.
+The effect are present, but the local community characteristics may influence their magnitude.
+It seems that at the highly diverse mid elevation, the effects of biotic factor is limited.
+Only one outstanding effect is of insects on biomass.
+Highest elevation has average species richness.
+But this elevation seems also to have the strongest and most frequently significant effects of biotic factors.
+This suggests that the strength of biotic factors may be related non linearly with plant species richness.
+
+Surprising is also similarity in the effects of insects and their predators.
+If insects are being controlled by their top predators we should see a negative effects of insects and positive effects of predators.
+This results says that in the plots where the insecticide was sprayed was the same as in the enclosure, where top predators were not present.
+When we were constantly re-setting insects community we disturbed natural processes of interaction forming between predators and insects.
+There also may be an additional.
+
+# The effects of site richness and productivity of the LRR of biomass and diversity.
+
+```{r, results='asis'}
+tab_model(final.biolrr.nl.mod, final.divlrr.nl.mod)
 ```
@@
 ```{r pressure, echo=FALSE, fig.width=10, fig.height=10}
-ggpubr::ggarrange(p1, p1.1,
-                  nrow = 2)
+ggpubr::ggarrange(BioPredPlot, DivPredPlot, nrow = 2,
+                  labels = c("A", "B"))
 ```
+
+A substantial amount of variation in biomass responses (LRRs) can be explained by baseline community characteristics. However in cases of

### 2023-07-21 code/elevation_descriptors.R

@@ -1,0 +1,57 @@
+# source('code/prepare_data.R')
+
+# RICHNESS # both sp_no and diversity
+# BIOMASS  # Biomass
+# DENSITY
+
+library(ggplot2)
+library(dplyr)
+
+BIOMASS <- BIOMASS %>%
+  mutate(elevation = ifelse(site == 'wanang', 200, ifelse(site == "numba", 750, 1900)))
+
+RICHNESS <- RICHNESS %>%
+  mutate(elevation = ifelse(site == 'wanang', 200, ifelse(site == "numba", 750, 1900)))
+
+DENSITY <- DENSITY %>%
+  mutate(elevation = ifelse(site == 'wanang', 200, ifelse(site == "numba", 750, 1900)))
+
+bio.plot <- ggplot(BIOMASS %>%
+                     filter(treatment == "c"), aes(x = elevation, y = tot_bio)) + 
+  stat_smooth(lwd = 2) + 
+  stat_summary(fun = 'mean',
+               fun.min = function(x) mean(x) - sd(x), 
+               fun.max = function(x) mean(x) + sd(x),
+               geom = 'pointrange', lwd = 1.2, cex = 1)+
+  theme_bw() + 
+  xlab("Elevation") + ylab('Total aboveground biomass')
+
+rich.plot <- ggplot(RICHNESS %>%
+                      filter(treatment == "c"), aes(x = elevation, y = sp_no )) + 
+  stat_smooth(lwd = 2) + 
+  stat_summary(fun = 'mean',
+               fun.min = function(x) mean(x) - sd(x), 
+               fun.max = function(x) mean(x) + sd(x),
+               geom = 'pointrange', lwd = 1.2, cex = 1)+
+  theme_bw() + 
+  xlab("Elevation") + ylab('Total number of woody species')
+
+div.plot <- ggplot(RICHNESS %>%
+                     filter(treatment == "c"), aes(x = elevation, y = simpson )) + 
+  stat_smooth(lwd = 2) + 
+  stat_summary(fun = 'mean',
+               fun.min = function(x) mean(x) - sd(x), 
+               fun.max = function(x) mean(x) + sd(x),
+               geom = 'pointrange', lwd = 1.2, cex = 1)+
+  theme_bw() + 
+  xlab("Elevation") + ylab("Simpson's diersity")
+
+dens.plot <- ggplot(DENSITY %>%
[…]

### 2023-07-21 code/exp1_dummy_coded_treatments.R

@@ -1,0 +1,125 @@
+source('code/log_ratio_extraction.R')
+
+RichLRR = RICHNESS.LR$lratio
+ConRicVal = RICHNESS.LR$cval
+DivLRR = DIVERSITY.LR$lratio
+ConDivVal = DIVERSITY.LR$cval
+DenLRR = DENSITY.LR$lratio
+ConDenVal = DENSITY.LR$cval
+
+LRR_DEP_TEST_DATA <- BIOMASS.LR %>%
+  rename(Site = 'site',
+         Garden = 'garden',
+         Treatment = 'treatment',
+         TrtBioVal = 'value',
+         ConBioVal = 'cval',
+         BioLRR = 'lratio')
+
+# LRR_DEP_TEST_DATA %>%
+#   dplyr::mutate(RichLRR = RICHNESS.LR$lratio,
+#                 ConRicVal = RICHNESS.LR$cval,
+#                 DivLRR = DIVERSITY.LR$lratio,
+#                 ConDivVal = DIVERSITY.LR$cval,
+#                 DenLRR = DENSITY.LR$lratio,
+#                 ConDenVal = DENSITY.LR$cval)
+
+LRR_DEP_TEST_DATA$RichLRR = RichLRR
+LRR_DEP_TEST_DATA$ConRicVal = ConRicVal
+LRR_DEP_TEST_DATA$DivLRR = DivLRR
+LRR_DEP_TEST_DATA$ConDivVal = ConDivVal
+LRR_DEP_TEST_DATA$DenLRR = DenLRR
+LRR_DEP_TEST_DATA$ConDenVal = ConDenVal
+LRR_DEP_TEST_DATA = LRR_DEP_TEST_DATA %>%
+  mutate(SiteTreatment = paste(Site, Treatment, sep = " "))
+
+# Prepare model data - h exluded
+moddata <- LRR_DEP_TEST_DATA %>% 
+  filter(Treatment != "h") %>%
+  mutate(Insects = ifelse(Treatment == "i", 1, 0),
+         Fungi = ifelse(Treatment == "f", 1, 0),
+         Predators = ifelse(Treatment == "p", 1, 0),
+         LogBioC = log(ConBioVal),
+         LogRicC = log(ConRicVal))
+
+# Dummy code the treatments
+moded
+
+biolrr.nl.mod0 <- lm(BioLRR ~
+                       LogBioC*Insects*Site +
+                       LogBioC*Predators*Site + 
+                       I(LogBioC^2)*Insects*Site + 
[…]

### 2023-07-21 code/lrr_and_community_chars_correlations.R

@@ -1,9 +1,9 @@
 source('code/log_ratio_extraction.R')
 
-RICHNESS.LR
-BIOMASS.LR
-DIVERSITY.LR
-DENSITY.LR
+# RICHNESS.LR
+# BIOMASS.LR
+# DIVERSITY.LR
+# DENSITY.LR
 
 RichLRR = RICHNESS.LR$lratio
 ConRicVal = RICHNESS.LR$cval
@@ -332,7 +332,11 @@ summary(divlrr.ric.mod0)
 ##############################################
 # richness does not explain LRR in diversity
 
-# More complex models with two desscriptors
+
+#############################################################################
+###########   BIOMASS
+#############################################################################
+# More complex models with two descriptors ---
 biolrr.mod0 <- lm(BioLRR ~ lcbv * Treatment*Site + lcrv*Treatment*Site, 
                   data = LRR_DEP_TEST_DATA %>% 
                     filter(Treatment != "h"))
@@ -429,34 +433,300 @@ p2 <- plot_model(biolrr.mod0, type="pred", terms = c("lcbv", "Site"))
 p3 <- plot_model(biolrr.mod0, type="pred", terms = c("lcbv", "Treatment"))
 
 
+#############################################################################
+###########   Diversity  
+#############################################################################
+
+# More complicated models for diversity ----
+divlrr.mod0 <- lm(DivLRR ~ lcbv * Treatment*Site + lcrv*Treatment*Site, 
+                  data = LRR_DEP_TEST_DATA %>% 
+                    filter(Treatment != "h"))
+
+# Alternative models
+divlrr.mod1 <- lm(DivLRR ~ lcbv * Treatment*Site + lcrv*Treatment,
+                  data = LRR_DEP_TEST_DATA %>%
+                    filter(Treatment != "h"))
+
+divlrr.mod2 <- lm(DivLRR ~ lcbv * Treatment*Site + lcrv*Site,
+                  data = LRR_DEP_TEST_DATA %>%
+                    filter(Treatment != "h"))
+
+divlrr.mod3 <- lm(DivLRR ~ lcbv * Treatment*Site + lcrv,
+                  data = LRR_DEP_TEST_DATA %>%
+                    filter(Treatment != "h"))
+
+divlrr.mod4 <- lm(DivLRR ~ lcbv * Treatment*Site,
[…]

### 2023-07-21 code/predictAndVisualize_function.R

@@ -1,89 +1,4 @@
 # Predictions and visualizations!
-# predictAndVisualize <- function (model,
-#                                  deltaData,
-#                                  traitnm,
-#                                  n=100,
-#                                  titlestring = "",
-#                                  ylabstring = "",
-#                                  xlabstring = ""
-# ) {
-# 
-#   other_traits <- names(deltaData)[!(names(deltaData) %in% c(traitnm,
-#                                                              "fsite",
-#                                                              "site",
-#                                                              "garden",
-#                                                              "treatment",
-#                                                              "unified_names",
-#                                                              "pg.cat",
-#                                                              "comparison"))]
-# 
-#   # Create a value gradient for the focal variable.
-#   # focal.column = seq(from = min(deltaData[, traitnm], na.rm = T),
-#   #                    to = max(deltaData[, traitnm], na.rm = T),
-#   #                    length = n)
-# 
-#   focal.column.w = seq(from = min(deltaData[deltaData$site == "wanang",
-#                                             traitnm], na.rm = T),
-#                        to = max(deltaData[deltaData$site == "wanang",
-#                                           traitnm], na.rm = T),
-#                        length = n)
-#   focal.column.n = seq(from = min(deltaData[deltaData$site == "numba", traitnm], na.rm = T),
-#                        to = max(deltaData[deltaData$site == "numba",
-#                                           traitnm], na.rm = T),
-#                        length = n)
-#   focal.column.y = seq(from = min(deltaData[deltaData$site == "yawan", traitnm], na.rm = T),
-#                        to = max(deltaData[deltaData$site == "yawan",
-#                                           traitnm], na.rm = T),
-#                        length = n)
-# 
-#   # Assume average values for the rest.
-#   other.vals.w <- t(apply(deltaData[deltaData$site == "wanang",
-#                                     other_traits], 2, mean, na.rm = T))
-#   other.vals.n <- t(apply(deltaData[deltaData$site == "numba",
-#                                     other_traits], 2, mean, na.rm = T))
-#   other.vals.y <- t(apply(deltaData[deltaData$site == "yawan",
[…]

### 2023-08-03 ms1_supplement.Rmd
Moving the main analysis to another markdown.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results='hide')
+knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results='hide', fig.align = 'center')
 ```
@@
 source('code/elevation_descriptors.R')
+source('code/log_ratio_extraction.R')
 ```
 
-# Introduction
+## Introduction
 
@@
 
-```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 1.
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 2.
 Magnitude and direction of fungi, insects and their predators on biomass, richness, diversity and density of woody plants along elevation gradient: Wanang (200 m a.s.l.); Numba (750 m a.s.l.); Yawan (1900 m a.s.l.).
@@
 For example, the effect of fungi on biomass is lacking and seems exactly as the biomass patterns at each elevation.
-Effects of fungi, insects and predators on diversity measures seem to be inversely proportional to the patterns in diversity and density for three elevations.
+Effects of fungi, insects and predators on diversity measures seem to be inversely proportional to the elevational patterns in diversity and density.
 
@@
 But this elevation seems also to have the strongest and most frequently significant effects of biotic factors.
-This suggests that the strength of biotic factors may be related non linearly with plant species richness.
+This suggests that the strength of biotic factors may be related non-linearly with plant species richness.
 
@@
 If insects are being controlled by their top predators we should see a negative effects of insects and positive effects of predators.
-This results says that in the plots where the insecticide was sprayed was the same as in the enclosure, where top predators were not present.
+This pattern means that in the plots where the insecticide was sprayed the outcome for plant community was the same as in the enclosure, where top predators were not present.
 When we were constantly re-setting insects community we disturbed natural processes of interaction forming between predators and insects.
-There also may be an additional.
 
-# The effects of site richness and productivity of the LRR of biomass and diversity.
+## The effects of site richness and productivity of the LRR of biomass and diversity.
+
+I investigated how well the actual observed effects of biotic factors can be explained by species richness and biomass of the woody plant community at the site level.
+Within each garden i used richness and biomass of the control plot as a predictor of the effect.
+These relationships can differ for treatments and sites, and can also be non-linear, all of which was allowed while defining models.
+Full models were then subjected to step-wise backward selection based on the AIC criterion to produce final models (Tab 1.).
+
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+    Table 1.
+Models of richness and biomass and their relationships with LRR of studied treatments at different sites.
+Quadratic terms were allowed and slope and intercepts were allowed to differ between the sites and treatments and also between treatments within each site.
+The model sare results of backward selection process based on tha AIC criterion.
+</p>
 
 ```{r, results='asis'}
-tab_model(final.biolrr.nl.mod, final.divlrr.nl.mod)
+tab_model(final.biolrr.nl.mod, final.riclrr.nl.mod,final.divlrr.nl.mod, final.denlrr.nl.mod)
 ```
 
-```{r pressure, echo=FALSE, fig.width=10, fig.height=10}
-ggpubr::ggarrange(BioPredPlot, DivPredPlot, nrow = 2,
-                  labels = c("A", "B"))
+Based on the results we can explain much of the variance of the biomass effects of the treatments on biomass change (adjusted $R^2 = 0.615$).
+The least amount of variance can be explained for the diversity change (adjusted $R^2 = 0.296$).
+This agrees with the rather weak effects of biotic factors on biomass but more diverse and clear effects on diversity, richness and density.
+It is possible that the biomass is determined by the local conditions, but the assembly process is mostly affected by biotic factors.
+
+```{r pressure, echo=FALSE, fig.width=10, fig.height=15}
+ggpubr::ggarrange(BioPredPlot, RicPredPlot, DivPredPlot, DenPredPlot, 
+                  BioPredPlotR, RicPredPlotR, DivPredPlotR, DenPredPlotR,
+                  nrow = 4, ncol = 2,
+                  labels = c("A", "B", "C", "D",
+                             "E", "F", "G", "H"),
+                   common.legend = TRUE, legend="bottom")
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+    Figure 1. Predictions of the models for LRR form the Table 1.
+</p>
+
+Let us look at similar models, but for each individual site, to compare the $R^2$ values.
+We are interested in the adjusted $R2$ because we want to evaluate the predictive ability of the richness and biomass.
+Therefore, I excluded the treatment, used linear and quadratic terms and did not perform model selection.
+Whenever we see a large positive $R^2$ we can suspect strong bottom-up control.
+Low to negative adjusted $R^2$ values indicate top-down action
+
+```{r}
+
+# ggplot(r2data, aes(x = Site, y = AdjustedR2, fill = Site)) +
+#   geom_bar(stat = 'identity') + 
+#   coord_flip() +
+#   facet_grid(~Descriptor) + 
+#   theme_bw()
+
+ggplot(r2data, aes(x = Site, y = AdjR2, fill = Treatement)) +
+  geom_bar(stat = 'identity') + 
+  # coord_flip() + 
+  facet_grid(vars(Descriptors), vars(Treatement)) + theme_bw() + 
+  geom_hline(yintercept = c(-0.75, -0.5, 0, 0.5, 0.75), 
+             lty=2)
+
 ```
 
-A substantial amount of variation in biomass responses (LRRs) can be explained by baseline community characteristics. However in cases of
+The adjusted $R^2$ varies a lot for different treatments, sites and descriptors.
+The value of the adjusted $R^2$ could be interpreted as a measure of the level of bottom-up and top-down control: negative values indicate that adding diversity and richness only makes the model worse - this is a strong indicator of top-down effects.
+In contrast, high positive values are indicative to bottom-up control: we can predict well changes in descriptors depending on the baseline community characteristics.
+
+Fungi effects on diversity shifts from bottom-up at low to top-down at high elevation.
+Effect of fungi on biomass gets stronger with elevation.
+At the mid elevation insects have weak effect on richness, but stronger on diversity.
+
+Some of the strongest top-down effects of insects on richness and diversity were found in Wanang.
+Effect of insect on diversity decreases with elevation, but changes non-monotonously for richness density and biomass, showing peaks in bottom-up control at mid-elevation.
+
+For diversity there is a clear compensation of herbivores for fungi at the low elevations, that shifts at the high elevation.
+
+Predators have stronger effects at the higher elevations except for plant biomass, which is low at mid and high elevations.
+
+All of the community characteristics are interrelated.
+For example biomass is related to richness and also to the identity of the species involved (ref), and increases with the complexity of a plant community (Cardinale et al 2007).
+We see some strong indications that insects determine richness and diversity at low elevations, density and biomass at high elevation, and at the mid elevation they only affect diversity in a significant way.
+Changes in richness and diversity can happen in two ways: biotic factors can either be filters for species, or they can change competitive outcomes at between species.
+Unfortunately the end results may be the same for these two processes.
+
+What we can investigate though is what actually was more important in responding to a given treatment.
+Was it species identity (which basically means that we need to look at each species individually - bad) or was it some trait values.
+Basically what we want to do here is to see whether response to a treatment is species specific or trait specific.
+We could only use species that are present in at least 3 gardens at all three elevations, or within each elevation do the analysis, which is more plausible.
+We will look whether certain species are being excluded, or whether species composition response can be better explained by changes in plant trait composition.
+We will perform RDA or CCA on species id's and on species traits instead.
+
+These results suggest that:
+* biomass is largely bottom up controlled, but the effects of biotic factors: insects, fungi and predators are increasingly more important at the higher elevations.
+* community assembly processes seem to be more strongly affected by biotic factors.
+* some of the strongest top-down effects on diversity and richness were found at the lowest elevation for insects, and highest elevation for fungi.
+* there is a general weak bottom-up control at the highest elevation.
+
+Lets us investigate the effect of individual treatments on the community assembly process.
+
+## Effect of biotic factors on the community assembly.
+
+First we want to know whether some treatments resulted in a statistically different communities.
+We build RDA models for each elevation.
+We evaluated which processes species competition or the species turnover were mostly affected, and we used Raup-Crick to evaluate which biotic factors can improve predictability of the successional process.
+
+### RDA
+
+```{r, fig.height=10, fig.width=10}
+source('code/cx_rda_analysis.R')
+## 3.2 Plot the ordination ----
+ggpubr::ggarrange(plot.list[["w"]]$sites.ord,
+                  plot.list[["w"]]$species.ord,
+                  plot.list[["n"]]$sites.ord,
+                  plot.list[["n"]]$species.ord,
+                  plot.list[["y"]]$sites.ord,
+                  plot.list[["y"]]$species.ord, ncol = 2, nrow = 3)
+```
+
+Plant communities treated with insecticide are different from the control plots at all elevations. Fungicide treated communities are differenet from the control only at the high elevation.
+
+```{r, results='asis'}
+options(knitr.kable.NA = '')
+knitr::kable(anova(ord.list[["w"]], by="terms"))
+knitr::kable(anova(ord.list[["n"]], by="terms"))
+knitr::kable(anova(ord.list[["y"]], by="terms"))
+```
+
+### B-C
+
+### Raup-Crick
+
+## Are traits better predictors of the species success?

### 2023-08-03 code/cx_rda_analysis.R

@@ -1,0 +1,129 @@
+# RDA analysis
+source('code/log_ratio_extraction.R')
+
+library(vegan)
+library(cowplot)
+library(ggplot2)
+
+# 1. Load the data
+COMMDATA <- t(COMMDATA)
+COMMVARS
+
+# Select only i,c,p,f
+COMM.SEL <- COMMDATA[grepl("i|c$|p|f|_h", rownames(COMMDATA)), ]
+VARS.SEL <- COMMVARS[grepl("i|c$|p|f|_h", COMMVARS$plotid), ]
+
+treats.to.consider <- c("i","c","h","f","p")
+
+# Dummy-coding of the treatments
+COMMVARS <- COMMVARS %>%
+  mutate(f = ifelse(treatment == "f", 1, 0),
+         h = ifelse(treatment == "h", 1, 0),
+         p = ifelse(treatment == "p", 1, 0),
+         c = ifelse(treatment == "c", 1, 0),
+         i = ifelse(treatment == "i", 1, 0))
+
+# RDA at each elevation
+
+## 3.1 Analyses for each elevation
+ord.list <- list()
+plot.list <- list()
+
+for(site.id in substr(unique(COMMVARS$site),1,1)){
+  
+  # print(site.id)
+  
+  SUB.VARS <- COMMVARS %>%
+    filter(grepl(paste('^', site.id, sep = ""), site) & treatment %in% treats.to.consider)
+  
+  SUB.DATA <- COMMDATA[SUB.VARS$plotid, ]
+  SUB.DATA <- decostand(SUB.DATA, method = 'hellinger')
+  # Perform RDA and ANOVA and print the results
+  rdamod <- rda(as.data.frame(SUB.DATA)~f+h+p+i+Condition(garden), 
+                data = SUB.VARS)
+  
+  ord.list[[site.id]] <- rdamod
+  
+  h <- how(blocks = SUB.VARS$garden)
+  
+  # print(anova(rdamod, permutations = h, by = "terms"))
+  
+  # Goodness
+  gdnsof <- goodness(rdamod, model = "CCA", proportional = T)
+  gof2ax <- gdnsof[,1] + gdnsof[,2]
+  cgof <- gof2ax[gof2ax > 0.2]
+  cgof <- sort(cgof, decreasing = T)
+  sign.names <- names(cgof[!is.na(cgof) & !is.infinite(cgof)])
+  
+  colors <- as.numeric(as.factor(SUB.VARS$treatment))
+  
[…]

### 2023-08-03 code/lrr_and_community_chars_correlations.R

@@ -156,11 +156,11 @@ summary(lrrm1)
 #   geom_point()
 
 # Correlation between LRRs ---- 
-library(corrplot)
-cor.dat <- as.matrix(LRR_DEP_TEST_DATA[, c("BioLRR","RichLRR","DivLRR","DenLRR")])
-M <- cor(cor.dat)
-M.test <- cor.mtest(cor.dat)
-corrplot::corrplot(M, p.mat = M.test$p)
+# library(corrplot)
+# cor.dat <- as.matrix(LRR_DEP_TEST_DATA[, c("BioLRR","RichLRR","DivLRR","DenLRR")])
+# M <- cor(cor.dat)
+# M.test <- cor.mtest(cor.dat)
+# corrplot::corrplot(M, p.mat = M.test$p)
 
 ## Biomass is correlated to density. Three groups of models: ----
 # 1. BIOLRR ~ biomass_C + richenss_C
@@ -225,12 +225,12 @@ divLRR.ric <- ggplot(LRR_DEP_TEST_DATA %>%
   facet_wrap(~ Treatment) + 
   theme_bw()
 
-ggpubr::ggarrange(bioLRR.bio,
-                  bioLRR.ric,
-                  ricLRR.bio,
-                  ricLRR.ric,
-                  divLRR.bio,
-                  divLRR.ric)
+# ggpubr::ggarrange(bioLRR.bio,
+#                   bioLRR.ric,
+#                   ricLRR.bio,
+#                   ricLRR.ric,
+#                   divLRR.bio,
+#                   divLRR.ric)
 
 # Make some transformed columns
 LRR_DEP_TEST_DATA$lcbv <- log(LRR_DEP_TEST_DATA$ConBioVal)
@@ -337,189 +337,189 @@ summary(divlrr.ric.mod0)
 ###########   BIOMASS
 #############################################################################
 # More complex models with two descriptors ---
-biolrr.mod0 <- lm(BioLRR ~ lcbv * Treatment*Site + lcrv*Treatment*Site, 
-                  data = LRR_DEP_TEST_DATA %>% 
-                    filter(Treatment != "h"))
-
-# Alternative models
-biolrr.mod1 <- lm(BioLRR ~ lcbv * Treatment*Site + lcrv*Treatment,
-                  data = LRR_DEP_TEST_DATA %>%
-                    filter(Treatment != "h"))
-
-biolrr.mod2 <- lm(BioLRR ~ lcbv * Treatment*Site + lcrv*Site,
[…]

### 2023-08-08 ms1_supplement.Rmd
Raup-Crick done, beta prat and multi site variance comparison

@@
 
-ggplot(r2data, aes(x = Site, y = AdjR2, fill = Treatement)) +
+# ggplot(r2data,
+#        aes(x = Site, y = AdjR2, fill = Treatement)) +
+#   geom_bar(stat = 'identity') +
+#   # coord_flip() +
+#   facet_grid(vars(Descriptors), vars(Treatement)) + theme_bw() +
+#   geom_hline(yintercept =  0,
+#              lty=2)
+
+ggplot(r2data %>%
+         mutate(fDescriptors = factor(Descriptors,
+                                       labels = c("Biomass LRR",
+                                                  "Diversity LRR",
+                                                  "Richness LRR",
+                                                  "Density LRR"),
+                                       levels = c("BioLRR",
+                                                  "DivLRR",
+                                                  "RichLRR",
+                                                  "DenLRR"))), 
+       aes(x = Site, y = AdjR2, fill= Treatement)) +
   geom_bar(stat = 'identity') + 
   # coord_flip() + 
-  facet_grid(vars(Descriptors), vars(Treatement)) + theme_bw() + 
-  geom_hline(yintercept = c(-0.75, -0.5, 0, 0.5, 0.75), 
+  facet_grid(vars(fDescriptors), vars(Treatement)) + theme_bw() + 
+  geom_hline(yintercept =  0,
              lty=2)
@@
 
+It is obvious from the previous analyses that studied biotic factors have limited effect on the plant biomass, but have various effects on species richness and diversity.
 First we want to know whether some treatments resulted in a statistically different communities.
@@
                   plot.list[["y"]]$sites.ord,
-                  plot.list[["y"]]$species.ord, ncol = 2, nrow = 3)
+                  plot.list[["y"]]$species.ord, ncol = 2, nrow = 3,
+                  labels = c("Wanang","", "Numba","","Yawan",""))
 ```
 
-Plant communities treated with insecticide are different from the control plots at all elevations. Fungicide treated communities are differenet from the control only at the high elevation.
+Plant communities treated with insecticide are different from the control plots at all elevations.
+Fungicide treated communities are different from the control only at the high elevation.
+There are a few species that we can see from the diagrams that respond to treatements: *Pipturus argenteus* increase in abundance at the insecticide treated plots in Wanang and Yawan, *Trema orientale* in Numba, and *Piper recessum* and *Macaranga strigosa* to fungicide treatment in Yawan.
 
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Table X. Results of the Monte Carlo permutation to evaluate significance of each treatment on the plant community composition. RDA anlayses are conditioned on the garden variable.
+</p>
 ```{r, results='asis'}
 options(knitr.kable.NA = '')
-knitr::kable(anova(ord.list[["w"]], by="terms"))
-knitr::kable(anova(ord.list[["n"]], by="terms"))
-knitr::kable(anova(ord.list[["y"]], by="terms"))
+knitr::kable(anova(ord.list[["w"]], by="terms"), caption = "Wanang")
+knitr::kable(anova(ord.list[["n"]], by="terms"), caption = "Numba")
+knitr::kable(anova(ord.list[["y"]], by="terms"), caption = "Yawan")
 ```
+
+RDA tries to find a general trends, which may be difficult to evaluate because of the high variation in species composition between replications (individual experimental gardens) within a site.
+
+```{r}
+source('code/multi_site_beta.R')
+
+ggplot(vardf_full, aes(y = JaccardVar, x = Treatment, fill = Treatment))+
+  geom_col()+facet_grid(~Site)+theme_bw()+
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Figure 6. Variance in species composition (Jaccard distance) for all pairwise comparisons between control sites at each elevation and treatment.
+</p>
+
+Variation between the sites reflects relative species richness at each elevation.
+Average dissimilarity is equally high for all sites and it is always more than 95%.
+We see that in Wanang and Yawan the variance war reduced by all treatments, but qualitatively different pattern is seen in Numba, where variance for the predator treatment was higher than in the control treatment.
 
@@
 
+In this section I will investigate whether some of studied treatments were able to reduce this dissimilarity.
+Further I brake it down into two components of beta diversity: gradient and balanced components (ref).
+Balanced component tells us how much of the dissimilarity is due to species turnover, whereas the gradient component is related to gradient changes in abundance - this would be related to changes in diversity (e.g.
+dominance structure changes).
+
+```{r}
+source('code/cx_beta_part.R')
+```
+
+```{r, fig.width=8, fig.height=6}
+beta.part.plot2
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Figure 7.
+All pairwise dissimilarities (Bray-Curtis) in species composition between given treatment plots at each elevation.
+Dissimilarities are broken down into two components: responsible for the species turnover (balanced component), and for changes in species dominance structures (gradient components).
+Starst indicate that a given average was significantly different from that at the control plot ($\alpha < 0.05$).
+To compare the averages the beta regression was used as the values for B-c are always between 0 and 1.
+</p>
+
+The figure shows different processes being affected by different treatments.
+In Wanang, high dissimilarity was reduced by insects, by shifts from species turnover that was reduced towards more significant changes in species composition.
+No visible changes in dissimilarity were found in Numba.
+However, there for additional herbivores, predators, and most strongly insects, species turnover increased significantly at the cost of changes in dominance structure.
+The same treatments in Yawan caused reduction in species turnover that were reflected in general decrease in species dissimilarity, because were not equally significantly balanced by increase in changes in dominance structure.
+
+The above analysis suggests that biotic factors may have different effects at different elevations.
+These effects can be related to the baseline species richness and variance in community composition.
+At the mid elevation, presence insects reduce species turnover and increase dominance shifts.
+In Yawan, where biomass accumulates more slowly, insect increase turnover that results in higher dissimilarity between plant communities.
+Reduction in species turnover could have two fold effects, If it is increased it may cause divergence in the compositions between sites, but the same may apply to reduced species turnover, i.e., smaller pool of species may be by chance different from site to site, thus propagating differences even further.
+To investigate this issue, we performed a simulation analysis using the Raup-Crick index, that measures the predictability of the assembly process.
+
 ### Raup-Crick
 
+```{r}
+all.pairs_max_accuracy <- read.csv("data/raup_crick/all.pairs_max_accuracy.csv")
+all.pairs_max_accuracy$fsite <- factor(all.pairs_max_accuracy$fsite, 
+                                       levels = c("w","n","y"), labels = c("Wanang",
+                                                                           "Numba",
+                                                                           "Yawan"))
+
+test.df <- data.frame()
+
+all.pairs_max_accuracy$fsite <- factor(all.pairs_max_accuracy$fsite, levels = c("Wanang",
+                                                  "Numba","Yawan"))
+
+for (st in unique(all.pairs_max_accuracy$fsite)){
+  for (cp in unique(all.pairs_max_accuracy$comp)){
+
+    print(paste(st, cp))
+
+    subdat <- all.pairs_max_accuracy %>%
+      filter(fsite == st & comp == cp)
+
+    mod <- nlme::lme(RC ~ treatment,
+                     random = ~1|comp.id,
+                     data = subdat)
+
+    # print(t.test(RC ~ treatment, data = subdat, paired = TRUE))
+
+    test.df <- rbind(test.df, data.frame(fsite = st,
+                                         comp = cp,
+                                         group1 = tolower(substr(cp,1,1)),
+                                         group2 = tolower(substr(cp,6,6)),
+                                         test.res = summary(mod)$tTable[2,5]))
+
+    # print(summary(mod))
+
+  }
+}
+
+test.df <- test.df %>%
+  mutate(label = ifelse(test.res < 0.05, "*"," "))
+test.df$fsite <- factor(test.df$fsite, levels = c("Wanang",
+                                                  "Numba",
+                                                  "Yawan"))
+
+ggplot(all.pairs_max_accuracy, aes(x = treatment, y = RC))+
+  # geom_boxplot()+
+  ylim(c(-1.25,1.5))+
+  stat_summary(fun.y = mean, na.rm = T,
+               geom = "point", size = 4)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  theme_bw()+
+  ggtitle(paste("Accuracy = 0.001"))+
+  geom_hline(yintercept = 0, lty = 2)+
+  geom_hline(yintercept = 1, lty = 3)+
+  geom_hline(yintercept = -1, lty = 3)+
+  facet_grid(vars(fsite), vars(comp), scales = "free")+
+  ggpubr::stat_pvalue_manual(data = test.df[test.df$label == "*",], label = "label",
+                     y.position = 1.25, 
+                     label.size = 6)
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Figure X. Differences in Raup-Crick values between the control plot and treatment plots at each elevation.
+</p>
+
+In Wanang and more strongly in Yawan we see a deterministic divergence in community composition at the control plots.
+In Wanang only additional herbivores and predator treatments were able to shift this process towards more random.
+In Yawan all treatments were able to do this.
+IN Numba, however, baseline community was rather random, only additional herbivores were able to push it towards deterministic divergence, possibly through reducing richness of the plant community.
+
+Ok. So we know that the biotic factors have significant effect on the community composition, some of which can be better explained by the baseline characteristics of local plant community.
+
 ## Are traits better predictors of the species success?
+
+## Conclusions
+
+The surprising result that all treatments have similar effects, and only Numba shows some qualitatively different patterns in these effects.
+In Wanang, where species richness in the lowest, insects increased richness and fungi increased richness and density of pioneer woody plants.
+Nearly identical situation.
+
+Some of the effects are independent of the baseline community, but some of them are.
+
+We did not achieve an expected separation of effects in case of predator and insecticide treatment.
+This may be related to some additional mechanisms within the exclosures.
+All treatments, herbivore addition, insecticide treatmetn and predator exclosure had similar effects on plant community composition.

### 2023-08-08 code/cx_beta_part.R

@@ -1,0 +1,203 @@
+library(betapart)
+library(lsmeans)
+library(multcomp)
+
+beta.df <- data.frame()
+
+# siteid <- "^w"
+# treat <- "_c\\b"
+
+COMMDATA <- t(COMMDATA)
+
+for(siteid in c("^w","^y","^n")){
+  for(treat in c("_c\\b","_f","_h","_i","_p")){
+    
+    # Subset the data
+    nms <- rownames(COMMDATA)
+    sel.nms <- nms[grepl(siteid, nms) & grepl(treat, nms)]
+    site.treat.comm <- COMMDATA[sel.nms,]
+    
+    # Calculate beta diversity
+    subbeta <- beta.pair.abund(site.treat.comm)
+    
+    # Add to a data frame
+    balanced.df <- data.frame(site = siteid,
+                              treat = treat,
+                              vals = as.numeric(subbeta$beta.bray.bal),
+                              component = "balanced")
+    
+    gradient.df <- data.frame(site = siteid,
+                              treat = treat,
+                              vals = as.numeric(subbeta$beta.bray.gra),
+                              component = "gradient")
+    
+    bray.df <- data.frame(site = siteid,
+                          treat = treat,
+                          vals = as.numeric(subbeta$beta.bray),
+                          component = "bray")
+    
+    sub.df <- rbind(balanced.df, gradient.df, bray.df)
+    
+    # Merge the data-set
+    beta.df <- rbind(beta.df,sub.df)
+    
+  }
+}
+
+# beta.df
+
+# Fix the site order
+beta.df$fsite <- factor(beta.df$site, levels = c("^w","^n","^y"),
+                        labels = c("w","n","y"))
+
+# Transform 0s to 0.001 and 1s to 0.999
+beta.df$trans.vals = ifelse(beta.df$vals == 1, 0.999999,
+                            ifelse(beta.df$vals == 0, 0.000001, beta.df$vals))
+
+# Tests for individual facets
+cld.dat = data.frame()
+prws.comp <- list()
+
+# CLD data for comparisons between elevations
+for (comp in unique(beta.df$component)){
+  for (trt in unique(beta.df$treat)){
+    
+    test.df <- beta.df %>%
+      filter(treat == trt,
+             component == comp)
+    
+    mod1 <- betareg::betareg(trans.vals ~ fsite, data = test.df)
+    
+    lsmod1 <- lsmeans(mod1,
+                      pairwise ~ fsite,
+                      adjust="tukey")
+    
[…]

### 2023-08-08 code/multi_site_beta.R

@@ -1,0 +1,92 @@
+# Multi-site beta diversity
+
+# Upload community composition for each garden within each site.
+source('code/prepare_data.R')
+
+# Controls
+ycdf <- COMMDATA[,grepl("yawan.*_c", colnames(COMMDATA))]
+ycdfnz <- ycdf[rowSums(ycdf) != 0,]
+
+ncdf <- COMMDATA[,grepl("numba.*_c", colnames(COMMDATA))]
+ncdfnz <- ncdf[rowSums(ncdf) != 0,]
+
+wcdf <- COMMDATA[,grepl("wanang.*_c", colnames(COMMDATA))]
+wcdfnz <- wcdf[rowSums(wcdf) != 0,]
+
+# Insects
+yidf <- COMMDATA[,grepl("yawan.*_i", colnames(COMMDATA))]
+yidfnz <- yidf[rowSums(yidf) != 0,]
+
+nidf <- COMMDATA[,grepl("numba.*_i", colnames(COMMDATA))]
+nidfnz <- nidf[rowSums(nidf) != 0,]
+
+widf <- COMMDATA[,grepl("wanang.*_i", colnames(COMMDATA))]
+widfnz <- widf[rowSums(widf) != 0,]
+
+
+# Predators
+ypdf <- COMMDATA[,grepl("yawan.*_p", colnames(COMMDATA))]
+ypdfnz <- ypdf[rowSums(ypdf) != 0,]
+
+npdf <- COMMDATA[,grepl("numba.*_p", colnames(COMMDATA))]
+npdfnz <- npdf[rowSums(npdf) != 0,]
+
+wpdf <- COMMDATA[,grepl("wanang.*_p", colnames(COMMDATA))]
+wpdfnz <- wpdf[rowSums(wpdf) != 0,]
+
+
+# Fungi
+yfdf <- COMMDATA[,grepl("yawan.*_f", colnames(COMMDATA))]
+yfdfnz <- yfdf[rowSums(yfdf) != 0,]
+
+nfdf <- COMMDATA[,grepl("numba.*_f", colnames(COMMDATA))]
+nfdfnz <- nfdf[rowSums(nfdf) != 0,]
+
+wfdf <- COMMDATA[,grepl("wanang.*_f", colnames(COMMDATA))]
+wfdfnz <- wfdf[rowSums(wfdf) != 0,]
+
+
+
+
+
+
+library(BAT)
[…]

### 2023-08-29 ms1_supplement.Rmd
Breaking code into components and integrating in the markdown file.

--- 2023-07-21 ms1_supplement.Rmd
@@
 ```{r, echo = FALSE, warning=FALSE, message=FALSE, results='hide'}
+source('code/log_ratio_extraction.R')
 source('code/lrr_and_community_chars_correlations.R')
 source('code/elevation_descriptors.R')
+source("code/pgls_data_prep.R")
 ```
 
-# Introduction
+## Elevational patterns
 
-Aim of our study was to evaluate to what extent pathogenic fungi, insect herbivores, and their predators shape community composition of pioneer stages of tropical forest regeneration.
-Further we were interested in the effect of natural elevation gradient on the direction and strength of these effects.
+The aim of our study was to evaluate the extent to which pathogenic fungi, insect herbivores, and their predators shape the biomass, richness, diversity, and community composition of pioneer stages in tropical forest regeneration along a natural elevation gradient.
 
-Looking at biomass, diversity (number of species and Simpson's index values), and density of woody plant species.
-Biomass did not change much with elevation, but diversity and density peaked at the mid elevation.
+The biomass of the control plots did not change significantly with elevation.
+However, richness, diversity, and density peaked at the mid-elevation.
+Surprisingly, the low elevation had the lowest number of pioneer woody plant species.
 
-```{r, fig.cap="Figure 1. Characteristics of three studied sites: biomass, richness, diversity (Simpson's index) and density of woody plants at the control sites."}
+```{r}
 
@@
 ```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+"Figure 1. Characteristics of three studied sites: biomass, richness, diversity (Simpson's index) and density of woody plants at the control sites."
+</p>
 
-Now lets look at the effects that the biotic factors had on the vegetation characteristics.
+Now, let's examine the effects that biotic factors had on the characteristics of vegetation.
 
-```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 1.
+```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.width= 8, fig.height=8}
+
+source('code/cx_lrr_panel.R')
+lrr_panel
+
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+    Figure 2.
 Magnitude and direction of fungi, insects and their predators on biomass, richness, diversity and density of woody plants along elevation gradient: Wanang (200 m a.s.l.); Numba (750 m a.s.l.); Yawan (1900 m a.s.l.).
@@
 Grey points represent empirical values.
-Letters indicate statistically significant ($\\alpha$ = 0.05) differences for pairwise comparisons between elevations within a treatment-descriptor combination with Tukey correction for multiplicity.
+Letters indicate statistically significant ($\alpha$ = 0.05) differences for pairwise comparisons between elevations within a treatment-descriptor combination with Tukey correction for multiplicity.
 For better readability letters are given only in cases where any of the mean were either different form each other or different from zero.
 Zero effect are indicated with a dashed line.
-Stars in the upper index indicate significant ($\\alpha$ = 0.05) differences of the effect mean from zero (significance of the effect at a given site)", fig.width= 8, fig.height=8}
+Stars in the upper index indicate significant ($\alpha$ = 0.05) differences of the effect mean from zero (significance of the effect at a given site)
+</p>
 
-# bio, dens, div, rich
-panel_data <- data.frame()
+There are a few interesting patterns emerging from this comparison.
+First is the similarity in the impacts of insects and their predators on vegetation.
+If insects are controlled by their higher-level predators, we would expect to observe negative effects of insects and positive effects of predators on vegetation.
+However, empirical patterns indicate that the plant community's response to top predators exclusion resembled that of insecticide treatment.
+Because of the substantial influx of insects into the experimental plots, we were unable to completely eliminate insects from the area.
+Rather, we induced repeated insect population reductions.
+It's plausible that the recurrent eradication of the insect community may have disrupted the natural processes of interaction between predators and insects, which likely play a crucial role in shaping the ecosystem's reaction to the insecticide treatment.
 
-dats <- list(biomass = BIOMASS.LR, 
-             denssity = DENSITY.LR, 
-             diversity = DIVERSITY.LR, 
-             richness = RICHNESS.LR)
+Secondly, patterns in richness, diversity, and density appear to be linked to patterns in the effects of treatments.
+For instance, the impact of fungi on richness and density of woody plants appear to be inversely related to the elevation-based patterns in diversity.
+We specifically observe the following relationship: high richness is associated with a weak effect, while low richness is linked to a strong effect.
 
-# Stack the data and rename descriptor names
-for (file in names(dats)){
-  # print(file)
-  dat <- dats[[file]]
-  dat$descriptor <- ifelse(grepl("bio", file), "Biomass",
-                           ifelse(grepl("dens", file), "Woody plant density",
-                                  ifelse(grepl("div", file),"Diversity", "Richness")))
-  dat <- dat[, c("site", "garden", "treatment", "lratio","descriptor")]
-  panel_data <- rbind(dat, panel_data)
-}
+This observation could imply a pronounced bottom-up control on the strength of these effects.
+While the effects are indeed present, their magnitude appears to be influenced by the specific characteristics of the local community.
+Notably, at the mid-elevation with high plant diversity, the impact of biotic factors are weak, except for the significant effect of insects on biomass, not observed at low and high elevations.
+The highest elevation, with average species richness, also exhibits the most potent and consistently significant effects of biotic factors.
+This suggests a possible non-linear relationship between plant species richness and the strength of biotic factors.
 
-# Make a renamed factor variable for the treatments
+## The effects of site richness and productivity of the LRR of biomass and diversity.
 
-panel_data$ftreatment <- panel_data$treatment
+To explore the potential bottom-up forcing of effect sizes, I constructed nonlinear models.
+These models aimed to explain the log response ratios for biomass, richness, diversity, and density solely based on the potential productivity (biomass) and species influx (richness) of the garden.
+Those basal characteristics can be sourced from either the control plot or all treatment plots within a garden.
+In the former case, we derive a 'natural' species richness and productivity, which results from the interaction of all biotic factors in a naturally restricted area equivalent to the plot size.
+Alternatively, considering the entire garden provides a larger sample of potential species richness and potential productivity.
+In the following analysis I took the second approach.
+I calculated models individually for each site and compared them using the adjusted $R^2$ value to assess the predictive capability of total richness and average biomass at the garden level.
+I incorporated linear and quadratic terms without performing any model selection.
+The adjusted $R^2$ value can be interpreted as a measure of the extent of bottom-up and top-down control: negative values suggest that enhancing diversity and richness deteriorates the model - a strong indicator of top-down effects.
+Conversely, significantly positive values indicate bottom-up control: effectively predicting changes in descriptors based on the underlying community characteristics.
 
-# Fix treatment names
-panel_data$ftreatment[panel_data$ftreatment == "i"] <- "Insects"
-panel_data$ftreatment[panel_data$ftreatment == "p"] <- "Birds, bats and ants"
-panel_data$ftreatment[panel_data$ftreatment == "f"] <- "Fungi"
-panel_data$ftreatment[panel_data$ftreatment == "h"] <- "Increased herbivory"
-panel_data$ftreatment <- factor(panel_data$ftreatment,
-                           levels = c("Fungi",
-                                      "Insects",
-                                      "Birds, bats and ants",
-                                      "Increased herbivory"
-                                      ))
+```{r}
+#Treatment panels
+ggplot(r2data %>%
+         mutate(fDescriptors = factor(Descriptors,
+                                       labels = c("Biomass LRR",
+                                                  "Diversity LRR",
+                                                  "Richness LRR",
+                                                  "Density LRR"),
+                                       levels = c("BioLRR",
+                                                  "DivLRR",
+                                                  "RichLRR",
+                                                  "DenLRR"))), 
+       aes(x = Site, y = AdjR2, fill= Treatement)) +
+  geom_bar(stat = 'identity') + 
+  # coord_flip() + 
+  facet_grid(vars(fDescriptors), vars(Treatement)) + theme_bw() + 
+  geom_hline(yintercept =  0,
+             lty=2)
 
-# Set the transparency value
-aph <- 0.5
+# Site panels
+# ggplot(r2data %>%
+#          mutate(fDescriptors = factor(Descriptors,
+#                                        labels = c("Biomass LRR",
+#                                                   "Diversity LRR",
+#                                                   "Richness LRR",
+#                                                   "Density LRR"),
+#                                        levels = c("BioLRR",
+#                                                   "DivLRR",
+#                                                   "RichLRR",
+#                                                   "DenLRR"))), 
+#        aes(x = Treatement, y = AdjR2, fill= Site)) +
+#   geom_bar(stat = 'identity') + 
+#   # coord_flip() + 
+#   facet_grid(vars(fDescriptors), vars(Site)) + theme_bw() + 
+#   geom_hline(yintercept =  0,
+#              lty=2)
 
-# Create a factor variable for sites
-panel_data$fsite <- factor(panel_data$site, labels = c("Low",
-                                      "Mid",
-                                      "High"))
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+    Figure 3.
+Adjusted $R^2$ values of quadratic model of log response ratios (LRR) of pioneer plant biomass, diversity, richness and density to selected biotic factors: f- fungi; i - insects; p - predators.
+Natural logarithm of biomass and richness were used as only predictors of the LRR's.
+Quadratic terms were also introduced to allow for nonlinear responses.
+</p>
 
-# Display panel plot
-ggplot(panel_data, aes(x = fsite, y = lratio))+
-  geom_jitter(width = 0.1, alpha = 0.1, size = 2)+
+The adjusted $R^2$ demonstrates significant variation across different treatments, sites, and descriptors.
+In terms of diversity, the impact of fungi shifts from top-down influence at low elevations to bottom-up effects at higher elevations.
+Notably, the effect of fungi on biomass is most pronounced at higher elevations.
+The influence on richness exhibits strong indications of top-down control at both low and high elevations.
+Conversely, at mid elevations, fungi display a weak effect on richness.
 
-  # Group means
-  stat_summary(fun = mean, na.rm = T,
-               geom = "point", size = 2,
-               alpha = aph)+
+Interestingly, we discover limited top-down effects of insects, except for density at mid elevations and richness at high elevations.
+The influence of insects on richness exhibits an elevation-dependent increase.
+Notably, the adjusted $R^2$ values for the effect on biomass, richness, and density in Wanang are surprisingly high (bottom-up effect).
+However, it's crucial to exercise caution when interpreting parameter estimates due to the small sample size of 6 gardens in Wanang and only 5 parameters for estimation.
+Nevertheless, it's remarkable that a quadratic plane fits exceptionally well to these six data points.
+Concerning diversity, a noticeable compensation effect between herbivores and fungi is evident at low elevations, which undergoes a shift at higher elevations.
+Predators demonstrate more pronounced effects at greater elevations, except for plant biomass, which remains low at mid and high elevations.
 
-  stat_summary(fun.data = "mean_cl_boot",
-               alpha = aph) +
+To explain the Log Response Ratio (LRR) of biomass, diversity, richness, and density in relation to potential richness and productivity, I developed an extensive model that incorporated all relevant interactions.
+This model incorporated both linear and quadratic terms for both richness and productivity and interaction of these terms with site and treatment.
+Within each garden, I utilized the total richness and average biomass values at the garden level as predictors of the effect.
+Following this, the comprehensive models underwent a step-wise backward selection procedure, guided by the AIC criterion, to derive the final models (as presented in Table 1).
 
-  # Annotations
-  geom_text(data = panel_data %>%
-              group_by(ftreatment, fsite) %>%
-              summarise(mlr = mean(lratio)),
-            aes(x = fsite, y = mlr + 1.5, group = ftreatment,
-                label = lrr_panel_labs),
-            size = 3)+
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Table 1: Models depicting the relationships between richness, biomass, and their interactions with Log Response Ratio (LRR) for the investigated treatments across distinct sites.
+The inclusion of quadratic terms was permitted, with variations in slopes and intercepts accounted for across sites and treatments, as well as between treatments within each site.
+The presented models are the outcomes of a backward selection procedure driven by the AIC criterion.
+</p>
+```{r, results='asis'}
+tab_model(final.biolrr.nl.mod, final.riclrr.nl.mod,final.divlrr.nl.mod, final.denlrr.nl.mod)
+```
 
-  theme_bw()+
-  theme(strip.text = element_text(size =5.5),
-        axis.text.x = element_text(angle = 0,
-                                   hjust = 0.5))+
-  geom_hline(yintercept = 0, lty = 2)+
-  ylab("Log response ratio")+xlab("")+
-  facet_grid(descriptor ~ ftreatment)
+The outcomes allow us to account for a substantial portion of the variance in treatment effects on biomass change (adjusted $R^2 = 0.367$).
+Conversely, the lowest proportion of variance can be elucidated for diversity change (adjusted $R^2 = 0.296$).
+This discrepancy suggests considerable impacts of biotic factors on richness, whereas effects on biomass, density, and diversity are more confined.
+It appears that local conditions play a determining role in biomass, while biotic factors predominantly influence the assembly process.
+
+```{r pressure, echo=FALSE, fig.width=10, fig.height=15}
+ggpubr::ggarrange(BioPredPlot, RicPredPlot, DivPredPlot, DenPredPlot,
+                  BioPredPlotR, RicPredPlotR, DivPredPlotR, DenPredPlotR,
+                  nrow = 4, ncol = 2,
+                  labels = c("A", "B", "C", "D",
+                             "E", "F", "G", "H"),
+                   common.legend = TRUE, legend="bottom")
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+    Figure 4. Predictions of the models for LRR form the Table 1.
+</p>
+
+All plant community characteristics are interconnected.
+For instance, biomass is associated not only with richness but also with species composition.
+It tends to increase with the complexity of a plant community (Cardinale et al.
+2007).
+
+There are two non-mutually exclusive ways in which biotic factors affect plant communities: they either serve as filters for species or alter competitive outcomes among species.
+However, in our experiment, distinguishing between these two processes' final outcomes might be challenging without tracking species influx and mortality with a reasonable temporal resolution.
+
+Just to stay cnsostent I compared the most complex models with the simplest models to have an idea how good of predictors site and treatments are.
+
+```{r, results='asis'}
+knitr::kable(biolrr.comp, caption =  "Response: Biomass effects")
+knitr::kable(riclrr.comp, caption =  "Response: Richness effects")
+knitr::kable(divlrr.comp, caption =  "Response: Diversity effects")
+knitr::kable(denlrr.comp, caption =  "Response: Density effect")
 
@@
 
-If we look at the biomass plots and a the effect plots we might be struck by a close relationship between some of these patterns.
-For example, the effect of fungi on biomass is lacking and seems exactly as the biomass patterns at each elevation.
-Effects of fungi, insects and predators on diversity measures seem to be inversely proportional to the patterns in diversity and density for three elevations.
+# Treatment responses on individual species
 
-This may suggest a strong bottom-up control of the effect strength.
-The effect are present, but the local community characteristics may influence their magnitude.
-It seems that at the highly diverse mid elevation, the effects of biotic factor is limited.
-Only one outstanding effect is of insects on biomass.
-Highest elevation has average species richness.
-But this elevation seems also to have the strongest and most frequently significant effects of biotic factors.
-This suggests that the strength of biotic factors may be related non linearly with plant species richness.
+The previous analyses clearly indicate that the studied biotic factors exert a limited influence on plant biomass at the community level.
+However, they manifest diverse impacts on species richness and diversity.
+Our primary objective is to identify potential differences between treatment plots and controls concerning community composition.
 
-Surprising is also similarity in the effects of insects and their predators.
-If insects are being controlled by their top predators we should see a negative effects of insects and positive effects of predators.
-This results says that in the plots where the insecticide was sprayed was the same as in the enclosure, where top predators were not present.
-When we were constantly re-setting insects community we disturbed natural processes of interaction forming between predators and insects.
-There also may be an additional.
+We collected traits linked to growth rate and resource acquisition, encompassing specific leaf area (SLA), water content, biomass, and herbivore damage.
+In this section, our focus is on plant species identity and traits, assessing their utility in predicting individual species performance within specific plots.
+I explore the primary predictor of treatment response: whether it is species identity, necessitating individual species-level analysis (which might be impractical), or species traits, which allows for elevation-based comparisons.
+This inquiry is pursued through a range of methodologies, including:
 
-# The effects of site richness and productivity of the LRR of biomass and diversity.
+Redundancy Analysis (RDA) to discern species exhibiting directional responses to treatments.
+
+Correlation of RDA axes with plant species traits.
+Rather than considering plant IDs, we examine their traits to identify whether certain traits are correlated with the treatment effect vectors—vectors of change identified in the RDA.
+
+Employing Random Forest to establish the significance of traits in predicting categorical plant responses to treatments.
+
+Through this, I aim to predict whether a species, based on its traits, will be lost, gained, or retained across various treatment plots.
+To achieve this, we utilize the Random Forest algorithm to explore the importance of traits in predicting species outcomes.
+We assess prediction accuracy using the "classifier_RF" approach.
+This analysis provides insights into the effectiveness of the trait set (including the most influential traits) in predicting outcomes for individual species.
+
+Analyzing Community Weighted Means (CWMs) for trait responses across the entire community.
+
+Applying Phylogenetic Generalized Least Squares (PGLS) models to evaluate changes in plant species biomass based on their trait values in response to treatments.
+
+### 1. RDA
+
+```{r, fig.height=10, fig.width=10}
+source('code/cx_rda_analysis.R')
+## 3.2 Plot the ordination ----
+ggpubr::ggarrange(plot.list[["w"]]$sites.ord,
+                  plot.list[["w"]]$species.ord,
+                  plot.list[["n"]]$sites.ord,
+                  plot.list[["n"]]$species.ord,
+                  plot.list[["y"]]$sites.ord,
+                  plot.list[["y"]]$species.ord, ncol = 2, nrow = 3,
+                  labels = c("Wanang","", "Numba","","Yawan",""))
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+    Figure 5. RDA plots with red arrows indicating statistical significance of treatments.
+</p>
+
+Plant communities subjected to insecticide treatments differ from the control plots across all elevations.
+On the other hand, communities treated with fungicide exhibit variations from the control plots solely at higher elevation.
+Several species show discernible responses to these treatments as observed in the diagrams:
+* *Pipturus argenteus* experiences an increase in abundance within insecticide-treated plots in Wanang and Yawan.
+* *Trema orientale* exhibits this trend in Numba.
+* *Piper recessum* and *Macaranga strigosa* demonstrate a response to fungicide treatment specifically in Yawan.
+
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Table 2. Results of the Monte Carlo permutation to evaluate significance of each treatment on the plant community composition. RDA analyses are conditioned on the garden variable.
+</p>
+```{r, results='asis'}
+options(knitr.kable.NA = '')
+knitr::kable(anova(ord.list[["w"]], by="terms"), caption = "Wanang")
+knitr::kable(anova(ord.list[["n"]], by="terms"), caption = "Numba")
+knitr::kable(anova(ord.list[["y"]], by="terms"), caption = "Yawan")
+```
+
+As we tried to show in the previous sections t's possible that certain alterations in plant community composition could be attributed to initial variations in richness rather than the influence of a specific factor.
+Biotic factors could affect richness that has effect on plant species dynamics and result in changes in species composition.
+
+When we introduce the conditional impact of richness and biomass instead of garden identity, we observe the loss of significance in treatment effects.
+This observation could imply that insects, predators, and fungi exert a more pronounced influence on the assembly process compared to inter-specific plant interactions.
+
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Table 3.
+Results of the Monte Carlo permutation to evaluate significance of each treatment on the plant community composition.
+RDA analyses are conditioned on the garden species richness and average plot biomass.
+</p>
+```{r, results='asis'}
+options(knitr.kable.NA = '')
+knitr::kable(anova(result.list[["w"]], by="terms"), caption = "Wanang")
+knitr::kable(anova(result.list[["n"]], by="terms"), caption = "Numba")
+knitr::kable(anova(result.list[["y"]], by="terms"), caption = "Yawan")
+```
+
+### 2. Can traits predict change in composition?
+
+For each treatment I correlated the position of each species along an 'effect gradient' - a vector that point to same direction as a treatment, and correlated it with individual species tratits.
+SLA doesn’t predict RDA position along C->I, C->H, C->P, C -> F axes, at any elevation.
+Community composition change under a treatment may be independent of SLA - the key indicator of functionally important aspects of photosynthetic capacity, and is positively correlated with ecosystem productivity (White et al., 2000, Madani et al., 2017, Anderson et al., 2020).
+SLA reflects the abilities of light capture and adaption to environments of plants (Gao, Wang, and Zhang 2022)].
+The regression analyses are corrected for phylogeny using the phylogenetic covariance matrix.
 
 ```{r, results='asis'}
-tab_model(final.biolrr.nl.mod, final.divlrr.nl.mod)
+
+source('code/cx_trait_ord.R')
+
+m1 <- traitPredsCommComp(treatment = "f",
+                             ordination = result.list,
+                             st = "wanang")
+m2 <- traitPredsCommComp(treatment = "f",
+                             ordination = result.list,
+                             st = "numba")
+m3 <- traitPredsCommComp(treatment = "f",
+                             ordination = result.list,
+                             st = "yawan")
+m4 <- traitPredsCommComp(treatment = "i",
+                             ordination = result.list,
+                             st = "wanang")
+m5 <- traitPredsCommComp(treatment = "i",
+                             ordination = result.list,
+                             st = "numba")
+m6 <- traitPredsCommComp(treatment = "i",
+                             ordination = result.list,
+                             st = "yawan")
+
+tab_model(m1$model,
+          m2$model,
+          m3$model,
+          dv.labels = c(m1$title,
+                        m2$title,
+                        m3$title))
+
+tab_model(m4$model,
+          m5$model,
+          m6$model,
+          dv.labels = c(m4$title,
+                        m5$title,
+                        m6$title))
+
 ```
 
-```{r pressure, echo=FALSE, fig.width=10, fig.height=10}
-ggpubr::ggarrange(BioPredPlot, DivPredPlot, nrow = 2,
-                  labels = c("A", "B"))
+In Numba, both Specific Leaf Area (SLA) and water content are significant predictors of a species' position along the vector of treatment effects.
+However, in Yawan, only water content holds significance in this regard.
+Otherwise, traits demonstrate limited predictive power for the ordination process.
+RDA tries to find a general trends, which may be difficult to evaluate because of the high variation in species composition between replications (individual experimental gardens) within a site.
+
+```{r}
+source('code/multi_site_beta.R')
+
+ggplot(vardf_full, aes(y = JaccardVar, x = Treatment, fill = Treatment))+
+  geom_col()+facet_grid(~Site)+theme_bw()+
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Figure 6. Variance in species composition (Jaccard distance) for all pairwise comparisons between control sites at each elevation and treatment.
+</p>
+
+The variation observed among the different sites is influenced, to some extent, by the relative species richness present at each elevation.
+The average Bray-Curtis within treatment, within site dissimilarity holds a consistently high value of around 95%.
+Notably, in Wanang and Yawan, all treatments led to a reduction in variance.
+However, a distinct pattern emerges in Numba, where the variance under the predator treatment surpassed that of the control treatment.
+
+Given the potential influence of insects on the assembly process and subsequent alterations in community composition, we will later delve into the components of B-C dissimilarity and explore the assembly process itself using null models.
+
+### 3. Random forest
+
+First we look whether some of the traits can help us understand wether some plants will be:
+* missing at the treatment plot but present at the control plot (lost);
+* present at the treatment plot but missing from the control plot (gained)
+* present in both treatment and control plot (stayed).
+
+```{r}
+source('code/cx_random_forest.R')
 ```
 
-A substantial amount of variation in biomass responses (LRRs) can be explained by baseline community characteristics. However in cases of
+```{r, figures-side, fig.show="hold", out.width="50%"}
+vipP <- varImpPlot(prfres$classifier_for_varImpPlot,
+           main = 'Variable Importance for the predator exclusion')
+# barplot(height = c(vipP), names.arg = rownames(vipP))
+varImpPlot(irfres$classifier_for_varImpPlot,
+           main = 'Variable Importance for the insect exclusion')
+varImpPlot(frfres$classifier_for_varImpPlot,
+           main = 'Variable Importance for the fungi exclusion')
+varImpPlot(hrfres$classifier_for_varImpPlot,
+           main = 'Variable Importance for the herbivore addition')
+
+# MDSplot(prfres$classifier_for_varImpPlot, prfres$train_data$pg.cat, 
+#         main = "Predator")
+# MDSplot(irfres$classifier_for_varImpPlot, irfres$train_data$pg.cat,
+#         main = "Insecticide")
+# MDSplot(frfres$classifier_for_varImpPlot, frfres$train_data$pg.cat,
+#         main = "Fungicide")
+# MDSplot(hrfres$classifier_for_varImpPlot, hrfres$train_data$pg.cat,
+#         main = "Herbivory")
+```
+
+```{r}
+
+pcm <- prfres$rf_confusion_matrix
+icm <- irfres$rf_confusion_matrix
+fcm <- frfres$rf_confusion_matrix
+hcm <- hrfres$rf_confusion_matrix
+
+pacc <- sum(diag(pcm))/sum(pcm)
+iacc <- sum(diag(icm))/sum(icm)
+facc <- sum(diag(fcm))/sum(fcm)
+hacc <- sum(diag(hcm))/sum(hcm)
+
+```
+
+With the variables we have collected, accurately predicting the fate of individual species remains challenging.
+The prediction accuracies are as follows: `r iacc` for the insecticide treatment; `r facc` for the fungicide treatment; `r pacc` for the predator treatment; and `r hacc` for the herbivore addition treatment.
+Notably, for all predictions, total biomass emerges as the most significant predictor of a species' fate.
+Conversely, SLA and the number of stems are the least influential predictors.
+To illustrate this division, we employ a single regression tree.
+
+### 4. Whole community trait response (CWMs)
+
+The preceding analysis focused on the individual level, aiming to anticipate whether we can predict a species' fate based on its traits.
+This had a moderately satisfactory level of predictive ability, with minimal discrepancies in prediction accuracy across the various treatments.
+
+Alternatively, I explore whether any of the treatments impacted the community-weighted mean for traits.
+This approach delves into understanding how traits at the community level were affected by the treatments.
+
+```{r, fig.cap='Community weighted mean vs number of species at a site for the studied treatments at three elevations. Size of a point indicates total plant biomass of that plot.'}
+
+source('code/cx_cwm.R')
+ggpubr::ggarrange(p1,p1.wc)
+
+```
+
+The Community Weighted Mean of Specific Leaf Area (CWM SLA) exhibited higher values in plots with lower diversity.
+This implies that species with relatively higher SLA were the dominant contributors to these communities.
+This pattern is also evident for water content and remains consistent regardless of the treatment or site.
+This trend appears to be quite widespread, suggesting that biotic factors might exert a limited impact on the trait composition itself.
+Instead, their influence seems to primarily affect species richness, which subsequently affects other aspects of the plant community's characteristics.
+
+```{r, results='asis'}
+# Test for the relationship and the differences
+require(sjPlot)
+cwm.lm <- lm(log(cwm.sla)~log(sp.rich)*treatment + site, data = siteSLAdf)
+cwm.lm.nosite <- lm(log(cwm.sla)~log(sp.rich)*treatment, data = siteSLAdf)
+cwm.lm.nosite.noint <- lm(log(cwm.sla)~log(sp.rich)+treatment, data = siteSLAdf)
+cwm.lm.onlysp <- lm(log(cwm.sla)~log(sp.rich), data = siteSLAdf)
+
+# cwm.glm <- glm(cwm.sla~sp.rich*treatment + site, data = siteSLAdf,
+#              family = gaussian(link = 'log'))
+# cwm.glm.nosite <- glm(cwm.sla~sp.rich*treatment, data = siteSLAdf,
+#                     family = gaussian(link = 'log'))
+# cwm.glm.nosite.noint <- glm(cwm.sla~sp.rich+treatment, data = siteSLAdf, 
+#                           family = gaussian(link = 'log'))
+# cwm.glm.onlysp <- glm(cwm.sla~sp.rich, data = siteSLAdf, family = gaussian(link = 'log'))
+
+# tab_model(cwm.lm.nosite.noint)
+
+# anova(cwm.lm.nosite, cwm.lm.onlysp)
+knitr::kable(AIC(cwm.lm, cwm.lm.nosite, cwm.lm.nosite.noint, cwm.lm.onlysp))
+# knitr::kable(AIC(cwm.glm, cwm.glm.nosite, cwm.glm.nosite.noint, cwm.glm.onlysp,
+                 # cwm.lm, cwm.lm.nosite, cwm.lm.nosite.noint, cwm.lm.onlysp))
+
+# tab_model(cwm.glm)
+tab_model(cwm.lm.onlysp)
+
+# plot(cwm.glm)
+# plot(cwm.lm)
+
+# there are no differences
+
+```
+
+### 5. Individual species performance baed on taits
+
+In control plots with increasing elevation, we hypothesized that Specific Leaf Area (SLA), serving as a proxy for growth rate, will progressively become a stronger predictor of various aspects of species' success.
+This includes factors such as fitness, dominance, and changes in biomass in response to treatments.
+More specifically, we expect that plants with higher SLA will demonstrate an enhanced capability to dominate the community, as reflected by higher total biomass in scenarios where natural enemies have been removed.
+In these situations, we assume that interspecific plant competition will be primarily influenced by differences in growth rates.
+
+For a more comprehensive measurement of a species' response to a treatment, change in biomass (delta biomass) might be a superior metric.
+This measurement excludes species that were either lost or gained in response to treatments.
+However, the prior analysis has indicated that these plants tend to be relatively small, often in poor condition or newly established in a plot.
+Several challenges complicate the interpretation of this relationship.
+To mitigate these issues, incorporating covariates into the model of biomass change can offer some improvement.
+
+1.
+**Richness**: The Specific Leaf Area (SLA) of a species might be influenced by the characteristics of the community it inhabits, as indicated by both Community Weighted Mean (CWM) traits and their dependency on species richness.
+However, at the individual level, we do not observe any alteration in the SLA of individual species in response to the richness of the plot.
+This implies that the increased CWM SLA in low richness plots is driven by the dominance of high SLA plants rather than changes in the SLA of the dominants.
+This aligns with the absence of discernible differences in community composition, that would be captured if changes in species composition were more based on changes in dominance structure.
+
+It's worth noting that the lack of significant treatment effects in the RDA analysis could potentially be attributed to substantial variation between gardens, especially evident in the control sites with high variability.
+To address this, incorporating the identity of an experimental plot as a co-variate in the model could be beneficial.
+
+```{r}
+
+# SLA of selected species at different treatments.
+
+# Sla of selected species vs species richness
+
+```
+
+2.
+**Change in SLA**: From our data, it is apparent that certain treatments can lead to variations in the average Specific Leaf Area (SLA) of the same species.
+This change in SLA has the potential to impact the alteration in biomass, or conversely, changes in biomass might influence SLA.
+Importantly, this dynamic relationship extends to the two other traits as well.
+Therefore, it's crucial to consider all relevant traits when assessing these interactions.
+
+```{r}
+
+# Change in SLA vs change in Biomass
+
+```
+
+3.
+**Abundance**: Biomass/$\Delta$Biomass ratios might be reduced when there is a higher density of individuals from the same species in the vicinity.
+Consequently, I believe it's important to incorporate the influence of co-specifics’ abundance in our analysis.
+
+4.
+**Total biomass**: Furthermore, the productivity of a site (total biomass) could also impact the change in biomass; for instance, the change in biomass might be less pronounced for smaller plants and more significant for larger plants.
+Regarding the interpretation of these measures.
+
+Delta BIO quantifies how dominance within the community has changed due to the treatment.
+If there was a plant that was dominant in the control condition and its biomass experienced a substantial reduction in response to the treatment, this would signify that the treatment had a notable effect on diminishing the dominance of that specific plant in the community.
+In essence, it indicates a shift in the plant's importance or influence within the community due to the treatment's impact on its biomass.
+
+Taking the aforementioned considerations into consideration, we have formulated the final statistical model.
+In this model, the response variable is $\Delta$Biomass.
+We assess its dependence on SLA while incorporating the corresponding site richness and abundance of co-specifics as covariates.
+To examine the influence of treatments on the relationship between $\Delta$Biomass and SLA, we introduce the interaction between SLA and the treatment variable.
+This model is replicated for each site, and we evaluate its performance to compare the predictive power of SLA for biomass changes across elevations.
+In order to maintain phylogenetic correlations, we include a quadratic term.
+
+Furthermore, I explore nonlinear relationships between biomass log-ratios and species traits.
+The inclusion of these nonlinear relationships allows for the consideration of optimal trait values.
+All these insights lead us to define the comprehensive structure of the final model.
+
+$\Delta$Biomass ~ SLA $\times$ site + SLA^2 $\times$ site + 
+water.content $\times$ site + water.content^2 $\times$ site + 
+perc.herb $\times$ site  + perc.herb^2 $\times$ site + 
+richness + abundance + $\Delta$SLA + 
+(Phylogenetic structure).
+
+This model needs to be fit independently for each treatment to calculate the respective changes in SLA and biomass.
+
+There is some evidence pointing towards changes in total biomass being influenced by SLA.
+Both water content and log-response ratios of SLA significantly impact changes in total biomass.
+However, it's important to note that these changes could potentially be driven by a strong correlation in a single site, Wanang.
+Specifically, higher water content leads to positive changes in species biomass, while a substantial increase in SLA of a species correlates with a reduction in that species' biomass.
+Notably, none of the interactions with site were found to be significant.
+
+In terms of random effects, they might not be necessary given the paired comparisons of control vs.
+treatment within the same garden.
+Another option would be to incorporate the site as a random factor, which could lead to the elimination of all interaction terms.
+
+It is important to acknowledge that if we analyze the data separately for each elevation, the results might differ, especially if we solely rely on the variable lsla.d.
+More comprehensive models, such as those encompassing the effects of species richness and other traits, as discussed earlier, could potentially alter this relationship.
+
+```{r, results='asis'}
+
+source('code/cx_sla_pgls.R')
+
+# Plot
+ggpubr::ggarrange(pslap +
+                    scale_y_continuous(limits = c(-5, 10)),
+                  fwaterp + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  hwaterp + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1, legend = F)
+ggpubr::ggarrange(psitep + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  fsitep + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  hsitep + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1,legend = F)
+ggpubr::ggarrange(prichp + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  frichp + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1, legend = F)
+ggpubr::ggarrange(plabup + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  ilabup + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  hlabup + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1, legend = F)
+ggpubr::ggarrange(plrrslap +
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  hlrrslap + 
+                    scale_y_continuous(limits = c(-5, 10)), 
+                  nrow = 1, common.legend = T)
+
+```
+
+```{r, results = 'asis'}
+
+tab_model(p.mod, i.mod, h.mod, f.mod,
+          pred.labels = c("Intercept", 
+                          "Log[SLA]-linear", 
+                          "Numba", 
+                          "Yawan",
+                          "Log[SLA]-quadratic", 
+                          "Water content - linear", 
+                          "Water content - quadratic",
+                          "Herbivory [%] - linear",
+                          "Herbivory [%] - quadratic",
+                          "Species richness",
+                          "Abundance",
+                          "LRR SLA",
+                          "Log[SLA]-linear : Numba",
+                          "Log[SLA]-linear : Yawan",
+                          "Log[SLA]-quadratic : Numba",
+                          "Log[SLA]-quadratic : Yawan",
+                          "Water content-linear : Numba",
+                          "Water content-linear : Yawan",
+                          "Water content-quadratic : Numba",
+                          "Water content-quadratic : Yawan",
+                          "Herbivory-linear : Numba",
+                          "Herbivory-linear : Yawan",
+                          "Herbivory-quadratic : Numba",
+                          "Herbivory-quadratic : Yawan"),
+          dv.labels = c("Predator", "Insecticide", "Extra herbivory", "Fungicide"),
+          string.pred = "Coeffcient",
+          string.ci = "Conf. Int (95%)",
+          string.p = "P-Value")
+
+```
+
+## Effect of biotic factors on the community assembly.
+
+We employed beta-diversity partitioning techniques to assess the predominant influences on processes, namely species competition and species turnover.
+Subsequently, we delve into the enhancement of predictability for the successional process by evaluating the impact of biotic factors using the Raup-Crick index.
+
+### 1. B-C
+
+In this section, I will explore whether any of the studied treatments have the capacity to mitigate this dissimilarity.
+Moreover, I will dissect this dissimilarity into its two constituent components of beta diversity, known as the gradient and balanced components (as referenced).
+The balanced component provides insight into the extent of dissimilarity attributed to species turnover, while the gradient component pertains to shifts in abundance gradients.
+This latter aspect is connected to alterations in diversity, such as changes in the dominance structure.
+
+```{r}
+source('code/cx_beta_part.R')
+```
+
+```{r, fig.width=8, fig.height=6}
+beta.part.plot2
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Figure 7.
+All pairwise dissimilarities (Bray-Curtis) in species composition between given treatment plots at each elevation.
+Dissimilarities are broken down into two components: responsible for the species turnover (balanced component), and for changes in species dominance structures (gradient components).
+Starst indicate that a given average was significantly different from that at the control plot ($\alpha < 0.05$).
+To compare the averages the beta regression was used as the values for B-c are always between 0 and 1.
+</p>
+
+The figure illustrates distinct impacts on various processes attributed to different treatments.
+In Wanang, high dissimilarity is mitigated by insects, leading to a transition from significant species turnover towards more substantial shifts in species composition.
+Conversely, in Numba, no apparent alterations in dissimilarity were observed.
+However, the introduction of additional herbivores, predators, and particularly insects led to a noteworthy increase in species turnover while simultaneously causing changes in dominance structure.
+These same treatments in Yawan led to a decrease in species turnover, contributing to a general reduction in species dissimilarity.
+However, these reductions were not equally balanced by significant changes in dominance structure.
+
+The comprehensive analysis implies that biotic factors can yield varying effects at different elevations.
+These effects may be correlated with the initial species richness and the community composition's variance.
+At mid elevations, the presence of insects reduces species turnover and promotes shifts in dominance.
+In Yawan, where biomass accumulation is slower, insects increase turnover, subsequently resulting in heightened dissimilarity between plant communities.
+Alterations in species turnover can have dual consequences: if it increases, it could drive divergence in compositions between sites; conversely, reduced species turnover might also magnify differences by chance, due to the inherently smaller pool of species being distinct from site to site, thus amplifying dissimilarities.
+
+To further investigate this matter, we conducted a simulation analysis employing the Raup-Crick index, which gauges the predictability of the assembly process.
+
+### 2. Raup-Crick
+
+```{r}
+source('code/cx_rcplot.R')
+
+rcplot
+
+```
+<p style="font-family: times, serif; font-size:11pt; font-style:italic">
+Figure X. Differences in Raup-Crick values between the control plot and treatment plots at each elevation.
+</p>
+
+In Wanang and even more notably in Yawan, we observe a deterministic divergence in community composition within the control plots.
+In Wanang, only the additional herbivores and predator treatments managed to shift this process towards a more random pattern.
+In Yawan, all treatments achieved this outcome.
+In Numba, on the other hand, the baseline community displayed a relatively random pattern, and only the addition of herbivores managed to nudge it towards deterministic divergence, potentially by reducing the richness of the plant community.
+
+With this understanding, we now recognize that biotic factors exert a significant influence on community composition.
+These alterations are contingent upon the baseline community richness and the productivity of the site.
+Biotic factors impact the assembly process by modifying the richness of pioneer communities at both the highest and lowest elevations, albeit to a lesser extent in the latter.
+At the mid elevation, the considerable variation in community composition appears to attenuate the effects of biotic factors.
+
+Our findings highlight that biotic factors indeed can significantly affect richness, and these effects subsequently trigger shifts in community composition.
+We are now interested in understanding whether specific traits undergo filtration or perform better under these conditions.
+Furthermore, we aim to investigate which traits emerge after filtration and how they are subsequently suppressed.
+
+## Discussion
+
+The unexpected outcome reveals that all treatments yield comparable effects, with Numba displaying some distinct patterns in these effects.
+In Wanang, where species richness is the lowest, insects led to an increase in richness, while fungi prompted enhancements in both richness and the density of pioneer woody plants.
+These situations are nearly identical in nature.
+
+Surprisingly, the anticipated differentiation in effects between predator and insecticide treatments was not achieved.
+This could potentially be linked to additional mechanisms operating within the exclosure setups.
+Strikingly, all treatments—herbivore addition, insecticide treatment, and predator exclosure—exerted analogous effects on plant community composition.

### 2023-08-29 code/cx_cwm.R

@@ -1,0 +1,105 @@
+# SLA values of individual plant species for each treatment.
+
+# I am using the data.woody.brut data-set from the pgls_data_prep script. It has
+# filtered woody palnt names, cleaned sla.d values and limited sla range.
+
+# Define CWM function
+cwmFun <- function(sla.d.m2kg1, tot_bio){
+  # summarize function: CWM for sites
+  return(mean(sla.d.m2kg1*(tot_bio/sum(tot_bio, na.rm = T)), na.rm = T))
+}
+
+# siteSLA <- data.woody.brut %>%
+#   filter(unified_names %in% trait.dat[trait.dat$Life.form == "woody", ]$dsName) %>%
+#   mutate(tot_bio = ifelse(is.infinite(tot_bio), NA, tot_bio),
+#          sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1)) %>%
+#   group_by(site, garden, treatment) %>%
+#   summarise(cwm.sla = cwmFun(sla.d.m2kg1, n.tot_bio),
+#             sp.rich = n())
+
+siteSLA <- data.woody.brut %>%
+  filter(treatment %in% c('i','c','p','h', 'f')) %>%
+  mutate(tot_bio = ifelse(is.infinite(tot_bio), NA, tot_bio),
+         sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1),
+         water.cont = ifelse(is.infinite(water.cont), NA, water.cont)) %>%
+  group_by(site, garden, treatment) %>%
+  summarise(cwm.sla = cwmFun(sla.d.m2kg1, n.tot_bio),
+            cwm.wc = cwmFun(water.cont, n.tot_bio),
+            sp.rich = n(),
+            total.bio = sum(n.tot_bio))
+
+siteSLAdf <- as.data.frame(siteSLA)
+
+rownames(siteSLAdf) <- paste(substr(siteSLA$site,1,1), 
+                             siteSLA$garden, 
+                             "_", 
+                             siteSLA$treatment, sep = "")
+
+# Plot of CWM vs species richness
+require(scales)
+
+
+p1.wc <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.wc))+
+  geom_point(aes(size = total.bio, color = treatment)) + 
+  stat_smooth(method = "lm")+
+  scale_x_continuous(trans = 'log2',
+                     breaks = trans_breaks("log2", function(x) 2^x),
[…]

### 2023-08-29 code/cx_lrr_panel.R

@@ -1,0 +1,70 @@
+# bio, dens, div, rich
+panel_data <- data.frame()
+
+dats <- list(biomass = BIOMASS.LR, 
+             denssity = DENSITY.LR, 
+             diversity = DIVERSITY.LR, 
+             richness = RICHNESS.LR)
+
+# Stack the data and rename descriptor names
+for (file in names(dats)){
+  # print(file)
+  dat <- dats[[file]]
+  dat$descriptor <- ifelse(grepl("bio", file), "Biomass",
+                           ifelse(grepl("dens", file), "Woody plant density",
+                                  ifelse(grepl("div", file),"Diversity", "Richness")))
+  dat <- dat[, c("site", "garden", "treatment", "lratio","descriptor")]
+  panel_data <- rbind(dat, panel_data)
+}
+
+# Make a renamed factor variable for the treatments
+
+panel_data$ftreatment <- panel_data$treatment
+
+# Fix treatment names
+panel_data$ftreatment[panel_data$ftreatment == "i"] <- "Insects"
+panel_data$ftreatment[panel_data$ftreatment == "p"] <- "Birds, bats and ants"
+panel_data$ftreatment[panel_data$ftreatment == "f"] <- "Fungi"
+panel_data$ftreatment[panel_data$ftreatment == "h"] <- "Increased herbivory"
+panel_data$ftreatment <- factor(panel_data$ftreatment,
+                                levels = c("Fungi",
+                                           "Insects",
+                                           "Birds, bats and ants",
+                                           "Increased herbivory"
+                                ))
+
+# Set the transparency value
+aph <- 0.5
+
+# Create a factor variable for sites
+panel_data$fsite <- factor(panel_data$site, labels = c("Low",
+                                                       "Mid",
+                                                       "High"))
+
+# Display panel plot
+lrr_panel <- ggplot(panel_data, aes(x = fsite, y = lratio))+
+  geom_jitter(width = 0.1, alpha = 0.1, size = 2)+
+  
+  # Group means
+  stat_summary(fun = mean, na.rm = T,
+               geom = "point", size = 2,
+               alpha = aph)+
+  
+  stat_summary(fun.data = "mean_cl_boot",
+               alpha = aph) + 
+  
+  # Annotations
+  geom_text(data = panel_data %>%
+              group_by(ftreatment, fsite) %>%
+              summarise(mlr = mean(lratio)),
[…]

### 2023-08-29 code/cx_random_forest.R

@@ -1,0 +1,116 @@
+library(caTools)
+library(C50)
+library(randomForest)
+
+# Add richness
+richdat <- data %>%
+  group_by(site, garden, treatment) %>%
+  summarise(richness = n()) %>%
+  mutate(sgt = paste(site,garden,treatment, sep = "_"))
+
+data.woody.brut <- data.woody.brut %>%
+  mutate(sgt = paste(site,garden,treatment, sep = "_"))
+
+dwbRich <- merge(data.woody.brut, richdat, by = "sgt")
+
+# Filter and add estimated leaf area
+dwbRichF <- dwbRich %>%
+  mutate(est.leaf.area = num_leaf_weight * sla.w.m2kg1,
+         lsla.d = log(sla.d.m2kg1),
+         ltb = log(as.numeric(tot_bio)),
+         labu = log(as.numeric(no_stems+1))) %>%
+  dplyr::select(site.x, garden.x, treatment.x, 
+                unified_names, tot_bio, no_stems, 
+                perc.herb, water.cont, est.leaf.area, sla.d.m2kg1,
+                richness, ltb, lsla.d, labu) %>%
+  filter(water.cont > 0)
+
+names(dwbRichF) <- c("site", "garden", "treatment", "unified_names","tot_bio",
+                     "no_stems", "perc.herb", "water.cont", "est.leaf.area",
+                     "sla.d.m2kg1", "richness", "ltb","lsla.d","labu")
+
+
+# Random Forest
+# Loading package
+
+
+rfAnalysis <- function(lgsData) {
+  
+  # rfData <- lgsData %>%
+  #   mutate(pg.cat = as.factor(pg.cat),
+  #          lest.leaf.area = log(est.leaf.area),
+  #          lstem = log(no_stems)) %>%
+  #   dplyr::select(pg.cat, lstem, perc.herb, water.cont,
+  #                 lest.leaf.area, richness, ltb, lsla.d, treatment)
+  
+  rfData <- lgsData %>%
+    mutate(pg.cat = as.factor(pg.cat),
+           lest.leaf.area = log(est.leaf.area),
+           lstem = log(no_stems)) %>%
+    dplyr::select(pg.cat, site, lstem, perc.herb, water.cont,
+                  lest.leaf.area, richness, ltb, lsla.d, treatment)
+  
+  # Try standardized, non-correlated variables
+  # rfDataRF <- rfData %>%
[…]

### 2023-08-29 code/cx_rcplot.R

@@ -1,0 +1,57 @@
+all.pairs_max_accuracy <- read.csv("data/raup_crick/all.pairs_max_accuracy.csv")
+all.pairs_max_accuracy$fsite <- factor(all.pairs_max_accuracy$fsite, 
+                                       levels = c("w","n","y"), labels = c("Wanang",
+                                                                           "Numba",
+                                                                           "Yawan"))
+
+test.df <- data.frame()
+
+all.pairs_max_accuracy$fsite <- factor(all.pairs_max_accuracy$fsite, levels = c("Wanang",
+                                                                                "Numba","Yawan"))
+
+for (st in unique(all.pairs_max_accuracy$fsite)){
+  for (cp in unique(all.pairs_max_accuracy$comp)){
+    
+    print(paste(st, cp))
+    
+    subdat <- all.pairs_max_accuracy %>%
+      filter(fsite == st & comp == cp)
+    
+    mod <- nlme::lme(RC ~ treatment,
+                     random = ~1|comp.id,
+                     data = subdat)
+    
+    # print(t.test(RC ~ treatment, data = subdat, paired = TRUE))
+    
+    test.df <- rbind(test.df, data.frame(fsite = st,
+                                         comp = cp,
+                                         group1 = tolower(substr(cp,1,1)),
+                                         group2 = tolower(substr(cp,6,6)),
+                                         test.res = summary(mod)$tTable[2,5]))
+    
+    # print(summary(mod))
+    
+  }
+}
+
+test.df <- test.df %>%
+  mutate(label = ifelse(test.res < 0.05, "*"," "))
+test.df$fsite <- factor(test.df$fsite, levels = c("Wanang",
+                                                  "Numba",
+                                                  "Yawan"))
+
+rcplot <- ggplot(all.pairs_max_accuracy, aes(x = treatment, y = RC))+
+  # geom_boxplot()+
+  ylim(c(-1.25,1.5))+
+  stat_summary(fun.y = mean, na.rm = T,
+               geom = "point", size = 4)+
+  stat_summary(fun.data = "mean_cl_boot")+
+  theme_bw()+
+  ggtitle(paste("Accuracy = 0.001"))+
+  geom_hline(yintercept = 0, lty = 2)+
+  geom_hline(yintercept = 1, lty = 3)+
+  geom_hline(yintercept = -1, lty = 3)+
+  facet_grid(vars(fsite), vars(comp), scales = "free")+
[…]

### 2023-08-29 code/cx_rda_analysis.R

@@ -1,16 +1,24 @@
 # RDA analysis
-source('code/log_ratio_extraction.R')
+# source('code/log_ratio_extraction.R')
 
 library(vegan)
 library(cowplot)
 library(ggplot2)
 
 # 1. Load the data
-COMMDATA <- t(COMMDATA)
-COMMVARS
+COMMDATAt <- t(COMMDATA)
+# COMMVARS
 
+# Add variables for richness, and biomass 
+# (BUT DOES IT NEEDS TO BE DONE ONLY FOR CONTROL? Or FULL species richness and average biomass?)
+COMMVARS$richness <- rowSums(COMMDATAt>0)[COMMVARS$plotid]
+COMMVARS$biomass <- rowSums(COMMDATAt)[COMMVARS$plotid]
+
+COMMVARS$logR <- log(COMMVARS$richness)
+COMMVARS$logB <- log(COMMVARS$biomass)
+ 
 # Select only i,c,p,f
-COMM.SEL <- COMMDATA[grepl("i|c$|p|f|_h", rownames(COMMDATA)), ]
+COMM.SEL <- COMMDATAt[grepl("i|c$|p|f|_h", rownames(COMMDATAt)), ]
 VARS.SEL <- COMMVARS[grepl("i|c$|p|f|_h", COMMVARS$plotid), ]
 
 treats.to.consider <- c("i","c","h","f","p")
@@ -28,6 +36,20 @@ COMMVARS <- COMMVARS %>%
 ## 3.1 Analyses for each elevation
 ord.list <- list()
 plot.list <- list()
+result.list <- list()
+# How to approach it? Differnt model candidates
+# rdamod0 <- rda(as.data.frame(SUB.DATA)~f+h+p+i+Condition(garden + logR + logB),
+#               data = SUB.VARS)
+# rdamod1 <- rda(as.data.frame(SUB.DATA)~f+h+p+i+Condition(logR + logB),
+#               data = SUB.VARS)
+# rdamod2 <- rda(as.data.frame(SUB.DATA)~f+h+p+i+logR+logB+Condition(garden),
+#                data = SUB.VARS)
+# rdamod3 <- rda(as.data.frame(SUB.DATA)~f+h+p+i+logR+logB,
+#                data = SUB.VARS)
+# anova(rdamod0, by='terms')
+# anova(rdamod1, by='terms')
+# anova(rdamod2, by='terms')
+# anova(rdamod3, by='terms')
 
[…]

### 2023-08-29 code/cx_sla_pgls.R

@@ -1,0 +1,137 @@
+
+subsetVars <- c("c", "i")
+source('code/get_data_ready_for_the_sla_analyses.R')
+source('code/model_definitions_pgls.R')
+i.mod <- q.herb.q.water
+i.dD <- deltaData
+
+subsetVars <- c("c", "f")
+source('code/get_data_ready_for_the_sla_analyses.R')
+source('code/model_definitions_pgls.R')
+f.mod <- just.water.linear.noint
+f.dD <- deltaData
+
+subsetVars <- c("c", "h")
+source('code/get_data_ready_for_the_sla_analyses.R')
+source('code/model_definitions_pgls.R')
+h.mod <- just.water.linear
+h.dD <- deltaData
+
+subsetVars <- c("c", "p")
+source('code/get_data_ready_for_the_sla_analyses.R')
+source('code/model_definitions_pgls.R')
+p.mod <- full.model
+p.dD <- deltaData
+
+source("code/predictAndVisualize_function.R")
+
+# Predator --------------------------
+pslap <- predictAndVisualize(p.mod, 
+                             p.dD,
+                             "lsla.d",
+                             n = 100,
+                             titlestring = "Predator - Specific Leaf Area",
+                             ylabstring = "Log-response ratio of total biomass.",
+                             xlabstring = "Log(SLA)")
+
+psitep <- predictAndVisualize(p.mod, 
+                              p.dD,
+                              "fsite",
+                              n = 100,
+                              titlestring = "Predator - Site",
+                              ylabstring = "Log-response ratio of total biomass.",
+                              xlabstring = "Sites")
+
+plabup <- predictAndVisualize(p.mod, 
+                              p.dD,
+                              "labu",
+                              n = 100,
+                              titlestring = "Predator - Abundance",
+                              ylabstring = "Log-response ratio of total biomass.",
+                              xlabstring = "Log(stem number)")
+prichp <- predictAndVisualize(p.mod, 
+                              p.dD,
+                              "richness",
+                              n = 100,
+                              titlestring = "Predator - Richness",
+                              ylabstring = "Log-response ratio of total biomass.",
+                              xlabstring = "Number od species")
+plrrslap <- predictAndVisualize(p.mod, 
+                                p.dD,
+                                "lrrsla",
+                                n = 100,
+                                titlestring = "Predator - LRR SLA",
+                                ylabstring = "Log-response ratio of total biomass.",
+                                xlabstring = "LRR SLA")
+
+
[…]

### 2023-08-29 code/cx_trait_ord.R

@@ -1,0 +1,128 @@
+# Calculate position of species and site along a given treatment vector.
+
+traitPredsCommComp <- function (data = data.woody.brut,
+                                ordination = ord.list,
+                                phylogeny = myTree.rand.poly,
+                                treatment = "f", st = "wanang") {
+  
+  surfData <- data %>%
+    filter(site == st) %>%
+    group_by(unified_names) %>%
+    summarise(lsla.d = log(mean(sla.d, na.rm=T)),
+              water.cont = mean(water.cont, na.rm = T),
+              herbivory = mean(perc.herb, na.rm = T))
+  
+  # Add row names to sla data
+  surfDataDf <- as.data.frame(surfData)
+  rownames(surfDataDf) <- surfDataDf$unified_names
+  
+  # Get the treatment vector.
+  trtVcts <- ordination[[substr(st, 1, 1)]]$CCA$biplot[treatment, c(1,2)]
+  
+  # Get species vector.
+  all.specs <- ordination[[substr(st, 1, 1)]]$CCA$v[, c(1,2)]
+  usednms <- surfDataDf$unified_names[surfDataDf$unified_names %in% rownames(all.specs)]
+  specVcts <- all.specs[usednms,]
+  
+  # Projection of species onto treatment vector.
+  VdotU <- ((specVcts %*% trtVcts))
+  uNormSq <- (trtVcts %*% trtVcts)
+  projUV <- (VdotU/as.numeric(uNormSq)) %*% t(as.matrix(trtVcts))
+  
+  # Get norm for each projection vector.
+  positions <- apply(projUV, 1, function(x) sqrt(x %*% x))
+  
+  # Add positions to the data frame.
+  modDat <- data.frame(value = c(surfDataDf[usednms, ]$lsla.d,
+                                 surfDataDf[usednms, ]$water.cont,
+                                 surfDataDf[usednms, ]$herbivory),
+                       label = rep(c("SLA", "Water Content", "Herbivory"), 
+                                   each = length(positions)),
+                       pos = positions[usednms])
+  
+  testDat <- data.frame(sla = surfDataDf[usednms, ]$lsla.d,
+                        water = surfDataDf[usednms, ]$water.cont,
+                        herb = surfDataDf[usednms, ]$herbivory,
+                        pos = positions[usednms],
+                        unified_names = usednms)
+  
+  
+  
[…]

### 2023-08-29 code/lrr_and_community_chars_correlations.R

@@ -1,9 +1,10 @@
-source('code/log_ratio_extraction.R')
+# source('code/log_ratio_extraction.R')
 
-# RICHNESS.LR
-# BIOMASS.LR
-# DIVERSITY.LR
-# DENSITY.LR
+# It needst to produce:
+# final.biolrr.nl.mod, final.riclrr.nl.mod,final.divlrr.nl.mod, final.denlrr.nl.mod
+# BioPredPlot, RicPredPlot, DivPredPlot, DenPredPlot,
+# BioPredPlotR, RicPredPlotR, DivPredPlotR, DenPredPlotR
+# r2data
 
 RichLRR = RICHNESS.LR$lratio
 ConRicVal = RICHNESS.LR$cval
@@ -20,14 +21,6 @@ LRR_DEP_TEST_DATA <- BIOMASS.LR %>%
          ConBioVal = 'cval',
          BioLRR = 'lratio')
 
-# LRR_DEP_TEST_DATA %>%
-#   dplyr::mutate(RichLRR = RICHNESS.LR$lratio,
-#                 ConRicVal = RICHNESS.LR$cval,
-#                 DivLRR = DIVERSITY.LR$lratio,
-#                 ConDivVal = DIVERSITY.LR$cval,
-#                 DenLRR = DENSITY.LR$lratio,
-#                 ConDenVal = DENSITY.LR$cval)
-
 LRR_DEP_TEST_DATA$RichLRR = RichLRR
 LRR_DEP_TEST_DATA$ConRicVal = ConRicVal
 LRR_DEP_TEST_DATA$DivLRR = DivLRR
@@ -38,548 +31,95 @@ LRR_DEP_TEST_DATA = LRR_DEP_TEST_DATA %>%
   mutate(SiteTreatment = paste(Site, Treatment, sep = " "))
 
 library(ggplot2)
+library(stringr)
+library(sjPlot)
+library(MASS)
+library(mgcv)
+library(tidyr)
+library(caret)
+library(MASS)
 
 responses <- c("BioLRR","RichLRR","DivLRR","DenLRR")
 predictors <- c("ConBioVal","ConRicVal","ConDivVal","ConDenVal")
 
-# responses <- c("BioLRR")
-# predictors <- c("ConBioVal","ConRicVal")
+# Calculate different bottom-up indicators: sum of all species within a garden
+# And average biomass across all plots.
+new_bu_vars <- as.data.frame(t(COMMDATA))
[…]

### 2023-08-29 code/predictAndVisualize_function.R

@@ -61,7 +61,7 @@ predictAndVisualize <- function (model,
                              levels = c("wanang", "numba", "yawan"))
     
     # Predict values for the newly created data.
-    predvals <- predictSE.gls(model, newdata = preddata)
+    predvals <- AICcmodavg::predictSE.gls(model, newdata = preddata)
     preddata$predvals <- predvals$fit
     preddata$predse <- predvals$se.fit
     
@@ -83,7 +83,7 @@ predictAndVisualize <- function (model,
     
   } else {
     preddata <- deltaData
-    predvals <- predictSE.gls(model, newdata = deltaData)
+    predvals <- AICcmodavg::predictSE.gls(model, newdata = deltaData)
     preddata$predvals <- predvals$fit
     preddata$predse <- predvals$se.fit
     

### 2024-03-26 ms1_data_analysis.Rmd
Breaking code into components and integrating in the markdown file.
Latest version of the manuscript, code and cleanup.

@@
 
-First we analyze only traits ignoring location. Later we can mak an attempt to see wether importance of
+First we analyze only traits ignoring location. Later we can make an attempt to see wether importance of
 
@@
 ## SLA
+
+### Are there differences in trait values between elevations?
+
+```{r}
+
+logit <- function(p){
+  return(log(p/(1-p)))
+}
+
+trait_data <- data.woody.brut %>%
+         mutate(fsite = factor(site, levels=c("wanang","numba","yawan")),
+                perc_wc = water.cont/wet_leafg) %>%
+         filter(treatment == "c",
+                perc_wc > 0,
+                unified_names != "piper_aduncum")
+
+sla_plot <- ggplot(trait_data, aes(y=log(sla.d.m2kg1), x = fsite)) +
+  geom_jitter(alpha = 0.3,pch=19) +
+  theme_classic() + ggtitle("SLA")
+wc_plot <- ggplot(trait_data, aes(y=logit(perc_wc), x = fsite)) +
+  geom_jitter(alpha = 0.3,pch=19) +
+  theme_classic()+ ggtitle("Logit[proportional water content]")
+
+jpeg('figures/sla_wc_elevation.jpg')
+ggpubr::ggarrange(sla_plot, wc_plot)
+dev.off()
+```
+
+```{r}
+# test with pgls model
+
+trait_data_tl <- trait_data %>%
+  filter(unified_names %in% myTree.rand.poly$tip.label)
+
+trait_elevation_model <- gls(
+    log(sla.d.m2kg1) ~ fsite,
+    data = trait_data_tl,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+trait_elevation_model2 <- gls(
+    logit(perc_wc) ~ fsite,
+    data = trait_data_tl,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# summary(trait_elevation_model)
+# plot(trait_elevation_model) # better fit for log
+
+posthoc_test <- predictmeans::predictmeans(trait_elevation_model, "fsite", adj='tukey')$mean_table$Table
+
+posthoc_test2 <- predictmeans::predictmeans(trait_elevation_model2, "fsite", adj='tukey')$mean_table$Table
+
+sla_ph <- ggplot(as.data.frame(posthoc_test), aes(x = fsite, y = Mean)) +
+  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.05) +
+  geom_point()
+
+wc_ph <- ggplot(as.data.frame(posthoc_test2), aes(x = fsite, y = Mean)) +
+  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.05) +
+  geom_point()
+
+ggpubr::ggarrange(sla_ph, wc_ph, labels = c("SLA", "WC"))
+
+```
 
@@
 
+```{r}
+source('code/model_definitions_pgls_no_herb.R')
+
+moddata_w <- moddata[moddata$site == "wanang", ]
+moddata_n <- moddata[moddata$site == "numba", ]
+moddata_y <- moddata[moddata$site == "yawan", ]
+
+null_w <- gls(
+    lrrtb ~ 1,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_w <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_n <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_y <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# Compare MSE only models with quadratic and 
+wanang_mse <- mean(just_traits_w$residuals^2)
+numba_mse <- mean(just_traits_n$residuals^2)
+yawan_mse <- mean(just_traits_y$residuals^2)
+
+# MSE for models with traits and all other descriptors
+full_model_w <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+full_model_n <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+full_model_y <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# MSE for models without the traits but all other descriptors
+desc_model_w <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+desc_model_n <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+desc_model_y <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# Compare MSE only models with quadratic and 
+wanang_full_mse <- mean(full_model_w$residuals^2)
+numba_full_mse <- mean(full_model_n$residuals^2)
+yawan_full_mse <- mean(full_model_y$residuals^2)
+
+# Compare MSE only models with desc
+wanang_desc_mse <- mean(desc_model_w$residuals^2)
+numba_desc_mse <- mean(desc_model_n$residuals^2)
+yawan_desc_mse <- mean(desc_model_y$residuals^2)
+
+mse_data_i = data.frame(Site = factor(c("Wanang", "Numba", "Yawan",
+                    "Wanang", "Numba", "Yawan",
+                    "Wanang", "Numba", "Yawan"),levels = c("Wanang", "Numba", "Yawan")),
+           MSE = c(wanang_mse, numba_mse,yawan_mse,                          
+                   wanang_full_mse, numba_full_mse, yawan_full_mse,
+                   wanang_desc_mse, numba_desc_mse, yawan_desc_mse),
+           Model = c("Traits","Traits","Traits",
+                     "Traits+Descriptors","Traits+Descriptors","Traits+Descriptors",
+                     "Descriptors","Descriptors","Descriptors"),
+           Treatment = "Insecticide")
+
+ggplot(mse_data_i, aes(x = Site, y = MSE, fill = Model))+
+  geom_bar(stat = 'identity', position = 'dodge')+
+  theme_bw() + ggtitle("Insecticide")
+
+```
+
+```{r}
+# 
+# 
+# just_traits_w_l <- gls(
+#     lrrtb ~ sla.d.m2kg1 + water.cont,
+#     data = moddata_w,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# just_sla_w <- gls(
+#     lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2),
+#     data = moddata_w,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# just_sla_w_l <- gls(
+#     lrrtb ~ sla.d.m2kg1,
+#     data = moddata_w,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# just_wc_w <- gls(
+#     lrrtb ~ water.cont + I(water.cont^2),
+#     data = moddata_w,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# just_wc_w_l <- gls(
+#     lrrtb ~ water.cont,
+#     data = moddata_w,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# 
+# # Traits are better
+# anova(null_w, just_traits_w)
+# 
+# # Quadratic + two traits is better
+# anova(just_traits_w, just_traits_w_l)
+# anova(just_traits_w, just_sla_w)
+# anova(just_traits_w, just_wc_w)
+# anova(just_traits_w, just_sla_w_l)
+# anova(just_traits_w, just_wc_w_l)
+# 
+# RsquareAdj(just_traits_w)
+# 
+# sjPlot::tab_model(just_traits_w,
+#                   just_traits_w_l)
+# 
+#
+
+```
+
+```{r}
+
+# All sites
+# 
+# null_traits <- gls(
+#     lrrtb ~ 1,
+#     data = moddata,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   ) 
+# 
+# just_traits.q.no_int <- gls(
+#     lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+#     data = moddata,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# 
+# just_traits.l.no_int <- gls(
+#     lrrtb ~ sla.d.m2kg1 + water.cont,
+#     data = moddata,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# just_traits.q <- gls(
+#     lrrtb ~ sla.d.m2kg1 * site + 
+#       I(sla.d.m2kg1^2) * site + 
+#       water.cont * site + 
+#       I(water.cont^2) * site,
+#     data = moddata,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# 
+# just_traits.l <- gls(
+#     lrrtb ~ sla.d.m2kg1 * site + water.cont *  site,
+#     data = moddata,
+#     method = "ML",
+#     correlation = corMartins(2.718282,
+#                              phy = myTree.rand.poly,
+#                              form = ~unified_names,
+#                              fixed = T),
+#     na.action = na.omit
+#   )
+# 
+# just_sla.q
+# 
+# sjPlot::tab_model(just_traits)
+
+```
+
 ```{r, results='asis'}
-
-source('code/model_definitions_pgls.R')
-
 # Are traits improving the null model
@@
 # 1. Full model (quadratic, and site interactions) vs no traits
-# anova(no.traits, full.model)            # full model is better
+anova(no.traits, full.model)            # full model is better
 
@@
 # 2.1 Remove SLA
-# anova(full.model, q.herb.q.water) # no-difference, keep model without the SLA a simpler model?
-# 2.2 Remove herbivory
-# anova(full.model, q.sla.q.water.noint) # Model without herbivory is significantly worse
+anova(full.model, just.water.quadratic) # full_model
+
 # 2.3 Remove water content?
-# anova(full.model, q.sla.q.herb) # Model without water is significantly worse
-
-# Herbivory and water are the most important!!!
-
-# 3. Reduce model even further - 1 trait models.
-# anova(q.herb.q.water, just.sla.quadratic)   # Two trait model better
-# anova(q.herb.q.water, just.herb.quadratic)  # Two trait model better
-# anova(q.herb.q.water, just.water.quadratic) # Two trait model better
+anova(full.model, just.sla.quadratic) # No significant difference
 
 # 4. Quadratic vs linear
-# anova(q.herb.q.water, l.herb.l.water) # Quadratic model is still better
+anova(just.sla.quadratic, just.sla.linear) # Quadratic model is still better
 
 # 5. Interaction reduction
-# anova(q.herb.q.water, q.herb.q.water.noint) # Model with interaction is better
-# anova(q.herb.q.water, just.water.quadratic.noint) # Model with interaction is better
-# anova(q.herb.q.water, just.water.linear.noint)
+anova(just.sla.quadratic, just.sla.quadratic.noint) # Model with interaction is better
 
@@
 # tab_model(just.water.quadratic.noint)
-i.mod <- q.herb.q.water
+i.mod <- just.sla.quadratic
 i.dD <- deltaData
@@
 
+```{r}
+
+source('code/model_definitions_pgls_no_herb.R')
+
+moddata_w <- moddata[moddata$site == "wanang", ]
+moddata_n <- moddata[moddata$site == "numba", ]
+moddata_y <- moddata[moddata$site == "yawan", ]
+
+null_w <- gls(
+    lrrtb ~ 1,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_w <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_n <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_y <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# Compare MSE only models with quadratic and 
+wanang_mse <- mean(just_traits_w$residuals^2)
+numba_mse <- mean(just_traits_n$residuals^2)
+yawan_mse <- mean(just_traits_y$residuals^2)
+
+# MSE for models with traits and all other descriptors
+full_model_w <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+full_model_n <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+full_model_y <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# MSE for models without the traits but all other descriptors
+desc_model_w <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+desc_model_n <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+desc_model_y <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# Compare MSE only models with quadratic and 
+wanang_full_mse <- mean(full_model_w$residuals^2)
+numba_full_mse <- mean(full_model_n$residuals^2)
+yawan_full_mse <- mean(full_model_y$residuals^2)
+
+# Compare MSE only models with desc
+wanang_desc_mse <- mean(desc_model_w$residuals^2)
+numba_desc_mse <- mean(desc_model_n$residuals^2)
+yawan_desc_mse <- mean(desc_model_y$residuals^2)
+
+mse_data_f = data.frame(Site = factor(c("Wanang", "Numba", "Yawan",
+                    "Wanang", "Numba", "Yawan",
+                    "Wanang", "Numba", "Yawan"),levels = c("Wanang", "Numba", "Yawan")),
+           MSE = c(wanang_mse, numba_mse,yawan_mse,                          
+                   wanang_full_mse, numba_full_mse, yawan_full_mse,
+                   wanang_desc_mse, numba_desc_mse, yawan_desc_mse),
+           Model = c("Traits","Traits","Traits",
+                     "Traits+Descriptors","Traits+Descriptors","Traits+Descriptors",
+                     "Descriptors","Descriptors","Descriptors"),
+           Treatment = "Fungi")
+
+ggplot(mse_data_f, aes(x = Site, y = MSE, fill = Model))+
+  geom_bar(stat = 'identity', position = 'dodge')+
+  theme_bw() + ggtitle("Fungi")
+
+```
+
 ```{r, results='asis'}
 
-source('code/model_definitions_pgls.R')
+source('code/model_definitions_pgls_no_herb.R')
 
@@
 # 1. Full model (quadratic, and site interactions) vs no traits
-# anova(no.traits, full.model)            # full model is better
-
-# 2. Removal of individual traits
-
-# 2.1 Remove SLA
-# anova(full.model, q.herb.q.water) # without sla is slightly better (AIC)
-
-# 2.2 Remove herbivory
-# anova(full.model, q.sla.q.water) # Model without herbivory is not better
-
-# Both q.herb.q.water and q.sla.q.water are simpler, but which one is better?
-# anova(full.model, q.sla.q.herb) # Model without water is significantly worse
-
-# 2.3 Remove water content?
-# anova(full.model, q.sla.q.herb) # Model with herbivory is much better!
-# Water is important!!!
-
-# anova(q.herb.q.water, q.sla.q.water) # q.sla.q.water is better based on AIC
-
-# 3. Reduce model even further - 1 trait models.
-# anova(q.sla.q.water, just.sla.quadratic)  # q.sla.q.water is still better
-# anova(q.sla.q.water, just.herb.quadratic) # q.sla.q.water is still better
-# anova(q.sla.q.water, just.water.quadratic) # model with just water is as good as q.sla.q.water
-
-# model with just water is not different form the full model, and it has lower AIC
-
-# Seems like they all stay!
-
-# 4. Quadratic vs linear
-# anova(just.water.quadratic, just.water.linear) # Linear is simpler model still better
-
-# 5. Interaction reduction
-# anova(just.water.linear, just.water.linear.noint) # Interaction does not bring anything to the model
+anova(no.traits, full.model)            
+# model with traits does not have a significantly higher likelihood.
 
@@
 # tab_model(just.water.linear.noint)
-f.mod <- just.water.linear.noint
+f.mod <- no.traits
 f.dD <- deltaData
@@
 
+```{r}
+
+source('code/model_definitions_pgls_no_herb.R')
+
+moddata_w <- moddata[moddata$site == "wanang", ]
+moddata_n <- moddata[moddata$site == "numba", ]
+moddata_y <- moddata[moddata$site == "yawan", ]
+
+null_w <- gls(
+    lrrtb ~ 1,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_w <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_n <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_y <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# Compare MSE only models with quadratic and 
+wanang_mse <- mean(just_traits_w$residuals^2)
+numba_mse <- mean(just_traits_n$residuals^2)
+yawan_mse <- mean(just_traits_y$residuals^2)
+
+# MSE for models with traits and all other descriptors
+full_model_w <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+full_model_n <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+full_model_y <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# MSE for models without the traits but all other descriptors
+desc_model_w <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+desc_model_n <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+desc_model_y <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# Compare MSE only models with quadratic and 
+wanang_full_mse <- mean(full_model_w$residuals^2)
+numba_full_mse <- mean(full_model_n$residuals^2)
+yawan_full_mse <- mean(full_model_y$residuals^2)
+
+# Compare MSE only models with desc
+wanang_desc_mse <- mean(desc_model_w$residuals^2)
+numba_desc_mse <- mean(desc_model_n$residuals^2)
+yawan_desc_mse <- mean(desc_model_y$residuals^2)
+
+mse_data_h = data.frame(Site = factor(c("Wanang", "Numba", "Yawan",
+                    "Wanang", "Numba", "Yawan",
+                    "Wanang", "Numba", "Yawan"),levels = c("Wanang", "Numba", "Yawan")),
+           MSE = c(wanang_mse, numba_mse,yawan_mse,                          
+                   wanang_full_mse, numba_full_mse, yawan_full_mse,
+                   wanang_desc_mse, numba_desc_mse, yawan_desc_mse),
+           Model = c("Traits","Traits","Traits",
+                     "Traits+Descriptors","Traits+Descriptors","Traits+Descriptors",
+                     "Descriptors","Descriptors","Descriptors"),
+           Treatment = "Herbivore")
+
+ggplot(mse_data_h, aes(x = Site, y = MSE, fill = Model))+
+  geom_bar(stat = 'identity', position = 'dodge')+
+  theme_bw() + ggtitle("Herbivore")
+
+```
+
 ```{r, results='asis'}
 
-source('code/model_definitions_pgls.R')
+source('code/model_definitions_pgls_no_herb.R')
 
@@
 # tab_model(just.water.linear)
-h.mod <- just.water.linear
+h.mod <- no.traits
 h.dD <- deltaData
@@
 
+```{r}
+
+source('code/model_definitions_pgls_no_herb.R')
+
+moddata_w <- moddata[moddata$site == "wanang", ]
+moddata_n <- moddata[moddata$site == "numba", ]
+moddata_y <- moddata[moddata$site == "yawan", ]
+
+null_w <- gls(
+    lrrtb ~ 1,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_w <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_n <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+just_traits_y <- gls(
+    lrrtb ~ sla.d.m2kg1 + I(sla.d.m2kg1^2) + water.cont + I(water.cont^2),
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# Compare MSE only models with quadratic and 
+wanang_mse <- mean(just_traits_w$residuals^2)
+numba_mse <- mean(just_traits_n$residuals^2)
+yawan_mse <- mean(just_traits_y$residuals^2)
+
+# MSE for models with traits and all other descriptors
+full_model_w <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+full_model_n <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+full_model_y <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) +
+      richness + labu + lrrsla + lrrwtrc,
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# MSE for models without the traits but all other descriptors
+desc_model_w <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_w,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+desc_model_n <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_n,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+desc_model_y <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc,
+    data = moddata_y,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+
+# Compare MSE only models with quadratic and 
+wanang_full_mse <- mean(full_model_w$residuals^2)
+numba_full_mse <- mean(full_model_n$residuals^2)
+yawan_full_mse <- mean(full_model_y$residuals^2)
+
+# Compare MSE only models with desc
+wanang_desc_mse <- mean(desc_model_w$residuals^2)
+numba_desc_mse <- mean(desc_model_n$residuals^2)
+yawan_desc_mse <- mean(desc_model_y$residuals^2)
+
+mse_data_p = data.frame(Site = factor(c("Wanang", "Numba", "Yawan",
+                    "Wanang", "Numba", "Yawan",
+                    "Wanang", "Numba", "Yawan"),levels = c("Wanang", "Numba", "Yawan")),
+           MSE = c(wanang_mse, numba_mse,yawan_mse,                          
+                   wanang_full_mse, numba_full_mse, yawan_full_mse,
+                   wanang_desc_mse, numba_desc_mse, yawan_desc_mse),
+           Model = c("Traits","Traits","Traits",
+                     "Traits+Descriptors","Traits+Descriptors","Traits+Descriptors",
+                     "Descriptors","Descriptors","Descriptors"),
+           Treatment = "Predator")
+
+ggplot(mse_data_p, aes(x = Site, y = MSE, fill = Model))+
+  geom_bar(stat = 'identity', position = 'dodge')+
+  theme_bw() + ggtitle("Predator")
+
+```
+
+```{r}
+
+mse_data_full <- rbind(mse_data_p, mse_data_f, mse_data_i, mse_data_h)
+
+ggplot(mse_data_full, aes(x = Site, y = 1/MSE, fill = Model))+
+  geom_bar(stat = 'identity', position = 'dodge')+
+  theme_bw() + facet_wrap(~Treatment)
+
+# ggplot(mse_data_full, aes(x = Site, y = MSE, fill = Treatment))+
+#   geom_bar(stat = 'identity', position = 'dodge')+
+#   theme_bw() + facet_wrap(~Model, scales='free')
+
+```
+
 ```{r, results='asis'}
 
-source('code/model_definitions_pgls.R')
+source('code/model_definitions_pgls_no_herb.R')
 
@@
 # 2.1 Remove SLA
-# anova(full.model, q.herb.q.water) # Full model is significanlty better
+anova(full.model, just.water.quadratic) # Full model is significanlty better
 
 # 2.2 Remove herbivory
-# anova(full.model, q.sla.q.water.noint) # Full model still better
-
-# 2.3 Remove water content?
-# anova(full.model, q.sla.q.herb) # Full model still better
-
-# 3. Reduce model even further - 1 trait models.
-# anova(full.model, just.sla.quadratic)
-# anova(full.model, just.herb.quadratic)
-# anova(full.model, just.water.quadratic)
-
-# Seems like they all stay!
+anova(full.model, just.sla.quadratic) # Full model still better
 
 # 4. Quadratic vs linear
-# anova(full.model, all.linear) # Full model still better
+anova(full.model, all.linear) # Full model still better
 
 # 5. Interaction reduction
-# anova(full.model, full.model.noint) # Full model still better
+anova(full.model, full.model.noint) # Full model still better
 ```
@@
                           "Water content - quadratic",
-                          "Herbivory [%] - linear",
-                          "Herbivory [%] - quadratic",
                           "Species richness",
                           "Abundance",
-                          "LRR SLA",
+                          "Plasticity in SLA",
+                          "Plasticity in WTR.C",
                           "Log[SLA]-linear : Numba",
@@
                           "Water content-quadratic : Numba",
-                          "Water content-quadratic : Yawan",
-                          "Herbivory-linear : Numba",
-                          "Herbivory-linear : Yawan",
-                          "Herbivory-quadratic : Numba",
-                          "Herbivory-quadratic : Yawan"),
-          dv.labels = c("Predator", "Insecticide", "Extra herbivory", "Fungicide"),
-          string.pred = "Coeffcient",
+                          "Water content-quadratic : Yawan"),
           string.ci = "Conf. Int (95%)",
           string.p = "P-Value")
+
+# tab_model(p.mod, i.mod, h.mod, f.mod,
+#           pred.labels = c("Intercept", 
+#                           "Log[SLA]-linear", 
+#                           "Numba", 
+#                           "Yawan",
+#                           "Log[SLA]-quadratic", 
+#                           "Water content - linear", 
+#                           "Water content - quadratic",
+#                           "Herbivory [%] - linear",
+#                           "Herbivory [%] - quadratic",
+#                           "Species richness",
+#                           "Abundance",
+#                           "LRR SLA",
+#                           "Log[SLA]-linear : Numba",
+#                           "Log[SLA]-linear : Yawan",
+#                           "Log[SLA]-quadratic : Numba",
+#                           "Log[SLA]-quadratic : Yawan",
+#                           "Water content-linear : Numba",
+#                           "Water content-linear : Yawan",
+#                           "Water content-quadratic : Numba",
+#                           "Water content-quadratic : Yawan",
+#                           "Herbivory-linear : Numba",
+#                           "Herbivory-linear : Yawan",
+#                           "Herbivory-quadratic : Numba",
+#                           "Herbivory-quadratic : Yawan"),
+#           dv.labels = c("Predator", "Insecticide", "Extra herbivory", "Fungicide"),
+#           string.pred = "Coeffcient",
+#           string.ci = "Conf. Int (95%)",
+#           string.p = "P-Value")
 
@@
                     n = 100,
-                    titlestring = "Predator - Specific Leaf Area",
+                    titlestring = "Predators",
                     ylabstring = "Log-response ratio of total biomass.",
@@
 
-# Focal traits
-
-# Co-variates
-psitep <- predictAndVisualize(p.mod, 
+pwaterp <- predictAndVisualize(p.mod, 
                     p.dD,
-                    "fsite",
-                    n = 100,
-                    titlestring = "Predator - Site",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Sites")
-
-plabup <- predictAndVisualize(p.mod, 
-                    p.dD,
-                    "labu",
-                    n = 100,
-                    titlestring = "Predator - Abundance",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Log(stem number)")
-prichp <- predictAndVisualize(p.mod, 
-                    p.dD,
-                    "richness",
-                    n = 100,
-                    titlestring = "Predator - Richness",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Number od species")
-plrrslap <- predictAndVisualize(p.mod, 
-                    p.dD,
-                    "lrrsla",
-                    n = 100,
-                    titlestring = "Predator - LRR SLA",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "LRR SLA")
-
-# Insecticide-----------------------
-ilabup <- predictAndVisualize(i.mod, 
-                    i.dD,
-                    "labu",
-                    n = 100,
-                    titlestring = "Insecticide - Abundance",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Log(stem number)")
-
-# Extra herbivory-------------------
-hsitep <- predictAndVisualize(h.mod, 
-                    h.dD,
-                    "fsite",
-                    n = 100,
-                    titlestring = "Herbivory - Site",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Sites")
-
-hwaterp <- predictAndVisualize(h.mod, 
-                    h.dD,
                     "water.cont",
                     n = 100,
-                    titlestring = "Herbivory - Water Content",
+                    titlestring = "Predators",
                     ylabstring = "Log-response ratio of total biomass.",
@@
 
-hlabup <- predictAndVisualize(h.mod, 
-                    h.dD,
-                    "labu",
+islap <- predictAndVisualize(i.mod, 
+                    i.dD,
+                    "lsla.d",
                     n = 100,
-                    titlestring = "Herbivory - Abundance",
+                    titlestring = "Insecticide",
                     ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Log(stem number)")
-
-hlrrslap <- predictAndVisualize(h.mod, 
-                    h.dD,
-                    "lrrsla",
-                    n = 100,
-                    titlestring = "Herbivory - LRR SLA",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "LRR SLA")
-
-# Fungicide
-
-fwaterp <- predictAndVisualize(f.mod, 
-                    f.dD,
-                    "water.cont",
-                    n = 100,
-                    titlestring = "Fungicide - Water Content",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Water content")
-
-fsitep <- predictAndVisualize(f.mod, 
-                    f.dD,
-                    "fsite",
-                    n = 100,
-                    titlestring = "Fungicide - Site",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Site")
-
-frichp <- predictAndVisualize(f.mod, 
-                    f.dD,
-                    "richness",
-                    n = 100,
-                    titlestring = "Fungicide - Richness",
-                    ylabstring = "Log-response ratio of total biomass.",
-                    xlabstring = "Number of species per plot")
-
-# Plot
-ggpubr::ggarrange(pslap +
-                    scale_y_continuous(limits = c(-5, 10)),
-                  fwaterp + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  hwaterp + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  nrow = 1, legend = F)
-ggpubr::ggarrange(psitep + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  fsitep + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  hsitep + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  nrow = 1,legend = F)
-ggpubr::ggarrange(prichp + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  frichp + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  nrow = 1, legend = F)
-ggpubr::ggarrange(plabup + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  ilabup + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  hlabup + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  nrow = 1, legend = F)
-ggpubr::ggarrange(plrrslap +
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  hlrrslap + 
-                    scale_y_continuous(limits = c(-5, 10)), 
-                  nrow = 1, common.legend = T)
+                    xlabstring = "Log(SLA)")
+
+# tiff(filename="figures/traits_pgls.tiff",
+#      height=2600,
+#      width=3*2600,
+#      units='px',
+#      res=800,compression='lzw')
+# ggpubr::ggarrange(pslap +
+#                     scale_y_continuous(limits = c(-5, 10)),
+#                   pwaterp +
+#                     scale_y_continuous(limits = c(-5, 10)),
+#                   islap +
+#                     scale_y_continuous(limits = c(-5, 10)),
+#                   nrow = 1, legend = F)
+# dev.off()
+
+# Focal traits
+
+# # Co-variates
+# psitep <- predictAndVisualize(p.mod, 
+#                     p.dD,
+#                     "fsite",
+#                     n = 100,
+#                     titlestring = "Predator - Site",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Sites")
+# 
+# plabup <- predictAndVisualize(p.mod, 
+#                     p.dD,
+#                     "labu",
+#                     n = 100,
+#                     titlestring = "Predator - Abundance",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Log(stem number)")
+# prichp <- predictAndVisualize(p.mod, 
+#                     p.dD,
+#                     "richness",
+#                     n = 100,
+#                     titlestring = "Predator - Richness",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Number od species")
+# plrrslap <- predictAndVisualize(p.mod, 
+#                     p.dD,
+#                     "lrrsla",
+#                     n = 100,
+#                     titlestring = "Predator - LRR SLA",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "LRR SLA")
+# 
+# 
+# # Insecticide-----------------------
+# ilabup <- predictAndVisualize(i.mod, 
+#                     i.dD,
+#                     "labu",
+#                     n = 100,
+#                     titlestring = "Insecticide - Abundance",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Log(stem number)")
+# 
+# 
+# 
+# # Extra herbivory-------------------
+# hsitep <- predictAndVisualize(h.mod, 
+#                     h.dD,
+#                     "fsite",
+#                     n = 100,
+#                     titlestring = "Herbivory - Site",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Sites")
+# 
+# hwaterp <- predictAndVisualize(h.mod, 
+#                     h.dD,
+#                     "water.cont",
+#                     n = 100,
+#                     titlestring = "Herbivory - Water Content",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Water content")
+# 
+# 
+# hlabup <- predictAndVisualize(h.mod, 
+#                     h.dD,
+#                     "labu",
+#                     n = 100,
+#                     titlestring = "Herbivory - Abundance",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Log(stem number)")
+# 
+# hlrrslap <- predictAndVisualize(h.mod, 
+#                     h.dD,
+#                     "lrrsla",
+#                     n = 100,
+#                     titlestring = "Herbivory - LRR SLA",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "LRR SLA")
+# 
+# # Fungicide
+# 
+# fwaterp <- predictAndVisualize(f.mod, 
+#                     f.dD,
+#                     "water.cont",
+#                     n = 100,
+#                     titlestring = "Fungicide - Water Content",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Water content")
+# 
+# fsitep <- predictAndVisualize(f.mod, 
+#                     f.dD,
+#                     "fsite",
+#                     n = 100,
+#                     titlestring = "Fungicide - Site",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Site")
+# 
+# frichp <- predictAndVisualize(f.mod, 
+#                     f.dD,
+#                     "richness",
+#                     n = 100,
+#                     titlestring = "Fungicide - Richness",
+#                     ylabstring = "Log-response ratio of total biomass.",
+#                     xlabstring = "Number of species per plot")
+# 
+# # Plot
+# ggpubr::ggarrange(pslap +
+#                     scale_y_continuous(limits = c(-5, 10)),
+#                   fwaterp + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   hwaterp + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   nrow = 1, legend = F)
+# ggpubr::ggarrange(psitep + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   fsitep + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   hsitep + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   nrow = 1,legend = F)
+# ggpubr::ggarrange(prichp + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   frichp + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   nrow = 1, legend = F)
+# ggpubr::ggarrange(plabup + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   ilabup + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   hlabup + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   nrow = 1, legend = F)
+# ggpubr::ggarrange(plrrslap +
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   hlrrslap + 
+#                     scale_y_continuous(limits = c(-5, 10)), 
+#                   nrow = 1, common.legend = T)
 

### 2024-03-26 code/cx_beta_part.R

@@ -183,21 +183,34 @@ pairwise.comp.res.rf <- pairwise.comp.res %>%
                                "Fungi",
                                ifelse(y_loc_trt == '_h', "Herbivores",
                                       ifelse(y_loc_trt == '_i', "Insecticide",'Predator')))),
-         component = ifelse(component == 'balanced', "Species turnover (balanced component)",
-                            ifelse(component == "gradient", "Dominance shifts (gradient component)",
-                                   "Bray-Curtis dissimilarty")))
+         component = ifelse(component == 'balanced', "Species turnover",
+                            ifelse(component == "gradient", "Dominance shifts",
+                                   "Bray-Curtis Ind.")))
 
+colors <- c("gray30", "gray","gray","red","gray",
+            "gray30", "gray","gray","gray","gray",
+            "gray30", "gray","red","red","gray",
+            "gray30", "gray","gray","red","gray",
+            "gray30", "gray","gray","red","red",
+            "gray30", "gray","gray","gray","gray",
+            "gray30", "gray","gray","gray","gray",
+            "gray30", "gray","red","red","red",
+            "gray30", "gray","red","red","red")
 
 # Alternative boxplot
 beta.part.plot2 <- ggplot(beta.df.rf, aes(x = treat, y = vals))+
   # geom_jitter(width = 0.1, alpha = 0.1, color = "navyblue")+
   stat_summary(fun.y = mean, na.rm = T,
-               geom = "point", size = 3, col = "gray30")+
-  stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
-  geom_text(data = pairwise.comp.res.rf,
-            aes(x = y_loc_trt, y = y_loc + 0.15,
-                label = label), cex = 6,  col = "red")+
+               geom = "point", size = 5, col = colors)+
+  stat_summary(fun.data = "mean_cl_boot", col = colors, lwd = 1.1)+
+  # geom_text(data = pairwise.comp.res.rf,
+  #           aes(x = y_loc_trt, y = y_loc + 0.15,
+  #               label = label), cex = 6,  col = "red")+
   facet_grid(component ~ fsite, scales = 'free',
              labeller = label_wrap_gen(25)) + 
-  theme_bw()+ theme(axis.text.x = element_text(angle=90, hjust =1 ))
+  theme_bw()+ theme(axis.text.x = element_text(angle=90, hjust =1 )) + 
+  xlab("") + ylab("")
+
[…]

### 2024-03-26 code/get_data_ready_for_the_sla_analyses.R

@@ -85,9 +85,11 @@ for (st in unique(lgsData$site)){
     if (subsetVars[2] == "h"){
       cDat$lrrtb <- log(tDat$tot_bio/cDat$tot_bio)
       cDat$lrrsla <- log(tDat$sla.d.m2kg1/cDat$sla.d.m2kg1)
+      cDat$lrrwtrc <- log(tDat$water.cont/cDat$water.cont)
     } else{
       cDat$lrrtb <- log(cDat$tot_bio/tDat$tot_bio)
       cDat$lrrsla <- log(cDat$sla.d.m2kg1/tDat$sla.d.m2kg1)
+      cDat$lrrwtrc <- log(cDat$water.cont/tDat$water.cont)
     }
     
     deltaData <- rbind(deltaData, cDat)

### 2024-03-26 code/log_ratio_extraction.R

@@ -65,3 +65,4 @@ lrr_test_df <- rbind(biotest_table,
                        divtest_table,
                        richtest_table,
                        denstest_table)
+

### 2024-03-26 code/model_definitions_pgls_no_herb.R

@@ -1,0 +1,513 @@
+moddata <- deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label)
+
+modelDefinitions <-  function () {
+  
+  # Runs all of model definitions
+  
+  # NULL MODEL
+  null.model <- gls(
+    lrrtb ~ 1,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  # NO TRAITS
+  no.traits <- gls(
+    lrrtb ~ richness + labu + lrrsla + lrrwtrc + fsite,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  # FULL MODEL
+  full.model <- gls(
+    lrrtb ~ lsla.d*fsite+ I(lsla.d^2)*fsite+water.cont*fsite+I(water.cont^2)*fsite+
+      richness+labu+lrrsla+lrrwtrc+fsite,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  # FULL MODEL
+  full.model.noint <- gls(
+    lrrtb ~ lsla.d + I(lsla.d^2) + water.cont + I(water.cont^2) + 
+      richness+labu+lrrsla+lrrwtrc+fsite,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  all.linear <- gls(
+    lrrtb ~ lsla.d*fsite + water.cont*fsite + 
+      richness + labu + lrrsla + lrrwtrc + fsite,
+    data = moddata,
+    method = "ML",
+    correlation = corMartins(2.718282,
+                             phy = myTree.rand.poly,
+                             form = ~unified_names,
+                             fixed = T),
+    na.action = na.omit
+  )
+  
+  all.linear.noint <- gls(
+    lrrtb ~ lsla.d + water.cont+
[…]

### 2024-03-26 code/raup_crick_analysis_v2.R

@@ -1,12 +1,13 @@
 rm(list = ls())
 
 # Start with Wanang C and I
-source("code/data_processing.R")
+# source("code/data_processing.R")
+source("code/prepare_data.R")
 COMMDATA <- t(COMMDATA)
 
 # This script performs RC analysis on the experimental elevation plots.
 # For each plot comparison within the same treatment I constrained the
-# plant data to species and biomass observed within the SITE (but not garden).
+# plant data to species and biomass observed within the SITE (not garden, not only gardern).
 # This included all treatments to obtain the most reliable species list for
 # random sampling.
 
@@ -19,9 +20,12 @@ getPlantMat <- function (COMMDATA, site) {
   return(COMMDATA[c1, ])
 }
 
+
+siteCode="wanang_g1_c"
+
 ### Assemble function
 assemble <- function (siteCode,
-                      plant_mat, # subset (either within garden/or within)
+                      plant_mat, # sampling community (either within garden/site/treatment)
                       accuracy = 0.001){
   
   # Samples random community with constraints defined by the given site code:
@@ -176,7 +180,7 @@ pb <- txtProgressBar(min = 0,      # Minimum value of the progress bar
                      style = 3,    # Progress bar style (also available style = 1 and style = 2)
                      width = 50,   # Progress bar width. Defaults to getOption("width")
                      char = "=")
-
+row = 1
 for(row in 1:dim(all.pairs)[1]){
   
   site1 <- all.pairs[row, 1]
@@ -279,6 +283,7 @@ for (st in unique(all.pairs_max_accuracy$fsite)){
                      data = subdat)
 
     # print(t.test(RC ~ treatment, data = subdat, paired = TRUE))
+    print(summary(mod))
     
     test.df <- rbind(test.df, data.frame(fsite = st,
                                          comp = cp,
@@ -311,3 +316,22 @@ ggplot(all.pairs_max_accuracy, aes(x = treatment, y = RC))+
   ggpubr::stat_pvalue_manual(data = test.df[test.df$label == "*",], label = "label",
                      y.position = 1.25, 
                      label.size = 6)
+
+
+
+
+ggplot(all.pairs_max_accuracy, aes(x = treatment, y = RC))+
+  # geom_boxplot()+
+  ylim(c(-1.25,1.5))+
+  stat_summary(fun.y = mean, na.rm = T,
[…]
